### 2019-04-10 Manuscript/Manuscript.tex
add initial .tex materials

\documentclass[11pt,letterpaper]{article}

%\usepackage{fontspec}
%\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[normalem]{ulem}
\usepackage[left]{lineno}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{color}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}
\usepackage{threeparttable}
\doublespacing

\raggedright
\textwidth = 6.5 in
\textheight = 8.25 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.5 in
\parskip = 0.1 in
\parindent = 0.2in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Remove brackets from numbering in List of References
%\makeatletter
%\renewcommand{\@biblabel}[1]{\quad#1.}
%\makeatother

% Self defined commands
\newcommand{\degC}{$^{\circ}$C\xspace}
\newcommand{\dC}{$\delta^{13}$C\xspace}
\newcommand{\dO}{$\delta^{18}$O\xspace}
\newcommand{\SrSr}{$^{87}$Sr/$^{86}$Sr\xspace}
\newcommand{\permil}{\textperthousand\xspace}
\newcommand{\dsil}{$d$\xspace}
\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
%

\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{Park et al., in preparation for XXX}
\rhead{\thepage}

\begin{document}

\begin{flushleft}
{\Large \textbf{GEOCLIM Modern}}

Yuem Park\textsuperscript{1},
Nicholas L. Swanson-Hysell\textsuperscript{1}

\bigskip
\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA

\textsuperscript{2} XXX
\bigskip

\end{flushleft}

\noindent\textit{This article is in preparation for XXX}

\linenumbers

\section*{ABSTRACT \label{sec:ABSTRACT}}

XXX

\section*{INTRODUCTION \label{sec:INTRODUCTION}}

XXX

\section*{METHODS}

XXX

\section*{RESULTS}

XXX

\section*{DISCUSSION}

XXX

\section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}

XXX

\clearpage
\newpage

\section*{TABLES}

XXX

\clearpage
\newpage

\section*{FIGURES}

XXX

\clearpage
\newpage
\footnotesize

\singlespacing

\bibliographystyle{gsabull}
\bibliography{References}

\end{document}

### 2019-06-11 Manuscript/Data_Repository/Data_Repository.tex
add template for data repository

--- 2019-04-10 Manuscript/Manuscript.tex
@@
 \usepackage[left]{lineno}
-\usepackage{booktabs}
 \usepackage{changepage}
@@
 \usepackage{setspace}
-\usepackage{array}
+\usepackage{}
 \usepackage{fancyhdr}
@@
 \usepackage{xspace}
-\usepackage[hidelinks]{hyperref}
+\usepackage{threeparttable}
+\usepackage{color,colortbl}
+\usepackage{url}
+%\usepackage[hidelinks]{hyperref}
 \urlstyle{same}
-\usepackage{threeparttable}
 \doublespacing
@@
 \parskip = 0.1 in
-\parindent = 0.2in
+\parindent = 0.1in
 
@@
 \newcommand{\dsil}{$d$\xspace}
-\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{DR\arabic{figure}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{DR\arabic{table}}
+
+\definecolor{Yellow}{rgb}{1,1,0.35}
 %
@@
 \fancyhf{}
-\lhead{Park et al., in preparation for XXX}
+\lhead{Park et al., in review for XXX}
 \rhead{\thepage}
@@
 \begin{flushleft}
-{\Large \textbf{GEOCLIM Modern}}
-
+{\Large \textbf{Data Repository for ``GEOCLIM Modern''}}
+\\
 Yuem Park\textsuperscript{1},
-Nicholas L. Swanson-Hysell\textsuperscript{1}
-
+\\
 \bigskip
 \textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
-
-\textsuperscript{2} XXX
+\\
 \bigskip
@@
 
-\noindent\textit{This article is in preparation for XXX}
-
 \linenumbers
 
-\section*{ABSTRACT \label{sec:ABSTRACT}}
+This document accompanies the discussion contained in the main text.
+All the Python code used for this study, as well as the associated data tables not included in this document, can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
-XXX
+\clearpage
 
-\section*{INTRODUCTION \label{sec:INTRODUCTION}}
-
-XXX
-
-\section*{METHODS}
-
-XXX
-
-\section*{RESULTS}
-
-XXX
-
-\section*{DISCUSSION}
-
-XXX
-
-\section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}
+\section*{XXX}
 
@@
 \clearpage
-\newpage
-
-\section*{TABLES}
-
-XXX
-
-\clearpage
-\newpage
-
-\section*{FIGURES}
-
-XXX
-
-\clearpage
-\newpage
-\footnotesize
 
 \singlespacing
+
+\newpage
 

### 2019-06-11 Manuscript/Manuscript.tex

@@
 % Self defined commands
-\newcommand{\degC}{$^{\circ}$C\xspace}
+\newcommand{\degreesC}{\textdegree C\xspace}
+\newcommand{\degrees}{\textdegree\xspace}
 \newcommand{\dC}{$\delta^{13}$C\xspace}
@@
 \newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
+\newcommand{\pCOtwo}{\textit{p}CO$_{2}$\xspace}
+\newcommand{\COtwo}{CO$_{2}$\xspace}
 %
@@
 Yuem Park\textsuperscript{1},
-Nicholas L. Swanson-Hysell\textsuperscript{1}
+Nicholas L. Swanson-Hysell\textsuperscript{1},
+Pierre Maffre\textsuperscript{2},
+Yves Godd\'eris\textsuperscript{2}
 
@@
 
-XXX
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
+Imbalances between the magnitude of these source and sink fluxes will cause \pCOtwo to change until the silicate weathering negative feedback (in which elevated \pCOtwo leads to higher temperatures and a more intense hydrological cycle that enhances chemical weathering of silicate rocks and therefore the consumption of \COtwo, and vice versa) stabilizes \pCOtwo at a new steady state value \citep{Walker1981a}.
+One way to perturb steady state \pCOtwo is to change global weatherability - the sum of factors other than climate that contribute to the chemical weathering of rocks and associated \COtwo consumption, such as lithology \citep{Gaillardet1999a, Dessert2003a}, topography \citep{Maher2014a, Maffre2018a}, and paleolatitude \citep{Swanson-Hysell2017a}.
+\COtwo input from the solid Earth, assuming that it remains constant, can be removed via silicate weathering at a lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
+
+Mafic rocks have a higher weatherability than felsic rocks because of both the higher concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) in mafic lithologies, as well as their faster weathering rates in identical conditions \citep{Dessert2001a, Dessert2003a}.
+Soil shielding is another important factor that affects weatherability - in low-relief regions, regolith tends to be thicker, which can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock is dependent only on the uplift rate \citep{Gabet2009a, Maher2014a}.
+In this framework, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions where temperature and runoff is high, should exert a strong control on global weatherability and long-term climate.
+Recent work has suggested that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a}, the Cenozoic \citep{Jagoutz2016a}, and even the Neoproterozoic \citep{Park2018a} played an important role in initiating glacial climate states at those times.
+Furthermore, a comparison between the paleolatitudinal position of all major Phanerozoic arc-continent collisions and the latitudinal extent of continental ice sheets reveals a strong correlation between the extent of glaciation and arc-continent collisions in the tropics \citep{Macdonald2019a}.
+On the other hand, \citet{Park2019a} found no correlation between the extent of glaciation and large igneous province area in the tropics, which was explained by the fact that large igneous provinces, although mafic, are often emplaced into low relief areas where soil shielding is relatively high.
+
+However, quantifying the increase in global weatherability associated with these tropical arc-continent collisions or large igneous province emplacements and the consequent decrease in steady state \pCOtwo is challenging for the past, primarily due to the large uncertainties on the global context within which these events took place.
+For instance, in order to even begin to model global weathering fluxes and \pCOtwo, one would need to have constraints on the paleogeographic distribution of land masses, the spatial distribution of temperature, runoff, topography, and lithology on these land masses, and the carbon input flux coming from the degassing of the solid Earth at the time of these events.
+While the paleogeographic distribution of the continents is broadly well-constrained for recent Earth history (e.g.
+\citealp{Torsvik2016a}) and temperature and runoff can be estimated using a global climate model, constraints on the other parameters are poor, particularly as we go deeper into Earth's history.
+
+However, the growth of the Malay Archipelago (the islands of Indonesia and surrounding regions, Fig.
+X) as Australia migrates northward represents an episode of arc-continent collision occurring in the tropics today.
+Given that the global context within which this event is occurring is well-constrained, the growth of the Malay Archipelago presents an opportunity to quantify the increase in global weatherability and consequent decrease in steady state \pCOtwo associated with it, and evaluate the feasibility of similar tropical arc-continent collisions initiating glacial climate states throughout Earth's history \citep{Macdonald2019a}.
 
@@
 
-XXX
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the Malay Archipelago, we use the spatially resolved GEOCLIM Earth system model.
+At its core, GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell to obtain steady state \pCOtwo.
+In previous versions of the model, the silicate weathering rate for each continental grid cell was a function of temperature, runoff, and a calibration constant only (e.g.
+\citealp{Godderis2014a}).
+However, more recent versions of the model implement a dependence on...
+
+\begin{itemize}
+    \item GEOCLIM and GFDL models
+    \item added lithology (GLiM)
+    \item calibration
+    \item tested scenarios
+\end{itemize}
+
+The global lithological map database GLiM of Hartmann and Moosdorf was used to develop a
+
+The lithological categories of the global lithological map database GLiM of \cite{Hartmann2012a} were used and grouped into 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic.
+These values were assigned at the 0.5\textdegree resolution and were then downsampled using a mode downsampling scheme to 1\textdegree resolution
 
@@
 
+\section{Supplemental Materials}
+Experimental determinations of the activation energy (E$_a$) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
+However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (41.8 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42 kJ/mol; \citealp{Carroll2005a}.
+We use the values of 42 kJ/mol in our model runs.
+
+%note that these values are for 25ºC To
+
+% http://dx.doi.org/10.1016/B0-08-043751-6/05075-1 Brantley2003a
+% https://doi.org/10.1016/0016-7037(93)90431-U Knauss1993a
+% http://dx.doi.org/10.1016/j.chemgeo.2004.12.008 Carroll2005a
+
 \end{document}

### 2019-06-12 Manuscript/Manuscript.tex

@@
 Pierre Maffre\textsuperscript{2},
+Francis Macdonald\textsuperscript{3},
 Yves Godd\'eris\textsuperscript{2}
@@
 \section*{INTRODUCTION \label{sec:INTRODUCTION}}
+
+The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes.
+The confluence of steep topography, a warm and wet tropical climate and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a,Hartmann2014a}.
+There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between the Sunda-Banda arc system \citep{Molnar2015a}.
+
+Given that the region is a modern-day hotspot of chemical weathering and \COtwo consumption, what effect has this growth had on Earth's climate state and what would \pCOtwo on Earth be if the region wasn't there at all?
+Assessing the effect of the growth of the Indonesia region on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
+As \COtwo sinks are removed and \COtwo rises, increases in temperature and runoff cause other weathering sinks to increase.
+The steady-state \COtwo of Earth is set by the pCO$_2$ level at which the sinks of \COtwo are equal to the sources.
+Based on paleogeographic reconstructions of arc-continent collisions through Earth history, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and ind
 
@@
 \section{Supplemental Materials}
+
 Experimental determinations of the activation energy (E$_a$) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
-However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (41.8 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42 kJ/mol; \citealp{Carroll2005a}.
+However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5$\pm$1.7 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1 kJ/mol; \citealp{Carroll2005a}.
 We use the values of 42 kJ/mol in our model runs.
-
-%note that these values are for 25ºC To
 

### 2019-06-17 Manuscript/Data_Repository/Data_Repository.tex
Updates from Overleaf

@@
 Yuem Park\textsuperscript{1},
+Nicholas L. Swanson-Hysell\textsuperscript{1},
+Pierre Maffre\textsuperscript{2},
+Yves Godd\'eris\textsuperscript{2}
 \\
@@
 
-\clearpage
+\section*{Regolith Component}
 
-\section*{XXX}
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=0.5\textwidth]{../Figures/regolith_schematic.pdf}
+	\caption{A schematic representation of the regolith component in GEOCLIM at steady state.
+A rock particle leaves the unweathered bedrock with production rate $P$, and transits vertically through a regolith of height $h$.
+As the rock particle transits through the regolith, some fraction of the weatherable phases in the rock particle ($x$) are dissolved and leaves the regolith column as weathered phases ($W$).
+The fraction of the weatherable phases in the rock particle that are not dissolved after transiting through the regolith column leaves it via erosion ($E$).}
+	\label{fig:regolith_schematic}
+\end{center}
+\end{figure}
+
+The regolith component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (Fig.
+\ref{fig:regolith_schematic}).
+The component is based upon the theoretical model of \citet{Gabet2009a}, in which material enters the regolith through the downward propagation of the weathering front, and leaves the regolith through physical erosion at the land surface and through the transport of dissolved phases by chemical weathering.
+
+The fundamental processes within the regolith component are described by the following system of three equations:
+
+\begin{equation}
+    \frac{dh}{dt} = P - E
+    \label{eq:1}
+\end{equation}
+
+\begin{equation}
+    \frac{\partial x}{\partial t} = -P \frac{\partial x}{\partial z} - K \tau^{\sigma}x
+    \label{eq:2}
+\end{equation}
+
+\begin{equation}
+    \frac{\partial \tau}{\partial t} = -P \frac{\partial \tau}{\partial z} + 1
+    \label{eq:3}
+\end{equation}
+
+\noindent
+Equation \ref{eq:1} is simply a statement of mass conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P$ is the regolith production rate (m/yr), and $E$ is the physical erosion rate (m/yr).
+Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as time in the model ($t$, yr) progresses, where $z$ is the height above the base of the regolith (m).
+$K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless).
+Equation \ref{eq:3} describes how the time that a given rock particle has spent in the regolith changes as time in the model progresses.
+
+The net weathering rate in the regolith column ($W$, m/yr) can then be calculated with:
+
+\begin{equation}
+    W = \int_{0}^{h} K \tau^{\sigma} x\;dz
+    \label{eq:4}
+\end{equation}
+
+The steady state solution of the regolith component is obtained by setting the left hand side of Equations \ref{eq:1}, \ref{eq:2} and \ref{eq:3} to zero.
+
+\citet{Heimsath1997a} found that the rate at which material is entrained into the physically disturbed portion of the regolith (often called soil production) declines exponentially with increasing depth of this disturbed layer.
+The regolith model that we use does not differentiate the physically disturbed portion of the regolith, but given that the extent of weathering in the regolith below the physically disturbed layer has been found to covary with the rate of soil production \citep{Burke2007a}, we assume that regolith production behaves similarly to soil production:
+
+\begin{equation}
+    P = P_{0}\;f(h)
+    \label{eq:5}
+\end{equation}
+
+\begin{equation}
+    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:6}
+\end{equation}
+
+\begin{equation}
+    f(h) = e^{-\frac{-h}{d_{0}}}
+    \label{eq:7}
+\end{equation}
+
+\noindent
+$P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
+In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the universal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
+$f(h)$ is the `soil production function,' which describes how regolith production decreases as the thickness of the regolith increases.
+In Equation \ref{eq:7}, we use an exponential equation to represent the soil production function as in \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
+
+The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the dissolution rate on climate:
+
+\begin{equation}
+    K = k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:8}
+\end{equation}
+
+\noindent
+Equation \ref{eq:8} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citet{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modified the dependence of dissolution rate on runoff only (yr/m).
+
+Finally, the erosion rate in the regolith component of GEOCLIM uses the parameterization of \citet{Maffre2018a}, which combines the small-scale incision model of \citet{Davy2000a} with the large-scale BQART model of \citet{Syvitski2007a}:
+
+\begin{equation}
+    E = k_{e}\;q^{m}\;s^{n}
+    \label{eq:9}
+\end{equation}
+
+\noindent
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$), $s$ is the slope (m/m), and $m$ and $n$ are adjustable exponents.
+
+\section*{Choice of Parameters}
 
 XXX
+
+Experimental determinations of the activation energy (E$_a$) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
+However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5$\pm$1.7 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1 kJ/mol; \citealp{Carroll2005a}.
+We use the values of 42 kJ/mol in our model runs.
 

### 2019-06-17 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 As \COtwo sinks are removed and \COtwo rises, increases in temperature and runoff cause other weathering sinks to increase.
-The steady-state \COtwo of Earth is set by the pCO$_2$ level at which the sinks of \COtwo are equal to the sources.
+The steady-state \COtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
 Based on paleogeographic reconstructions of arc-continent collisions through Earth history, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and ind
+
+XXXX
 
@@
 
-\section{Supplemental Materials}
-
-Experimental determinations of the activation energy (E$_a$) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
-However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5$\pm$1.7 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1 kJ/mol; \citealp{Carroll2005a}.
-We use the values of 42 kJ/mol in our model runs.
-
-% http://dx.doi.org/10.1016/B0-08-043751-6/05075-1 Brantley2003a
-% https://doi.org/10.1016/0016-7037(93)90431-U Knauss1993a
-% http://dx.doi.org/10.1016/j.chemgeo.2004.12.008 Carroll2005a
-
 \end{document}

### 2019-06-18 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 Pierre Maffre\textsuperscript{2},
-Francis Macdonald\textsuperscript{3},
+Francis A. Macdonald\textsuperscript{3},
 Yves Godd\'eris\textsuperscript{2}
@@
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the Malay Archipelago, we use the spatially resolved GEOCLIM Earth system model.
-At its core, GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell to obtain steady state \pCOtwo.
-In previous versions of the model, the silicate weathering rate for each continental grid cell was a function of temperature, runoff, and a calibration constant only (e.g.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the Malay Archipelago, we use the spatially resolved GEOCLIM Earth system model \citep{Godderis2014a, Godderis2017c}, which iteratively couples box models of surface biogeochemical cycles to climate model output.
+Estimates of fluxes of species between carbon, alkalinity, phosphorous, and oxygen reservoirs in the ocean and atmosphere are used to calculate \pCOtwo, which is then compared against climate model output computed at various \pCOtwo levels to estimate temperature and runoff at the current \pCOtwo in GEOCLIM.
+These new temperature and runoff fields are then fed back into the biogeochemical box models to update \pCOtwo, for which new temperature and runoff fields are again estimated.
+This iteration is continued until a steady state is achieved in all reservoirs.
+
+In this study, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
+These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
+1860 at a 1\degrees $\times$ 1\degrees resolution.
+In the `1860 control' experiment, forcing agents representative of conditions ca.
+1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
+\pCOtwo in 1860 is assumed to be 286~ppm.
+In the `+1\%/yr to 2$\times$' experiment, initial conditions are taken from the `1860 control' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
+\pCOtwo is then held constant until the end of the 280 year experiment.
+All non-\COtwo forcing agents are held constant.
+The `+1\%/yr to 4$\times$' experiment is identical to the `+1\%/yr to 2$\times$' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
+We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
+The primary strength of using these three GFDL CM2.0 experiments is that all non-\COtwo forcing agents are held constant, allowing the effect of \pCOtwo on climate to be isolated.
+
+The silicate weathering component of GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell.
+In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions (e.g.
 \citealp{Godderis2014a}).
-However, more recent versions of the model implement a dependence on...
+However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see Data Repository).
+While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of arc-continent collisions in the Malay Archipelago on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
+We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
+16 lithologic categories are represented in GLiM, which we simplify into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+XX; see Data Repository).
+Each continental grid cell is then assigned one of these lithologic categories.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in the EarthChem Portal (\url{www.earthchem.org/portal}).
+The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of any metamorphic or siliciclastic sedimentary rock is strongly dependent on the protolith.
+
+and maturity for sediments
+
+XXX
 
@@
     \item tested scenarios
+    \item in discussion, talk about GFDL limitations (equilibrated by last 100 years?, not changing paleogeography, low resolution, model is quite old)
 \end{itemize}
@@
 
-XXX
+\begin{itemize}
+    \item lithologic mask
+\end{itemize}
 

### 2019-06-18 Manuscript/Data_Repository/Data_Repository.tex

@@
 
-\section*{Regolith Component}
+\section*{Silicate Weathering Component}
 
@@
 	\includegraphics[width=0.5\textwidth]{../Figures/regolith_schematic.pdf}
-	\caption{A schematic representation of the regolith component in GEOCLIM at steady state.
+	\caption{A schematic representation of the silicate weathering in GEOCLIM at steady state.
 A rock particle leaves the unweathered bedrock with production rate $P$, and transits vertically through a regolith of height $h$.
@@
 
-The regolith component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (Fig.
+The silicate weathering of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (Fig.
 \ref{fig:regolith_schematic}).
@@
 
-The fundamental processes within the regolith component are described by the following system of three equations:
+The fundamental processes within the silicate weathering are described by the following system of three equations:
 
@@
 
-The steady state solution of the regolith component is obtained by setting the left hand side of Equations \ref{eq:1}, \ref{eq:2} and \ref{eq:3} to zero.
+The steady state solution of the silicate weathering is obtained by setting the left hand side of Equations \ref{eq:1}, \ref{eq:2} and \ref{eq:3} to zero.
 
@@
 
-Finally, the erosion rate in the regolith component of GEOCLIM uses the parameterization of \citet{Maffre2018a}, which combines the small-scale incision model of \citet{Davy2000a} with the large-scale BQART model of \citet{Syvitski2007a}:
+Finally, the erosion rate in the silicate weathering of GEOCLIM uses the parameterization of \citet{Maffre2018a}, which combines the small-scale incision model of \citet{Davy2000a} with the large-scale BQART model of \citet{Syvitski2007a}:
 
@@
 
+\section*{Lithology Implementation}
+
+XXX
+
+going from GLiM to our lithologic classes.
+
 \clearpage

### 2019-06-20 Manuscript/Data_Repository/Data_Repository.tex
Updates from Overleaf

@@
 \usepackage{changepage}
+\usepackage{sidecap}
 \usepackage{rotating}
@@
 
-This document accompanies the discussion contained in the main text.
-All the Python code used for this study, as well as the associated data tables not included in this document, can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These repository materials provide details on the model framework used tin this study. All of the Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
@@
 
-\begin{figure}[h!]
+\begin{SCfigure}[0.6][h!]
 \begin{center}
-	\includegraphics[width=0.5\textwidth]{../Figures/regolith_schematic.pdf}
-	\caption{A schematic representation of the silicate weathering in GEOCLIM at steady state.
+	\includegraphics[width=0.6\textwidth]{../Figures/regolith_schematic.pdf}
+	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady state.
 A rock particle leaves the unweathered bedrock with production rate $P$, and transits vertically through a regolith of height $h$.
-As the rock particle transits through the regolith, some fraction of the weatherable phases in the rock particle ($x$) are dissolved and leaves the regolith column as weathered phases ($W$).
-The fraction of the weatherable phases in the rock particle that are not dissolved after transiting through the regolith column leaves it via erosion ($E$).}
+As the particle transits upwards, some fraction of the calcium and magnesium in the particle ($x$) leaves the regolith column as solutes ($W$) due to chemical weathering.
+The fraction of calcium and magnesium remaining in the weatherable minerals after transiting through the regolith column leaves via physical erosion ($E$) along with other residual weathering products.}
 	\label{fig:regolith_schematic}
 \end{center}
-\end{figure}
+\end{SCfigure}
 
@@
 \ref{fig:regolith_schematic}).
+The usage of regolith here defines it as the chemically weathered profile where the base of the regolith is the contact with unweathered bedrock.
 The component is based upon the theoretical model of \citet{Gabet2009a}, in which material enters the regolith through the downward propagation of the weathering front, and leaves the regolith through physical erosion at the land surface and through the transport of dissolved phases by chemical weathering.
@@
 
+XXX
+
+Add Ca+Mg concentrations
+
 \section*{Choice of Parameters}

### 2019-06-20 Manuscript/Manuscript.tex

@@
 
+\subsection*{GEOCLIM}
+
 To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the Malay Archipelago, we use the spatially resolved GEOCLIM Earth system model \citep{Godderis2014a, Godderis2017c}, which iteratively couples box models of surface biogeochemical cycles to climate model output.
@@
 The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of any metamorphic or siliciclastic sedimentary rock is strongly dependent on the protolith.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
+The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
 
-and maturity for sediments
+\subsection*{Parameter Calibration}
+
+The silicate weathering component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (see Data Repository).
+As the rock particle transits through the regolith, some fraction of the weatherable phases in the rock particle are dissolved and leaves the regolith column as weathered phases.
+The fraction of the weatherable phases in the rock particle that are not dissolved after transiting through the regolith column leaves it via erosion.
+Central to this model is the dissolution rate constant, which describes the rate at which weatherable phases in the rock particle are dissolved as it transits through the regolith:
+
+\begin{equation}
+    k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}\tau^{\sigma}
+    \label{eq:1}
+\end{equation}
+
+$q$ is the runoff, $E_{a}$ is the activation energy, $R$ is the universal gas constant, $T$ is the temperature, $\tau$ is the time that a given rock particle has spent in the regolith, and $k_{d}$, $k_{w}$, and $\sigma$ are calibration constants.
+
+clean up equation - add units?
+
+allbcc vs lith map
 

### 2019-06-24 Manuscript/Data_Repository/Data_Repository.tex

@@
 
-These repository materials provide details on the model framework used tin this study. All of the Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These repository materials provide details on the model framework used in this study. All of the Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
@@
 	\includegraphics[width=0.6\textwidth]{../Figures/regolith_schematic.pdf}
-	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady state.
+	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state.
 A rock particle leaves the unweathered bedrock with production rate $P$, and transits vertically through a regolith of height $h$.
-As the particle transits upwards, some fraction of the calcium and magnesium in the particle ($x$) leaves the regolith column as solutes ($W$) due to chemical weathering.
-The fraction of calcium and magnesium remaining in the weatherable minerals after transiting through the regolith column leaves via physical erosion ($E$) along with other residual weathering products.}
+Regolith production and physical erosion ($E$) are equal at steady-state.
+As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).}
 	\label{fig:regolith_schematic}
@@
 
-The silicate weathering of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (Fig.
-\ref{fig:regolith_schematic}).
-The usage of regolith here defines it as the chemically weathered profile where the base of the regolith is the contact with unweathered bedrock.
-The component is based upon the theoretical model of \citet{Gabet2009a}, in which material enters the regolith through the downward propagation of the weathering front, and leaves the regolith through physical erosion at the land surface and through the transport of dissolved phases by chemical weathering.
-
-The fundamental processes within the silicate weathering are described by the following system of three equations:
+The silicate weathering portion of the GEOCLIM model implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
+We refer to this chemically weathered profile as regolith where the base of the regolith is the contact with unweathered bedrock.
+In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering as a particle transit through the regolith or as a physically weathered particle once it reaches the top.
+We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}.
+The transient time-varying versions of this regolith model is described by the following system of three equations which form the basis of DynSoil:
 
@@
 \noindent
-Equation \ref{eq:1} is simply a statement of mass conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P$ is the regolith production rate (m/yr), and $E$ is the physical erosion rate (m/yr).
-Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as time in the model ($t$, yr) progresses, where $z$ is the height above the base of the regolith (m).
-$K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless).
+Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P$ is the regolith production rate (m/yr), and $E$ is the physical erosion rate (m/yr).
+Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth (where $z$ is the height above the base of the regolith; m).
+$K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless) which implements a time-dependence.
 Equation \ref{eq:3} describes how the time that a given rock particle has spent in the regolith changes as time in the model progresses.
@@
 
-The steady state solution of the silicate weathering is obtained by setting the left hand side of Equations \ref{eq:1}, \ref{eq:2} and \ref{eq:3} to zero.
+The erosion rate (E; m/yr) in the silicate weathering module of GEOCLIM follows one of the approaches taken in \citet{Maffre2018a} where it is parameterized based on runoff (q; m/yr) and slope (s; m/m):
 
-\citet{Heimsath1997a} found that the rate at which material is entrained into the physically disturbed portion of the regolith (often called soil production) declines exponentially with increasing depth of this disturbed layer.
-The regolith model that we use does not differentiate the physically disturbed portion of the regolith, but given that the extent of weathering in the regolith below the physically disturbed layer has been found to covary with the rate of soil production \citep{Burke2007a}, we assume that regolith production behaves similarly to soil production:
+\begin{equation}
+    E = k_{e}\;q^{m}\;s^{n}
+    \label{eq:9}
+\end{equation}
+
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
+This formulation is similar to the large-scale BQART model of \citet{Syvitski2007a}, but without there being an implemented temperature dependence and by replacing elevation with slope such that it more closely follows the stream power law functional form \citep{Davy2000a}.
+This form and these exponent values are supported by compilations, such as that in \citet{Lague2014a}, although \citet{Lague2014a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
+
+The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and the soil production function:
 
@@
 
-\begin{equation}
-    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
-    \label{eq:6}
-\end{equation}
+In this implementation of DynSoil, we use an exponential soil production function as is commonly applied in the literature (e.g.
+\citealp{Gabet2009a}) and follows the form as in \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
+Here, $f(h)$ is the `soil production function,' which describes how regolith production decreases as the thickness of the regolith increases.
 
@@
 
-\noindent
 $P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
 In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the universal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
-$f(h)$ is the `soil production function,' which describes how regolith production decreases as the thickness of the regolith increases.
-In Equation \ref{eq:7}, we use an exponential equation to represent the soil production function as in \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
 
-The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the dissolution rate on climate:
+\begin{equation}
+    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:6}
+\end{equation}
+
+%\citet{Heimsath1997a} found that the rate at which material is entrained into the physically disturbed portion of the regolith (often called soil production) declines exponentially with increasing depth of this disturbed layer.
+The regolith model that we use does not differentiate the physically disturbed portion of the regolith, but given that the extent of weathering in the regolith below the physically disturbed layer has been found to covary with the rate of soil production \citep{Burke2007a}, we assume that regolith production behaves similarly to soil production:
+
+The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the chemical weathering on climate:
 
@@
 \noindent
-Equation \ref{eq:8} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citet{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modified the dependence of dissolution rate on runoff only (yr/m).
+Equation \ref{eq:8} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
+As described below, in our implementation of the model, we vary $k_{d}$, $k_{w}$ and $\sigma$ and compare the resulting calculated chemical weathering fluxes to data compilations.
+We take varying parameter combinations that result in the better fits between the modeled fluxes and the data and use them for the GEOCLIM simulations.
 
-Finally, the erosion rate in the silicate weathering of GEOCLIM uses the parameterization of \citet{Maffre2018a}, which combines the small-scale incision model of \citet{Davy2000a} with the large-scale BQART model of \citet{Syvitski2007a}:
+For our application of DynSoil, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
+The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations where $x_s$ is the proportion of primary phases remaining at the top of the column of regolith:
 
-\begin{equation}
-    E = k_{e}\;q^{m}\;s^{n}
-    \label{eq:9}
+\begin{equation}\label{eq:dynsoil_ss}
+\left\{\ 
+\begin{aligned}
+& h        \ =\   f^{-1}\left( \frac{P_o}{E} \right)                                                 \\
+& z(x)     \ =\   E \cdot\! \left( \frac{ (\sigma+1) \log(\frac{1}{x}) }{ K } \right) ^ \frac{1}{\sigma+1}                                                                  \\
+& x_s      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{h}{E}\right)^{\sigma+1}   \right)   \\
+& W        \ =\   (1-x_s) \, E                                                                           \\
+\end{aligned}
+\right.
 \end{equation}
-
-\noindent
-$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$), $s$ is the slope (m/m), and $m$ and $n$ are adjustable exponents.
 

### 2019-06-27 Manuscript/Data_Repository/Data_Repository.tex
Updates from Overleaf

@@
 
-These repository materials provide details on the model framework used in this study. All of the Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These repository materials provide details on the model framework used in this study. Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
@@
 
-The erosion rate (E; m/yr) in the silicate weathering module of GEOCLIM follows one of the approaches taken in \citet{Maffre2018a} where it is parameterized based on runoff (q; m/yr) and slope (s; m/m):
+The erosion rate (E; m/yr) in the silicate weathering module of GEOCLIM follows one of the approaches taken in \citet{Godderis2017c} and \citet{Maffre2018a} where it is parameterized based on runoff (q; m/yr) and slope (s; m/m):
 
@@
 For our application of DynSoil, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
-The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations where $x_s$ is the proportion of primary phases remaining at the top of the column of regolith:
+The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations.
 
@@
 \begin{aligned}
-& h        \ =\   f^{-1}\left( \frac{P_o}{E} \right)                                                 \\
-& z(x)     \ =\   E \cdot\! \left( \frac{ (\sigma+1) \log(\frac{1}{x}) }{ K } \right) ^ \frac{1}{\sigma+1}                                                                  \\
-& x_s      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{h}{E}\right)^{\sigma+1}   \right)   \\
-& W        \ =\   (1-x_s) \, E                                                                           \\
+& h   \ =\    \max\left(  \;  0  \;,\;  d_o \log\left(\frac{P_o}{E}\right)  \right)                                                 \\
+& x(z)      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{z}{E}\right)^{\sigma+1}   \right)   \\
+& W        \ =\   (1-x(h)) \, E  \ =\   (1-x_s) \, E                                                                         \\
 \end{aligned}
@@
 
-XXX
+\noindent $x(z)$ gives the curve of primary phases varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
+Setting z equal to regolith thickness (h) gives $x_s$ which is the proportion of primary phases remaining at the top of the column of regolith.
+This value controls the overall integrated chemical weathering rate out of the regolith.
+
+To calculate the flux of Ca and Mg associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
+\ref{fig:regolith_schematic}).
+Previous implementations of GEOCLIM have taken bulk continental crust values across all land.
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithological resolved version of the model.
+To do so, we grouped
 

### 2019-08-21 Manuscript/Manuscript.tex
Updates from Overleaf

@@
+% Letters
+% Letters are short reports of original research focused on an outstanding finding whose importance means that it will be of interest to scientists in other fields.
+
+% They do not normally exceed 4 pages of Nature and, as a guideline, allow up to 30 references.
+They begin with a fully referenced paragraph, ideally of about 200 words, but certainly no more than 300 words, aimed at readers in other disciplines.
+This paragraph starts with a 2-3 sentence basic introduction to the field; followed by a one-sentence statement of the main conclusions starting 'Here we show' or equivalent phrase; and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards.
+
+% Please refer to our annotated example to see how the summary paragraph for a Letter should be constructed.
+
+% The rest of the text is typically about 1,500 words long.
+Any discussion at the end of the text should be as succinct as possible, not repeating previous summary/introduction material, to briefly convey the general relevance of the work.
+
+% Letters typically have 3 or 4 small display items (figures or tables).
+
+% Word counts refer to the text of the paper. References, title, author list and acknowledgements do not have to be included in total word counts.
+
 \documentclass[11pt,letterpaper]{article}
@@
 \fancyhf{}
-\lhead{Park et al., in preparation for XXX}
+\lhead{Park et al., in preparation}
 \rhead{\thepage}
@@
 \begin{flushleft}
-{\Large \textbf{GEOCLIM Modern}}
+{\Large \textbf{Growth of the broader Indonesia archipelago as a driver for Neogene cooling}}
 
@@
 Nicholas L. Swanson-Hysell\textsuperscript{1},
-Pierre Maffre\textsuperscript{2},
-Francis A. Macdonald\textsuperscript{3},
-Yves Godd\'eris\textsuperscript{2}
+Pierre Maffre\textsuperscript{1},
+Francis A. Macdonald\textsuperscript{2},
+Eliel A. Anttila\textsuperscript{2},
+Yves Godd\'eris\textsuperscript{3}
 
@@
 
-\textsuperscript{2} XXX
+\textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
+
+\textsuperscript{3} G\'eosciences Environnement Toulouse, CNRS--Universit\'e Paul Sabatier - IRD, Toulouse, France
+
 \bigskip
@@
 
-\noindent\textit{This article is in preparation for XXX}
-
 \linenumbers
 
-\section*{ABSTRACT \label{sec:ABSTRACT}}
+The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a, Hartmann2014a}.
+There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a}; Fig.
+\ref{Fig:indo_growth}).
+This collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics.
 
-XXX
-
-\section*{INTRODUCTION \label{sec:INTRODUCTION}}
-
-The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a,Hartmann2014a}.
-There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between the Sunda-Banda arc system \citep{Molnar2015a}.
-
-Given that the region is a modern-day hotspot of chemical weathering and \COtwo consumption, what effect has this growth had on Earth's climate state and what would \pCOtwo on Earth be if the region wasn't there at all?
-Assessing the effect of the growth of the Indonesia region on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
-As \COtwo sinks are removed and \COtwo rises, increases in temperature and runoff cause other weathering sinks to increase.
-The steady-state \COtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
-Based on paleogeographic reconstructions of arc-continent collisions through Earth history, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and ind
-
-XXXX
+Given that the region is a modern-day hot spot of chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be on Earth if the region was not there at all?
+Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
+There has been significant planetary cooling since the Mid-Miocene Climatic Optimum leading to the onset of northern hemisphere glaciation, in addition to Antarctic glaciation, over the past few million years \citep{Shackleton1984a}.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a} providing an explanation for the delay between southern hemisphere and bipolar glaciation.
+We seek to evaluate the hypothesis that the growth of the broader Indonesian archipelago has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
 
 Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-Imbalances between the magnitude of these source and sink fluxes will cause \pCOtwo to change until the silicate weathering negative feedback (in which elevated \pCOtwo leads to higher temperatures and a more intense hydrological cycle that enhances chemical weathering of silicate rocks and therefore the consumption of \COtwo, and vice versa) stabilizes \pCOtwo at a new steady state value \citep{Walker1981a}.
+However, assessing the effect of the growth of the broader Indonesia archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
+As \COtwo sinks are removed and \COtwo rises, there will be increases in temperature and invigorated hydrological cycling that causes other weathering sinks to increase.
+The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
+The magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
+
+%Imbalances between the magnitude of these source and sink fluxes will cause \pCOtwo to change until the silicate weathering negative feedback (in which elevated \pCOtwo leads to higher temperatures and a more intense hydrological cycle that enhances chemical weathering of silicate rocks and therefore the consumption of \COtwo, and vice versa) stabilizes \pCOtwo at a new steady state value \citep{Walker1981a}.
 One way to perturb steady state \pCOtwo is to change global weatherability - the sum of factors other than climate that contribute to the chemical weathering of rocks and associated \COtwo consumption, such as lithology \citep{Gaillardet1999a, Dessert2003a}, topography \citep{Maher2014a, Maffre2018a}, and paleolatitude \citep{Swanson-Hysell2017a}.
@@
 
-Mafic rocks have a higher weatherability than felsic rocks because of both the higher concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) in mafic lithologies, as well as their faster weathering rates in identical conditions \citep{Dessert2001a, Dessert2003a}.
-Soil shielding is another important factor that affects weatherability - in low-relief regions, regolith tends to be thicker, which can lead to a supply-limited weathering regime in which the weathering rate of the underlying bedrock is dependent only on the uplift rate \citep{Gabet2009a, Maher2014a}.
-In this framework, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions where temperature and runoff is high, should exert a strong control on global weatherability and long-term climate.
+Topography, climate, and lithology all play important roles in modulating chemical weathering.
+In low-relief regions, low physical erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \citep{Gabet2009a, West2012a, Maher2014a}.
+In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \citep{Lasaga1994a, West2012a}.
+Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance the total chemical weathering rate \citep{West2012a, Maher2014a}.
+In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}.
+These factors taken together are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability and initiating glacial climate over the past 520~m.y.
+\citep{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \citep{Park2019a}.
+
+%higher weatherability than felsic rocks because of both the higher concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) in mafic lithologies, as well as their faster weathering rates in identical conditions \citep{Dessert2001a, Dessert2003a}.
+Soil shielding is another important factor that affects weatherability - in low-relief regions, regolith tends to be thicker, which can lead to  In this framework, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions where temperature and runoff is high, should exert a strong control on global weatherability and long-term climate.
 Recent work has suggested that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a}, the Cenozoic \citep{Jagoutz2016a}, and even the Neoproterozoic \citep{Park2018a} played an important role in initiating glacial climate states at those times.
@@
 
-However, quantifying the increase in global weatherability associated with these tropical arc-continent collisions or large igneous province emplacements and the consequent decrease in steady state \pCOtwo is challenging for the past, primarily due to the large uncertainties on the global context within which these events took place.
+%However, quantifying the increase in global weatherability associated with these tropical arc-continent collisions or large igneous province emplacements and the consequent decrease in steady state \pCOtwo is challenging for the past, primarily due to the large uncertainties on the global context within which these events took place.
 For instance, in order to even begin to model global weathering fluxes and \pCOtwo, one would need to have constraints on the paleogeographic distribution of land masses, the spatial distribution of temperature, runoff, topography, and lithology on these land masses, and the carbon input flux coming from the degassing of the solid Earth at the time of these events.
@@
 
-However, the growth of the Malay Archipelago (the islands of Indonesia and surrounding regions, Fig.
+%However, the growth of the Malay Archipelago (the islands of Indonesia and surrounding regions, Fig.
 X) as Australia migrates northward represents an episode of arc-continent collision occurring in the tropics today.
@@
 
-\section*{METHODS}
+% \section*{METHODS}
 
-\subsection*{GEOCLIM}
+% \subsection*{Paleoshoreline data}
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the Malay Archipelago, we use the spatially resolved GEOCLIM Earth system model \citep{Godderis2014a, Godderis2017c}, which iteratively couples box models of surface biogeochemical cycles to climate model output.
+We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia archipelago over the past 20~m.y.
+Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
+Carbonate strata and thin-bedded siliciclastic strata were interpreted to represent a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as evidence for a subaerial environment.
+Taken together, these constraints delineate the position of shorelines.
+For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norvick2003a, Cloos2005a, Gold2017a}.
+While there are difference in data sources and paleoshoreline outlines from \cite{Molnar2015a}, the calculated area at 5~Ma is quite similar.
+
+While there has been an increase in land area associated with arc-continent collision, there has also been a decrease in land area associated with subsidence of the Sunda Shelf \citep{Sarr2019a}.
+Although the majority of the Sunda Shelf is currently submerged, large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Halls2002a}.
+Similarities in terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
+During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence \citep{Morley2013a} and recent subsidence is interpreted to be a result of dynamic topography, triggered by subduction-related mantle flow \citep{Sarr2019a}.
+In the Miocene, this region was likely a mixture of crystalline lithologies and siliciclastic sediments similar to eastern Borneo.
+The relatively low-relief of the region contrasts with the high-relief and mafic lithologies of the broader Indonesia archipelago.
+We incorporate land area associated with the subsided Sunda Shelf into our modeling.
+
+%From Macdonald text:
+%A critical issue for the paleogeographic  analysis of Indonesia is how to deal with the Sunda Shelf.
+Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (38).
+From a biogeographical standpoint, similarities between the terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
+During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
+Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially-enclosed inland seas and extensive brackish or saline wetland environments.
+Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley 2013).
+Basement highs (e.g.
+the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al, 2007).
+Although most paleogeographic reconstructions of the region incorporate some degree of constant exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al, 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
+We test the sensitivity with two separate scenarios in which the Sunda Shelf is either variably exposed or entirely inundated from 20 Ma to present.
+
+%The areas of Indonesia and New Guinea has substantially increased over this time period with an acceleration over the past 5 million years.
+
+%We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics.
+We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.
+
+% \subsection*{GEOCLIM}
+
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to climate models run at varying \pCOtwo levels.
+We use the DynSoil implementation of the \cite{Gabet2009a} regolith development model which integrates a climatic dependence on the chemical weathering using the formulation of \cite{West2012a}.
+Changes in silicate weathering lead to changes in \pCOtwo with the climate models being used to estimate temperature and runoff fields which then modulates the silicate weathering
+
+then iteratively
+
+%which iteratively couples box models of surface biogeochemical cycles to climate model output.
 Estimates of fluxes of species between carbon, alkalinity, phosphorous, and oxygen reservoirs in the ocean and atmosphere are used to calculate \pCOtwo, which is then compared against climate model output computed at various \pCOtwo levels to estimate temperature and runoff at the current \pCOtwo in GEOCLIM.
@@
 
-XXX
+After the Miocene Climatic Optimum, a cooling trend began ca.
+15 Ma that accelerated over the past 5 Ma.
+Many hypotheses have been proposed to explain this cooling trend....
+One of these hypotheses is that
+
+Discuss source term. Can cite: https://doi.org/10.1093/gji/ggw277 related to spreading rate.
 
@@
 
-\begin{itemize}
-    \item lithologic mask
-\end{itemize}
+\begin{figure}
+    \centering
+    \includegraphics{Manuscript/Figures/regolith_schematic.pdf}
+    \caption{Growth of the broader Indonesia archipelago.
+Past shorelines are shown in A with resultant area estimates summarized in B.
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \cite{Zachos2008a}.}
+    \label{fig:my_label}
+\end{figure}
 
@@
 \end{document}
+
+Text from Francis draft on shoreline analysis:
+
+Neogene cooling and the emergence of Indonesia and New Guinea
+
+Abstract
+The cooling trend from 15 Ma to present coincides with arc-continent collision and exhumation in Indonesia and New Guinea.
+To assess relationships between the emergence of highly weatherable arc terranes with tectonic topography in the tropics, an increase in global weatherability, geochemical records, and climate change, we use geological and paleoshoreline data quantify the changes in the area of volcanic and ultramafic rocks in Indonesia and New Guinea over the past 20 Myrs.
+We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics.
+We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.
+Arc-continent collision has resulted in the exhumation of both the East Sulawesi and Irian Jaya-New Guinea ophiolites, which are two of the three largest preserved ophiolites on Earth.
+We suggest that cooling from 15 to 10 Ma was due in large part to the subaerial exposure of these ophiolites and increase in global weatherability, and that the more recent acceleration in cooling was due to increased exhumation rates in Indonesia and New Guinea.
+Both a decrease of mid ocean ridge spreading rates and changes in the composition and amount of riverine input is necessary to explain both global cooling and geochemical proxy data.
+The data presented herein helps disentangle these effects.
+
+Significance statement
+	The tectonic drivers for the Neogene cooling trend from ~15 Ma to present are unknown.
+Here we quantify paleoshoreline data in Indonesia and New Guinea over the Neogene to assess the hypothesis that cooling was caused by changes in paleogeography in the warm wet tropics.
+These data document the emergence of Indonesia over this interval and are consistent with changes in global weatherability associated with arc-continent collision in the  tropics as a major driver of long-term climate change.
+
+\body
+After the Miocene Climatic Optimum (MCO), a cooling trend began at ~15 Ma, and accelerated over the past 5 Ma (1-4) (Fig.
+1).
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling (5, 6), perhaps associated with increased organic carbon burial (7, 8) or the closure of the  Central American Seaway (9), an increase in the strength of the Walker Circulation associated with the growth of Indonesia (10), feedbacks with terrestrial vegetation (11), a decrease in volcanic outgassing (12), an increase in basaltic watersheds in the tropics (10, 13), uplift in the Himalayas (14), and arc-continent collision in the tropics (15).
+Although changes in circulation and vegetation may have contributed to these signals on hundreds of thousands to million year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 15 Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback (16-20).
+On geological timescales, atmospheric CO2 levels can vary due to changes in volcanic CO2 flux, but also due to changes in global weatherability — the cumulative factors that affect chemical weathering aside from climate (20, 21).
+Global weatherability is the product of variables such as lithology, tectonic uplift rates, and paleolatitude (20).
+An increase in global weatherability can result in cooling as the consumption of carbon through silicate weathering will match that of volcanic input at lower concentrations of atmospheric CO2 (20).
+Present-day global CO2-consumption estimates emphasize the importance of highly weatherable areas: ~10-20% of land area is responsible for ~50-75% of CO2 consumption through silicate weathering and associated carbonate precipitation (22, 23).
+These highly weatherable areas (e.g.
+Indonesia and New Guinea) are mainly situated in the warm, wet tropics with significant topographic relief, and have bedrock composed of Ca- and Mg-rich mafic and ultramafic lithologies, which have faster dissolution rates than felsic rocks (20, 23, 24).
+As a result of the intertropical convergence zone (ITCZ), and its migration through the year, annual rainfall on Earth is greatest within ~15º of the equator resulting in the tropical rain belt.
+Given that hemispherically asymmetric forcings are relatively small, the position of the ITCZ is interpreted to be stable even with large changes in forcing resulting from changing CO2 levels (25).
+In modern climate data, the region within 10º of either side of the equator corresponds with precipitation greater than 1.5 m/yr and the region within 15º corresponds with precipitation greater than 1.0 m/yr.
+Because of the strong relationship between runoff and silicate weathering (26), paleogeographic changes in tropics, particularly in tropical regions with tectonic topography and abundant volcanic rocks, like Indonesia and New Guinea, should have an outsized effect on the Earth’s long-term climate state (10, 15).
+Recently, it was proposed that arc-continent collision in the tropics drove the cooling trend over the past 15 million years (15, 27).
+This hypothesis builds off of the suggestion of Reusch and Maasch (28) that arc-continent collisions could accomplish both a decrease in outgassing by terminating arc volcanism (footnote: However, arc-continent collision generally results in subduction flip with renewal of arc magmatism (29)), and an increase in silicate weathering by exhuming soluble Ca- and Mg-rich silicate rocks.
+Jagoutz et al.
+(27) applied this process specifically to the closure of the Neo-Tethys ocean in the tropics and suggested that arc-continent collisions at approximately 80 and 50 Ma could account for the long-term, two-phased cooling trend from the Cretaceous to the E-O boundary.
+Importantly, Jagoutz et al.
+(27) combined the Himalayan uplift hypothesis (14) with the basalt drift hypothesis (13, 30), emphasizing that topography, precipitation in the tropical rain belt, and lithology are the key variables in global weatherability.
+Macdonald et al.
+(15) compiled paleolatitude data on ophiolite bearing sutures through the Phanerozoic and concluded that the extent of arc-continent collisions tropics correlates with extent of continental glaciation, and specifically, that arc-continent collisions in Indonesia and New Guinea were associated with cooling over the past 15 Myr.
+Due to rugged tectonic topography (Fig.
+2A), tropical temperatures and precipitation rates (Fig.
+2B), and a predominance of highly soluble mafic and ultramafic rocks (Fig.
+3A), watersheds in Indonesia and New Guinea have the highest rates of silicate weathering in the world accounting for ~9% of the modern global carbon sink (23, 26).
+Although there are significant monsoonal and orographic effects in Indonesia, and particularly in New Guinea (Fig.
+2A), the high mean annual rainfall sill results in rivers with the highest average silicate weathering rates (26).
+Recognizing the importance of the paleogeography of this region, Molnar and Cronin (10) used paleoshoreline data to estimate that the area of subaerially exposed landmass in Indonesia and New Guinea increased by 60% over the last 5 Myrs.
+Here we quantify the changes in exposed area and rock types in the tropical region of Indonesia and New Guinea over the past 20 Myrs in a paleogeographic model.
+These results allow us to assess the changes of mafic and ultramafic rocks in the warm wet tropics of Indonesia and New Guinea throughout the Neogene and test the hypothesis that the subaerial emergence of this region associated with arc-continent collision caused an increase in global weatherability and cooling.
+These data can further be incorporated into future coupled climate and weathering models to assess the effect of changing paleogeography and global weatherabilty on long-term climate state.
+
+Tectonic Evolution
+During Neogene arc-continent collision between the Australian-Indian and Eurasian plates and intervening arc terranes (31-33), Indonesia and New Guinea emerged as several large supra-subduction ophiolites and associated forearcs were exhumed in the warm, wet tropics.
+Some of these ophiolites formed immediately prior to collision and others have an older history.
+The oldest ophiolites preserved in the Indonesia orogenic system are situated in the northwest and associated with the easternmost Paleo-Tethyan sutures (34).
+On the Malay Peninsula, the Bentong-Raub belt (suture 1; Fig.
+3) demarcates the eastern margin of Cimmerian ribbon continent Sibumasu, which collided with the Suhothai arc in the latest Permian to early Triassic (34).
+Thermochronology suggests that the  Malay Peninsula was first exhumed during the mid-Eocene to early Oligocene (35).
+This region became a significant sediment source for the Crocker Fan on the NW margin of Borneo during the Oligocene (36), as well as to the east of Malay Peninsula during the Neogene, and has been largely exposed over the last 20 Ma (37, 38).
+In Sumatra, the Wolya arc (suture 2; Fig.
+3) formed at 95 Ma and is overlain by Campanian 83.5-70.6 Ma radiolarian cherts (39), which were uplifted and eroded from 60-40 Ma (40).
+These ophiolites were entrained in the Banda Arc, which has experienced magmatism and accretion throughout most of the Cenozoic as the Australian plate subducts beneath it (37, 41).
+From ~20-10 Ma, the majority of Sumatra was submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20 Ma (37, 41).
+Most of Sumatra was submerged until after 5 Ma with the arrival of the leading edge of continental Australia (37, 41).
+The Meratus belt (sutures 3 and 4; Fig.
+3), which extends from Java (42) through eastern Borneo (43) to Palawan (44), marks the Late Cretaceous collision of the Banda blocks and Argo.
+After this collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed (45), which we refer to here as the Great Indonesian arc.
+In southern Borneo, sedimentological data in the adjacent Barito Basin suggests that ophiolites in the Meratus Mountains were not emergent until the Late Miocene (46).
+In northern Borneo and Palawan, collision occurred in the Late Oligocene to Early Miocene, resulting in exhumation of the Palawan Ophiolite Complex (44, 46).
+Over 10 km of Neogene sediment fills the Sarawak Basin and NW Borneo trough, indicative of extensive exhumation and erosion in northern Borneo (36).
+Thermochronological data suggest extremely rapid uplift (7 km/Myr) and exhumation during the Late Miocene and Early Pliocene, resulting in the recent formation of Mt.
+Kinabalu, the tallest peak in Borneo (47).
+The East Sulawesi ophiolite (suture 5; Fig.
+3) was likely generated in a supra-subduction to backarc setting in the Great Indonesian arc (45, 48, 49) [but see (50) for an alternative view], detached from its metamorphic sole in the late Oligocene (32-28 Ma; 51), and subsequently thrust to the east above a west dipping slab.
+During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing additional ophiolite emplacement in west and southeast Sulawesi, including Buton (52, 53).
+Uplift was diachronous, not effecting units is western Sulawesi until the Middle Miocene (~13 Ma) (54).
+Exhumation is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins (52, 54).
+Collision of the Sula-Banggai block with the east margin of Sulawesi began in the Late Miocene and uplift accelerated in the Early Pliocene (5.2 to 3.8 Ma) and is associated with a major pulse of sedimentation in adjacent basins (55, 56).
+Off the northeastern margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes mélange (55).
+Thus, although the East Sulawesi ophiolite was trapped/emplaced onto the composite Sundaland margin (i.e.
+a fragment of previously accreted Australian crust) between 36 and 23 Ma, mafic and ultramafic rocks appear to have not been substantially subaerially exposed in the south and west arms until after 15 Ma, and on the east arm until after 5 Ma.
+Thermochronology indicates that exhumation accelerated in the Late Miocene to Pliocene (52-54, 57).
+North Sulawesi, where active volcanoes extend through the Sangihe Arc into the southern Philippines, was not substantially emergent until the Pliocene (58).
+Along the Papuan Peninsula of Eastern New Guinea, Oligocene arc-continent collision between Australian continental fragments and the Inner Melanesian arc exhumed the Papuan Ultramafic Belt (Fig.
+3; suture 6) above a north dipping slab (59, 60).
+In the Central Range of New Guinea, a second set of arc-continent collisions in New Guinea began in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62).
+This resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt (Fig.
+3; suture 7).
+The Middle Miocene (16–14 Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the PNG ophiolite belt (61, 63).
+This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
+12 Ma (59).
+Mountain building began in the Late Miocene, ca.
+8-7 Ma (59, 62) with 3-4 km of denudation between 8-5 Ma (64), but major relief was not generated until the Pliocene (65).
+Exhumation of the Central Range accelerated over the past 5 Myrs due to slab-breakoff and buoyant uplift under New Guinea (61).
+Although the Papuan Ultramafic Belt was generated and obducted earlier than the ophiolites in the Central Range (60, 66), it was also exhumed very rapidly over the past 10 Ma (62).
+After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin (62).
+Over the past 4 Myr, this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi (67), Halmahera (68, 69), and northern New Guinea (70) (Fig.
+3; sutures 8-10).
+In the Bird’s Head region, this change in plate motion led to CCW rotation and the exhumation of several core complexes (71).
+The Philippine ophiolites have been grouped in four distinct belts (72).
+We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone (Fig.
+3; sutures 11 and 12).
+Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments (73).
+The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present (74).
+In addition to collision, arc magmatism may have contributed more significantly to crustal growth in the Philippines (75).
+The youngest ophiolites of Indonesia formed in the Banda Sea starting at ca.
+15 Ma (33).
+The ongoing arc-continent collision became emergent in Seram and Timor (Fig.
+3; sutures 13 and 14) over the last 5 Myr, as the Banda arc began to abut the Banda embayment emplacing the Timor nappe (49, 76, 77).
+
+Paleoshorelines
+Geological maps and paleoshoreline data were used to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea.
+Following Molnar and Cronin (10), we analyzed the area changes of islands that are greater than ~200 km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda shelf.
+Larger islands were further divided into regions based upon their position on microplates as defined by Matthews, et al.
+(78).
+Lithological maps were modified from Hartmann and Moosdorf (79) and then supplemented with regional geological maps.
+From these maps, the percentage of ultramafic and volcanic rocks relative to other rock types was calculated in QGIS for each island and microplate region.
+Generally, the paleoshoreline reconstructions within the paleogeographic models coincide with sub-Epoch boundaries, i.e.
+Early Miocene = 23-16 Ma, Middle Miocene = 15.99-11.61 Ma, Late Miocene = 11.6-5.4 Ma, Pliocene =5.39-3.8 Ma.
+However, some additional shorelines have been added where additional age information is present to refine the timing of major changes in area.
+Miocene to present stratigraphic columns were compiled from each region to develop an age model for various paleoenvironmental indicators.
+In general, carbonate strata and thin-bedded siliciclastic strata were assumed to be deposited in a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as evidence for subaerial environment.
+Following previous work on paleoshorelines and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma), Middle Miocene (15 Ma), and Early Miocene (20 Ma).
+We used geological maps and paleoshoreline data to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea.
+Using Gplates, paleoshoreline shapefiles were incorporated in the paleogeographic model of Matthews, et al.
+(78).
+Although there are some differences between this reconstruction and the recent reconstruction of Hall (33), these do not entail significant changes in paleolatitude.
+Many new sources for paleoshoreline data have become available since Molnar & Cronin (10).
+Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by Nugraha and Hall (80), and in New Guinea, we followed Gold et al.
+(81),  Cloos et al.
+(61), Norvick (2003) and derivative work by Harrington et al.
+(82), and Nichols and Hall (1991).
+We also compiled our own stratigraphies for the region to provide further paleoenvironmental context.
+Although the outlines of our paleoshorelines are very different from those of Molnar & Cronin (10) our calculated area for 5 Ma is not substantially different (Fig.
+X).
+A critical issue for the paleogeographic  analysis of Indonesia is how to deal with the Sunda Shelf.
+Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (38).
+From a biogeographical standpoint, similarities between the terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
+During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
+Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially-enclosed inland seas and extensive brackish or saline wetland environments.
+Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley 2013).
+Basement highs (e.g.
+the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al, 2007).
+Although most paleogeographic reconstructions of the region incorporate some degree of constant exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al, 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
+We test the sensitivity with two separate scenarios in which the Sunda Shelf is either variably exposed or entirely inundated from 20 Ma to present.
+Independent of the Sunda Shelf, our studies suggest that the volume of sediment eroded from Indonesia and New Guinea has exponentially increased over the past 15 Ma, particularly in New Guinea.DISCUSS CHANGES FOR EACH REGION.
+
+References
+1.	Zachos J, Pagani M, Sloan L, Thomas E, & Billups K (2001) Trends, Rhythms, and Aberrations in the Global Climate 65 Ma to present. Science 292:686-693.
+2.	Cramer B, Miller K, Barrett P, & Wright J (2011) Late Cretaceous–Neogene trends in deep ocean temperature and continental ice volume: Reconciling records of benthic foraminiferal geochemistry (δ18O and Mg/Ca) with sea level history.
+Journal of Geophysical Research: Oceans 116(C12).
+3.	Herbert TD, et al. (2016) Late Miocene global cooling and the rise of modern ecosystems. Nature Geoscience 9(11):843-847.
+4.	Zachos JC, Dickens GR, & Zeebe REJN (2008) An early Cenozoic perspective on greenhouse warming and carbon-cycle dynamics. 451(7176):279.
+5.	Shevenell AE, Kennett JP, & Lea DW (2004) Middle Miocene southern ocean cooling and Antarctic cryosphere expansion. Science 305(5691):1766-1770.
+6.	Holbourn A, Kuhnt W, Kochhann KG, Andersen N, & Sebastian Meier KJG (2015) Global perturbation of the carbon cycle at the onset of the Miocene Climatic Optimum. 43(2):123-126.
+7.	Vincent E & Berger W (1985) Carbon dioxide and polar cooling in the Miocene: The Monterey hypothesis.
+AGU Geophysical Mongograph Series.
+The Carbon Cycle and Atmospheric CO2: Natural Variations from Archean to Present 32:455-468.
+8.	Flower BP & Kennett JPJG (1993) Relations between Monterey Formation deposition and middle Miocene global cooling: Naples Beach section, California. 21(10):877-880.
+9.	Haug GH & Tiedemann RJN (1998) Effect of the formation of the Isthmus of Panama on Atlantic Ocean thermohaline circulation. 393(6686):673.
+10.	Molnar P & Cronin TW (2015) Growth of the Maritime Continent and its possible contribution to recurring Ice Ages. Paleoceanography 30(3):196-225.
+11.	Cerling TE, et al. (1997) Global vegetation change through the Miocene/Pliocene boundary. Nature 389(6647):153-158.
+12.	Berner RA, Lasaga AC, & Garrels RM (1983) The carbonate-silicate geochemical cycle and its effect on atmospheric carbon dioxide over the past 100 million years.
+American Journal of Science 283:641-683.
+13.	Kent DV & Muttoni G (2013) Modulation of Late Cretaceous and Cenozoic climate by variable drawdown of atmospheric pCO2 from weathering of basaltic provinces on continents drifting through the equatorial humid belt.
+Climate of the Past 9(2):525.
+14.	Raymo ME, Ruddiman WF, & Froelich PN (1988) Influence of late Cenozoic mountain building on ocean geochemical cycles. Geology 16(7):649-653.
+15.	Macdonald FA, Swanson-Hysell NL, Park Y, & Lissiecki LE (2019) Arc-continent collisions in the tropics set the Earth’s climate state. Science 364:181-184.
+16.	Walker JCG, Hays PB, & Kasting JF (1981) A negative feedback mechanism for the long-term stabilization of the earth's surface temperature. Geophysical Research Letters 86:9776-9782.
+17.	Raymo ME (1991) Geochemical evidence supporting TC Chamberlin's theory of glaciation. Geology 19(4):344-347.
+18.	Berner RA (1994) Geocarb II: a revised model of atmospheric CO2 over Phanerozoic time. American Journal of Science 294:56-91.
+19.	Berner RA & Caldeira K (1997) The need for mass balance and feedback in the geochemical carbon cycle. Geology 25(10):955-956.
+20.	Kump LR & Arthur MA (1997) Global chemical erosion during the Cenozoic: Weatherability balances the budgets. Tectonic Uplift and Climate Change,  (Springer), pp 399-426.
+21.	Maher K & Chamberlain C (2014) Hydrologic regulation of chemical weathering and the geologic carbon cycle. Science 343(6178):1502-1504.
+22.	Hartmann J, Moosdorf N, Lauerwald R, Hinderer M, & West AJ (2014) Global chemical weathering and associated P-release—The role of lithology, temperature and soil properties.
+Chemical geology 363:145-163.
+23.	Dessert C, Dupré B, Gaillardet J, François LM, & Allegre CJ (2003) Basalt weathering laws and the impact of basalt weathering on the global carbon cycle. Chemical Geology 202(3):257-273.
+24.	Ibarra DE, et al. (2016) Differential weathering of basaltic and granitic catchments from concentration–discharge relationships. 190:265-293.
+25.	Donohoe A, Voigt AJCEP, & Mechanisms (2017) Why Future Shifts in Tropical Precipitation Will Likely Be Small: The Location of the Tropical Rain Belt and the Hemispheric Contrast of Energy Input to the Atmosphere.115-137.
+26.	Gaillardet J, Dupré B, Louvat P, & Allegre C (1999) Global silicate weathering and CO 2 consumption rates deduced from the chemistry of large rivers. Chemical geology 159(1):3-30.
+27.	Jagoutz O, Macdonald FA, & Royden L (2016) Low-latitude arc–continent collision as a driver for global cooling. Proceedings of the National Academy of Sciences 113(18):4935-4940.
+28.	Reusch DN & Maasch KA (1998) The transition from arc volcanism to exhumation, weathering of young Ca, Mg, Sr silicates, and CO2 drawdown. Oxford Monographs on Geology and Geophysics 39:261-276.
+29.	McKenzie DP (1969) Speculations on the consequences and causes of plate motions. Geophysical Journal International 18(1):1-32.
+30.	Kent DV & Muttoni G (2008) Equatorial convergence of India and early Cenozoic climate trends. Proceedings of the National Academy of Sciences 105(42):16065-16070.
+31.	Hall R (1996) Reconstructing Cenozoic SE Asia. Geological Society, London, Special Publications 106(1):153-184.
+32.	Hamilton WB (1979) Tectonics of the Indonesian region (US Govt. Print. Off.).
+33.	Hall R (2017) Southeast Asia: New Views of the Geology of the Malay Archipelago. Annual Review of Earth and Planetary Sciences 45(1).
+34.	Metcalfe I (2013) Gondwana dispersion and Asian accretion: tectonic and palaeogeographic evolution of eastern Tethys. Journal of Asian Earth Sciences 66:1-33.
+35.	Cottam MA, Hall R, & Ghani AAJJoAES (2013) Late Cretaceous and Cenozoic tectonics of the Malay Peninsula constrained by thermochronology. 76:241-257.
+36.	Hall R, van Hattum MW, & Spakman WJT (2008) Impact of India–Asia collision on SE Asia: the record in Borneo. 451(1-4):366-389.
+37.	Hall R (2013) The palaeogeography of Sundaland and Wallacea since the Late Jurassic. Journal of Limnology 72(s2):1.
+38.	Hall R & Nichols GJGS, London, Special Publications (2002) Cenozoic sedimentation and tectonics in Borneo: climatic influences on orogenesis. 191(1):5-22.
+39.	Pedersen R, Searle M, Carter A, & Bandopadhyay P (2010) U–Pb zircon age of the Andaman ophiolite: implications for the beginning of subduction beneath the Andaman–Sumatra arc.
+Journal of the Geological Society 167(6):1105-1112.
+40.	Allen R, et al. (2008) New constraints on the sedimentation and uplift history of the Andaman-Nicobar accretionary prism, South Andaman Island. 436:223.
+41.	Hall R (2009) The Eurasian SE Asian margin as a modern example of an accretionary orogen. Geological Society, London, Special Publications 318(1):351-372.
+42.	Wakita K (2000) Cretaceous accretionary–collision complexes in central Indonesia. Journal of Asian Earth Sciences 18(6):739-749.
+43.	Monnier C, et al. (1999) Extensional to compressive Mesozoic magmatism at the SE Eurasia margin as recorded from the Meratus ophiolite (SE Borneo, Indonesia). Geodinamica Acta 12(1):43-55.
+44.	Yumul Jr GP, Dimalanta CB, Marquez EJ, & Queaño KL (2009) Onland signatures of the Palawan microcontinental block and Philippine mobile belt collision and crustal growth process: a review.
+Journal of Asian Earth Sciences 34(5):610-623.
+45.	Hall R (2002) Cenozoic geological and plate tectonic evolution of SE Asia and the SW Pacific: computer-based reconstructions, model and animations. Journal of Asian Earth Sciences 20(4):353-431.
+46.	Witts D, Hall R, Nichols G, & Morley R (2012) A new depositional and provenance model for the Tanjung Formation, Barito Basin, SE Kalimantan, Indonesia. Journal of Asian Earth Sciences 56:77-104.
+47.	Cottam M, et al. (2013) Neogene rock uplift and erosion in northern Borneo: evidence from the Kinabalu granite, Mount Kinabalu.2011-2130.
+48.	Monnier C, Girardeau J, Maury RC, & Cotten J (1995) Back-arc basin origin for the East Sulawesi ophiolite (eastern Indonesia). Geology 23(9):851-854.
+49.	Harris R (2006) Rise and fall of the Eastern Great Indonesian arc recorded by the assembly, dispersion and accretion of the Banda Terrane, Timor. Gondwana Research 10(3-4):207-231.
+50.	Kadarusman A, Miyashita S, Maruyama S, Parkinson CD, & Ishikawa A (2004) Petrology, geochemistry and paleogeographic reconstruction of the East Sulawesi Ophiolite, Indonesia.
+Tectonophysics 392(1):55-83.
+51.	Parkinson C (1998) Emplacement of the East Sulawesi Ophiolite: evidence from subophiolite metamorphic rocks. Journal of Asian Earth Sciences 16(1):13-28.
+52.	Bergman SC, Coffield DQ, Talbot JP, & Garrard RA (1996) Tertiary tectonic and magmatic evolution of western Sulawesi and the Makassar Strait, Indonesia: evidence for a Miocene continent-continent collision.
+Geological Society, London, Special Publications 106(1):391-429.
+53.	Smith RB & Silver EA (1991) Geology of a Miocene collision complex, Buton, eastern Indonesia. Geological Society of America Bulletin 103(5):660-678.
+54.	van Leeuwen TM, Susanto ES, Maryanto S, & Hadiwisastra S (2010) Tectonostratigraphic evolution of Cenozoic marginal basin and continental margin successions in the Bone Mountains, Southwest Sulawesi, Indonesia.
+Journal of Asian Earth Sciences 38(6):233-254.
+55.	Davies I (1990) Geological and exploration review of the Tomori PSC, Eastern Indonesia.
+56.	Villeneuve M, et al.
+(2000) Continental block collision in the eastern arm of Sulawesi (Indonesia).
+Structure and geodynamic interpretation.
+Comptes Rendus de l'Académie des Sciences-Series IIA-Earth and Planetary Science 330(5):371-378.
+57.	Hennig J, Hall R, Forster MA, Kohn BP, & Lister GS (2017) Rapid cooling and exhumation as a consequence of extension and crustal thinning: Inferences from the Late Miocene to Pliocene Palu Metamorphic Complex, Sulawesi, Indonesia.
+Tectonophysics 712:600-622.
+58.	van Leeuwen TM & Muhardjo (2005) Stratigraphy and tectonic setting of the Cretaceous and Paleogene volcanic-sedimentary successions in northwest Sulawesi, Indonesia: implications for the Cenozoic evolution of Western and Northern Sulawesi.
+Journal of Asian Earth Sciences 25(3):481-511.
+59.	van Ufford AQ & Cloos M (2005) Cenozoic tectonics of New Guinea. AAPG bulletin 89(1):119-140.
+60.	Davies H & Jaques A (1984) Emplacement of ophiolite in Papua New Guinea. Geological Society, London, Special Publications 13(1):341-349.
+61.	Cloos M, et al. (2005) Collisional delamination in New Guinea: The geotectonics of subducting slab breakoff. Geological Society of America Special Papers 400:1-51.
+62.	Baldwin SL, Fitzgerald PG, & Webb LE (2012) Tectonics of the New Guinea region. Annual Review of Earth and Planetary Sciences 40.
+63.	Visser WA & Hermes JJ (1962) Geological results of the exploration for oil in Netherlands New Guinea: carried out by the'Nederlandsche Nieuw Guinee Petroleum Maatschappij'1935-1960 (Staatsdrukkerij-en Uitgeverijbedrijf).
+64.	Crowhurst P, Hill K, Foster D, & Bennett A (1996) Thermochronological and geochemical constraints on the tectonic evolution of northern Papua New Guinea.
+Geological Society, London, Special Publications 106(1):525-537.
+65.	Weiland RJ & Cloos M (1996) Pliocene-Pleistocene asymmetric unroofing of the Irian fold belt, Irian Jaya, Indonesia: Apatite fission-track thermochronology.
+Geological Society of America Bulletin 108(11):1438-1449.
+66.	Davies HL & Smith IE (1971) Geology of eastern Papua. Geological Society of America Bulletin 82(12):3299-3312.
+67.	Ali J, Hall R, & Baker S (2001) Palaeomagnetic data from a Mesozoic Philippine Sea plate ophiolite on Obi Island, eastern Indonesia. Journal of Asian Earth Sciences 19(4):535-546.
+68.	Hall R, Audley-Charles M, Banner F, Hidayat S, & Tobing S (1988) Basement rocks of the Halmahera region, eastern Indonesia: a Late Cretaceous–early Tertiary arc and fore-arc.
+Journal of the Geological Society 145(1):65-84.
+69.	Ballantyne P (1992) Petrology and geochemistry of the plutonic rocks of the Halmahera ophiolite, eastern Indonesia, an analogue of modern oceanic forearcs.
+Geological Society, London, Special Publications 60(1):179-202.
+70.	Monnier C, et al.
+(1999) Petrology and geochemistry of the Cyclops ophiolites (Irian Jaya, East Indonesia): consequences for the Cenozoic evolution of the north Australian margin.
+Mineralogy and Petrology 65(1-2):1-28.
+71.	Sapin F, Pubellier M, Ringenbach J-C, & Bailly VJJoSG (2009) Alternating thin versus thick-skinned decollements, example in a fast tectonic setting: The Misool–Onin–Kumawa Ridge (West Papua).
+31(4):444-459.
+72.	Yumul GP (2007) Westward younging disposition of Philippine ophiolites and its implication for arc evolution. Island Arc 16(2):306-317.
+73.	Pubellier M, et al. (1991) The Mindanao collision zone: a soft collision event within a continuous Neogene strike-slip setting. Journal of Southeast Asian Earth Sciences 6(3-4):239-248.
+74.	Yumul Jr GP, Dimalanta CB, Tamayo Jr RA, & Faustino-Eslava DV (2013) Geological features of a collision zone marker: the Antique Ophiolite Complex (Western Panay, Philippines).
+Journal of Asian Earth Sciences 65:53-63.
+75.	Dimalanta C & Yumul G (2003) Magmatic and amagmatic contributions to crustal growth of an island-arc system: The Philippine example. International Geology Review 45(10):922-935.
+76.	Pownall JM, Forster MA, Hall R, & Watkinson IM (2017) Tectonometamorphic evolution of Seram and Ambon, eastern Indonesia: Insights from 40Ar/39Ar geochronology. Gondwana Research 44:35-53.
+77.	Monnier C, et al. (2003) Dynamics and age of formation of the Seram-Ambon ophiolites (Central Indonesia). Bulletin de la Société géologique de France 174(6):529-543.
+78.	Matthews KJ, et al. (2016) Global plate boundary evolution and kinematics since the late Paleozoic. Global and Planetary Change 146:226-250.
+79.	Hartmann J & Moosdorf N (2012) The new global lithological map database GLiM: A representation of rock properties at the Earth surface. Geochemistry, Geophysics, Geosystems 13(12).
+80.	Nugraha AMS & Hall R (2018) Late Cenozoic palaeogeography of Sulawesi, Indonesia. Palaeogeography, Palaeoclimatology, Palaeoecology 490:191-209.
+81.	Gold DP, White LT, Gunawan I, BouDagher-Fadel MJM, & Geology P (2017) Relative sea-level change in western New Guinea recorded by regional biostratigraphic data. 86:1133-1158.
+82.	Harrington L, Zahirovic S, Flament N, Müller RDJE, & Letters PS (2017) The role of deep Earth dynamics in driving the flooding and emergence of New Guinea since the Jurassic. 479:273-283.
+
+Figure Legends
+Fig. 1. Oxygen isotopic record of Neogene cooling. Benthic foraminifera 18O stack in black from Zachos et al. (4), with grey envelope of bottom water temperature estimates from Cramer et al. (2).
+
+Fig. 2. A) Slope map, showing active volcanoes, trenches and areas with high topography. B) Precipitation map over Indonesia (ADD NOAA reference).
+
+Fig. 3. A) Distribution of bedrock, mafic and ultramafic rock, of SE Asia from Hartmann and Moosdorf (79). Plate outlines from Mathews et al. (2016). B) Paleoshorelines shown for 20, 15, 10, and 5 Ma.
+
+Fig. 4. Paleogeogrpahic reconstructions of subaerially exposed area of Indonesia at 20, 15, 10, and 5 Ma.
+
+Fig. 5. Calculated change in total area, area in the tropics, and area of ultramafic and volcanic rocks in the tropics over that last 20 Ma in Indonesia and New Guinea.
+
+Fig. 6. Strontium and Osmium isotope model.
+
+Fig. 1
+
+Fig. 2

### 2019-08-21 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \newcommand{\permil}{\textperthousand\xspace}
-\newcommand{\dsil}{$d$\xspace}
 \newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
@@
 \begin{flushleft}
-{\Large \textbf{Growth of the broader Indonesia archipelago as a driver for Neogene cooling}}
+{\Large \textbf{Growth of the broader Indonesia Archipelago as a driver for Neogene cooling}}
 
@@
 
-The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes.
+The broader Indonesia Archipelago has an out-sized influence on modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a, Hartmann2014a}.
@@
 Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-However, assessing the effect of the growth of the broader Indonesia archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
+However, assessing the effect of the growth of the broader Indonesia Archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
 As \COtwo sinks are removed and \COtwo rises, there will be increases in temperature and invigorated hydrological cycling that causes other weathering sinks to increase.
@@
 
-We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia archipelago over the past 20~m.y.
+We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia Archipelago over the past 20~m.y.
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
@@
 In the Miocene, this region was likely a mixture of crystalline lithologies and siliciclastic sediments similar to eastern Borneo.
-The relatively low-relief of the region contrasts with the high-relief and mafic lithologies of the broader Indonesia archipelago.
+The relatively low-relief of the region contrasts with the high-relief and mafic lithologies of the broader Indonesia Archipelago.
 We incorporate land area associated with the subsided Sunda Shelf into our modeling.
@@
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to climate models run at varying \pCOtwo levels.
-We use the DynSoil implementation of the \cite{Gabet2009a} regolith development model which integrates a climatic dependence on the chemical weathering using the formulation of \cite{West2012a}.
-Changes in silicate weathering lead to changes in \pCOtwo with the climate models being used to estimate temperature and runoff fields which then modulates the silicate weathering
-
-then iteratively
-
-%which iteratively couples box models of surface biogeochemical cycles to climate model output.
-Estimates of fluxes of species between carbon, alkalinity, phosphorous, and oxygen reservoirs in the ocean and atmosphere are used to calculate \pCOtwo, which is then compared against climate model output computed at various \pCOtwo levels to estimate temperature and runoff at the current \pCOtwo in GEOCLIM.
-These new temperature and runoff fields are then fed back into the biogeochemical box models to update \pCOtwo, for which new temperature and runoff fields are again estimated.
-This iteration is continued until a steady state is achieved in all reservoirs.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
+We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
 
@@
 \citealp{Godderis2014a}).
-However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see Data Repository).
+However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see Supplementary Information).
+
 While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of arc-continent collisions in the Malay Archipelago on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 16 lithologic categories are represented in GLiM, which we simplify into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
-XX; see Data Repository).
+XX; see Supplementary Information).
 Each continental grid cell is then assigned one of these lithologic categories.
@@
 
-\subsection*{Parameter Calibration}
+% \subsection*{Parameter Calibration}
 
-The silicate weathering component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (see Data Repository).
+The silicate weathering component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (see Supplementary Information).
 As the rock particle transits through the regolith, some fraction of the weatherable phases in the rock particle are dissolved and leaves the regolith column as weathered phases.
@@
 \begin{itemize}
-    \item GEOCLIM and GFDL models
-    \item added lithology (GLiM)
     \item calibration
@@
     \includegraphics{Manuscript/Figures/regolith_schematic.pdf}
-    \caption{Growth of the broader Indonesia archipelago.
+    \caption{Growth of the broader Indonesia Archipelago.
 Past shorelines are shown in A with resultant area estimates summarized in B.
@@
 
+\section*{RANDOM STUFF}
+
+\subsection*{Text from Francis draft on shoreline analysis}
+
 \singlespacing
@@
 \end{document}
-
-Text from Francis draft on shoreline analysis:
 
@@
 Fig. 6. Strontium and Osmium isotope model.
-
-Fig. 1
-
-Fig. 2

### 2019-08-22 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 % Word counts refer to the text of the paper. References, title, author list and acknowledgements do not have to be included in total word counts.
+
+%%%%%%%%%%%%
+% GAMEPLAN %
+%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%
+%% MAIN TEXT %%
+%%%%%%%%%%%%%%%
+
+% introduction to the question of IA
+% topography, climate, and lithology are important
+% GEOCLIM model
+% dependence on slope and lithology
+% calibration (various feasible combinations of parameters)
+% temperature and runoff from GFDL
+% tested scenarios, including explanation of the paleoshoreline data
+% IA is important, traps are less important
+% limitations with approach
+    % GFDL equilibrated by last 100 years?
+    % not changing paleogeography
+    % low resolution
+    % model is quite old
+    % kinetics are not changing between lithologies
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% MAIN TEXT FIGURES %%
+%%%%%%%%%%%%%%%%%%%%%%%
+
+% paleoshorelines
+% steady state pCO2 for all scenarios
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SUPPLEMENTARY INFORMATION %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% details of DynSoil and GEOCLIM
+% details of parameter exploration/calibration
+    % basin data used for parameter exploration
+% details of lithology implementation
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SUPPLMENTARY INFORMATION FIGURES %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% regolith model schematic
+% maps of all scenarios
 
@@
 
-Given that the region is a modern-day hot spot of chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be on Earth if the region was not there at all?
+Given that the region is a modern-day hot spot of chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be if the region was not there at all?
 Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
@@
 
-%Imbalances between the magnitude of these source and sink fluxes will cause \pCOtwo to change until the silicate weathering negative feedback (in which elevated \pCOtwo leads to higher temperatures and a more intense hydrological cycle that enhances chemical weathering of silicate rocks and therefore the consumption of \COtwo, and vice versa) stabilizes \pCOtwo at a new steady state value \citep{Walker1981a}.
-One way to perturb steady state \pCOtwo is to change global weatherability - the sum of factors other than climate that contribute to the chemical weathering of rocks and associated \COtwo consumption, such as lithology \citep{Gaillardet1999a, Dessert2003a}, topography \citep{Maher2014a, Maffre2018a}, and paleolatitude \citep{Swanson-Hysell2017a}.
-\COtwo input from the solid Earth, assuming that it remains constant, can be removed via silicate weathering at a lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
-
 Topography, climate, and lithology all play important roles in modulating chemical weathering.
@@
 
-%higher weatherability than felsic rocks because of both the higher concentrations of Ca and Mg (that ultimately sequester carbon through precipitation as carbonate) in mafic lithologies, as well as their faster weathering rates in identical conditions \citep{Dessert2001a, Dessert2003a}.
-Soil shielding is another important factor that affects weatherability - in low-relief regions, regolith tends to be thicker, which can lead to  In this framework, processes that lead to continued exhumation of mafic lithologies and the creation of steep topography that minimizes soil shielding, particularly in tropical regions where temperature and runoff is high, should exert a strong control on global weatherability and long-term climate.
-Recent work has suggested that arc-continent collisions in the tropics during the Ordovician \citep{Swanson-Hysell2017a}, the Cenozoic \citep{Jagoutz2016a}, and even the Neoproterozoic \citep{Park2018a} played an important role in initiating glacial climate states at those times.
-Furthermore, a comparison between the paleolatitudinal position of all major Phanerozoic arc-continent collisions and the latitudinal extent of continental ice sheets reveals a strong correlation between the extent of glaciation and arc-continent collisions in the tropics \citep{Macdonald2019a}.
-On the other hand, \citet{Park2019a} found no correlation between the extent of glaciation and large igneous province area in the tropics, which was explained by the fact that large igneous provinces, although mafic, are often emplaced into low relief areas where soil shielding is relatively high.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
+We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
 
-%However, quantifying the increase in global weatherability associated with these tropical arc-continent collisions or large igneous province emplacements and the consequent decrease in steady state \pCOtwo is challenging for the past, primarily due to the large uncertainties on the global context within which these events took place.
-For instance, in order to even begin to model global weathering fluxes and \pCOtwo, one would need to have constraints on the paleogeographic distribution of land masses, the spatial distribution of temperature, runoff, topography, and lithology on these land masses, and the carbon input flux coming from the degassing of the solid Earth at the time of these events.
-While the paleogeographic distribution of the continents is broadly well-constrained for recent Earth history (e.g.
-\citealp{Torsvik2016a}) and temperature and runoff can be estimated using a global climate model, constraints on the other parameters are poor, particularly as we go deeper into Earth's history.
+The silicate weathering component of GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell.
+In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions (e.g.
+\citealp{Godderis2014a}).
+However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see SI).
+While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of arc-continent collisions in the Malay Archipelago on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
+We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
+16 lithologic categories are represented in GLiM, which we simplify into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+XX; see SI).
+Each continental grid cell is then assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in the EarthChem Portal (\url{www.earthchem.org/portal}).
+The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
+The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
 
-%However, the growth of the Malay Archipelago (the islands of Indonesia and surrounding regions, Fig.
-X) as Australia migrates northward represents an episode of arc-continent collision occurring in the tropics today.
-Given that the global context within which this event is occurring is well-constrained, the growth of the Malay Archipelago presents an opportunity to quantify the increase in global weatherability and consequent decrease in steady state \pCOtwo associated with it, and evaluate the feasibility of similar tropical arc-continent collisions initiating glacial climate states throughout Earth's history \citep{Macdonald2019a}.
+Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), four parameters in particular are poorly constrained in the existing literature: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
+As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
+Rather than prescribing values for these six parameters, we instead allowed these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
+For each of these combinations, we computed spatially-resolved long-term \COtwo consumption (i.e.
+Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
+Following the approach of \citet{Maffre2018a}, we then summed the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a} as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}, then calculated the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
+Finally, we eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
+This calibration method leaves 3,339 unique parameter combinations that produce total global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present day.
 
-% \section*{METHODS}
-
-% \subsection*{Paleoshoreline data}
+Having calibrated the  silicate weathering component of GEOCLIM, we now estimate the decrease in steady state \pCOtwo associated with increases in global weatherability.
+To do so, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
+These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
+1860 at a 1\degrees $\times$ 1\degrees resolution.
+In the `1860 control' experiment, forcing agents representative of conditions ca.
+1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
+\pCOtwo in 1860 is assumed to be 286~ppm.
+In the `+1\%/yr to 2$\times$' experiment, initial conditions are taken from the `1860 control' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
+\pCOtwo is then held constant until the end of the 280 year experiment.
+All non-\COtwo forcing agents are held constant.
+The `+1\%/yr to 4$\times$' experiment is identical to the `+1\%/yr to 2$\times$' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
+We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
+The primary strength of using these three GFDL CM2.0 experiments is that all non-\COtwo forcing agents are held constant, allowing the effect of \pCOtwo on climate to be isolated.
 
@@
 The relatively low-relief of the region contrasts with the high-relief and mafic lithologies of the broader Indonesia Archipelago.
+
+Given these uncertainties, we test three scenarios related to the growth of the Indonesia Archipelago.
+The "removed Indonesia Archipelago" represents an end-member scenario, in which all islands associated with
+
 We incorporate land area associated with the subsided Sunda Shelf into our modeling.
-
-%From Macdonald text:
-%A critical issue for the paleogeographic  analysis of Indonesia is how to deal with the Sunda Shelf.
-Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (38).
-From a biogeographical standpoint, similarities between the terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
-During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
-Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially-enclosed inland seas and extensive brackish or saline wetland environments.
-Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley 2013).
-Basement highs (e.g.
-the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al, 2007).
-Although most paleogeographic reconstructions of the region incorporate some degree of constant exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al, 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
-We test the sensitivity with two separate scenarios in which the Sunda Shelf is either variably exposed or entirely inundated from 20 Ma to present.
-
-%The areas of Indonesia and New Guinea has substantially increased over this time period with an acceleration over the past 5 million years.
-
-%We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics.
-We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.
-
-% \subsection*{GEOCLIM}
-
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
-We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
-
-In this study, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
-These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
-1860 at a 1\degrees $\times$ 1\degrees resolution.
-In the `1860 control' experiment, forcing agents representative of conditions ca.
-1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
-\pCOtwo in 1860 is assumed to be 286~ppm.
-In the `+1\%/yr to 2$\times$' experiment, initial conditions are taken from the `1860 control' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
-\pCOtwo is then held constant until the end of the 280 year experiment.
-All non-\COtwo forcing agents are held constant.
-The `+1\%/yr to 4$\times$' experiment is identical to the `+1\%/yr to 2$\times$' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
-We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
-The primary strength of using these three GFDL CM2.0 experiments is that all non-\COtwo forcing agents are held constant, allowing the effect of \pCOtwo on climate to be isolated.
-
-The silicate weathering component of GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell.
-In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions (e.g.
-\citealp{Godderis2014a}).
-However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see Supplementary Information).
-
-While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of arc-continent collisions in the Malay Archipelago on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
-We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
-16 lithologic categories are represented in GLiM, which we simplify into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
-XX; see Supplementary Information).
-Each continental grid cell is then assigned one of these lithologic categories.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in the EarthChem Portal (\url{www.earthchem.org/portal}).
-The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
-The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
-
-% \subsection*{Parameter Calibration}
-
-The silicate weathering component of GEOCLIM models the geochemical evolution of a rock particle as it leaves unweathered bedrock and transits through overlying regolith (see Supplementary Information).
-As the rock particle transits through the regolith, some fraction of the weatherable phases in the rock particle are dissolved and leaves the regolith column as weathered phases.
-The fraction of the weatherable phases in the rock particle that are not dissolved after transiting through the regolith column leaves it via erosion.
-Central to this model is the dissolution rate constant, which describes the rate at which weatherable phases in the rock particle are dissolved as it transits through the regolith:
-
-\begin{equation}
-    k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}\tau^{\sigma}
-    \label{eq:1}
-\end{equation}
-
-$q$ is the runoff, $E_{a}$ is the activation energy, $R$ is the universal gas constant, $T$ is the temperature, $\tau$ is the time that a given rock particle has spent in the regolith, and $k_{d}$, $k_{w}$, and $\sigma$ are calibration constants.
-
-clean up equation - add units?
-
-allbcc vs lith map
-
-XXX
-
-\begin{itemize}
-    \item calibration
-    \item tested scenarios
-    \item in discussion, talk about GFDL limitations (equilibrated by last 100 years?, not changing paleogeography, low resolution, model is quite old)
-\end{itemize}
-
-The global lithological map database GLiM of Hartmann and Moosdorf was used to develop a
-
-The lithological categories of the global lithological map database GLiM of \cite{Hartmann2012a} were used and grouped into 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic.
-These values were assigned at the 0.5\textdegree resolution and were then downsampled using a mode downsampling scheme to 1\textdegree resolution
-
-\section*{RESULTS}
-
-XXX
 
@@
     \centering
-    \includegraphics{Manuscript/Figures/regolith_schematic.pdf}
+    \includegraphics{Manuscript/Figures/shoreline_growth.pdf}
     \caption{Growth of the broader Indonesia Archipelago.
@@
 A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \cite{Zachos2008a}.}
-    \label{fig:my_label}
+    \label{fig:shoreline_growth}
 \end{figure}
@@
 \footnotesize
-
-\section*{RANDOM STUFF}
-
-\subsection*{Text from Francis draft on shoreline analysis}
 

### 2019-08-23 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 % details of lithology implementation
+    % including Sunda Shelf
 
@@
 There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a}; Fig.
-\ref{Fig:indo_growth}).
+\ref{fig:shoreline_growth}).
 This collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics.
@@
 Finally, we eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
-This calibration method leaves 3,339 unique parameter combinations that produce total global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present day.
+This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present day.
 
@@
 
-We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia Archipelago over the past 20~m.y.
+We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia Archipelago over the past 20~m.y (Fig.
+\ref{fig:shoreline_growth}).
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
@@
 
-Given these uncertainties, we test three scenarios related to the growth of the Indonesia Archipelago.
-The "removed Indonesia Archipelago" represents an end-member scenario, in which all islands associated with
+Given these uncertainties, we test three scenarios related to the growth of the Indonesia Archipelago (Fig.
+\ref{fig:scenario_pCO2}).
+The "removed Indonesia Archipelago" scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
+The "reduced Indonesia Archipelago" scenario represents our best estimate of the subaerial extent of islands in the area at ca.
+15~Ma, excluding the Sunda Shelf.
+The "reduced Indonesia Archipelago + Sunda Shelf" is identical to the "reduced Indonesia Archipelago" scenario with the addition of the Sunda Shelf.
+Given that no land in the Sunda Shelf is currently exposed, across the shelf we distribute felsic and siliciclastic lithologies, assign shallow slope values, and interpolate runoff across latitude bands (see SI).
 
-We incorporate land area associated with the subsided Sunda Shelf into our modeling.
+Our results indicate that
 
@@
     \centering
-    \includegraphics{Manuscript/Figures/shoreline_growth.pdf}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
     \caption{Growth of the broader Indonesia Archipelago.
@@
     \label{fig:shoreline_growth}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
+    \caption{Placeholder for steady state pCO2 figure.}
+    \label{fig:scenario_pCO2}
 \end{figure}

### 2019-08-26 Manuscript/Manuscript.tex
Updates from Overleaf

@@
     % including Sunda Shelf
+% details of paleoshoreline reconstructions
 
@@
 \begin{flushleft}
-{\Large \textbf{Growth of the broader Indonesia Archipelago as a driver for Neogene cooling}}
+{\Large \textbf{Emergence of Indonesia and New Guinea as a driver for Neogene cooling}}
 
@@
 
-The broader Indonesia Archipelago has an out-sized influence on modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a, Hartmann2014a}.
-There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a}; Fig.
+Indonesia and New Guinea have an outsized influence on modern chemical weathering fluxes.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Hartmann2014a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a, Hall2017a}; Fig.
 \ref{fig:shoreline_growth}).
-This collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics.
+The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citet{Macdonald2019a}.
 
-Given that the region is a modern-day hot spot of chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be if the region was not there at all?
+Given that the region is a modern-day hot spot of fresh water and sediment discharge \citep{Milliman2013a}, and chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has the emergence of Indonesia and New Guinea had on Earth's geochemical cycles and climate state?
+If this region was instead under water, would we have a glacial climate today?
 Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
-There has been significant planetary cooling since the Mid-Miocene Climatic Optimum leading to the onset of northern hemisphere glaciation, in addition to Antarctic glaciation, over the past few million years \citep{Shackleton1984a}.
+After the Miocene Climatic Optimum (MCO), a cooling trend began at $\sim$14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
+culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a} providing an explanation for the delay between southern hemisphere and bipolar glaciation.
-We seek to evaluate the hypothesis that the growth of the broader Indonesian archipelago has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
+We seek to evaluate the hypothesis that the increase of subaerially-exposed land area in Indonesia and New Guinea has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
 
 Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-However, assessing the effect of the growth of the broader Indonesia Archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
-As \COtwo sinks are removed and \COtwo rises, there will be increases in temperature and invigorated hydrological cycling that causes other weathering sinks to increase.
+However, assessing the effect of the growth of Indonesia and New Guinea on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
 The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
-The magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
+For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady state is achieved at a higher \pCOtwo.
+Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
 
@@
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the growth of the broader Indonesia Archipelago, we use the spatially resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the increase of subaerially-exposed land area in Indonesia and New Guinea, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
 We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
@@
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
-16 lithologic categories are represented in GLiM, which we simplify into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM simplified into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
 XX; see SI).
@@
 
-Having calibrated the  silicate weathering component of GEOCLIM, we now estimate the decrease in steady state \pCOtwo associated with increases in global weatherability.
+Having calibrated the  silicate weathering component of GEOCLIM, we estimate the decrease in steady state \pCOtwo associated with increases in global weatherability.
 To do so, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
@@
 \ref{fig:shoreline_growth}).
-Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
+Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
 Carbonate strata and thin-bedded siliciclastic strata were interpreted to represent a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as evidence for a subaerial environment.
 Taken together, these constraints delineate the position of shorelines.
-For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norvick2003a, Cloos2005a, Gold2017a}.
-While there are difference in data sources and paleoshoreline outlines from \cite{Molnar2015a}, the calculated area at 5~Ma is quite similar.
+For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norvick2003a, Harrington2017a, Gold2017a}.
+Although there are difference in data sources and paleoshoreline outlines from \cite{Molnar2015a}, the calculated area at 5~Ma is similar.
 
-While there has been an increase in land area associated with arc-continent collision, there has also been a decrease in land area associated with subsidence of the Sunda Shelf \citep{Sarr2019a}.
-Although the majority of the Sunda Shelf is currently submerged, large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Halls2002a}.
+Along with the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, there has also been a decrease in subaerially-exposed land area on the Sunda Shelf \citep{Sarr2019a}.
+Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}.
 Similarities in terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
-During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence \citep{Morley2013a} and recent subsidence is interpreted to be a result of dynamic topography, triggered by subduction-related mantle flow \citep{Sarr2019a}.
-In the Miocene, this region was likely a mixture of crystalline lithologies and siliciclastic sediments similar to eastern Borneo.
-The relatively low-relief of the region contrasts with the high-relief and mafic lithologies of the broader Indonesia Archipelago.
+During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence of the Sunda Shelf \citep{Morley2013a} and recent subsidence is interpreted to be a result of dynamic topography, triggered by subduction-related mantle flow \citep{Sarr2019a}.
+In the Miocene, this region exposed a mixture of felsic basement rocks and siliciclastic sediments \citet{Darmadi2007a}.
+The relatively low-relief and felsic lithologies of the region contrasts with the high-relief and mafic lithologies of most of Indonesia and New Guinea.
 
-Given these uncertainties, we test three scenarios related to the growth of the Indonesia Archipelago (Fig.
+To explore the sensitivity of global weatherability and climate state to the emergence of Indonesia and New Guinea, we test three scenarios using GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The "removed Indonesia Archipelago" scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
-The "reduced Indonesia Archipelago" scenario represents our best estimate of the subaerial extent of islands in the area at ca.
+The ``removed Indonesia and New Guinea'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
+The ``reduced Indonesia and New Guinea'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
 15~Ma, excluding the Sunda Shelf.
-The "reduced Indonesia Archipelago + Sunda Shelf" is identical to the "reduced Indonesia Archipelago" scenario with the addition of the Sunda Shelf.
+The ``reduced Indonesia and New Guinea + Sunda Shelf'' is identical to the ``reduced Indonesia and New Guinea'' scenario with the addition of the Sunda Shelf.
 Given that no land in the Sunda Shelf is currently exposed, across the shelf we distribute felsic and siliciclastic lithologies, assign shallow slope values, and interpolate runoff across latitude bands (see SI).
 
-Our results indicate that
+Of the 3,339 unique parameter combinations, the ``removed Indonesia and New Guinea'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced Indonesia and New Guinea'' scenario of 547-638~ppm, and the ``reduced Indonesia and New Guinea + Sunda Shelf'' of 533-611~ppm (Fig.
+\ref{fig:scenario_pCO2}).
+These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been \textgreater500~ppm.
+Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is minimal, given the low Mg+Ca of the lithologies that comprise the Sunda Shelf as well as its low slope.
 
-\section*{DISCUSSION}
+%%%
+% maybe place this at the end of the discussion
+The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an alternative hypothesis that has been put forth to explain the draw down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the sensitivity of global weatherability and climate state to the presence of these traps, we test three additional scenarios using GEOCLIM (Fig.
+\ref{fig:scenario_pCO2}).
+In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective traps are replaced with bulk continental crust.
+In the ``removed both traps'' scenario, all volcanics associated with both traps are replaced with bulk continental crust.
+These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three Indonesia and New Guinea scenarios.
+%%%
 
-After the Miocene Climatic Optimum, a cooling trend began ca.
-15 Ma that accelerated over the past 5 Ma.
-Many hypotheses have been proposed to explain this cooling trend....
-One of these hypotheses is that
+%%%
 
-Discuss source term. Can cite: https://doi.org/10.1093/gji/ggw277 related to spreading rate.
+After the MCO, a cooling trend began ca.
+14~Ma that accelerated over the past 4~m.y.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling \citep{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \citep{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \citep{Haug1998a}, an increase in the strength of the Walker Circulation \citep{Molnar2015a}, feedbacks with terrestrial vegetation \citep{Cerling1997a}(MAYBE ONE OF THE DINEZIO REFS), uplift in the Himalayas \citep{Raymo1988a}, or a decrease in volcanic outgassing \citep{Berner1983a}(ROWAN+ROWLEY).
+Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of Indonesia and New Guinea.
+
+Indonesia and New Guinea are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
+This paleogeographical feature originated in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62).
+In New Guinea, this resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt (Fig.
+3; suture 7).
+The Middle Miocene (16–14 Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the New Guinea ophiolite belt (61, 63).
+This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
+12 Ma (59).
+Mountain building began in the Late Miocene, ca.
+8-7 Ma (59, 62) with 3-4 km of denudation between 8-5 Ma (64).
+Exhumation of the Central Range accelerated over the past 4 Myrs due to slab-breakoff and buoyant uplift under New Guinea (61,65).
+After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin (62).
+Over the past 4 Myr, this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi (67), Halmahera (68, 69), and northern New Guinea (70) (Fig.
+3; sutures 8-10).
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4 Ma (van Ufford & Cloos, 2005), which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994; Abbott, 1995) and unroofing of basement-cored uplifts (Hill & Gleadow, 1989).
+
+Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an exceleration over the past 4 Ma.
+This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in New Guinea in the past 4 Ma.
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4 Ma that culminated in the initiation of the Northern Hemisphere ice sheets.
+Not only does the timing of paleogeographic changes in the tropics coincide with the cooling trend, but our coupled weathering climate model demonstrates that the magnitude of change in pCO2 due to the reconstructed paleogeographic changes is sufficient to explain climate change over the Neogene.
 
@@
 
-XXX
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their contributions to the paleoshoreline reconstructions. FAM and NLS-H thank NSF FRES grant #20191023 for support.
 
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
-    \caption{Growth of the broader Indonesia Archipelago.
+    \caption{Growth of Indonesia and New Guinea.
 Past shorelines are shown in A with resultant area estimates summarized in B.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
-    \caption{Placeholder for steady state pCO2 figure.}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2.pdf}
+    \caption{Box shows quartiles, whiskers show 1.5$\times$ the interquartile range, notch shows median with 95\% confidence interval.
+Each dot shows a parameter combination.
+Dashed line is PI \pCOtwo.}
     \label{fig:scenario_pCO2}

### 2019-08-27 Manuscript/Manuscript.tex

@@
 \ref{fig:shoreline_growth}).
-The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citet{Macdonald2019a}.
+The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citep{Macdonald2019a}.
 
@@
 Along with the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, there has also been a decrease in subaerially-exposed land area on the Sunda Shelf \citep{Sarr2019a}.
-Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}.
+Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}, potentially driving climate amplification on orbital timescales (ADD DENEZIO&TIERNY REFS).
 Similarities in terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
@@
 
-Of the 3,339 unique parameter combinations, the ``removed Indonesia and New Guinea'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced Indonesia and New Guinea'' scenario of 547-638~ppm, and the ``reduced Indonesia and New Guinea + Sunda Shelf'' of 533-611~ppm (Fig.
+Using the 3,339 unique parameter combinations, the ``removed Indonesia and New Guinea'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced Indonesia and New Guinea'' scenario of 547-638~ppm, and the ``reduced Indonesia and New Guinea + Sunda Shelf'' of 533-611~ppm (Fig.
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been \textgreater500~ppm.
-Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is minimal, given the low Mg+Ca of the lithologies that comprise the Sunda Shelf as well as its low slope.
+These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been $sim$500-700~ppm.
+Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca of the lithologies that comprise the Sunda Shelf as well as its low slope.
+
+These GEOCLIM \pCOtwo values resemble proxy-based estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea is enough to explain climate change over the Neogene.
+Furthermore, our paleoshoreline reconstructions (Fig.
+\ref{fig:shoreline_growth}) show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y.
+that culminated in the initiation of the Northern Hemisphere ice sheets.
+
+XXX
 
@@
 % maybe place this at the end of the discussion
-The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an alternative hypothesis that has been put forth to explain the draw down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
 To explore the sensitivity of global weatherability and climate state to the presence of these traps, we test three additional scenarios using GEOCLIM (Fig.
@@
 Indonesia and New Guinea are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
-This paleogeographical feature originated in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62).
-In New Guinea, this resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt (Fig.
-3; suture 7).
+In New Guinea, this paleogeographical feature originated in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62), which resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt.
 The Middle Miocene (16–14 Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the New Guinea ophiolite belt (61, 63).

### 2019-08-27 Manuscript/Manuscript.tex
Updates from Overleaf

@@
     % not changing paleogeography
+    % source term. Can cite: Rowan2016a related to spreading rate.
     % low resolution
-    % model is quite old
+    % GFDL model is quite old
     % kinetics are not changing between lithologies
@@
 culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
-The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a} providing an explanation for the delay between southern hemisphere and bipolar glaciation.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}.
+The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
 We seek to evaluate the hypothesis that the increase of subaerially-exposed land area in Indonesia and New Guinea has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
@@
 Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-However, assessing the effect of the growth of Indonesia and New Guinea on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
+However, assessing the effect of the emergence of Indonesia and New Guinea on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
 The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
-For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady state is achieved at a higher \pCOtwo.
-Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
+For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady-state is achieved at a higher \pCOtwo.
+Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower steady-state \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
 
@@
 In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}.
-These factors taken together are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability and initiating glacial climate over the past 520~m.y.
+These factors are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo and initiating glacial climate over the past 520~m.y.
 \citep{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \citep{Park2019a}.
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability from the increase of subaerially-exposed land area in Indonesia and New Guinea, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in Indonesia and New Guinea, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
 We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
@@
 However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see SI).
-While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of arc-continent collisions in the Malay Archipelago on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
+While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM simplified into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
-XX; see SI).
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM represented by 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+SX).
 Each continental grid cell is then assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in the EarthChem Portal (\url{www.earthchem.org/portal}).
-The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
@@
 Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
-Following the approach of \citet{Maffre2018a}, we then summed the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a} as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}, then calculated the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
-Finally, we eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
-This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present day.
+Following the approach of \citet{Maffre2018a}, we then summed the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+We then calculated the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
+We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
+This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
 
-Having calibrated the  silicate weathering component of GEOCLIM, we estimate the decrease in steady state \pCOtwo associated with increases in global weatherability.
+Having calibrated the silicate weathering component of GEOCLIM, we estimate the decrease in steady-state \pCOtwo associated with increases in global weatherability.
 To do so, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
@@
 
-We use geological and paleoshoreline data to quantify the changes in the area of the broader Indonesia Archipelago over the past 20~m.y (Fig.
+We use geological and paleoshoreline data to quantify the changes in the area of Indonesia and New Guinea over the past 20~m.y (Fig.
 \ref{fig:shoreline_growth}).
@@
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been $sim$500-700~ppm.
+These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been $\sim$500-700~ppm.
 Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca of the lithologies that comprise the Sunda Shelf as well as its low slope.
+These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea is itself sufficient to explain climate change over the Neogene.
 
-These GEOCLIM \pCOtwo values resemble proxy-based estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea is enough to explain climate change over the Neogene.
-Furthermore, our paleoshoreline reconstructions (Fig.
-\ref{fig:shoreline_growth}) show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y.
-that culminated in the initiation of the Northern Hemisphere ice sheets.
+Indonesia and New Guinea are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
+In New Guinea, these paleogeographic features originated in the Middle Miocene with the collision between the Melanesian Arc and the distal margin of Australia \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which resulted in the exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
+The Middle Miocene (16-14~Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the New Guinea ophiolite belt \citep{Cloos2005a}.
+This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
+12~Ma \citep{vanUfford2005a}.
+Exhumation of the Central Range accelerated over the past 4~m.y.
+due to slab-breakoff and buoyant uplift under New Guinea \citep{Cloos2005a}.
+After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin \citep{Baldwin2012a}.
+Over the past 4~m.y., this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi, Halmahera, and northern New Guinea \citep{Baldwin2012a}.
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~m.y.
+\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill & Gleadow, 1989).
 
-XXX
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013a}.
+In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+(Audley-Charles, 1986).
+
+Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an acceleration over the past 4 Ma.
+This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda arc and New Guinea in the past 4 Ma.
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4 Ma that culminated in the initiation of the Northern Hemisphere ice sheets.
+Not only does the timing of paleogeographic changes in the tropics coincide with the cooling trend, but our coupled weathering climate model demonstrates that the magnitude of change in pCO2 due to the reconstructed paleogeographic changes is sufficient to explain climate change over the Neogene.
+
+However, it is important to acknowledge the limitations of the approach taken in this study.
+First, the tested scenarios only simulate changes in the location of the shorelines in Indonesia and New Guinea and the lithology exposed in the Deccan and Ethiopian Traps.
+Changes in paleogeography that are not accounted for in our simulations may have resulted in changes in oceanic and atmospheric circulation \citep{Haug1998a, Shevenell2004a, Holbourn2015a, Molnar2015a} that contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales.
+However, the global extent and multi-million year timescale of the long-term cooling trend since 14~Ma suggests that the primary driver for this cooling is changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+In this way, changes in the flux of \COtwo outgassing via changes in spreading rates and ridge lengths \citep{Rowan2016a} and changes in weatherability via mountain building and the movement of land masses into/out of areas of high temperature and runoff are likely more important paleogeography-related factors in terms of modulating Neogene cooling than changes in oceanic and atmospheric circulation.
+Second, the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a low spatial resolution (2.0\degrees $\times$ 2.5\degrees) and is based upon physics and geochemical cycles that have been superceded by more recent climate models.
+Using an updated climate model at higher resolution may result in runoff and temperature fields with different sensitivities to paleogeography and \pCOtwo.
+Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, it does not yet implement lithology-dependent kinetics of mineral dissolution.
+Nevertheless, relative to felsic lithologies, mafic lithologies more dominantly comprise of minerals with faster kinetics (e.g.
+olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea.
 
@@
 % maybe place this at the end of the discussion
-The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-To explore the sensitivity of global weatherability and climate state to the presence of these traps, we test three additional scenarios using GEOCLIM (Fig.
+The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective traps are replaced with bulk continental crust.
-In the ``removed both traps'' scenario, all volcanics associated with both traps are replaced with bulk continental crust.
+In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with bulk continental crust composition.
+In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with bulk continental crust.
 These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three Indonesia and New Guinea scenarios.
@@
 
-Indonesia and New Guinea are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
-In New Guinea, this paleogeographical feature originated in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62), which resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt.
-The Middle Miocene (16–14 Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the New Guinea ophiolite belt (61, 63).
-This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
-12 Ma (59).
-Mountain building began in the Late Miocene, ca.
-8-7 Ma (59, 62) with 3-4 km of denudation between 8-5 Ma (64).
-Exhumation of the Central Range accelerated over the past 4 Myrs due to slab-breakoff and buoyant uplift under New Guinea (61,65).
-After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin (62).
-Over the past 4 Myr, this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi (67), Halmahera (68, 69), and northern New Guinea (70) (Fig.
-3; sutures 8-10).
-In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4 Ma (van Ufford & Cloos, 2005), which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994; Abbott, 1995) and unroofing of basement-cored uplifts (Hill & Gleadow, 1989).
-
-Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an exceleration over the past 4 Ma.
-This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in New Guinea in the past 4 Ma.
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4 Ma that culminated in the initiation of the Northern Hemisphere ice sheets.
-Not only does the timing of paleogeographic changes in the tropics coincide with the cooling trend, but our coupled weathering climate model demonstrates that the magnitude of change in pCO2 due to the reconstructed paleogeographic changes is sufficient to explain climate change over the Neogene.
+XX
 
@@
 
-We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their contributions to the paleoshoreline reconstructions. FAM and NLS-H thank NSF FRES grant #20191023 for support.
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their contributions to the paleoshoreline reconstructions.
+Collaborative research between NLS-H and YG was initially supported by the France-Berkeley Fund.
+FAM and NLS-H thank NSF FRES grant #20191023 for support.
 
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
-    \caption{Growth of Indonesia and New Guinea.
+    \caption{The emergence of Indonesia and New Guinea.
 Past shorelines are shown in A with resultant area estimates summarized in B.
@@
 1).
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling (5, 6), perhaps associated with increased organic carbon burial (7, 8) or the closure of the  Central American Seaway (9), an increase in the strength of the Walker Circulation associated with the growth of Indonesia (10), feedbacks with terrestrial vegetation (11), a decrease in volcanic outgassing (12), an increase in basaltic watersheds in the tropics (10, 13), uplift in the Himalayas (14), and arc-continent collision in the tropics (15).
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling (5, 6), perhaps associated with increased organic carbon burial (7, 8) or the closure of the  Central American Seaway (9), an increase in the strength of the Walker Circulation associated with the emergence of Indonesia (10), feedbacks with terrestrial vegetation (11), a decrease in volcanic outgassing (12), an increase in basaltic watersheds in the tropics (10, 13), uplift in the Himalayas (14), and arc-continent collision in the tropics (15).
 Although changes in circulation and vegetation may have contributed to these signals on hundreds of thousands to million year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 15 Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback (16-20).

### 2019-08-28 Manuscript/Manuscript.tex

@@
 
-Indonesia and New Guinea have an outsized influence on modern chemical weathering fluxes.
+Indonesia and New Guinea (I\&NG) have an outsized influence on modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a, Hall2017a}; Fig.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system (\citealp{Molnar2015a, Hall2017a}; Fig.
 \ref{fig:shoreline_growth}).
-The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citep{Macdonald2019a}.
+The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of I\&NG and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citep{Macdonald2019a}.
 
-Given that the region is a modern-day hot spot of fresh water and sediment discharge \citep{Milliman2013a}, and chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has the emergence of Indonesia and New Guinea had on Earth's geochemical cycles and climate state?
+Given that the region is a modern-day hot spot of fresh water and sediment discharge \citep{Milliman2013a} and chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has the emergence of I\&NG had on Earth's geochemical cycles and climate state?
 If this region was instead under water, would we have a glacial climate today?
@@
 The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
-We seek to evaluate the hypothesis that the increase of subaerially-exposed land area in Indonesia and New Guinea has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
+We seek to evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
 
 Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-However, assessing the effect of the emergence of Indonesia and New Guinea on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
+However, assessing the effect of the emergence of I\&NG on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
 The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
@@
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in Indonesia and New Guinea, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
+To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
 We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
@@
 However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see SI).
-While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca and Mg concentration in arc rocks relative to other lithologies must also be considered.
+While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM represented by 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
 SX).
-Each continental grid cell is then assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
+Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
 The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
@@
 As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-Rather than prescribing values for these six parameters, we instead allowed these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
-For each of these combinations, we computed spatially-resolved long-term \COtwo consumption (i.e.
+Rather than prescribing values for these six parameters, we instead allow these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
+For each of these combinations, we compute spatially-resolved long-term \COtwo consumption (i.e.
 Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
-Following the approach of \citet{Maffre2018a}, we then summed the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
-We then calculated the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
+Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
 We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
@@
 1860 at a 1\degrees $\times$ 1\degrees resolution.
-In the `1860 control' experiment, forcing agents representative of conditions ca.
+In the ``1860 control'' experiment, forcing agents representative of conditions ca.
 1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
 \pCOtwo in 1860 is assumed to be 286~ppm.
-In the `+1\%/yr to 2$\times$' experiment, initial conditions are taken from the `1860 control' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
+In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
 \pCOtwo is then held constant until the end of the 280 year experiment.
 All non-\COtwo forcing agents are held constant.
-The `+1\%/yr to 4$\times$' experiment is identical to the `+1\%/yr to 2$\times$' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
+The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
 We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
@@
 
-We use geological and paleoshoreline data to quantify the changes in the area of Indonesia and New Guinea over the past 20~m.y (Fig.
+We use geological and paleoshoreline data to quantify the changes in the area of I\&NG over the past 20~m.y.
+(Fig.
 \ref{fig:shoreline_growth}).
@@
 Along with the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, there has also been a decrease in subaerially-exposed land area on the Sunda Shelf \citep{Sarr2019a}.
-Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}, potentially driving climate amplification on orbital timescales (ADD DENEZIO&TIERNY REFS).
+Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}, potentially driving climate amplification on orbital timescales \citep{DiNezio2018a}.
 Similarities in terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
@@
 
-To explore the sensitivity of global weatherability and climate state to the emergence of Indonesia and New Guinea, we test three scenarios using GEOCLIM (Fig.
+To explore the sensitivity of global weatherability and climate state to the emergence of I\&NG, we test three scenarios using GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The ``removed Indonesia and New Guinea'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
-The ``reduced Indonesia and New Guinea'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
+The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
+The ``reduced I\&NG'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
 15~Ma, excluding the Sunda Shelf.
-The ``reduced Indonesia and New Guinea + Sunda Shelf'' is identical to the ``reduced Indonesia and New Guinea'' scenario with the addition of the Sunda Shelf.
+The ``reduced I\&NG + Sunda Shelf'' is identical to the ``reduced I\&NG'' scenario with the addition of the Sunda Shelf.
 Given that no land in the Sunda Shelf is currently exposed, across the shelf we distribute felsic and siliciclastic lithologies, assign shallow slope values, and interpolate runoff across latitude bands (see SI).
 
-Using the 3,339 unique parameter combinations, the ``removed Indonesia and New Guinea'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced Indonesia and New Guinea'' scenario of 547-638~ppm, and the ``reduced Indonesia and New Guinea + Sunda Shelf'' of 533-611~ppm (Fig.
+Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced I\&NG'' scenario of 547-638~ppm, and the ``reduced I\&NG + Sunda Shelf'' of 533-611~ppm (Fig.
 \ref{fig:scenario_pCO2}).
 These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been $\sim$500-700~ppm.
-Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca of the lithologies that comprise the Sunda Shelf as well as its low slope.
-These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea is itself sufficient to explain climate change over the Neogene.
+Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca concentration of the lithologies that comprise the Sunda Shelf as well as its low slope.
+These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is itself sufficient to explain climate change over the Neogene.
 
-Indonesia and New Guinea are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
+I\&NG are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
 In New Guinea, these paleogeographic features originated in the Middle Miocene with the collision between the Melanesian Arc and the distal margin of Australia \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which resulted in the exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
@@
 
-Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an acceleration over the past 4 Ma.
-This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda arc and New Guinea in the past 4 Ma.
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4 Ma that culminated in the initiation of the Northern Hemisphere ice sheets.
-Not only does the timing of paleogeographic changes in the tropics coincide with the cooling trend, but our coupled weathering climate model demonstrates that the magnitude of change in pCO2 due to the reconstructed paleogeographic changes is sufficient to explain climate change over the Neogene.
+Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
+This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda Arc and New Guinea in the past 4~m.y.
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
+Not only does the timing of the increase in subaerially-exposed land area of I\&NG coincide with the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
 
 However, it is important to acknowledge the limitations of the approach taken in this study.
-First, the tested scenarios only simulate changes in the location of the shorelines in Indonesia and New Guinea and the lithology exposed in the Deccan and Ethiopian Traps.
+First, the tested scenarios only simulate changes in the location of the shorelines in I\&NG and the lithology exposed in the Deccan and Ethiopian Traps.
 Changes in paleogeography that are not accounted for in our simulations may have resulted in changes in oceanic and atmospheric circulation \citep{Haug1998a, Shevenell2004a, Holbourn2015a, Molnar2015a} that contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales.
@@
 
+%%% RELATIVELY COMPLETE TEXT ABOVE - BELOW ARE SNIPPETS THAT COULD BE INCORPORATED IN
+
 %%%
@@
 \ref{fig:scenario_pCO2}).
-In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with bulk continental crust composition.
-In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with bulk continental crust.
-These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three Indonesia and New Guinea scenarios.
+In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with the Ca+Mg concentration of bulk continental crust.
+In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with the Ca+Mg concentration of bulk continental crust.
+These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
 %%%
@@
 Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of Indonesia and New Guinea.
-
-XX
 

### 2019-08-28 Manuscript/Manuscript.tex

@@
 We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
-We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1\times10$^{12}$~mol/yr or \textgreater1\times10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
+We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1$\times$10$^{12}$~mol/yr or \textgreater1$\times$10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
 This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
@@
 Taken together, these constraints delineate the position of shorelines.
-For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norvick2003a, Harrington2017a, Gold2017a}.
+For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norwick2003a, Harrington2017a, Gold2017a}.
 Although there are difference in data sources and paleoshoreline outlines from \cite{Molnar2015a}, the calculated area at 5~Ma is similar.
@@
 In the Miocene, this region exposed a mixture of felsic basement rocks and siliciclastic sediments \citet{Darmadi2007a}.
-The relatively low-relief and felsic lithologies of the region contrasts with the high-relief and mafic lithologies of most of Indonesia and New Guinea.
+The relatively low-relief and felsic lithologies of the region contrasts with the high-relief and mafic lithologies of most of I\&NG.
 
@@
 In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~m.y.
-\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill & Gleadow, 1989).
+\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill \& Gleadow, 1989).
 
@@
 
+The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM (Fig.
+\ref{fig:scenario_pCO2}).
+In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with the Ca+Mg concentration of bulk continental crust.
+In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with the Ca+Mg concentration of bulk continental crust.
+These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
+This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in low-latitude igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is significantly smaller than that of high-relief I\&NG.
+
 However, it is important to acknowledge the limitations of the approach taken in this study.
@@
 Nevertheless, relative to felsic lithologies, mafic lithologies more dominantly comprise of minerals with faster kinetics (e.g.
-olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of Indonesia and New Guinea.
+olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
 
@@
 %%%
-% maybe place this at the end of the discussion
-The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM (Fig.
-\ref{fig:scenario_pCO2}).
-In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with the Ca+Mg concentration of bulk continental crust.
-In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with the Ca+Mg concentration of bulk continental crust.
-These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
-%%%
 
-%%%
-
-After the MCO, a cooling trend began ca.
+%After the MCO, a cooling trend began ca.
 14~Ma that accelerated over the past 4~m.y.
@@
 Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
-Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of Indonesia and New Guinea.
+Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of I\&NG.
 
@@
 \end{document}
-
-Neogene cooling and the emergence of Indonesia and New Guinea
-
-Abstract
-The cooling trend from 15 Ma to present coincides with arc-continent collision and exhumation in Indonesia and New Guinea.
-To assess relationships between the emergence of highly weatherable arc terranes with tectonic topography in the tropics, an increase in global weatherability, geochemical records, and climate change, we use geological and paleoshoreline data quantify the changes in the area of volcanic and ultramafic rocks in Indonesia and New Guinea over the past 20 Myrs.
-We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics.
-We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.
-Arc-continent collision has resulted in the exhumation of both the East Sulawesi and Irian Jaya-New Guinea ophiolites, which are two of the three largest preserved ophiolites on Earth.
-We suggest that cooling from 15 to 10 Ma was due in large part to the subaerial exposure of these ophiolites and increase in global weatherability, and that the more recent acceleration in cooling was due to increased exhumation rates in Indonesia and New Guinea.
-Both a decrease of mid ocean ridge spreading rates and changes in the composition and amount of riverine input is necessary to explain both global cooling and geochemical proxy data.
-The data presented herein helps disentangle these effects.
-
-Significance statement
-	The tectonic drivers for the Neogene cooling trend from ~15 Ma to present are unknown.
-Here we quantify paleoshoreline data in Indonesia and New Guinea over the Neogene to assess the hypothesis that cooling was caused by changes in paleogeography in the warm wet tropics.
-These data document the emergence of Indonesia over this interval and are consistent with changes in global weatherability associated with arc-continent collision in the  tropics as a major driver of long-term climate change.
-
-\body
-After the Miocene Climatic Optimum (MCO), a cooling trend began at ~15 Ma, and accelerated over the past 5 Ma (1-4) (Fig.
-1).
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling (5, 6), perhaps associated with increased organic carbon burial (7, 8) or the closure of the  Central American Seaway (9), an increase in the strength of the Walker Circulation associated with the emergence of Indonesia (10), feedbacks with terrestrial vegetation (11), a decrease in volcanic outgassing (12), an increase in basaltic watersheds in the tropics (10, 13), uplift in the Himalayas (14), and arc-continent collision in the tropics (15).
-Although changes in circulation and vegetation may have contributed to these signals on hundreds of thousands to million year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 15 Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback (16-20).
-On geological timescales, atmospheric CO2 levels can vary due to changes in volcanic CO2 flux, but also due to changes in global weatherability — the cumulative factors that affect chemical weathering aside from climate (20, 21).
-Global weatherability is the product of variables such as lithology, tectonic uplift rates, and paleolatitude (20).
-An increase in global weatherability can result in cooling as the consumption of carbon through silicate weathering will match that of volcanic input at lower concentrations of atmospheric CO2 (20).
-Present-day global CO2-consumption estimates emphasize the importance of highly weatherable areas: ~10-20% of land area is responsible for ~50-75% of CO2 consumption through silicate weathering and associated carbonate precipitation (22, 23).
-These highly weatherable areas (e.g.
-Indonesia and New Guinea) are mainly situated in the warm, wet tropics with significant topographic relief, and have bedrock composed of Ca- and Mg-rich mafic and ultramafic lithologies, which have faster dissolution rates than felsic rocks (20, 23, 24).
-As a result of the intertropical convergence zone (ITCZ), and its migration through the year, annual rainfall on Earth is greatest within ~15º of the equator resulting in the tropical rain belt.
-Given that hemispherically asymmetric forcings are relatively small, the position of the ITCZ is interpreted to be stable even with large changes in forcing resulting from changing CO2 levels (25).
-In modern climate data, the region within 10º of either side of the equator corresponds with precipitation greater than 1.5 m/yr and the region within 15º corresponds with precipitation greater than 1.0 m/yr.
-Because of the strong relationship between runoff and silicate weathering (26), paleogeographic changes in tropics, particularly in tropical regions with tectonic topography and abundant volcanic rocks, like Indonesia and New Guinea, should have an outsized effect on the Earth’s long-term climate state (10, 15).
-Recently, it was proposed that arc-continent collision in the tropics drove the cooling trend over the past 15 million years (15, 27).
-This hypothesis builds off of the suggestion of Reusch and Maasch (28) that arc-continent collisions could accomplish both a decrease in outgassing by terminating arc volcanism (footnote: However, arc-continent collision generally results in subduction flip with renewal of arc magmatism (29)), and an increase in silicate weathering by exhuming soluble Ca- and Mg-rich silicate rocks.
-Jagoutz et al.
-(27) applied this process specifically to the closure of the Neo-Tethys ocean in the tropics and suggested that arc-continent collisions at approximately 80 and 50 Ma could account for the long-term, two-phased cooling trend from the Cretaceous to the E-O boundary.
-Importantly, Jagoutz et al.
-(27) combined the Himalayan uplift hypothesis (14) with the basalt drift hypothesis (13, 30), emphasizing that topography, precipitation in the tropical rain belt, and lithology are the key variables in global weatherability.
-Macdonald et al.
-(15) compiled paleolatitude data on ophiolite bearing sutures through the Phanerozoic and concluded that the extent of arc-continent collisions tropics correlates with extent of continental glaciation, and specifically, that arc-continent collisions in Indonesia and New Guinea were associated with cooling over the past 15 Myr.
-Due to rugged tectonic topography (Fig.
-2A), tropical temperatures and precipitation rates (Fig.
-2B), and a predominance of highly soluble mafic and ultramafic rocks (Fig.
-3A), watersheds in Indonesia and New Guinea have the highest rates of silicate weathering in the world accounting for ~9% of the modern global carbon sink (23, 26).
-Although there are significant monsoonal and orographic effects in Indonesia, and particularly in New Guinea (Fig.
-2A), the high mean annual rainfall sill results in rivers with the highest average silicate weathering rates (26).
-Recognizing the importance of the paleogeography of this region, Molnar and Cronin (10) used paleoshoreline data to estimate that the area of subaerially exposed landmass in Indonesia and New Guinea increased by 60% over the last 5 Myrs.
-Here we quantify the changes in exposed area and rock types in the tropical region of Indonesia and New Guinea over the past 20 Myrs in a paleogeographic model.
-These results allow us to assess the changes of mafic and ultramafic rocks in the warm wet tropics of Indonesia and New Guinea throughout the Neogene and test the hypothesis that the subaerial emergence of this region associated with arc-continent collision caused an increase in global weatherability and cooling.
-These data can further be incorporated into future coupled climate and weathering models to assess the effect of changing paleogeography and global weatherabilty on long-term climate state.
-
-Tectonic Evolution
-During Neogene arc-continent collision between the Australian-Indian and Eurasian plates and intervening arc terranes (31-33), Indonesia and New Guinea emerged as several large supra-subduction ophiolites and associated forearcs were exhumed in the warm, wet tropics.
-Some of these ophiolites formed immediately prior to collision and others have an older history.
-The oldest ophiolites preserved in the Indonesia orogenic system are situated in the northwest and associated with the easternmost Paleo-Tethyan sutures (34).
-On the Malay Peninsula, the Bentong-Raub belt (suture 1; Fig.
-3) demarcates the eastern margin of Cimmerian ribbon continent Sibumasu, which collided with the Suhothai arc in the latest Permian to early Triassic (34).
-Thermochronology suggests that the  Malay Peninsula was first exhumed during the mid-Eocene to early Oligocene (35).
-This region became a significant sediment source for the Crocker Fan on the NW margin of Borneo during the Oligocene (36), as well as to the east of Malay Peninsula during the Neogene, and has been largely exposed over the last 20 Ma (37, 38).
-In Sumatra, the Wolya arc (suture 2; Fig.
-3) formed at 95 Ma and is overlain by Campanian 83.5-70.6 Ma radiolarian cherts (39), which were uplifted and eroded from 60-40 Ma (40).
-These ophiolites were entrained in the Banda Arc, which has experienced magmatism and accretion throughout most of the Cenozoic as the Australian plate subducts beneath it (37, 41).
-From ~20-10 Ma, the majority of Sumatra was submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20 Ma (37, 41).
-Most of Sumatra was submerged until after 5 Ma with the arrival of the leading edge of continental Australia (37, 41).
-The Meratus belt (sutures 3 and 4; Fig.
-3), which extends from Java (42) through eastern Borneo (43) to Palawan (44), marks the Late Cretaceous collision of the Banda blocks and Argo.
-After this collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed (45), which we refer to here as the Great Indonesian arc.
-In southern Borneo, sedimentological data in the adjacent Barito Basin suggests that ophiolites in the Meratus Mountains were not emergent until the Late Miocene (46).
-In northern Borneo and Palawan, collision occurred in the Late Oligocene to Early Miocene, resulting in exhumation of the Palawan Ophiolite Complex (44, 46).
-Over 10 km of Neogene sediment fills the Sarawak Basin and NW Borneo trough, indicative of extensive exhumation and erosion in northern Borneo (36).
-Thermochronological data suggest extremely rapid uplift (7 km/Myr) and exhumation during the Late Miocene and Early Pliocene, resulting in the recent formation of Mt.
-Kinabalu, the tallest peak in Borneo (47).
-The East Sulawesi ophiolite (suture 5; Fig.
-3) was likely generated in a supra-subduction to backarc setting in the Great Indonesian arc (45, 48, 49) [but see (50) for an alternative view], detached from its metamorphic sole in the late Oligocene (32-28 Ma; 51), and subsequently thrust to the east above a west dipping slab.
-During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing additional ophiolite emplacement in west and southeast Sulawesi, including Buton (52, 53).
-Uplift was diachronous, not effecting units is western Sulawesi until the Middle Miocene (~13 Ma) (54).
-Exhumation is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins (52, 54).
-Collision of the Sula-Banggai block with the east margin of Sulawesi began in the Late Miocene and uplift accelerated in the Early Pliocene (5.2 to 3.8 Ma) and is associated with a major pulse of sedimentation in adjacent basins (55, 56).
-Off the northeastern margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes mélange (55).
-Thus, although the East Sulawesi ophiolite was trapped/emplaced onto the composite Sundaland margin (i.e.
-a fragment of previously accreted Australian crust) between 36 and 23 Ma, mafic and ultramafic rocks appear to have not been substantially subaerially exposed in the south and west arms until after 15 Ma, and on the east arm until after 5 Ma.
-Thermochronology indicates that exhumation accelerated in the Late Miocene to Pliocene (52-54, 57).
-North Sulawesi, where active volcanoes extend through the Sangihe Arc into the southern Philippines, was not substantially emergent until the Pliocene (58).
-Along the Papuan Peninsula of Eastern New Guinea, Oligocene arc-continent collision between Australian continental fragments and the Inner Melanesian arc exhumed the Papuan Ultramafic Belt (Fig.
-3; suture 6) above a north dipping slab (59, 60).
-In the Central Range of New Guinea, a second set of arc-continent collisions in New Guinea began in the Middle Miocene with the collision between the Outer Melanesian arc and the distal margin of Australia (59, 61, 62).
-This resulted in the exhumation of the 1440 km long Irian-Marum-April ophiolite belt (Fig.
-3; suture 7).
-The Middle Miocene (16–14 Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the PNG ophiolite belt (61, 63).
-This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
-12 Ma (59).
-Mountain building began in the Late Miocene, ca.
-8-7 Ma (59, 62) with 3-4 km of denudation between 8-5 Ma (64), but major relief was not generated until the Pliocene (65).
-Exhumation of the Central Range accelerated over the past 5 Myrs due to slab-breakoff and buoyant uplift under New Guinea (61).
-Although the Papuan Ultramafic Belt was generated and obducted earlier than the ophiolites in the Central Range (60, 66), it was also exhumed very rapidly over the past 10 Ma (62).
-After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin (62).
-Over the past 4 Myr, this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi (67), Halmahera (68, 69), and northern New Guinea (70) (Fig.
-3; sutures 8-10).
-In the Bird’s Head region, this change in plate motion led to CCW rotation and the exhumation of several core complexes (71).
-The Philippine ophiolites have been grouped in four distinct belts (72).
-We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone (Fig.
-3; sutures 11 and 12).
-Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments (73).
-The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present (74).
-In addition to collision, arc magmatism may have contributed more significantly to crustal growth in the Philippines (75).
-The youngest ophiolites of Indonesia formed in the Banda Sea starting at ca.
-15 Ma (33).
-The ongoing arc-continent collision became emergent in Seram and Timor (Fig.
-3; sutures 13 and 14) over the last 5 Myr, as the Banda arc began to abut the Banda embayment emplacing the Timor nappe (49, 76, 77).
-
-Paleoshorelines
-Geological maps and paleoshoreline data were used to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea.
-Following Molnar and Cronin (10), we analyzed the area changes of islands that are greater than ~200 km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda shelf.
-Larger islands were further divided into regions based upon their position on microplates as defined by Matthews, et al.
-(78).
-Lithological maps were modified from Hartmann and Moosdorf (79) and then supplemented with regional geological maps.
-From these maps, the percentage of ultramafic and volcanic rocks relative to other rock types was calculated in QGIS for each island and microplate region.
-Generally, the paleoshoreline reconstructions within the paleogeographic models coincide with sub-Epoch boundaries, i.e.
-Early Miocene = 23-16 Ma, Middle Miocene = 15.99-11.61 Ma, Late Miocene = 11.6-5.4 Ma, Pliocene =5.39-3.8 Ma.
-However, some additional shorelines have been added where additional age information is present to refine the timing of major changes in area.
-Miocene to present stratigraphic columns were compiled from each region to develop an age model for various paleoenvironmental indicators.
-In general, carbonate strata and thin-bedded siliciclastic strata were assumed to be deposited in a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as evidence for subaerial environment.
-Following previous work on paleoshorelines and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma), Middle Miocene (15 Ma), and Early Miocene (20 Ma).
-We used geological maps and paleoshoreline data to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea.
-Using Gplates, paleoshoreline shapefiles were incorporated in the paleogeographic model of Matthews, et al.
-(78).
-Although there are some differences between this reconstruction and the recent reconstruction of Hall (33), these do not entail significant changes in paleolatitude.
-Many new sources for paleoshoreline data have become available since Molnar & Cronin (10).
-Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by Nugraha and Hall (80), and in New Guinea, we followed Gold et al.
-(81),  Cloos et al.
-(61), Norvick (2003) and derivative work by Harrington et al.
-(82), and Nichols and Hall (1991).
-We also compiled our own stratigraphies for the region to provide further paleoenvironmental context.
-Although the outlines of our paleoshorelines are very different from those of Molnar & Cronin (10) our calculated area for 5 Ma is not substantially different (Fig.
-X).
-A critical issue for the paleogeographic  analysis of Indonesia is how to deal with the Sunda Shelf.
-Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (38).
-From a biogeographical standpoint, similarities between the terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
-During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
-Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially-enclosed inland seas and extensive brackish or saline wetland environments.
-Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley 2013).
-Basement highs (e.g.
-the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al, 2007).
-Although most paleogeographic reconstructions of the region incorporate some degree of constant exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al, 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
-We test the sensitivity with two separate scenarios in which the Sunda Shelf is either variably exposed or entirely inundated from 20 Ma to present.
-Independent of the Sunda Shelf, our studies suggest that the volume of sediment eroded from Indonesia and New Guinea has exponentially increased over the past 15 Ma, particularly in New Guinea.DISCUSS CHANGES FOR EACH REGION.
-
-References
-1.	Zachos J, Pagani M, Sloan L, Thomas E, & Billups K (2001) Trends, Rhythms, and Aberrations in the Global Climate 65 Ma to present. Science 292:686-693.
-2.	Cramer B, Miller K, Barrett P, & Wright J (2011) Late Cretaceous–Neogene trends in deep ocean temperature and continental ice volume: Reconciling records of benthic foraminiferal geochemistry (δ18O and Mg/Ca) with sea level history.
-Journal of Geophysical Research: Oceans 116(C12).
-3.	Herbert TD, et al. (2016) Late Miocene global cooling and the rise of modern ecosystems. Nature Geoscience 9(11):843-847.
-4.	Zachos JC, Dickens GR, & Zeebe REJN (2008) An early Cenozoic perspective on greenhouse warming and carbon-cycle dynamics. 451(7176):279.
-5.	Shevenell AE, Kennett JP, & Lea DW (2004) Middle Miocene southern ocean cooling and Antarctic cryosphere expansion. Science 305(5691):1766-1770.
-6.	Holbourn A, Kuhnt W, Kochhann KG, Andersen N, & Sebastian Meier KJG (2015) Global perturbation of the carbon cycle at the onset of the Miocene Climatic Optimum. 43(2):123-126.
-7.	Vincent E & Berger W (1985) Carbon dioxide and polar cooling in the Miocene: The Monterey hypothesis.
-AGU Geophysical Mongograph Series.
-The Carbon Cycle and Atmospheric CO2: Natural Variations from Archean to Present 32:455-468.
-8.	Flower BP & Kennett JPJG (1993) Relations between Monterey Formation deposition and middle Miocene global cooling: Naples Beach section, California. 21(10):877-880.
-9.	Haug GH & Tiedemann RJN (1998) Effect of the formation of the Isthmus of Panama on Atlantic Ocean thermohaline circulation. 393(6686):673.
-10.	Molnar P & Cronin TW (2015) Growth of the Maritime Continent and its possible contribution to recurring Ice Ages. Paleoceanography 30(3):196-225.
-11.	Cerling TE, et al. (1997) Global vegetation change through the Miocene/Pliocene boundary. Nature 389(6647):153-158.
-12.	Berner RA, Lasaga AC, & Garrels RM (1983) The carbonate-silicate geochemical cycle and its effect on atmospheric carbon dioxide over the past 100 million years.
-American Journal of Science 283:641-683.
-13.	Kent DV & Muttoni G (2013) Modulation of Late Cretaceous and Cenozoic climate by variable drawdown of atmospheric pCO2 from weathering of basaltic provinces on continents drifting through the equatorial humid belt.
-Climate of the Past 9(2):525.
-14.	Raymo ME, Ruddiman WF, & Froelich PN (1988) Influence of late Cenozoic mountain building on ocean geochemical cycles. Geology 16(7):649-653.
-15.	Macdonald FA, Swanson-Hysell NL, Park Y, & Lissiecki LE (2019) Arc-continent collisions in the tropics set the Earth’s climate state. Science 364:181-184.
-16.	Walker JCG, Hays PB, & Kasting JF (1981) A negative feedback mechanism for the long-term stabilization of the earth's surface temperature. Geophysical Research Letters 86:9776-9782.
-17.	Raymo ME (1991) Geochemical evidence supporting TC Chamberlin's theory of glaciation. Geology 19(4):344-347.
-18.	Berner RA (1994) Geocarb II: a revised model of atmospheric CO2 over Phanerozoic time. American Journal of Science 294:56-91.
-19.	Berner RA & Caldeira K (1997) The need for mass balance and feedback in the geochemical carbon cycle. Geology 25(10):955-956.
-20.	Kump LR & Arthur MA (1997) Global chemical erosion during the Cenozoic: Weatherability balances the budgets. Tectonic Uplift and Climate Change,  (Springer), pp 399-426.
-21.	Maher K & Chamberlain C (2014) Hydrologic regulation of chemical weathering and the geologic carbon cycle. Science 343(6178):1502-1504.
-22.	Hartmann J, Moosdorf N, Lauerwald R, Hinderer M, & West AJ (2014) Global chemical weathering and associated P-release—The role of lithology, temperature and soil properties.
-Chemical geology 363:145-163.
-23.	Dessert C, Dupré B, Gaillardet J, François LM, & Allegre CJ (2003) Basalt weathering laws and the impact of basalt weathering on the global carbon cycle. Chemical Geology 202(3):257-273.
-24.	Ibarra DE, et al. (2016) Differential weathering of basaltic and granitic catchments from concentration–discharge relationships. 190:265-293.
-25.	Donohoe A, Voigt AJCEP, & Mechanisms (2017) Why Future Shifts in Tropical Precipitation Will Likely Be Small: The Location of the Tropical Rain Belt and the Hemispheric Contrast of Energy Input to the Atmosphere.115-137.
-26.	Gaillardet J, Dupré B, Louvat P, & Allegre C (1999) Global silicate weathering and CO 2 consumption rates deduced from the chemistry of large rivers. Chemical geology 159(1):3-30.
-27.	Jagoutz O, Macdonald FA, & Royden L (2016) Low-latitude arc–continent collision as a driver for global cooling. Proceedings of the National Academy of Sciences 113(18):4935-4940.
-28.	Reusch DN & Maasch KA (1998) The transition from arc volcanism to exhumation, weathering of young Ca, Mg, Sr silicates, and CO2 drawdown. Oxford Monographs on Geology and Geophysics 39:261-276.
-29.	McKenzie DP (1969) Speculations on the consequences and causes of plate motions. Geophysical Journal International 18(1):1-32.
-30.	Kent DV & Muttoni G (2008) Equatorial convergence of India and early Cenozoic climate trends. Proceedings of the National Academy of Sciences 105(42):16065-16070.
-31.	Hall R (1996) Reconstructing Cenozoic SE Asia. Geological Society, London, Special Publications 106(1):153-184.
-32.	Hamilton WB (1979) Tectonics of the Indonesian region (US Govt. Print. Off.).
-33.	Hall R (2017) Southeast Asia: New Views of the Geology of the Malay Archipelago. Annual Review of Earth and Planetary Sciences 45(1).
-34.	Metcalfe I (2013) Gondwana dispersion and Asian accretion: tectonic and palaeogeographic evolution of eastern Tethys. Journal of Asian Earth Sciences 66:1-33.
-35.	Cottam MA, Hall R, & Ghani AAJJoAES (2013) Late Cretaceous and Cenozoic tectonics of the Malay Peninsula constrained by thermochronology. 76:241-257.
-36.	Hall R, van Hattum MW, & Spakman WJT (2008) Impact of India–Asia collision on SE Asia: the record in Borneo. 451(1-4):366-389.
-37.	Hall R (2013) The palaeogeography of Sundaland and Wallacea since the Late Jurassic. Journal of Limnology 72(s2):1.
-38.	Hall R & Nichols GJGS, London, Special Publications (2002) Cenozoic sedimentation and tectonics in Borneo: climatic influences on orogenesis. 191(1):5-22.
-39.	Pedersen R, Searle M, Carter A, & Bandopadhyay P (2010) U–Pb zircon age of the Andaman ophiolite: implications for the beginning of subduction beneath the Andaman–Sumatra arc.
-Journal of the Geological Society 167(6):1105-1112.
-40.	Allen R, et al. (2008) New constraints on the sedimentation and uplift history of the Andaman-Nicobar accretionary prism, South Andaman Island. 436:223.
-41.	Hall R (2009) The Eurasian SE Asian margin as a modern example of an accretionary orogen. Geological Society, London, Special Publications 318(1):351-372.
-42.	Wakita K (2000) Cretaceous accretionary–collision complexes in central Indonesia. Journal of Asian Earth Sciences 18(6):739-749.
-43.	Monnier C, et al. (1999) Extensional to compressive Mesozoic magmatism at the SE Eurasia margin as recorded from the Meratus ophiolite (SE Borneo, Indonesia). Geodinamica Acta 12(1):43-55.
-44.	Yumul Jr GP, Dimalanta CB, Marquez EJ, & Queaño KL (2009) Onland signatures of the Palawan microcontinental block and Philippine mobile belt collision and crustal growth process: a review.
-Journal of Asian Earth Sciences 34(5):610-623.
-45.	Hall R (2002) Cenozoic geological and plate tectonic evolution of SE Asia and the SW Pacific: computer-based reconstructions, model and animations. Journal of Asian Earth Sciences 20(4):353-431.
-46.	Witts D, Hall R, Nichols G, & Morley R (2012) A new depositional and provenance model for the Tanjung Formation, Barito Basin, SE Kalimantan, Indonesia. Journal of Asian Earth Sciences 56:77-104.
-47.	Cottam M, et al. (2013) Neogene rock uplift and erosion in northern Borneo: evidence from the Kinabalu granite, Mount Kinabalu.2011-2130.
-48.	Monnier C, Girardeau J, Maury RC, & Cotten J (1995) Back-arc basin origin for the East Sulawesi ophiolite (eastern Indonesia). Geology 23(9):851-854.
-49.	Harris R (2006) Rise and fall of the Eastern Great Indonesian arc recorded by the assembly, dispersion and accretion of the Banda Terrane, Timor. Gondwana Research 10(3-4):207-231.
-50.	Kadarusman A, Miyashita S, Maruyama S, Parkinson CD, & Ishikawa A (2004) Petrology, geochemistry and paleogeographic reconstruction of the East Sulawesi Ophiolite, Indonesia.
-Tectonophysics 392(1):55-83.
-51.	Parkinson C (1998) Emplacement of the East Sulawesi Ophiolite: evidence from subophiolite metamorphic rocks. Journal of Asian Earth Sciences 16(1):13-28.
-52.	Bergman SC, Coffield DQ, Talbot JP, & Garrard RA (1996) Tertiary tectonic and magmatic evolution of western Sulawesi and the Makassar Strait, Indonesia: evidence for a Miocene continent-continent collision.
-Geological Society, London, Special Publications 106(1):391-429.
-53.	Smith RB & Silver EA (1991) Geology of a Miocene collision complex, Buton, eastern Indonesia. Geological Society of America Bulletin 103(5):660-678.
-54.	van Leeuwen TM, Susanto ES, Maryanto S, & Hadiwisastra S (2010) Tectonostratigraphic evolution of Cenozoic marginal basin and continental margin successions in the Bone Mountains, Southwest Sulawesi, Indonesia.
-Journal of Asian Earth Sciences 38(6):233-254.
-55.	Davies I (1990) Geological and exploration review of the Tomori PSC, Eastern Indonesia.
-56.	Villeneuve M, et al.
-(2000) Continental block collision in the eastern arm of Sulawesi (Indonesia).
-Structure and geodynamic interpretation.
-Comptes Rendus de l'Académie des Sciences-Series IIA-Earth and Planetary Science 330(5):371-378.
-57.	Hennig J, Hall R, Forster MA, Kohn BP, & Lister GS (2017) Rapid cooling and exhumation as a consequence of extension and crustal thinning: Inferences from the Late Miocene to Pliocene Palu Metamorphic Complex, Sulawesi, Indonesia.
-Tectonophysics 712:600-622.
-58.	van Leeuwen TM & Muhardjo (2005) Stratigraphy and tectonic setting of the Cretaceous and Paleogene volcanic-sedimentary successions in northwest Sulawesi, Indonesia: implications for the Cenozoic evolution of Western and Northern Sulawesi.
-Journal of Asian Earth Sciences 25(3):481-511.
-59.	van Ufford AQ & Cloos M (2005) Cenozoic tectonics of New Guinea. AAPG bulletin 89(1):119-140.
-60.	Davies H & Jaques A (1984) Emplacement of ophiolite in Papua New Guinea. Geological Society, London, Special Publications 13(1):341-349.
-61.	Cloos M, et al. (2005) Collisional delamination in New Guinea: The geotectonics of subducting slab breakoff. Geological Society of America Special Papers 400:1-51.
-62.	Baldwin SL, Fitzgerald PG, & Webb LE (2012) Tectonics of the New Guinea region. Annual Review of Earth and Planetary Sciences 40.
-63.	Visser WA & Hermes JJ (1962) Geological results of the exploration for oil in Netherlands New Guinea: carried out by the'Nederlandsche Nieuw Guinee Petroleum Maatschappij'1935-1960 (Staatsdrukkerij-en Uitgeverijbedrijf).
-64.	Crowhurst P, Hill K, Foster D, & Bennett A (1996) Thermochronological and geochemical constraints on the tectonic evolution of northern Papua New Guinea.
-Geological Society, London, Special Publications 106(1):525-537.
-65.	Weiland RJ & Cloos M (1996) Pliocene-Pleistocene asymmetric unroofing of the Irian fold belt, Irian Jaya, Indonesia: Apatite fission-track thermochronology.
-Geological Society of America Bulletin 108(11):1438-1449.
-66.	Davies HL & Smith IE (1971) Geology of eastern Papua. Geological Society of America Bulletin 82(12):3299-3312.
-67.	Ali J, Hall R, & Baker S (2001) Palaeomagnetic data from a Mesozoic Philippine Sea plate ophiolite on Obi Island, eastern Indonesia. Journal of Asian Earth Sciences 19(4):535-546.
-68.	Hall R, Audley-Charles M, Banner F, Hidayat S, & Tobing S (1988) Basement rocks of the Halmahera region, eastern Indonesia: a Late Cretaceous–early Tertiary arc and fore-arc.
-Journal of the Geological Society 145(1):65-84.
-69.	Ballantyne P (1992) Petrology and geochemistry of the plutonic rocks of the Halmahera ophiolite, eastern Indonesia, an analogue of modern oceanic forearcs.
-Geological Society, London, Special Publications 60(1):179-202.
-70.	Monnier C, et al.
-(1999) Petrology and geochemistry of the Cyclops ophiolites (Irian Jaya, East Indonesia): consequences for the Cenozoic evolution of the north Australian margin.
-Mineralogy and Petrology 65(1-2):1-28.
-71.	Sapin F, Pubellier M, Ringenbach J-C, & Bailly VJJoSG (2009) Alternating thin versus thick-skinned decollements, example in a fast tectonic setting: The Misool–Onin–Kumawa Ridge (West Papua).
-31(4):444-459.
-72.	Yumul GP (2007) Westward younging disposition of Philippine ophiolites and its implication for arc evolution. Island Arc 16(2):306-317.
-73.	Pubellier M, et al. (1991) The Mindanao collision zone: a soft collision event within a continuous Neogene strike-slip setting. Journal of Southeast Asian Earth Sciences 6(3-4):239-248.
-74.	Yumul Jr GP, Dimalanta CB, Tamayo Jr RA, & Faustino-Eslava DV (2013) Geological features of a collision zone marker: the Antique Ophiolite Complex (Western Panay, Philippines).
-Journal of Asian Earth Sciences 65:53-63.
-75.	Dimalanta C & Yumul G (2003) Magmatic and amagmatic contributions to crustal growth of an island-arc system: The Philippine example. International Geology Review 45(10):922-935.
-76.	Pownall JM, Forster MA, Hall R, & Watkinson IM (2017) Tectonometamorphic evolution of Seram and Ambon, eastern Indonesia: Insights from 40Ar/39Ar geochronology. Gondwana Research 44:35-53.
-77.	Monnier C, et al. (2003) Dynamics and age of formation of the Seram-Ambon ophiolites (Central Indonesia). Bulletin de la Société géologique de France 174(6):529-543.
-78.	Matthews KJ, et al. (2016) Global plate boundary evolution and kinematics since the late Paleozoic. Global and Planetary Change 146:226-250.
-79.	Hartmann J & Moosdorf N (2012) The new global lithological map database GLiM: A representation of rock properties at the Earth surface. Geochemistry, Geophysics, Geosystems 13(12).
-80.	Nugraha AMS & Hall R (2018) Late Cenozoic palaeogeography of Sulawesi, Indonesia. Palaeogeography, Palaeoclimatology, Palaeoecology 490:191-209.
-81.	Gold DP, White LT, Gunawan I, BouDagher-Fadel MJM, & Geology P (2017) Relative sea-level change in western New Guinea recorded by regional biostratigraphic data. 86:1133-1158.
-82.	Harrington L, Zahirovic S, Flament N, Müller RDJE, & Letters PS (2017) The role of deep Earth dynamics in driving the flooding and emergence of New Guinea since the Jurassic. 479:273-283.
-
-Figure Legends
-Fig. 1. Oxygen isotopic record of Neogene cooling. Benthic foraminifera 18O stack in black from Zachos et al. (4), with grey envelope of bottom water temperature estimates from Cramer et al. (2).
-
-Fig. 2. A) Slope map, showing active volcanoes, trenches and areas with high topography. B) Precipitation map over Indonesia (ADD NOAA reference).
-
-Fig. 3. A) Distribution of bedrock, mafic and ultramafic rock, of SE Asia from Hartmann and Moosdorf (79). Plate outlines from Mathews et al. (2016). B) Paleoshorelines shown for 20, 15, 10, and 5 Ma.
-
-Fig. 4. Paleogeogrpahic reconstructions of subaerially exposed area of Indonesia at 20, 15, 10, and 5 Ma.
-
-Fig. 5. Calculated change in total area, area in the tropics, and area of ultramafic and volcanic rocks in the tropics over that last 20 Ma in Indonesia and New Guinea.
-
-Fig. 6. Strontium and Osmium isotope model.

### 2019-08-29 Manuscript/Manuscript.tex

@@
 Collaborative research between NLS-H and YG was initially supported by the France-Berkeley Fund.
-FAM and NLS-H thank NSF FRES grant #20191023 for support.
-
-\clearpage
-\newpage
-
-\section*{TABLES}
-
-XXX
+FAM and NLS-H thank NSF FRES grant \#20191023 for support.
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
+    \includegraphics[width=0.9\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
     \caption{The emergence of Indonesia and New Guinea.
 Past shorelines are shown in A with resultant area estimates summarized in B.
-A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \cite{Zachos2008a}.}
+A significant increase in area over the past 4 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \citep{Zachos2008a}.}
     \label{fig:shoreline_growth}
@@
 
-\begin{figure}
+\begin{figure}[h!]
     \centering

### 2019-08-29 Manuscript/Manuscript.tex

@@
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
-SX).
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
 Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
@@
 The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM.
 
@@
 In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~m.y.
-\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill \& Gleadow, 1989).
+\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
 
@@
 In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
-(Audley-Charles, 1986).
+\citep{AudleyCharles1986a}.
 
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2.pdf}
-    \caption{Box shows quartiles, whiskers show 1.5$\times$ the interquartile range, notch shows median with 95\% confidence interval.
-Each dot shows a parameter combination.
-Dashed line is PI \pCOtwo.}
+    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
+For each of the six scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched measured \COtwo consumption in watersheds across the world during calibration (see SI).
+The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
+the interquartile range), and the notch represents the median with its 95\% confidence interval.
+The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
     \label{fig:scenario_pCO2}

### 2019-08-29 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

--- 2019-06-27 Manuscript/Data_Repository/Data_Repository.tex
@@
 \newcommand{\permil}{\textperthousand\xspace}
-\newcommand{\dsil}{$d$\xspace}
+\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
+\newcommand{\pCOtwo}{\textit{p}CO$_{2}$\xspace}
+\newcommand{\COtwo}{CO$_{2}$\xspace}
+%
 
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{DR\arabic{figure}}
+\renewcommand{\thefigure}{SI\arabic{figure}}
 \setcounter{table}{0}
-\renewcommand{\thetable}{DR\arabic{table}}
+\renewcommand{\thetable}{SI\arabic{table}}
 
@@
 \fancyhf{}
-\lhead{Park et al., in review for XXX}
+\lhead{Park et al., in preparation}
 \rhead{\thepage}
@@
 \begin{flushleft}
-{\Large \textbf{Data Repository for ``GEOCLIM Modern''}}
-\\
+{\Large \textbf{Supplementary Information for ``Emergence of Indonesia and New Guinea as a driver for Neogene cooling''}}
+
 Yuem Park\textsuperscript{1},
 Nicholas L. Swanson-Hysell\textsuperscript{1},
-Pierre Maffre\textsuperscript{2},
-Yves Godd\'eris\textsuperscript{2}
-\\
+Pierre Maffre\textsuperscript{1},
+Francis A. Macdonald\textsuperscript{2},
+Eliel A. Anttila\textsuperscript{2},
+Yves Godd\'eris\textsuperscript{3}
+
 \bigskip
 \textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
-\\
+
+\textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
+
+\textsuperscript{3} G\'eosciences Environnement Toulouse, CNRS--Universit\'e Paul Sabatier - IRD, Toulouse, France
+
 \bigskip
@@
 
-These repository materials provide details on the model framework used in this study. Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These supplementary information materials provide details on the model framework used in this study. Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
@@
 
-The silicate weathering portion of the GEOCLIM model implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
+The silicate weathering component of the GEOCLIM model implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
 We refer to this chemically weathered profile as regolith where the base of the regolith is the contact with unweathered bedrock.
@@
 
-%\citet{Heimsath1997a} found that the rate at which material is entrained into the physically disturbed portion of the regolith (often called soil production) declines exponentially with increasing depth of this disturbed layer.
-The regolith model that we use does not differentiate the physically disturbed portion of the regolith, but given that the extent of weathering in the regolith below the physically disturbed layer has been found to covary with the rate of soil production \citep{Burke2007a}, we assume that regolith production behaves similarly to soil production:
-
 The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the chemical weathering on climate:
@@
 Equation \ref{eq:8} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
-As described below, in our implementation of the model, we vary $k_{d}$, $k_{w}$ and $\sigma$ and compare the resulting calculated chemical weathering fluxes to data compilations.
-We take varying parameter combinations that result in the better fits between the modeled fluxes and the data and use them for the GEOCLIM simulations.
 
 For our application of DynSoil, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
-The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations.
+The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations:
 
@@
 
-\noindent $x(z)$ gives the curve of primary phases varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
-Setting z equal to regolith thickness (h) gives $x_s$ which is the proportion of primary phases remaining at the top of the column of regolith.
+\noindent
+$x(z)$ gives the curve of primary phases varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
+Setting $z$ equal to regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the column of regolith.
 This value controls the overall integrated chemical weathering rate out of the regolith.
+
+\section*{Implementation of Lithology}
 
@@
 Previous implementations of GEOCLIM have taken bulk continental crust values across all land.
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithological resolved version of the model.
-To do so, we grouped
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically resolved version of the model.
 
-Add Ca+Mg concentrations
+\begin{table}[h!]
+\begin{center}
+\resizebox{0.7\textwidth}{!}{
+    \begin{tabular}{cll|cl}
+    \textbf{GLiM} & \textbf{GLiM} & \textbf{GLiM} & \textbf{GEOCLIM} & \textbf{GEOCLIM}\\
+    \textbf{ID} & \textbf{code} & \textbf{classification} & \textbf{ID} & \textbf{classification}\\
+    \hline
+    1 & su & unconsolidated sediments & 6 & sediments \\
+    2 & vb & basic volcanic rocks & 4 & mafics \\
+    3 & ss & siliciclastic sedimentary rocks & 6 & sediments \\
+    4 & pb & basic plutonic rocks & 4 & mafics \\
+    5 & sm & mixed sedimentary rocks & 6 & sediments \\
+    6 & sc & carbonate sedimentary rocks & 5 & carbonates \\
+    7 & va & acid volcanic rocks & 2 & felsics \\
+    8 & mt & metamorphics & 1 & metamorphics \\
+    9 & pa & acid plutonic rocks & 2 & felsics \\
+    10 & vi & intermediate volcanic rocks & 3 & intermediates \\
+    11 & wb & water bodies & 0 & water/ice \\
+    12 & py & pyroclastics & 2 & felsics \\
+    13 & pi & intermediate plutonic rocks & 3 & intermediates \\
+    14 & ev & evaporites & 5 & carbonates \\
+    15 & nd & no data & 1 & metamorphics \\
+    16 & ig & ice and glaciers & 0 & water/ice \\
+    \end{tabular}
+}
+\vspace{10pt}
+\caption{Grouping of 16 lithologic categories in GLiM \citep{Hartmann2012a} to 6 broader categories for GEOCLIM.}
+\label{tab:GLiM_to_GEOCLIM}
+\end{center}
+\end{table}
+
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
+The raw data takes the form of polygon vectors, where each polygon is assigned one of 16 lithologic categories.
+We first group these 16 categories into 6 broader categories (metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment; Table \ref{tab:GLiM_to_GEOCLIM}).
+We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the most area of that pixel (i.e.
+the `mode' lithology; Fig.
+XX).
+
+%% FROM MAIN TEXT
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
+SX).
+Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
+The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
 
@@
 We use the values of 42 kJ/mol in our model runs.
-
-\section*{Lithology Implementation}
-
-XXX
-
-going from GLiM to our lithologic classes.
 

### 2019-08-31 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 
-The erosion rate (E; m/yr) in the silicate weathering module of GEOCLIM follows one of the approaches taken in \citet{Godderis2017c} and \citet{Maffre2018a} where it is parameterized based on runoff (q; m/yr) and slope (s; m/m):
-
-\begin{equation}
-    E = k_{e}\;q^{m}\;s^{n}
-    \label{eq:9}
-\end{equation}
-
-$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
-This formulation is similar to the large-scale BQART model of \citet{Syvitski2007a}, but without there being an implemented temperature dependence and by replacing elevation with slope such that it more closely follows the stream power law functional form \citep{Davy2000a}.
-This form and these exponent values are supported by compilations, such as that in \citet{Lague2014a}, although \citet{Lague2014a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
-
-The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and the soil production function:
+The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and the soil production function ($f(h)$):
 
@@
 
-In this implementation of DynSoil, we use an exponential soil production function as is commonly applied in the literature (e.g.
-\citealp{Gabet2009a}) and follows the form as in \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
-Here, $f(h)$ is the `soil production function,' which describes how regolith production decreases as the thickness of the regolith increases.
+\begin{equation}
+    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:6}
+\end{equation}
 
@@
 
+\noindent
 $P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
 In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the universal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
+$f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases.
+It takes an exponential form as is commonly applied in the literature (e.g.
+\citealp{Gabet2009a}).
+In Equation \ref{eq:7}, we follow \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
+
+Our implementation of the erosion rate follows one of the approaches taken in \citet{Godderis2017c} and \citet{Maffre2018a}, where it is parameterized based on runoff and slope ($s$; m/m):
 
 \begin{equation}
-    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
-    \label{eq:6}
+    E = k_{e}\;q^{m}\;s^{n}
+    \label{eq:8}
 \end{equation}
+
+\noindent
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
+This formulation is similar to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}.
+This form and these exponent values are supported by various compilations (e.g.
+\citealp{Lague2014a}) although \citet{Lague2014a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
 
@@
     K = k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
-    \label{eq:8}
+    \label{eq:9}
 \end{equation}
@@
 \noindent
-Equation \ref{eq:8} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
+Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
 
-For our application of DynSoil, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
+In this study, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
 The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations:
@@
 
-To calculate the flux of Ca and Mg associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
-\ref{fig:regolith_schematic}).
-Previous implementations of GEOCLIM have taken bulk continental crust values across all land.
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically resolved version of the model.
-
 \begin{table}[h!]
@@
 \resizebox{0.7\textwidth}{!}{
-    \begin{tabular}{cll|cl}
+    \begin{tabular}{cclcl}
     \textbf{GLiM} & \textbf{GLiM} & \textbf{GLiM} & \textbf{GEOCLIM} & \textbf{GEOCLIM}\\
     \textbf{ID} & \textbf{code} & \textbf{classification} & \textbf{ID} & \textbf{classification}\\
+    &&&& \\
     \hline
+    &&&& \\
     1 & su & unconsolidated sediments & 6 & sediments \\
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/world_lithology.jpg}
+    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM.}
+    \label{fig:world_lithology}
+\end{figure}
+
+To calculate the flux of Ca and Mg associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
+\ref{fig:regolith_schematic}).
+Previous implementations of GEOCLIM have taken bulk continental crust values across all land.
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically resolved version of the model.
+
 The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
@@
 We first group these 16 categories into 6 broader categories (metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment; Table \ref{tab:GLiM_to_GEOCLIM}).
-We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the most area of that pixel (i.e.
-the `mode' lithology; Fig.
-XX).
+We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the greatest area in that pixel (i.e.
+the `mode lithology'; Fig.
+\ref{fig:world_lithology}).
+However, to improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees.
+To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel that is covered by each of the 6 lithologic categories + water/ice is captured by the extra dimension.
+In this way, GEOCLIM is able to calculate the area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
+
+\section*{GEOCLIM Calibration}
+
+Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:8} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
+However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}.
+We use the value of 42~kJ/mol in our model runs.
+
+However, as discussed in the main text, $k_{d}$ (Equation \ref{eq:9}), $k_{w}$ (Equation \ref{eq:9}), $\sigma$ (Equation \ref{eq:4}), and $k_{rp}$ (Equation \ref{eq:8}) are poorly constrained.
+Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
+We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM (Table \ref{tab:parameter_combinations}).
+
+\begin{table}[h!]
+\begin{center}
+\resizebox{0.6\textwidth}{!}{
+    \begin{tabular}{cccccc}
+    \textbf{$k_{d}$} & \textbf{$k_{w}$} & \textbf{$\sigma$} & \textbf{$k_{rp}$} & \textbf{metamorphic$_{\text{Ca}+\text{Mg}}$} & \textbf{sediment$_{\text{Ca}+\text{Mg}}$}\\
+    unitless & unitless & unitless & unitless & mol/m$^{3}$ & mol/m$^{3}$ \\
+    &&&&& \\
+    \hline
+    &&&&& \\
+    1$\times$10$^{-5}$ & 1$\times$10$^{-3}$ & -0.4 & 1.2$\times$10$^{-3}$ & 1500 & 500 \\
+    2$\times$10$^{-5}$ & 2$\times$10$^{-3}$ & -0.2 & 2$\times$10$^{-3}$ & 2000 & 1000 \\
+    5$\times$10$^{-5}$ & 5$\times$10$^{-3}$ & -0.1 & 3$\times$10$^{-3}$ & 2500 & 1500 \\
+    1$\times$10$^{-4}$ & 1$\times$10$^{-2}$ & 0 & 5$\times$10$^{-3}$ & 3000 & 2000 \\
+    2$\times$10$^{-4}$ & 2$\times$10$^{-2}$ & 0.1 &  & 3500 & 2500 \\
+    5$\times$10$^{-4}$ & 5$\times$10$^{-2}$ & 0.3 &  & 4000 & 3000 \\
+    1$\times$10$^{-3}$ & 1$\times$10$^{-1}$ &  &  &  &  \\
+    2$\times$10$^{-3}$ & 2$\times$10$^{-1}$ &  &  &  &  \\
+    5$\times$10$^{-3}$ & 5$\times$10$^{-1}$ &  &  &  &  \\
+    1$\times$10$^{-2}$ & 1 &  &  &  &  \\
+    \end{tabular}
+}
+\vspace{10pt}
+\caption{Values tested for poorly constrained parameters in the silicate weathering component of GEOCLIM.
+Every permutation of the listed values were tested (except those permutations where the Ca+Mg concentration of the sediments are higher than that of the metamorphics), resulting in 78,000 unique parameter combinations.}
+\label{tab:parameter_combinations}
+\end{center}
+\end{table}
+
+We then compute spatially-resolved long-term \COtwo consumption (i.e.
+Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
+As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+80 watersheds in total are used in this study.
+We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:
+
+\begin{equation}
+    r^{2} = 1 - \frac{\sum\left[ \log_{10}(M_{i}) - \log_{10}(O_{i}) \right]^{2}}{\sum\left[ \log_{10}(O_{i}) - \overline{\log_{10}(O)} \right]^{2}}
+    \label{eq:10}
+\end{equation}
+
+\noindent
+$M_{i}$ is the modeled \COtwo consumption over watershed $i$, $O_{i}$ is the observed \COtwo consumption over watershed $i$, and $\overline{\log_{10}(O)}$ is the mean of the log of observed \COtwo consumption over all watersheds.
+
+% FIGURE: CO2 consumption vs r^2
+% FIGURE: CO2 consumption cross plot
 
 %% FROM MAIN TEXT
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (Fig.
-SX).
-Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
-The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
-The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM (see below).
+Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), four parameters in particular are poorly constrained in the existing literature: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
+As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
+Rather than prescribing values for these six parameters, we instead allow these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
+For each of these combinations, we compute spatially-resolved long-term \COtwo consumption (i.e.
+Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
+Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
+We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1$\times$10$^{12}$~mol/yr or \textgreater1$\times$10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
+This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
 
-\section*{Choice of Parameters}
+\section*{Climate Model}
 
-XXX
+xxx
 
-Experimental determinations of the activation energy (E$_a$) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
-However, multiple efforts to invert for E$_a$ in basaltic watersheds with varying temperature have yielded values (41.6$\pm$3.2 kJ/mol in \citealp{Li2016a}; 42.3 kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5$\pm$1.7 kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1 kJ/mol; \citealp{Carroll2005a}.
-We use the values of 42 kJ/mol in our model runs.
+\section*{Paleoshorelines}
+
+xxx
 

### 2019-08-31 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
-We seek to evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.
+To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, here we use Neogene paleoshoreline reconstructions and couple a weathering model to a climate model to estimate changes \pCOtwo with changes in tropical paleogeography.
 
@@
 12~Ma \citep{vanUfford2005a}.
+Arc-continent collision in New Guinea has progressed from west to east.
 Exhumation of the Central Range accelerated over the past 4~m.y.
-due to slab-breakoff and buoyant uplift under New Guinea \citep{Cloos2005a}.
-After arc-continent collision in New Guinea, a left-lateral strike-slip margin developed on the northern margin \citep{Baldwin2012a}.
-Over the past 4~m.y., this change in plate motion caused the westward extrusion and exhumation of the Cretaceous to Oligocene ophiolites of Obi, Halmahera, and northern New Guinea \citep{Baldwin2012a}.
+due to slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of  ophiolites both along the coast and on the islands Obi and Halmahera \citep{Baldwin2012a}.
 In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~m.y.
@@
 Collaborative research between NLS-H and YG was initially supported by the France-Berkeley Fund.
-FAM and NLS-H thank NSF FRES grant \#20191023 for support.
+FAM and NLS-H thank NSF FRES grant \#1926001 for support.
 

### 2019-09-03 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
-As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
 80 watersheds in total are used in this study.
@@
 
-% FIGURE: CO2 consumption vs r^2
-% FIGURE: CO2 consumption cross plot
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.pdf}
+    \caption{Modeled global \COtwo consumption vs.
+the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of the watersheds.
+Each point represents model output using one of the 78,000 parameter combinations.
+The right panel is zoomed to the plotting space with the highest $r^{2}$'s.}
+    \label{fig:W_vs_r2}
+\end{figure}
 
-%% FROM MAIN TEXT
-Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), four parameters in particular are poorly constrained in the existing literature: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
-As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-Rather than prescribing values for these six parameters, we instead allow these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
-For each of these combinations, we compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
-Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
-We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
-We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1$\times$10$^{12}$~mol/yr or \textgreater1$\times$10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
-This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
+    \caption{Modeled vs.
+observed \COtwo consumption in watersheds around the world.
+\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations.
+\textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.}
+    \label{fig:r2_cross_plot}
+\end{figure}
+
+% HYBAM description and dashed lines
+
+The majority of the original 78,000 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig.
+\ref{fig:W_vs_r2}).
+However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
+We therefore eliminate all parameter combinations that produce a $r^{2}\textless0.4$, which leaves 3,339 unique parameter combinations (Figs.
+\ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
+We note that these 3,339 combinations also all produce total global \COtwo consumption estimates that fall within the range of those estimated in the literature for the present-day.
+However, this is an expected outcome, given that total global \COtwo consumption is often estimated in the literature using the same measured watershed \COtwo consumption that we compare to.
 

### 2019-09-10 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 
-These supplementary information materials provide details on the model framework used in this study. Python code used for this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These supplementary information materials provide details on the model framework used in this study. Code associated with this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
 
@@
 The silicate weathering component of the GEOCLIM model implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
-We refer to this chemically weathered profile as regolith where the base of the regolith is the contact with unweathered bedrock.
-In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering as a particle transit through the regolith or as a physically weathered particle once it reaches the top.
+We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
+In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering as a particle transits through the regolith or as a physically weathered particle once it reaches the top.
 We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}.
-The transient time-varying versions of this regolith model is described by the following system of three equations which form the basis of DynSoil:
+The transient time-varying version of this regolith model is described by the following system of three equations which form the basis of DynSoil:
 
@@
 
-The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and the soil production function ($f(h)$):
+The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and a soil production function ($f(h)$):
 
@@
 \ref{fig:regolith_schematic}).
-Previous implementations of GEOCLIM have taken bulk continental crust values across all land.
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically resolved version of the model.
+Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land.
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically-resolved version of the model.
 
@@
 However, to improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees.
-To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel that is covered by each of the 6 lithologic categories + water/ice is captured by the extra dimension.
-In this way, GEOCLIM is able to calculate the area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
+To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel covered by each of the 6 lithologic categories (or water/ice) is captured by the extra dimension.
+In this way, we calculate an area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
 
@@
 We then compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
+Ca+Mg fluxes) using present-day runoff, temperature, and slope fields.
 As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.pdf}
-    \caption{Modeled global \COtwo consumption vs.
+    \caption{\textbf{A)} Modeled global \COtwo consumption vs.
 the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of the watersheds.
 Each point represents model output using one of the 78,000 parameter combinations.
-The right panel is zoomed to the plotting space with the highest $r^{2}$'s.}
+\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.}
     \label{fig:W_vs_r2}
@@
 Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations.
+The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines.
 \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.}
@@
 
-% HYBAM description and dashed lines
-
 The majority of the original 78,000 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig.
+\ref{fig:W_vs_r2}).
+Given that the coefficient of determination ($r^{2}$) is calculated using equation \ref{eq:10} rather than fitting a linear model, many of the combinations associated with particularly poor fits result in negative $r^{2}$ values (Fig.
 \ref{fig:W_vs_r2}).
@@
 
-xxx
+We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
+These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
+1860 at a 1\degrees $\times$ 1\degrees resolution.
+In the ``1860 control'' experiment, forcing agents representative of conditions ca.
+1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
+\pCOtwo in 1860 is assumed to be 286~ppm.
+In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
+\pCOtwo is then held constant until the end of the 280 year experiment.
+All non-\COtwo forcing agents are held constant.
+The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
+We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
+
+\section*{Indonesia and New Guinea Scenarios}
+
+% explanation of construction of lith maps: Sunda Shelf scenario also needs explanation of construction of runoff and slope fields.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Manuscript/Figures/ING_fracs.jpg}
+    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM.
+Each column represents a tested scenario.
+Each row represents a lithologic category.
+Solid black lines show present-day shorelines.}
+    \label{fig:ING_fracs}
+\end{figure}
 
@@
 
-xxx
+Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in I\&NG.
+Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$.
+We also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
+Larger islands were further divided into regions based upon their position on microplates as defined by \citep{Matthews2016a}.
+Miocene to present stratigraphic columns were compiled from each region to develop an age model for paleoenvironmental indicators.
+In general, carbonate strata and thin-bedded siliciclastic strata were assumed to be deposited in a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as indicators of a subaerial paleoenvironment.
+Following previous paleoshoreline compilations and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma) and Middle Miocene (15 Ma).
+The paleoshoreline reconstructions within the paleogeographic models broadly coincide with sub-Epoch boundaries, i.e.
+Early Miocene = 23-16 Ma, Middle Miocene = 15.99-11.61 Ma, Late Miocene = 11.6-5.4 Ma, Pliocene =5.39-3.8 Ma, but we recognize that the biostratigraphic resolution is an additional source of uncertainty.
+
+Several new sources for paleoshoreline data have become available since \citet{Molnar2015a}.
+Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and \citet{Hall2018a}, and in New Guinea, we followed \citet{Gold2017a} and \cite{Harrington2017a}.
+Although the outlines of our paleoshorelines are significantly different from those used for 5 Ma by \citet{Molnar2015a} our calculated area is comparable (Fig.
+1).
+
+\subsection*{Malay Peninsula and Sunda Shelf}
+
+Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20 Ma (Hall, 2013b; Hall and Nichols, 2002).
+Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (Hall and Nichols, 2002).
+Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
+During the Eocene through Early Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
+Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments.
+Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley, 2013).
+Basement highs (e.g.
+the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al., 2007).
+Although most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al., 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
+
+\subsection*{Borneo and Palawan}
+
+Our reconstructions of the paleoshorelines of Borneo and Palawan are largely informed by extant paleogeographic constructions detailed in Hall (2001, 2013a, b),  as well as geologic maps and local shoreline reconstructions found in van de Weerd and Armin (1992), Witts et al.
+(2012), Madon et al.
+(2013), and Kessler and Jong (2015).
+Comprised of Paleozoic to Mesozoic crustal components that were largely accreted to Sundaland by the late Cretaceous (Metcalfe, 2013), the southern and western portions of Borneo (SW Kalimantan) have been subaerially exposed throughout the Neogene (Hall, 2013a, b).
+A collision along the northern margin of Borneo associated with the initiation of rifting in the South China Sea resulted in the late Eocene uplift of the Central Range mountains (Hutchison, 1996), which provide sediment to basins along the Southern (Kalimantan) and northern (Sabah and Sarawak) coasts of the island.
+In the east, the Barito, Kutei, and Tarakan basins developed as a single area of subsidence (associated with the opening of the Makassar Straits in the Eocene) before the basins were isolated by Oligocene faulting and Miocene uplift (Witts et al., 2012).
+The Kutei Basin is characterized by eastward-prograding deltaic and shallow shelf deposits that have been steadily supplied with sediment from the Schwaner Mountains of SW Kalimantan and the Central Range (van de Weerd and Armin, 1992).
+North of the Kutei Basin, the deltaic deposits found in the Tarakan Basin similarly prograde eastward from the mid Miocene onward, fed by the northern drainages of the Central Range (Satyana et al., 1999).
+In SE Kalimantan, the Barito Basin is bounded to the east by the ophiolite-bearing Meratus Mountains.
+Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene (Witts et al., 2012).
+Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan (Hall et al., 2008; Ilao et al., 2018).
+The Sarawak Basin and NW Borneo trough offshore Sabah host >10 km-thick Neogene sediementary sequences, indicative of the extent and duration of exhumation in northern Borneo (Hall et al., 2008).
+The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan (Hall, 2013a), as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
+Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities (Hall et al., 2008).
+Early Miocene sediments in northern Sabah have a provenance from Palawan (Suggate et al., 2014), indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny.
+However, initiation of back-arc extension in the Sulu Sea around 19 Ma (Hall, 2013a) resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
+The 14 Ma Capoas Granite on Palawan and the 7.5 Ma Kinabalu Granite in Sabah are the products of crustal thinning related to pulses of regional backarc extension (Hall, 2013a).
+Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7km/Myr) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
+Kinabalu, the tallest peak in Borneo (Cottam et al., 2013).
+
+\subsection*{Sulawesi}
+
+For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of Nugraha and Hall (2018).
+Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms.
+These arms have high relief >3km that are separated by deep basins and broadly define Sulawesi’s seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi (Hamilton, 1979; Katili, 1978).
+Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system (Hall, 2017; Hennig et al., 2017a; Hennig et al., 2016; Parkinson, 1998b).
+After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed (Elburg and Foden, 1998; Polvé et al., 1997) that was connected to the paleo-Sunda Arc (Hall, 2002), also referred to as the Great Indonesian arc (Harris, 2006).
+The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (Monnier et al., 1995) [but see (Kadarusman et al., 2004) for an alternative view], and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32-28 Ma; Parkinson, 1998a) and thrust to the east above a west dipping slab (East Sulawesi block of Villeneuve et al., 2002).
+During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton (Bergman et al., 1996; Smith and Silver, 1991).
+
+The North Arm consists of Late Pliocene and younger volcanic rocks built upon Eocene to Early Miocene oceanic basalt, basaltic andesite, pelagic sediments, and metamorphic rocks (Elburg and Foden, 1998).
+Eocene to Early Miocene volcanic rocks formed above a NW dipping slab in an arc system that extended to West and South Sulawesi (van Leeuwen and Muhardjo, 2005).
+Gorontalo Bay is an extensional basin that formed in the Pliocene, and prior to that time, the East Sulawesi was attached to the North Arm of Sulawesi.
+After the Oligocene to Early Miocene accretion of the East Sulawesi ophiolite, the arc system stepped out to the SE forming a Neogene (23-16 Ma) volcanic belt on the East arm (Kadarusman et al., 2004).
+In the Late Miocene, SE-dipping subduction was initiated below the North Arm, which continues today at the North Sulawesi Trough.
+North Sulawesi was not substantially emergent until the Pliocene (van Leeuwen and Muhardjo, 2005).
+Active volcanoes extend from the North Arm through the Sangihe Arc into the southern Philippines.
+
+The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000 km2 (Monnier et al., 1995).
+The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, mélange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint (Monnier et al., 1995; Parkinson, 1998a; Silver et al., 1983).
+Locally, the structural thickness of the ophiolite exceeds 15 km with surface relief towering over 3 km (Kadarusman et al., 2004).
+Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93-32 Ma, clustering between 60 and 40 Ma (Parkinson, 1998, and references therein).
+K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36-23 Ma (Parkinson, 1998a; Villeneuve et al., 2002), which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) (Parkinson, 1998a; Silver et al., 1983); however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
+
+Collision of the Sula-Banggai block with the East margin of Sulawesi began in the latest Miocene and uplift accelerated in the early Pliocene (5.2 to 3.8 Ma) and is associated with a major pulse of sedimentation in adjacent basins (Davies, 1990; Villeneuve et al., 2000).
+Off the northeast margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes mélange (Davies, 1990).
+Thus, although the East Sulawesi ophiolite was trapped/emplaced onto the composite Sundaland margin (i.e.
+a fragment of previously accreted Australian crust) between 36 and 23 Ma, mafic and ultramafic rocks on the south and west arms appear to have not been substantially subaerially exposed in the south and west arms until after 15 Ma, and on the east arm until after 5 Ma.
+
+West Sulawesi rifted from Borneo during the Eocene forming the Makassar Straits back arc basin behind a southwest-facing arc (Polvé et al., 1997).
+Eroded fragments of ophiolite and extensive belts of volcanic rocks are preserved on the West and South Arms of Sulawesi (Bergman et al., 1996; van Leeuwen et al., 2010).
+During the Middle Miocene (ca.
+15-13 Ma), extensional faults in the Bone basin reversed, which was accompanied by uplift and erosion of the Bone Mountain ophiolite and Lamasi complex in West Sulawesi (Bergman et al., 1996; van Leeuwen et al., 2010).
+Shortening was likely due to the collision of the leading edge of the Buton-Tukang Besi block, which collided with Buton and the SE Arm of Sulawesi (Smith and Silver, 1991).
+Uplift and erosion is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins (Bergman et al., 1996; van Leeuwen et al., 2010).
+Uplift was diachronous, not effecting units below and to the west of the Lamasai ophiolite until the Middle Miocene (~13 Ma) (van Leeuwen et al., 2010).
+Alkali volcanism ensued at ca.
+11 Ma and is associated with a second phase of extension and exhumation.
+From the Late Miocene to Pliocene fission track ages imply deep exhumation (Bergman et al., 1996; Smith and Silver, 1991; van Leeuwen et al., 2010).
+
+Fission track ages from granitoids in central Sulawesi indicate rapid uplift of Central Sulawesi (200–700 m/Myr) starting at about 5 Ma associated with movement on the Palo-Koro fault (Bellier et al., 2006).
+The fault system also shows a normal component with rapid exhumation of rock west of the fault in western Sulawesi – all fission track dates are younger than 5 Ma (Bellier et al., 2006).
+Just east of the Palo-Koro fault, the Palu Metamorphic Complex was exhumed in the Late Miocene to early Pliocene in the north (~5.3 Ma) and later Pliocene in the south (~3.1–2.7 Ma) at rates of up to 400 m/Myr (Hennig et al., 2017b).
+
+\subsection*{New Guinea and Halmahera}
+
+Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of Cloos et al.
+(2005), Gold et al.
+(2017), Harrington et al.
+(2017), and Nichols and Hall (1991).
+These studies were all based to varying degrees on lithological distributions, biostratigraphic, borehole data, and tectonic models.
+We complimented these data with our stratigraphic compilations for the region to provide further paleoenvironmental context.
+
+Northern New Guinea, including the Melanesian Arc, was emplaced above the Australian plate during the Miocene (Baldwin et al., 2012; Cloos et al., 2005; Hamilton, 1979; van Ufford and Cloos, 2005).
+Two major ophiolite belts marking the suture—the Irian-Marum ophiolite belt (including the April ultramafics), and the Papuan Ultramafic Belt (PUB)—are preserved along the Central Range and Peninsular Range, respectively.
+South of the Irian-Marum ophiolites, the Ruffaer Metamorphic Complex constructed from the accretionary wedge, forms the spine of the Central Range (up to ~5 km elevation today).
+
+The Middle Miocene (16–14 Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt (Cloos et al., 2005; Visser and Hermes, 1962).
+This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
+12 Ma (van Ufford and Cloos, 2005).
+Mountain building began in Late Miocene time, ca.
+8-7 Ma (Baldwin et al., 2012; van Ufford and Cloos, 2005), but major relief was not generated until the Pliocene (Weiland and Cloos, 1996).
+Similarly, the Marum ophiolite was uplifted between 8-5 Ma with 3-4 km of denudation (Crowhurst et al., 1996).
+An estimated 80-100 km of shortening has been accommodated by deformation on the south side of the Central Range (Cloos et al., 2005; Hill and Gleadow, 1989).
+
+Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene (Davies and Jaques, 1984; Davies and Smith, 1971; van Ufford and Cloos, 2005).
+Miocene to present arc-continent collision in New Guinea has progressed from west to east.
+Exhumation of the Central Range accelerated over the past 4~m.y.
+due to slab-breakoff and buoyant uplift (Cloos et al., 2005).
+This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of ophiolites both along the coast (Monnier et al., 1999) and on the islands Obi (Ali et al., 2001) and Halmahera (Ballantyne, 1992; Hall et al., 1988).
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4 Ma (van Ufford and Cloos, 2005), which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill and Gleadow, 1989).
+
+In the Bird’s Neck of western New Guinea, the Lengguru foldbelt formed during Middle to Late Miocene CW rotation and obduction of the Weyland Terrane (Bailly et al., 2009).
+Recent CCW rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge (Sapin et al., 2009).
+
+\subsection*{Philippines}
+
+Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall (Hall, 1997), coupled with geologic maps of the archipelago (Geological Survey Division, 1963).
+The Philippine ophiolites have been grouped in four distinct belts (Yumul, 2007).
+We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone (Fig.
+3; sutures 11 and 12).
+Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments (Pubellier et al., 1991).
+The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present (Yumul Jr et al., 2013).
+In addition to collision, arc magmatism contributed significantly to crustal growth in the Philippines (Dimalanta and Yumul, 2003).
+
+\subsection*{Sunda-Banda Arc}
+
+Paleoshoreline reconstructions of the Sunda-Banda Arc system relied largely on the work of Hall (2013b).
+The Sunda-Banda Arc system is composed of thick sequences of mixed sedimentary rocks and basement intruded and overlain by igneous rocks derived from continual arc magmatism through much of the Cenozoic (Hall, 2017).
+Sumatra is the largest island within the Sunda-Banda Arc system that stretches from the Andaman Sea in the northwest to the Banda Sea in the east.
+The western portion of the Sumatra is underlain by the Wolya ophiolite, which was obducted in the Cretaceous and exhumed in the Eocene to Oligocene (Allen et al., 2008), whereas the southeast corner on the islands of Bangka and Belitung is composed of granite basement of the Sunda Shelf (Hall, 2009, 2013b).
+Java is composed of a complex of E-W-striking deformational and magmatic belts (Audley-Charles, 2004).
+The belts have a volcanic arc in the south and a continental shelf to the north with an adjacent sedimentary basin (Audley-Charles, 2004).
+Bali and Flores are mostly volcanic with some sedimentary cover (Hall, 2009).
+
+Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate (Harris, 2006).
+From ~20-10 Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20 Ma (Hall, 2009, 2013b).
+Six major islands were subaerially exposed before 5 Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores.
+However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5 Ma (Hall, 2009, 2013b).
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5 Ma, associated with slab roll-back and collision with the Australian continental margin (Audley-Charles, 2004; Hall, 2013b; Harris, 2006).
+In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations >1 km above sea level with estimated average uplift rates of 1.5 km/Myr (Audley-Charles, 1986).
+
+\section*{Limitations and future directions}
+
+As with any modeling study, there are limitations with the approach.
+We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
+Using an updated climate model at higher resolution may result in runoff and temperature fields with different spatial distribution and sensitivities to elevation.
+
+Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel.
+This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position.
+Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
+However, implementing high-resolution climate models with both varying shorelines and change in paleogeographic position for these scenarios is valuable future direction.
+Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.
+
+Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
+Relative to felsic lithologies, mafic lithologies contain minerals with faster dissolution kinetics (e.g.
+olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
 

### 2019-09-10 Manuscript/Manuscript.tex

@@
 \ref{fig:shoreline_growth}).
-The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of I\&NG and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics \citep{Macdonald2019a}.
+The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of I\&NG and the exhumation of large supra-subduction ophiolites and associated forearc sediments in the warm, wet tropics \citep{Macdonald2019a}.
 
@@
 Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
-After the Miocene Climatic Optimum (MCO), a cooling trend began at $\sim$14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
+After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
+14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
 culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
@@
 The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
-To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, here we use Neogene paleoshoreline reconstructions and couple a weathering model to a climate model to estimate changes \pCOtwo with changes in tropical paleogeography.
+To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, we use Neogene paleoshoreline reconstructions and couple a weathering model to climate model results to estimate the magnitude of changes in \pCOtwo that are coupled with changes in tropical paleogeography.
 
@@
 For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady-state is achieved at a higher \pCOtwo.
-Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower steady-state \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.
+Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower steady-state \pCOtwo when global weatherability is high \citep{Kump1997a}.
 
@@
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \citep{Lasaga1994a, West2012a}.
-Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance the total chemical weathering rate \citep{West2012a, Maher2014a}.
+Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance chemical weathering \citep{West2012a, Maher2014a}.
 In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}.
@@
 
-To estimate the decrease in steady state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to global climate models run at varying \pCOtwo levels.
-We use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to a global climate model run at varying \pCOtwo levels.
+Within the silicate weathering component of GEOCLIM, we use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
 
-The silicate weathering component of GEOCLIM calculates \COtwo consumption by silicate weathering for each continental grid cell.
+The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell.
 In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions (e.g.
@@
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}, with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+Given that the \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are within the mafic category.
 Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are then assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
-The weathering of carbonate grid cells does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on its protolith.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+The weathering of carbonate does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on protolith composition.
 The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
@@
 Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), four parameters in particular are poorly constrained in the existing literature: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
-As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-Rather than prescribing values for these six parameters, we instead allow these parameters to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
+As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are also not well known.
+Rather than prescribing values for these six parameters, we instead allow them to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
 For each of these combinations, we compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg consumption) using present-day runoff, temperature, and slope fields.
-Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+Ca+Mg fluxes) using present-day runoff, temperature, and slope fields.
+Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2018a}.
 We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
-We eliminate all parameter combinations that result in unreasonable total global \COtwo consumption (which we define to be \textless1$\times$10$^{12}$~mol/yr or \textgreater1$\times$10$^{13}$~mol/yr) and/or result in a low $r^{2}$ (which we define to be \textless0.4).
+We eliminate all parameter combinations that result in a low $r^{2}$ (defined as \textless0.4, see SI).
 This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
 
-Having calibrated the silicate weathering component of GEOCLIM, we estimate the decrease in steady-state \pCOtwo associated with increases in global weatherability.
-To do so, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
-These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
-1860 at a 1\degrees $\times$ 1\degrees resolution.
-In the ``1860 control'' experiment, forcing agents representative of conditions ca.
-1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium.
-\pCOtwo in 1860 is assumed to be 286~ppm.
-In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value.
-\pCOtwo is then held constant until the end of the 280 year experiment.
-All non-\COtwo forcing agents are held constant.
-The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
-We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
-The primary strength of using these three GFDL CM2.0 experiments is that all non-\COtwo forcing agents are held constant, allowing the effect of \pCOtwo on climate to be isolated.
+Implementing these parameter combinations in the silicate weathering component of GEOCLIM, we estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
+For the climate model component of GEOCLIM, we obtain modeled temperature and runoff fields at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; see SI).
+The primary strength of using these three GFDL CM2.0 experiments for this analysis is that all non-\COtwo forcing agents are held constant at values representative of conditions ca.
+1860, allowing the effect of \pCOtwo on climate to be isolated.
 
-We use geological and paleoshoreline data to quantify the changes in the area of I\&NG over the past 20~m.y.
+We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y.
 (Fig.
@@
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
-Carbonate strata and thin-bedded siliciclastic strata were interpreted to represent a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as evidence for a subaerial environment.
-Taken together, these constraints delineate the position of shorelines.
-For the island of Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and for New Guinea we followed work of \citet{Norwick2003a, Harrington2017a, Gold2017a}.
-Although there are difference in data sources and paleoshoreline outlines from \cite{Molnar2015a}, the calculated area at 5~Ma is similar.
+Terrestrial and marine sedimentary deposits were used to determine the position of paleoshorelines (details in SI).
+
+In New Guinea, emergence in the Middle Miocene is associated with the collision between the Melanesian Arc and the distal margin of Australia \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove the exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
+Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
+which has been intrepreted to be associated with slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has also caused major uplift over the past 4~m.y.
+\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013a}.
+In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+\citep{AudleyCharles1986a}.
+These drivers lead to progressive emergence, as seen in the paleoshoreline reconstructions, over the past 15 million years that accelerated following 5 Ma (Fig.
+\ref{fig:shoreline_growth}.
 
@@
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-Industrial \pCOtwo would have been $\sim$500-700~ppm.
+These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$500-700~ppm.
 Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca concentration of the lithologies that comprise the Sunda Shelf as well as its low slope.
@@
 
-I\&NG are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
-In New Guinea, these paleogeographic features originated in the Middle Miocene with the collision between the Melanesian Arc and the distal margin of Australia \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which resulted in the exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
-The Middle Miocene (16-14~Ma) basal Markats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the New Guinea ophiolite belt \citep{Cloos2005a}.
-This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
-12~Ma \citep{vanUfford2005a}.
-Arc-continent collision in New Guinea has progressed from west to east.
-Exhumation of the Central Range accelerated over the past 4~m.y.
-due to slab-breakoff and buoyant uplift \citep{Cloos2005a}.
-This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of  ophiolites both along the coast and on the islands Obi and Halmahera \citep{Baldwin2012a}.
-In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~m.y.
-\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
-
-Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013a}.
-In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
-\citep{AudleyCharles1986a}.
-
-Our paleoshoreline reconstructions show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
+Our  show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
 This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda Arc and New Guinea in the past 4~m.y.
@@
 
-The movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM (Fig.
+Enhance weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM that are now possible with given the spatial resolution of lithology (Fig.
 \ref{fig:scenario_pCO2}).
@@
 These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
-This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in low-latitude igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is significantly smaller than that of high-relief I\&NG.
+This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in these igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is significantly smaller than that of high-relief I\&NG.
 
-However, it is important to acknowledge the limitations of the approach taken in this study.
 First, the tested scenarios only simulate changes in the location of the shorelines in I\&NG and the lithology exposed in the Deccan and Ethiopian Traps.
@@
 In this way, changes in the flux of \COtwo outgassing via changes in spreading rates and ridge lengths \citep{Rowan2016a} and changes in weatherability via mountain building and the movement of land masses into/out of areas of high temperature and runoff are likely more important paleogeography-related factors in terms of modulating Neogene cooling than changes in oceanic and atmospheric circulation.
-Second, the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a low spatial resolution (2.0\degrees $\times$ 2.5\degrees) and is based upon physics and geochemical cycles that have been superceded by more recent climate models.
-Using an updated climate model at higher resolution may result in runoff and temperature fields with different sensitivities to paleogeography and \pCOtwo.
-Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, it does not yet implement lithology-dependent kinetics of mineral dissolution.
-Nevertheless, relative to felsic lithologies, mafic lithologies more dominantly comprise of minerals with faster kinetics (e.g.
-olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
+Second,
 
@@
 
-%%%
+%%%I\&NG are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
 
@@
 
+%%TC:ignore
 \section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}
@@
 FAM and NLS-H thank NSF FRES grant \#1926001 for support.
+%%TC:endignore
 
@@
     \includegraphics[width=0.9\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
-    \caption{The emergence of Indonesia and New Guinea.
-Past shorelines are shown in A with resultant area estimates summarized in B.
-A significant increase in area over the past 4 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \citep{Zachos2008a}.}
+    \caption{The emergence of Indonesia and New Guinea (I\&NG) from the mid-Miocene to present.
+Past shorelines are shown in A with associated area estimates summarized in B.
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \citep{Zachos2008a}.}
     \label{fig:shoreline_growth}

### 2019-09-11 Manuscript/Manuscript.tex

@@
 \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
-Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013a}.
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013b}.
 In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.

### 2019-09-11 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 Each point represents model output using one of the 78,000 parameter combinations.
-\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.}
+\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.
+The black line represents the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
+The black box represents the preferred range of those estimates.}
     \label{fig:W_vs_r2}
@@
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
-We note that these 3,339 combinations also all produce total global \COtwo consumption estimates that fall within the range of those estimated in the literature for the present-day.
-However, this is an expected outcome, given that total global \COtwo consumption is often estimated in the literature using the same measured watershed \COtwo consumption that we compare to.
+
+In our implementation, the mean global CO$_{2}$ consumption calculated from these 3,339 parameter combinations is 5.4$\times10^{12}$~mol/yr, with a standard deviation of 0.9$\times10^{12}$~mol/yr.
+These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
+\ref{fig:W_vs_r2}B).
 
@@
 
-We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \citep{Delworth2006a, Delworth2006b} for the climate model component of GEOCLIM.
+We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) for the climate model component of GEOCLIM.
 These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
-1860 at a 1\degrees $\times$ 1\degrees resolution.
+1860 at a 2.0\degrees $\times$ 2.5\degrees resolution.
 In the ``1860 control'' experiment, forcing agents representative of conditions ca.
@@
 The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value.
+\pCOtwo is then held constant for 160 years.
 We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
@@
 
-% explanation of construction of lith maps: Sunda Shelf scenario also needs explanation of construction of runoff and slope fields.
-
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=0.8\textwidth]{Manuscript/Figures/ING_fracs.jpg}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/ING_fracs.jpg}
     \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM.
@@
 \end{figure}
+
+We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshorelines}).
+To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig.
+\ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG.
+To generate the lithologic map used in the ``reduced I\&NG'' scenario (Fig.
+\ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15~Ma, then remove all land that falls outside of these bounds.
+For this scenario, we do not add any land (i.e.
+the Sunda Shelf) that is not exposed today to the lithologic map, and the slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for 15~Ma.
+
+The lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenario (Fig.
+\ref{fig:ING_fracs}) is identical to that used in the ``reduced I\&NG'' scenario, with the addition of the Sunda Shelf.
+However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed.
+As discussed in \textit{Paleoshorelines}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
+Basement highs between Malaysia/Sumatra and Borneo are typically granitic and bounded by basins dominated by shallow marine clastic fill.
+However, basement highs are less common in the Gulf of Thailand.
+We therefore assign all pixels of the Sunda Shelf to be 50\% felsic and 50\% sediment, except in the Gulf of Thailand where pixels are assigned to be 100\% sediment.
+For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.
+For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf.
+This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, but it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
 
@@
 In general, carbonate strata and thin-bedded siliciclastic strata were assumed to be deposited in a subaqueous marine environment, whereas evidence for exposure and terrestrial siliciclastic deposits were used as indicators of a subaerial paleoenvironment.
-Following previous paleoshoreline compilations and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma) and Middle Miocene (15 Ma).
+Following previous paleoshoreline compilations and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma), and Middle Miocene (15 Ma).
 The paleoshoreline reconstructions within the paleogeographic models broadly coincide with sub-Epoch boundaries, i.e.
@@
 Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene (Witts et al., 2012).
+
 Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan (Hall et al., 2008; Ilao et al., 2018).
@@
 
-\section*{Limitations and future directions}
+\section*{Limitations and Future Directions}
 

### 2019-09-12 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 This form and these exponent values are supported by various compilations (e.g.
-\citealp{Lague2014a}) although \citet{Lague2014a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
+\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
 
@@
 Following previous paleoshoreline compilations and these environmental indicators, paleoshorelines were outlined in QGIS to calculate areas for the Early Pliocene (5 Ma), Late Miocene (10 Ma), and Middle Miocene (15 Ma).
-The paleoshoreline reconstructions within the paleogeographic models broadly coincide with sub-Epoch boundaries, i.e.
-Early Miocene = 23-16 Ma, Middle Miocene = 15.99-11.61 Ma, Late Miocene = 11.6-5.4 Ma, Pliocene =5.39-3.8 Ma, but we recognize that the biostratigraphic resolution is an additional source of uncertainty.
+The paleoshoreline reconstructions within the paleogeographic models broadly coincide with sub-Epoch boundaries (i.e.
+Early Miocene = 23--16~Ma, Middle Miocene = 15.99--11.61~Ma, Late Miocene = 11.6--5.4~Ma, Pliocene =5.39--3.8 Ma) but we recognize that the biostratigraphic resolution is an additional source of uncertainty.
 
 Several new sources for paleoshoreline data have become available since \citet{Molnar2015a}.
-Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a} and \citet{Hall2018a}, and in New Guinea, we followed \citet{Gold2017a} and \cite{Harrington2017a}.
-Although the outlines of our paleoshorelines are significantly different from those used for 5 Ma by \citet{Molnar2015a} our calculated area is comparable (Fig.
-1).
+Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \cite{Harrington2017a}.
+Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable (Fig.
+\ref{fig:shoreline_growth}).
 
@@
 
-Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20 Ma (Hall, 2013b; Hall and Nichols, 2002).
-Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum (Hall and Nichols, 2002).
-Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene (Moss and Wilson, 1998).
-During the Eocene through Early Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf (Morley and Morley, 2013).
+Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2013b, Hall2002a}.
+Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}.
+Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
+During the Eocene through Early Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}.
 Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments.
-Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene (Morley and Morley, 2013).
+Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}.
 Basement highs (e.g.
-the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill (Darmadi et al., 2007).
-Although most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20 Ma onwards (Hall, 2013b; Madon et al., 2013), Molnar and Cronin (2015) omit the Sunda Shelf from their paleogeographic model for 5 Ma.
+the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}.
+Although most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, \citet{Molnar2015a} omit the Sunda Shelf from their paleogeographic model for 5~Ma.
 
@@
 
-Our reconstructions of the paleoshorelines of Borneo and Palawan are largely informed by extant paleogeographic constructions detailed in Hall (2001, 2013a, b),  as well as geologic maps and local shoreline reconstructions found in van de Weerd and Armin (1992), Witts et al.
-(2012), Madon et al.
-(2013), and Kessler and Jong (2015).
-Comprised of Paleozoic to Mesozoic crustal components that were largely accreted to Sundaland by the late Cretaceous (Metcalfe, 2013), the southern and western portions of Borneo (SW Kalimantan) have been subaerially exposed throughout the Neogene (Hall, 2013a, b).
-A collision along the northern margin of Borneo associated with the initiation of rifting in the South China Sea resulted in the late Eocene uplift of the Central Range mountains (Hutchison, 1996), which provide sediment to basins along the Southern (Kalimantan) and northern (Sabah and Sarawak) coasts of the island.
-In the east, the Barito, Kutei, and Tarakan basins developed as a single area of subsidence (associated with the opening of the Makassar Straits in the Eocene) before the basins were isolated by Oligocene faulting and Miocene uplift (Witts et al., 2012).
-The Kutei Basin is characterized by eastward-prograding deltaic and shallow shelf deposits that have been steadily supplied with sediment from the Schwaner Mountains of SW Kalimantan and the Central Range (van de Weerd and Armin, 1992).
-North of the Kutei Basin, the deltaic deposits found in the Tarakan Basin similarly prograde eastward from the mid Miocene onward, fed by the northern drainages of the Central Range (Satyana et al., 1999).
+Our reconstructions of the paleoshorelines of Borneo and Palawan are largely informed by extant paleogeographic constructions detailed in \citet{Hall2001a, Hall2013a, Hall2013b},  as well as geologic maps and local shoreline reconstructions found in \citet{vandeWeerd1992a, Witts2012a, Madon2013a, Kessler2015a}.
+Comprised of Paleozoic to Mesozoic crustal components that were largely accreted to Sundaland by the late Cretaceous \citep{Metcalfe2013a}, the southern and western portions of Borneo (SW Kalimantan) have been subaerially exposed throughout the Neogene \citep{Hall2013a, Hall2013b}.
+A collision along the northern margin of Borneo associated with the initiation of rifting in the South China Sea resulted in the late Eocene uplift of the Central Range mountains \citep{Hutchison1996a}, which provide sediment to basins along the Southern (Kalimantan) and northern (Sabah and Sarawak) coasts of the island.
+In the east, the Barito, Kutei, and Tarakan basins developed as a single area of subsidence (associated with the opening of the Makassar Straits in the Eocene) before the basins were isolated by Oligocene faulting and Miocene uplift \citep{Witts2012a}.
+The Kutei Basin is characterized by eastward-prograding deltaic and shallow shelf deposits that have been steadily supplied with sediment from the Schwaner Mountains of SW Kalimantan and the Central Range \citep{vandeWeerd1992a}.
+North of the Kutei Basin, the deltaic deposits found in the Tarakan Basin similarly prograde eastward from the mid Miocene onward, fed by the northern drainages of the Central Range \citep{Satyana1999a}.
 In SE Kalimantan, the Barito Basin is bounded to the east by the ophiolite-bearing Meratus Mountains.
-Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene (Witts et al., 2012).
+Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene \citep{Witts2012a}.
 
-Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan (Hall et al., 2008; Ilao et al., 2018).
-The Sarawak Basin and NW Borneo trough offshore Sabah host >10 km-thick Neogene sediementary sequences, indicative of the extent and duration of exhumation in northern Borneo (Hall et al., 2008).
-The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan (Hall, 2013a), as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
-Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities (Hall et al., 2008).
-Early Miocene sediments in northern Sabah have a provenance from Palawan (Suggate et al., 2014), indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny.
-However, initiation of back-arc extension in the Sulu Sea around 19 Ma (Hall, 2013a) resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
-The 14 Ma Capoas Granite on Palawan and the 7.5 Ma Kinabalu Granite in Sabah are the products of crustal thinning related to pulses of regional backarc extension (Hall, 2013a).
-Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7km/Myr) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
-Kinabalu, the tallest peak in Borneo (Cottam et al., 2013).
+Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \citep{Hall2008a, Ilao2018a}.
+The Sarawak Basin and NW Borneo trough offshore Sabah host $\textgreater$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}.
+The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \citep{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
+Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities \citep{Hall2008a}.
+Early Miocene sediments in northern Sabah have a provenance from Palawan \citep{Suggate2014a}, indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny.
+However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
+The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are the products of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
+Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
+Kinabalu, the tallest peak in Borneo \citep{Cottam2013a}.
 
@@
 
-For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of Nugraha and Hall (2018).
+For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of \citet{Nugraha2018a}.
 Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms.
-These arms have high relief >3km that are separated by deep basins and broadly define Sulawesi’s seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi (Hamilton, 1979; Katili, 1978).
-Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system (Hall, 2017; Hennig et al., 2017a; Hennig et al., 2016; Parkinson, 1998b).
-After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed (Elburg and Foden, 1998; Polvé et al., 1997) that was connected to the paleo-Sunda Arc (Hall, 2002), also referred to as the Great Indonesian arc (Harris, 2006).
-The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (Monnier et al., 1995) [but see (Kadarusman et al., 2004) for an alternative view], and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32-28 Ma; Parkinson, 1998a) and thrust to the east above a west dipping slab (East Sulawesi block of Villeneuve et al., 2002).
-During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton (Bergman et al., 1996; Smith and Silver, 1991).
+These arms have high relief $\textgreater$3~km that are separated by deep basins and broadly define Sulawesi’s seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
+Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
+After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.
+The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (\citealp{Monnier1995a}, but see \citealp{Kadarusman2004a} for an alternative view), and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block of \citealp{Villeneuve2001a}).
+During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.
 
-The North Arm consists of Late Pliocene and younger volcanic rocks built upon Eocene to Early Miocene oceanic basalt, basaltic andesite, pelagic sediments, and metamorphic rocks (Elburg and Foden, 1998).
-Eocene to Early Miocene volcanic rocks formed above a NW dipping slab in an arc system that extended to West and South Sulawesi (van Leeuwen and Muhardjo, 2005).
+The North Arm consists of Late Pliocene and younger volcanic rocks built upon Eocene to Early Miocene oceanic basalt, basaltic andesite, pelagic sediments, and metamorphic rocks \citep{Elburg1998a}.
+Eocene to Early Miocene volcanic rocks formed above a NW dipping slab in an arc system that extended to West and South Sulawesi \citep{vanLeeuwen2005a}.
 Gorontalo Bay is an extensional basin that formed in the Pliocene, and prior to that time, the East Sulawesi was attached to the North Arm of Sulawesi.
-After the Oligocene to Early Miocene accretion of the East Sulawesi ophiolite, the arc system stepped out to the SE forming a Neogene (23-16 Ma) volcanic belt on the East arm (Kadarusman et al., 2004).
+After the Oligocene to Early Miocene accretion of the East Sulawesi ophiolite, the arc system stepped out to the SE forming a Neogene (23--16~Ma) volcanic belt on the East arm \citep{Kadarusman2004a}.
 In the Late Miocene, SE-dipping subduction was initiated below the North Arm, which continues today at the North Sulawesi Trough.
-North Sulawesi was not substantially emergent until the Pliocene (van Leeuwen and Muhardjo, 2005).
+North Sulawesi was not substantially emergent until the Pliocene \citep{vanLeeuwen2005a}.
 Active volcanoes extend from the North Arm through the Sangihe Arc into the southern Philippines.
 
-The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000 km2 (Monnier et al., 1995).
-The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, mélange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint (Monnier et al., 1995; Parkinson, 1998a; Silver et al., 1983).
-Locally, the structural thickness of the ophiolite exceeds 15 km with surface relief towering over 3 km (Kadarusman et al., 2004).
-Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93-32 Ma, clustering between 60 and 40 Ma (Parkinson, 1998, and references therein).
-K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36-23 Ma (Parkinson, 1998a; Villeneuve et al., 2002), which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) (Parkinson, 1998a; Silver et al., 1983); however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
+The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000~km$^{2}$ \citep{Monnier1995a}.
+The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, m\'elange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint \citep{Silver1983a, Monnier1995a, Parkinson1998a}.
+Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief towering over 3~km \citep{Kadarusman2004a}.
+Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma (\citealp{Parkinson1998b} and references therein).
+K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36--23~Ma \citep{Parkinson1998a, Villeneuve2001a}, which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) \citep{Silver1983a, Parkinson1998a}; however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
 
-Collision of the Sula-Banggai block with the East margin of Sulawesi began in the latest Miocene and uplift accelerated in the early Pliocene (5.2 to 3.8 Ma) and is associated with a major pulse of sedimentation in adjacent basins (Davies, 1990; Villeneuve et al., 2000).
-Off the northeast margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes mélange (Davies, 1990).
+Collision of the Sula-Banggai block with the East margin of Sulawesi began in the latest Miocene and uplift accelerated in the early Pliocene (5.2~Ma to 3.8~Ma) and is associated with a major pulse of sedimentation in adjacent basins \citep{Davies1990a, Villeneuve2000a}.
+Off the northeast margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes m\'elange \citep{Davies1990a}.
 Thus, although the East Sulawesi ophiolite was trapped/emplaced onto the composite Sundaland margin (i.e.
-a fragment of previously accreted Australian crust) between 36 and 23 Ma, mafic and ultramafic rocks on the south and west arms appear to have not been substantially subaerially exposed in the south and west arms until after 15 Ma, and on the east arm until after 5 Ma.
+a fragment of previously accreted Australian crust) between 36~Ma and 23~Ma, mafic and ultramafic rocks on the south and west arms appear to have not been substantially subaerially exposed in the south and west arms until after 15~Ma, and on the east arm until after 5~Ma.
 
-West Sulawesi rifted from Borneo during the Eocene forming the Makassar Straits back arc basin behind a southwest-facing arc (Polvé et al., 1997).
-Eroded fragments of ophiolite and extensive belts of volcanic rocks are preserved on the West and South Arms of Sulawesi (Bergman et al., 1996; van Leeuwen et al., 2010).
+West Sulawesi rifted from Borneo during the Eocene forming the Makassar Straits back arc basin behind a southwest-facing arc \citep{Polve1997a}.
+Eroded fragments of ophiolite and extensive belts of volcanic rocks are preserved on the West and South Arms of Sulawesi \citep{Bergman1996a, vanLeeuwen2010a}.
 During the Middle Miocene (ca.
-15-13 Ma), extensional faults in the Bone basin reversed, which was accompanied by uplift and erosion of the Bone Mountain ophiolite and Lamasi complex in West Sulawesi (Bergman et al., 1996; van Leeuwen et al., 2010).
-Shortening was likely due to the collision of the leading edge of the Buton-Tukang Besi block, which collided with Buton and the SE Arm of Sulawesi (Smith and Silver, 1991).
-Uplift and erosion is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins (Bergman et al., 1996; van Leeuwen et al., 2010).
-Uplift was diachronous, not effecting units below and to the west of the Lamasai ophiolite until the Middle Miocene (~13 Ma) (van Leeuwen et al., 2010).
+15--13~Ma), extensional faults in the Bone basin reversed, which was accompanied by uplift and erosion of the Bone Mountain ophiolite and Lamasi complex in West Sulawesi \citep{Bergman1996a, vanLeeuwen2010a}.
+Shortening was likely due to the collision of the leading edge of the Buton-Tukang Besi block, which collided with Buton and the SE Arm of Sulawesi \citep{Smith1991a}.
+Uplift and erosion is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins \citep{Bergman1996a, vanLeeuwen2010a}.
+Uplift was diachronous, not effecting units below and to the west of the Lamasai ophiolite until the Middle Miocene (ca.
+13~Ma) \citep{vanLeeuwen2010a}.
 Alkali volcanism ensued at ca.
-11 Ma and is associated with a second phase of extension and exhumation.
-From the Late Miocene to Pliocene fission track ages imply deep exhumation (Bergman et al., 1996; Smith and Silver, 1991; van Leeuwen et al., 2010).
+11~Ma and is associated with a second phase of extension and exhumation.
+From the Late Miocene to Pliocene fission track ages imply deep exhumation \citep{Smith1991a, Bergman1996a, vanLeeuwen2010a}.
 
-Fission track ages from granitoids in central Sulawesi indicate rapid uplift of Central Sulawesi (200–700 m/Myr) starting at about 5 Ma associated with movement on the Palo-Koro fault (Bellier et al., 2006).
-The fault system also shows a normal component with rapid exhumation of rock west of the fault in western Sulawesi – all fission track dates are younger than 5 Ma (Bellier et al., 2006).
-Just east of the Palo-Koro fault, the Palu Metamorphic Complex was exhumed in the Late Miocene to early Pliocene in the north (~5.3 Ma) and later Pliocene in the south (~3.1–2.7 Ma) at rates of up to 400 m/Myr (Hennig et al., 2017b).
+Fission track ages from granitoids in central Sulawesi indicate rapid uplift of Central Sulawesi (200--700~m/m.y.) starting at about 5~Ma associated with movement on the Palo-Koro fault \citep{Bellier2006a}.
+The fault system also shows a normal component with rapid exhumation of rock west of the fault in western Sulawesi -- all fission track dates are younger than 5~Ma \citep{Bellier2006a}.
+Just east of the Palo-Koro fault, the Palu Metamorphic Complex was exhumed in the Late Miocene to early Pliocene in the north (ca.
+5.3~Ma) and later Pliocene in the south (ca.
+3.1--2.7~Ma) at rates of up to 400~m/m.y.
+\citep{Hennig2017b}.
 
@@
 
-Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of Cloos et al.
-(2005), Gold et al.
-(2017), Harrington et al.
-(2017), and Nichols and Hall (1991).
+Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of \citet{Nichols1991a}, \citet{Cloos2005a}, \citet{Gold2017a}, and \citet{Harrington2017a}.
 These studies were all based to varying degrees on lithological distributions, biostratigraphic, borehole data, and tectonic models.
@@
 
-Northern New Guinea, including the Melanesian Arc, was emplaced above the Australian plate during the Miocene (Baldwin et al., 2012; Cloos et al., 2005; Hamilton, 1979; van Ufford and Cloos, 2005).
-Two major ophiolite belts marking the suture—the Irian-Marum ophiolite belt (including the April ultramafics), and the Papuan Ultramafic Belt (PUB)—are preserved along the Central Range and Peninsular Range, respectively.
-South of the Irian-Marum ophiolites, the Ruffaer Metamorphic Complex constructed from the accretionary wedge, forms the spine of the Central Range (up to ~5 km elevation today).
+Northern New Guinea, including the Melanesian Arc, was emplaced above the Australian plate during the Miocene \citep{Hamilton1979a, Cloos2005a, vanUfford2005a, Baldwin2012a}.
+Two major ophiolite belts marking the suture -- the Irian-Marum ophiolite belt (including the April ultramafics), and the Papuan Ultramafic Belt (PUB) -- are preserved along the Central Range and Peninsular Range, respectively.
+South of the Irian-Marum ophiolites, the Ruffaer Metamorphic Complex constructed from the accretionary wedge, forms the spine of the Central Range (up to $\sim$5~km elevation today).
 
-The Middle Miocene (16–14 Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt (Cloos et al., 2005; Visser and Hermes, 1962).
+The Middle Miocene (16--14~Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt \citep{Visser1962a, Cloos2005a}.
 This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
-12 Ma (van Ufford and Cloos, 2005).
+12~Ma \citep{vanUfford2005a}.
 Mountain building began in Late Miocene time, ca.
-8-7 Ma (Baldwin et al., 2012; van Ufford and Cloos, 2005), but major relief was not generated until the Pliocene (Weiland and Cloos, 1996).
-Similarly, the Marum ophiolite was uplifted between 8-5 Ma with 3-4 km of denudation (Crowhurst et al., 1996).
-An estimated 80-100 km of shortening has been accommodated by deformation on the south side of the Central Range (Cloos et al., 2005; Hill and Gleadow, 1989).
+8--7~Ma \citep{vanUfford2005a, Baldwin2012a}, but major relief was not generated until the Pliocene \citep{Weiland1996a}.
+Similarly, the Marum ophiolite was uplifted between 8--5~Ma with 3--4~km of denudation \citep{Crowhurst1996a}.
+An estimated 80--100~km of shortening has been accommodated by deformation on the south side of the Central Range \citep{Hill1989a, Cloos2005a}.
 
-Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene (Davies and Jaques, 1984; Davies and Smith, 1971; van Ufford and Cloos, 2005).
+Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene \citep{Davies1971a, Davies1984a, vanUfford2005a}.
 Miocene to present arc-continent collision in New Guinea has progressed from west to east.
 Exhumation of the Central Range accelerated over the past 4~m.y.
-due to slab-breakoff and buoyant uplift (Cloos et al., 2005).
-This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of ophiolites both along the coast (Monnier et al., 1999) and on the islands Obi (Ali et al., 2001) and Halmahera (Ballantyne, 1992; Hall et al., 1988).
-In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4 Ma (van Ufford and Cloos, 2005), which is well dated with a change in provenance from continental to volcanic detritus (Abbott et al., 1994) and thermochronology (Hill and Gleadow, 1989).
+due to slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a}and Halmahera \citep{Hall1988a, Ballantyne1992a}.
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
 
-In the Bird’s Neck of western New Guinea, the Lengguru foldbelt formed during Middle to Late Miocene CW rotation and obduction of the Weyland Terrane (Bailly et al., 2009).
-Recent CCW rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge (Sapin et al., 2009).
+In the Bird’s Neck of western New Guinea, the Lengguru foldbelt formed during Middle to Late Miocene clockwise rotation and obduction of the Weyland Terrane \citep{Bailly2009a}.
+Recent counter-clockwise rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge \citep{Sapin2009a}.
 
@@
 
-Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall (Hall, 1997), coupled with geologic maps of the archipelago (Geological Survey Division, 1963).
+Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall \citep{Hall1997a}, coupled with geologic maps of the archipelago (Geological Survey Division, 1963).
 The Philippine ophiolites have been grouped in four distinct belts (Yumul, 2007).

### 2019-09-13 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable (Fig.
-\ref{fig:shoreline_growth}).
+1 in main text).
 
@@
 
-Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall \citep{Hall1997a}, coupled with geologic maps of the archipelago (Geological Survey Division, 1963).
-The Philippine ophiolites have been grouped in four distinct belts (Yumul, 2007).
-We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone (Fig.
-3; sutures 11 and 12).
-Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments (Pubellier et al., 1991).
-The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present (Yumul Jr et al., 2013).
-In addition to collision, arc magmatism contributed significantly to crustal growth in the Philippines (Dimalanta and Yumul, 2003).
+Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall \citep{Hall1997a}, coupled with geologic maps of the archipelago \citep{GeologicalSurveyDivision1963a}.
+The Philippine ophiolites have been grouped in four distinct belts \citep{Yumul2007a}.
+We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone.
+Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments \citep{Pubellier1991a}.
+The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present \citep{Yumul2013a}.
+In addition to collision, arc magmatism contributed significantly to crustal growth in the Philippines \citep{Dimalanta2006a}.
 
@@
 
-Paleoshoreline reconstructions of the Sunda-Banda Arc system relied largely on the work of Hall (2013b).
-The Sunda-Banda Arc system is composed of thick sequences of mixed sedimentary rocks and basement intruded and overlain by igneous rocks derived from continual arc magmatism through much of the Cenozoic (Hall, 2017).
+Paleoshoreline reconstructions of the Sunda-Banda Arc system relied largely on the work of \citet{Hall2013a}.
+The Sunda-Banda Arc system is composed of thick sequences of mixed sedimentary rocks and basement intruded and overlain by igneous rocks derived from continual arc magmatism through much of the Cenozoic \citep{Hall2017a}.
 Sumatra is the largest island within the Sunda-Banda Arc system that stretches from the Andaman Sea in the northwest to the Banda Sea in the east.
-The western portion of the Sumatra is underlain by the Wolya ophiolite, which was obducted in the Cretaceous and exhumed in the Eocene to Oligocene (Allen et al., 2008), whereas the southeast corner on the islands of Bangka and Belitung is composed of granite basement of the Sunda Shelf (Hall, 2009, 2013b).
-Java is composed of a complex of E-W-striking deformational and magmatic belts (Audley-Charles, 2004).
-The belts have a volcanic arc in the south and a continental shelf to the north with an adjacent sedimentary basin (Audley-Charles, 2004).
-Bali and Flores are mostly volcanic with some sedimentary cover (Hall, 2009).
+The western portion of the Sumatra is underlain by the Wolya ophiolite, which was obducted in the Cretaceous and exhumed in the Eocene to Oligocene \citep{Allen2008a}, whereas the southeast corner on the islands of Bangka and Belitung is composed of granite basement of the Sunda Shelf \citep{Hall2009a, Hall2013b}.
+Java is composed of a complex of E-W-striking deformational and magmatic belts \citep{AudleyCharles2004a}.
+The belts have a volcanic arc in the south and a continental shelf to the north with an adjacent sedimentary basin \citep{AudleyCharles2004a}.
+Bali and Flores are mostly volcanic with some sedimentary cover \citep{Hall2009a}.
 
-Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate (Harris, 2006).
-From ~20-10 Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20 Ma (Hall, 2009, 2013b).
-Six major islands were subaerially exposed before 5 Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores.
-However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5 Ma (Hall, 2009, 2013b).
-Most of the non-volcanic islands of the Outer Banda Arc emerged after 5 Ma, associated with slab roll-back and collision with the Australian continental margin (Audley-Charles, 2004; Hall, 2013b; Harris, 2006).
-In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations >1 km above sea level with estimated average uplift rates of 1.5 km/Myr (Audley-Charles, 1986).
+Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \citep{Harris2006a}.
+From ca.
+20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20~Ma \citep{Hall2009a, Hall2013b}.
+Six major islands were subaerially exposed before 5~Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores.
+However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5~Ma \citep{Hall2009a, Hall2013b}.
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}.
+In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $\textgreater$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+\citep{AudleyCharles1986a}.
 

### 2019-09-13 Manuscript/Manuscript.tex

@@
 \newcommand{\SrSr}{$^{87}$Sr/$^{86}$Sr\xspace}
+\newcommand{\OsOs}{$^{187}$Os/$^{188}$Os\xspace}
 \newcommand{\permil}{\textperthousand\xspace}
@@
 
-Indonesia and New Guinea (I\&NG) have an outsized influence on modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system (\citealp{Molnar2015a, Hall2017a}; Fig.
-\ref{fig:shoreline_growth}).
-The collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of I\&NG and the exhumation of large supra-subduction ophiolites and associated forearc sediments in the warm, wet tropics \citep{Macdonald2019a}.
+Indonesia and New Guinea (I\&NG), part of what is often referred to as the Maritime Continent, have an outsized contribution to modern chemical weathering fluxes.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system \citep{Molnar2015a, Hall2017a, Macdonald2019a}.
+Concurrently, after the Miocene Climatic Optimum (MCO), a cooling trend began ca.
+14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
+culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
+Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
+(Fig.
+\ref{fig:shoreline_growth}), then use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the magnitude of change in steady-state \pCOtwo associated with this emergence (Fig.
+\ref{fig:scenario_pCO2}).
+Not only does the inflection in the rate of increase in subaerially-exposed land area of I\&NG ca.
+4~Ma coincide with that in the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
+This result highlights the importance of highly weatherable regions in setting Earth's long-term climate state.
 
-Given that the region is a modern-day hot spot of fresh water and sediment discharge \citep{Milliman2013a} and chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has the emergence of I\&NG had on Earth's geochemical cycles and climate state?
-If this region was instead under water, would we have a glacial climate today?
-Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
+%Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
 After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
@@
 The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}, with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
-Given that the \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are within the mafic category.
 Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
 The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+We note that, given that \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are grouped into the mafic category.
+As a result, our Ca+Mg concentration assignment of the mafic category is likely an underestimate in regions of obducted ophiolites, which means that the estimated effectiveness of these regions on changing steady-state \pCOtwo is conservative.
 The weathering of carbonate does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
@@
 
-Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), four parameters in particular are poorly constrained in the existing literature: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
+Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
 As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are also not well known.
@@
 Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2018a}.
-We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins.
-We eliminate all parameter combinations that result in a low $r^{2}$ (defined as \textless0.4, see SI).
-This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins, and eliminate all parameter combinations that result in a low $r^{2}$.
+This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
 
-Implementing these parameter combinations in the silicate weathering component of GEOCLIM, we estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
+Having calibrated the silicate weathering component of GEOCLIM, we now use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
 For the climate model component of GEOCLIM, we obtain modeled temperature and runoff fields at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; see SI).
@@
 1860, allowing the effect of \pCOtwo on climate to be isolated.
-
-We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y.
-(Fig.
-\ref{fig:shoreline_growth}).
+For reconstructions of the paleoshoreline of I\&NG over the past 20~m.y., we use geologic data (Fig.
+\ref{fig:shoreline_growth}; see SI).
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
-Terrestrial and marine sedimentary deposits were used to determine the position of paleoshorelines (details in SI).
+Terrestrial and marine sedimentary deposits were used to determine the position of paleoshorelines.
 
@@
 Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
-which has been intrepreted to be associated with slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+which has been interpreted to be associated with slab-breakoff and buoyant uplift \citep{Cloos2005a}.
 In eastern New Guinea, progressive jamming of the north-dipping subduction zone has also caused major uplift over the past 4~m.y.
-\citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
+\citep{vanUfford2005a}, which is well dated based on a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
 Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013b}.
-In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
 \citep{AudleyCharles1986a}.
-These drivers lead to progressive emergence, as seen in the paleoshoreline reconstructions, over the past 15 million years that accelerated following 5 Ma (Fig.
-\ref{fig:shoreline_growth}.
-
-Along with the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, there has also been a decrease in subaerially-exposed land area on the Sunda Shelf \citep{Sarr2019a}.
-Large portions of this low-relief area of continental crust were likely emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}, potentially driving climate amplification on orbital timescales \citep{DiNezio2018a}.
-Similarities in terrestrial biotas of Borneo and mainland Southeast Asia suggest the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
-During the Eocene through early Miocene, the rifting of the South China Sea resulted in subsidence of the Sunda Shelf \citep{Morley2013a} and recent subsidence is interpreted to be a result of dynamic topography, triggered by subduction-related mantle flow \citep{Sarr2019a}.
-In the Miocene, this region exposed a mixture of felsic basement rocks and siliciclastic sediments \citet{Darmadi2007a}.
-The relatively low-relief and felsic lithologies of the region contrasts with the high-relief and mafic lithologies of most of I\&NG.
+These drivers lead to progressive emergence over the past 15~m.y.
+that accelerated following 5~Ma (Fig.
+\ref{fig:shoreline_growth}).
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
 
@@
 \ref{fig:scenario_pCO2}).
+The ``reduced I\&NG + Sunda Shelf'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
+15~Ma.
+Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations (see SI).
+While we modify the shorelines, the position of the tectonic blocks remain fixed since we use a climate model forced with modern paleogeography.
+While there has been motion of these tectonic blocks since 15~Ma, it has been minimal, and therefore this fixed scenario is a good approximation of the paleogeography (see SI).
+The ``reduced I\&NG'' scenario is identical to the ``reduced I\&NG + Sunda Shelf'' scenario, but excludes the Sunda Shelf.
 The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
-The ``reduced I\&NG'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
-15~Ma, excluding the Sunda Shelf.
-The ``reduced I\&NG + Sunda Shelf'' is identical to the ``reduced I\&NG'' scenario with the addition of the Sunda Shelf.
-Given that no land in the Sunda Shelf is currently exposed, across the shelf we distribute felsic and siliciclastic lithologies, assign shallow slope values, and interpolate runoff across latitude bands (see SI).
 
-Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598-729~ppm (5-95 percentile values), the ``reduced I\&NG'' scenario of 547-638~ppm, and the ``reduced I\&NG + Sunda Shelf'' of 533-611~ppm (Fig.
+Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598--729~ppm (5--95 percentile values), the ``reduced I\&NG'' scenario of 547--638~ppm, and the ``reduced I\&NG + Sunda Shelf'' of 533--611~ppm (Fig.
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$500-700~ppm.
-Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Mg+Ca concentration of the lithologies that comprise the Sunda Shelf as well as its low slope.
-These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is itself sufficient to explain climate change over the Neogene.
+These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$500--700~ppm.
+Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Ca+Mg concentration of the lithologies that comprise the Sunda Shelf as well as its low slope.
+These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is itself sufficient to explain the long-term cooling of Earth's climate over the Neogene.
 
-Our  show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
-This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda Arc and New Guinea in the past 4~m.y.
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
-Not only does the timing of the increase in subaerially-exposed land area of I\&NG coincide with the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
-
-Enhance weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is an hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-To explore the sensitivity of global weatherability and climate state to the presence of the surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM that are now possible with given the spatial resolution of lithology (Fig.
+Separately, enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the sensitivity of global weatherability and climate state to the presence of the modern-day surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM that are now possible with the spatial resolution of lithology (Fig.
 \ref{fig:scenario_pCO2}).
@@
 These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
-This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in these igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is significantly smaller than that of high-relief I\&NG.
+This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in these igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is smaller than that of the wetter and higher-relief I\&NG.
 
-First, the tested scenarios only simulate changes in the location of the shorelines in I\&NG and the lithology exposed in the Deccan and Ethiopian Traps.
-Changes in paleogeography that are not accounted for in our simulations may have resulted in changes in oceanic and atmospheric circulation \citep{Haug1998a, Shevenell2004a, Holbourn2015a, Molnar2015a} that contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales.
-However, the global extent and multi-million year timescale of the long-term cooling trend since 14~Ma suggests that the primary driver for this cooling is changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
-In this way, changes in the flux of \COtwo outgassing via changes in spreading rates and ridge lengths \citep{Rowan2016a} and changes in weatherability via mountain building and the movement of land masses into/out of areas of high temperature and runoff are likely more important paleogeography-related factors in terms of modulating Neogene cooling than changes in oceanic and atmospheric circulation.
-Second,
+Other hypotheses that have been proposed to explain cooling over the Neogene include changes in ocean circulation and upwelling \citep{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \citep{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \citep{Haug1998a}, an increase in the strength of the Walker Circulation \citep{Molnar2015a}, and feedbacks with terrestrial vegetation \citep{Cerling1997a}.
+Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the multi-million year timescale of the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Uplift in the Himalayas \citep{Raymo1988a} or a decrease in volcanic outgassing \citep{Berner1983a, Rowan2016a} have also been proposed as drivers for Neogene cooling.
+Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies.
+However, this rise in \SrSr since ca.
+35~Ma is punctuated with a distinct decrease in slope ca.
+15~Ma.
+This inflection has been attributed to exhumation and weathering of the relatively low \SrSr Lesser Himalaya tectonic unit and related strata \citep{Myrow2015a, Colleps2018a}, but could also be explained by exhumation and weathering of low \SrSr mafic and ultramafic lithologies in I\&NG, which began to emerge at this time.
+Furthermore, while global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, the fact that the inflection in the rate of increase in subaerially-exposed land area of I\&NG ca.
+4~Ma coincides with that in the cooling trend and that our coupled weathering-climate model estimates \pCOtwo change associated with emergence of I\&NG is sufficient to explain climate change over the Neogene, this study suggests that the global climate state is particularly sensitive to the tropical paleogeography of I\&NG.
 
 %%% RELATIVELY COMPLETE TEXT ABOVE - BELOW ARE SNIPPETS THAT COULD BE INCORPORATED IN
+
+%Our  show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
+This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda Arc and New Guinea in the past 4~m.y.
+This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
+Not only does the timing of the increase in subaerially-exposed land area of I\&NG coincide with the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
 

### 2019-09-13 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state.
-A rock particle leaves the unweathered bedrock with production rate $P$, and transits vertically through a regolith of height $h$.
-Regolith production and physical erosion ($E$) are equal at steady-state.
+A rock particle leaves the unweathered bedrock with production rate $P_{r}$, and transits vertically through a regolith of height $h$.
+Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
 As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).}
@@
 \begin{equation}
-    \frac{dh}{dt} = P - E
+    \frac{dh}{dt} = P_{r} - E_{p}
     \label{eq:1}
@@
 \begin{equation}
-    \frac{\partial x}{\partial t} = -P \frac{\partial x}{\partial z} - K \tau^{\sigma}x
+    \frac{\partial x}{\partial t} = -P_{r} \frac{\partial x}{\partial z} - K \tau^{\sigma}x
     \label{eq:2}
@@
 \begin{equation}
-    \frac{\partial \tau}{\partial t} = -P \frac{\partial \tau}{\partial z} + 1
+    \frac{\partial \tau}{\partial t} = -P_{r} \frac{\partial \tau}{\partial z} + 1
     \label{eq:3}
@@
 \noindent
-Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P$ is the regolith production rate (m/yr), and $E$ is the physical erosion rate (m/yr).
+Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
 Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth (where $z$ is the height above the base of the regolith; m).
@@
 \begin{equation}
-    P = P_{0}\;f(h)
+    P_{r} = P_{0}\;f(h)
     \label{eq:5}
@@
 \begin{equation}
-    E = k_{e}\;q^{m}\;s^{n}
+    E_{p} = k_{e}\;q^{m}\;s^{n}
     \label{eq:8}
@@
 \begin{aligned}
-& h   \ =\    \max\left(  \;  0  \;,\;  d_o \log\left(\frac{P_o}{E}\right)  \right)                                                 \\
-& x(z)      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{z}{E}\right)^{\sigma+1}   \right)   \\
-& W        \ =\   (1-x(h)) \, E  \ =\   (1-x_s) \, E                                                                         \\
+& h   \ =\    \max\left(  \;  0  \;,\;  d_{0} \log\left(\frac{P_{0}}{E_{p}}\right)  \right)                                                 \\
+& x(z)      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{z}{E_{p}}\right)^{\sigma+1}   \right)   \\
+& W        \ =\   (1-x(h)) \, E_{p}  \ =\   (1-x_s) \, E_{p}                                                                         \\
 \end{aligned}
@@
 
-However, as discussed in the main text, $k_{d}$ (Equation \ref{eq:9}), $k_{w}$ (Equation \ref{eq:9}), $\sigma$ (Equation \ref{eq:4}), and $k_{rp}$ (Equation \ref{eq:8}) are poorly constrained.
+Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$)
+
+However, the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
@@
 Relative to felsic lithologies, mafic lithologies contain minerals with faster dissolution kinetics (e.g.
-olivine and plagioclase; see SI), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
+olivine and plagioclase), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
+We also do not include the effects of reverse weathering \citep{Michalopoulos1995a}, which could modulate the relationship between liberation of Ca and Mg during silicate weathering and long-term \COtwo consumption.
 

### 2019-09-13 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-Indonesia and New Guinea (I\&NG), part of what is often referred to as the Maritime Continent, have an outsized contribution to modern chemical weathering fluxes.
+Indonesia and New Guinea (I\&NG) have an outsized contribution to modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
 There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system \citep{Molnar2015a, Hall2017a, Macdonald2019a}.
-Concurrently, after the Miocene Climatic Optimum (MCO), a cooling trend began ca.
-14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
-culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
+Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
+14~Ma and accelerated over the past 4~m.y.
+culminating with the onset of Northern Hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
 Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
 (Fig.
-\ref{fig:shoreline_growth}), then use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the magnitude of change in steady-state \pCOtwo associated with this emergence (Fig.
+\ref{fig:shoreline_growth}) and quantify the increase in land area.
+We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence (Fig.
 \ref{fig:scenario_pCO2}).
-Not only does the inflection in the rate of increase in subaerially-exposed land area of I\&NG ca.
-4~Ma coincide with that in the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
+We find that the magnitude of change in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
 This result highlights the importance of highly weatherable regions in setting Earth's long-term climate state.
-
-%Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
-After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
-14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
-culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
-The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}.
-The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
-To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, we use Neogene paleoshoreline reconstructions and couple a weathering model to climate model results to estimate the magnitude of changes in \pCOtwo that are coupled with changes in tropical paleogeography.
 
@@
 The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
-For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady-state is achieved at a higher \pCOtwo.
+For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
 Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower steady-state \pCOtwo when global weatherability is high \citep{Kump1997a}.
@@
 Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance chemical weathering \citep{West2012a, Maher2014a}.
-In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}.
+In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg \citep{Dessert2003a}.
 These factors are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo and initiating glacial climate over the past 520~m.y.
@@
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}, with the 16 lithologic categories represented in GLiM grouped into 6 broader categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
 Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO measurements on rocks of each of these lithologic categories in data compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
-We note that, given that \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are grouped into the mafic category.
-As a result, our Ca+Mg concentration assignment of the mafic category is likely an underestimate in regions of obducted ophiolites, which means that the estimated effectiveness of these regions on changing steady-state \pCOtwo is conservative.
+The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO data on rocks of each of these lithologic categories compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+Given that \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
+As a result, the Ca+Mg concentration of the mafic category is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative.
 The weathering of carbonate does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since the chemical composition of these two lithologic classes is strongly dependent on protolith composition.
-The chemical composition of siliciclastic sediment is further strongly dependent on its degree of chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during the calibration of the silicate weathering component of GEOCLIM.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during calibration of the silicate weathering component of GEOCLIM.
 
-Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$).
-As discussed above, the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are also not well known.
-Rather than prescribing values for these six parameters, we instead allow them to vary within reasonable bounds, resulting in a total of 78,000 unique parameter combinations.
-For each of these combinations, we compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg fluxes) using present-day runoff, temperature, and slope fields.
-Following the approach of \citet{Maffre2018a}, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin in the compilation of \citet{Moquet2018a}.
-We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins, and eliminate all parameter combinations that result in a low $r^{2}$.
-This calibration method leaves 3,339 unique parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
+Of the parameters used in the equations that govern the silicate weathering component of GEOCLIM, the values of four parameters that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age within the chemical weathering parameterization of \cite{West2012a} are poorly constrained (see SI).
+Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
+For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
+Following the approach of \citet{Maffre2018a}, we sum computed \COtwo consumption over large-scale watersheds from the global compilation of \citet{Gaillardet1999a} and smaller-scale watersheds of the Amazon Basin  \citep{Moquet2018a}.
+We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins, and eliminate all parameter combinations that result in low $r^{2}$.
+This approach leaves 3,339 parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
 
-Having calibrated the silicate weathering component of GEOCLIM, we now use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
-For the climate model component of GEOCLIM, we obtain modeled temperature and runoff fields at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; see SI).
-The primary strength of using these three GFDL CM2.0 experiments for this analysis is that all non-\COtwo forcing agents are held constant at values representative of conditions ca.
-1860, allowing the effect of \pCOtwo on climate to be isolated.
+Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
+For the climate model component of GEOCLIM, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; see SI).
+The primary strength of using these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
 For reconstructions of the paleoshoreline of I\&NG over the past 20~m.y., we use geologic data (Fig.
 \ref{fig:shoreline_growth}; see SI).
-Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
+Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
 Terrestrial and marine sedimentary deposits were used to determine the position of paleoshorelines.
 
-In New Guinea, emergence in the Middle Miocene is associated with the collision between the Melanesian Arc and the distal margin of Australia \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove the exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
+In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
 Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
-which has been interpreted to be associated with slab-breakoff and buoyant uplift \citep{Cloos2005a}.
-In eastern New Guinea, progressive jamming of the north-dipping subduction zone has also caused major uplift over the past 4~m.y.
-\citep{vanUfford2005a}, which is well dated based on a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
-Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013b}.
-In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations \textgreater1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
-\citep{AudleyCharles1986a}.
-These drivers lead to progressive emergence over the past 15~m.y.
+associated with slab-breakoff and buoyant uplift \citep{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \citep{vanUfford2005a}.
+Many of the islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013b}.
+These tectonic drivers lead to progressive emergence over the past 15~m.y.
 that accelerated following 5~Ma (Fig.
-\ref{fig:shoreline_growth}).
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
+\ref{fig:shoreline_growth}B).
+This trend mirrors broad cooling over the Neogene that resulted in the initiation of Northern Hemisphere ice sheets (Fig.
+\ref{fig:shoreline_growth}C).
 
@@
 Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations (see SI).
-While we modify the shorelines, the position of the tectonic blocks remain fixed since we use a climate model forced with modern paleogeography.
-While there has been motion of these tectonic blocks since 15~Ma, it has been minimal, and therefore this fixed scenario is a good approximation of the paleogeography (see SI).
+While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
+While there has been motion of these tectonic blocks since 15~Ma, it has been minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
 The ``reduced I\&NG'' scenario is identical to the ``reduced I\&NG + Sunda Shelf'' scenario, but excludes the Sunda Shelf.
-The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are entirely removed.
+The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are removed.
 
-Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598--729~ppm (5--95 percentile values), the ``reduced I\&NG'' scenario of 547--638~ppm, and the ``reduced I\&NG + Sunda Shelf'' of 533--611~ppm (Fig.
+Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598--729~ppm (5--95 percentile values), the ``reduced I\&NG'' scenario resulted in 547--638~ppm \pCOtwo, and the ``reduced I\&NG + Sunda Shelf'' resulted in 533--611~ppm \pCOtwo (Fig.
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed land area and tectonic topography associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$500--700~ppm.
-Notably, the effect of the Sunda Shelf on steady-state \pCOtwo is relatively small, given the low Ca+Mg concentration of the lithologies that comprise the Sunda Shelf as well as its low slope.
-These GEOCLIM \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is itself sufficient to explain the long-term cooling of Earth's climate over the Neogene.
+These results suggest that without the Neogene increase in subaerially-exposed mountains associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
+These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain the long-term cooling of Earth's climate over the Neogene.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}, such that these modeled values are consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice pCOtwo  .
 
-Separately, enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a hypothesis that has been put forth to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
-To explore the sensitivity of global weatherability and climate state to the presence of the modern-day surface exposure of basaltic lavas, we test three additional scenarios using GEOCLIM that are now possible with the spatial resolution of lithology (Fig.
+Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a distinct hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we test three additional scenarios using the lithologically-resolved implementation of GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-In the ``removed Deccan Traps'' and the ``removed Ethiopian Traps'' scenarios, all volcanics associated with the respective igneous provinces are replaced with the Ca+Mg concentration of bulk continental crust.
-In the ``removed both traps'' scenario, all volcanics associated with both provinces are replaced with the Ca+Mg concentration of bulk continental crust.
-These scenarios result in steady-state \pCOtwo's \textless500~ppm, distinctly lower than those for the three I\&NG scenarios.
-This result indicates that, while the presence of high Ca+Mg concentration mafic rocks in these igneous provinces provides a non-negligible contribution to modulating steady-state \pCOtwo, its contribution is smaller than that of the wetter and higher-relief I\&NG.
+We replace the  volcanics associated with the respective igneous provinces with the Ca+Mg concentration of bulk continental crust in the model.
+The resulting  \pCOtwo is \textless500~ppm indicating that, while the presence mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
 
-Other hypotheses that have been proposed to explain cooling over the Neogene include changes in ocean circulation and upwelling \citep{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \citep{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \citep{Haug1998a}, an increase in the strength of the Walker Circulation \citep{Molnar2015a}, and feedbacks with terrestrial vegetation \citep{Cerling1997a}.
-Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the multi-million year timescale of the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
-Uplift in the Himalayas \citep{Raymo1988a} or a decrease in volcanic outgassing \citep{Berner1983a, Rowan2016a} have also been proposed as drivers for Neogene cooling.
-Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies.
-However, this rise in \SrSr since ca.
-35~Ma is punctuated with a distinct decrease in slope ca.
-15~Ma.
-This inflection has been attributed to exhumation and weathering of the relatively low \SrSr Lesser Himalaya tectonic unit and related strata \citep{Myrow2015a, Colleps2018a}, but could also be explained by exhumation and weathering of low \SrSr mafic and ultramafic lithologies in I\&NG, which began to emerge at this time.
-Furthermore, while global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, the fact that the inflection in the rate of increase in subaerially-exposed land area of I\&NG ca.
-4~Ma coincides with that in the cooling trend and that our coupled weathering-climate model estimates \pCOtwo change associated with emergence of I\&NG is sufficient to explain climate change over the Neogene, this study suggests that the global climate state is particularly sensitive to the tropical paleogeography of I\&NG.
+Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence  \citep{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with I\&NG emergence \citep{Molnar2015a}.
+Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \citep{DeConto2008a}.
+However, the prolonged timescale of the cooling trend since 14~Ma (Fig.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Uplift in the Himalayas \citep{Raymo1988a} or a decrease in volcanic outgassing \citep{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic sea floor spreading to be relative uniform \citep{Rowan2016a}.
+Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies \citep{Raymo1992a}.
+This rise in \SrSr since ca.
+35~Ma is punctuated with a decrease in slope ca.
+15~Ma that has been attributed to exhumation of relatively low \SrSr Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}, but could also be partially driven by the emergence of low \SrSr mafic lithologies in I\&NG.
+While paleogeographic changes elsewhere on Earth would have also modulated geological carbon sources and sinks, the onset of Northern hemisphere glaciation coincides with increased emergence in I\&NG following 5~Ma.
+Our coupled weathering-climate model indicates that \pCOtwo change associated with emergence of I\&NG is sufficient to explain climate change over the Neogene, suggesting that the global climate state is particularly sensitive to tropical paleogeography.
 
 %%% RELATIVELY COMPLETE TEXT ABOVE - BELOW ARE SNIPPETS THAT COULD BE INCORPORATED IN
-
-%Our  show a broad increase in exposed land area over the Neogene and an acceleration over the past 4~m.y.
-This trend is due to progressive arc-continent collision from west to east that has culminated with increased exhumation in the Banda Arc and New Guinea in the past 4~m.y.
-This trend mirrors broad cooling over the Neogene and accelerated cooling over the past 4~m.y that culminated in the initiation of the Northern Hemisphere ice sheets.
-Not only does the timing of the increase in subaerially-exposed land area of I\&NG coincide with the cooling trend, but our coupled weathering-climate model demonstrates that the magnitude of change in \pCOtwo due to these changes is sufficient to explain climate change over the Neogene.
-
-%%%I\&NG are potent sinks of carbon because they contain large swaths of ophiolites that are exposed along E-W belts of tectonic topography in the warm wet tropics.
 
@@
 
+%Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
+After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
+14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
+culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}.
+The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
+To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, we use Neogene paleoshoreline reconstructions and couple a weathering model to climate model results to estimate the magnitude of changes in \pCOtwo that are coupled with changes in tropical paleogeography.
+
 %%TC:ignore
@@
 We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their contributions to the paleoshoreline reconstructions.
-Collaborative research between NLS-H and YG was initially supported by the France-Berkeley Fund.
-FAM and NLS-H thank NSF FRES grant \#1926001 for support.
+Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
+FAM and NLS-H thank NSF FRES grants \#1926001 and \#1925990for support.
 %%TC:endignore
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{Manuscript/Figures/shoreline_growth.pdf}
-    \caption{The emergence of Indonesia and New Guinea (I\&NG) from the mid-Miocene to present.
-Past shorelines are shown in A with associated area estimates summarized in B.
+    \includegraphics[width=0.9\textwidth]{Figures/shoreline_growth.pdf}
+    \caption{The emergence of Indonesia and New Guinea (I\&NG) from the mid-Miocene to present in the region referred to in climate science literature as the Maritime Continent.
+Past shorelines are shown in A with associated area estimates summarized in B (where I\&NG refers to the sum of all the subdivided regions).
 A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \citep{Zachos2008a}.}
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2.pdf}
+    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2.pdf}
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
@@
 
-\bibliographystyle{gsabull}
+\bibliographystyle{naturemag}
 \bibliography{References}

### 2019-09-14 Manuscript/Manuscript.tex

@@
 Indonesia and New Guinea (I\&NG) have an outsized contribution to modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system \citep{Molnar2015a, Hall2017a, Macdonald2019a}.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
 14~Ma and accelerated over the past 4~m.y.
-culminating with the onset of Northern Hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
+culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
 Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}.
-However, assessing the effect of the emergence of I\&NG on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes.
-The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources.
-For example, as \COtwo sinks are removed and \COtwo rises, temperature increases and the hydrological cycle is invigorated, which causes other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
-Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower steady-state \pCOtwo when global weatherability is high \citep{Kump1997a}.
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate rocks \cite{Kump2000a}.
+However, assessing the effect of the emergence of I\&NG on \pCOtwo levels cannot be determined simply through assessing the present-day chemical weathering fluxes.
+Steady-state \pCOtwo is set by the \pCOtwo level at which \COtwo sinks are equal to sources.
+For example, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
+Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.
 
 Topography, climate, and lithology all play important roles in modulating chemical weathering.
-In low-relief regions, low physical erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \citep{Gabet2009a, West2012a, Maher2014a}.
-In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \citep{Lasaga1994a, West2012a}.
-Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance chemical weathering \citep{West2012a, Maher2014a}.
-In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg \citep{Dessert2003a}.
-These factors are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo and initiating glacial climate over the past 520~m.y.
-\citep{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \citep{Park2019a}.
+In low-relief regions, low erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \cite{Gabet2009a, West2012a, Maher2014a}.
+In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
+Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance chemical weathering \cite{West2012a, Maher2014a}.
+In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}.
+These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo and initiating glacial climate over the past 520~m.y.
+\cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019a}.
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \citep{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to a global climate model run at varying \pCOtwo levels.
-Within the silicate weathering component of GEOCLIM, we use the DynSoil implementation of the \citet{Gabet2009a} regolith development model which integrates a climatic dependence on chemical weathering using the formulation of \cite{West2012a}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to a global climate model run at varying \pCOtwo levels.
+Within the silicate weathering component of GEOCLIM, we implement a regolith development model \cite{Gabet2009a} that integrates a climatic dependence on chemical weathering \cite{West2012a}.
 
 The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell.
-In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions (e.g.
-\citealp{Godderis2014a}).
-However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component based on work by \citet{Heimsath1997a}, \citet{Gabet2009a}, \citet{West2012a}, \citet{Carretier2014a}, \citet{Godderis2017b}, and \citet{Maffre2018a}, which introduces an additional dependence on topographic slope (see SI).
+In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
+However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces an additional dependence on topographic slope (see SI).
 While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
 We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) \cite{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
 Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
 The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO data on rocks of each of these lithologic categories compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
-Given that \citet{Hartmann2012a} does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
+Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
 As a result, the Ca+Mg concentration of the mafic category is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative.
@@
 
-Of the parameters used in the equations that govern the silicate weathering component of GEOCLIM, the values of four parameters that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age within the chemical weathering parameterization of \cite{West2012a} are poorly constrained (see SI).
+The values of four parameters within the silicate weathering component of GEOCLIM
+
+Of the parameters used in the equations that govern the silicate weathering component of GEOCLIM, the values of four parameters within the chemical weathering parameterization \cite{West2012a} that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age are poorly constrained (see SI).
 Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
 For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
-Following the approach of \citet{Maffre2018a}, we sum computed \COtwo consumption over large-scale watersheds from the global compilation of \citet{Gaillardet1999a} and smaller-scale watersheds of the Amazon Basin  \citep{Moquet2018a}.
+We sum computed \COtwo consumption over large-scale watersheds from a global compilation\cite{Gaillardet1999a} and watersheds of the Amazon Basin \cite{Moquet2018a}.
 We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins, and eliminate all parameter combinations that result in low $r^{2}$.
@@
 The primary strength of using these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
-For reconstructions of the paleoshoreline of I\&NG over the past 20~m.y., we use geologic data (Fig.
+For reconstructions of the paleoshoreline  We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 20~m.y.
+and include changes in areas of submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
 \ref{fig:shoreline_growth}; see SI).
-Following \citet{Molnar2015a}, we analyzed the area changes of islands that are greater than $\sim$200~km$^{2}$, and also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
-Terrestrial and marine sedimentary deposits were used to determine the position of paleoshorelines.
-
-In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \citep{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
+In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
 Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
-associated with slab-breakoff and buoyant uplift \citep{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \citep{vanUfford2005a}.
-Many of the islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{Harris2006a, Hall2013b}.
+associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}.
+Many of the islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{Harris2006a, Hall2013b}.
 These tectonic drivers lead to progressive emergence over the past 15~m.y.
@@
 These results suggest that without the Neogene increase in subaerially-exposed mountains associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \citep{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain the long-term cooling of Earth's climate over the Neogene.
-The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}, such that these modeled values are consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice pCOtwo  .
+These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain the long-term cooling of Earth's climate over the Neogene.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
+These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets as \pCOtwo subsequently decreased.
 
-Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a distinct hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \citep{Kent2008a, Kent2013a}.
+Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a distinct hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
 To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we test three additional scenarios using the lithologically-resolved implementation of GEOCLIM (Fig.
@@
 
-Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence  \citep{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with I\&NG emergence \citep{Molnar2015a}.
-Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \citep{DeConto2008a}.
+Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence  \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with I\&NG emergence \cite{Molnar2015a}.
+Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
 However, the prolonged timescale of the cooling trend since 14~Ma (Fig.
-\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
-Uplift in the Himalayas \citep{Raymo1988a} or a decrease in volcanic outgassing \citep{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic sea floor spreading to be relative uniform \citep{Rowan2016a}.
-Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies \citep{Raymo1992a}.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Uplift in the Himalayas \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic sea floor spreading to be relative uniform \cite{Rowan2016a}.
+Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
 This rise in \SrSr since ca.
 35~Ma is punctuated with a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}, but could also be partially driven by the emergence of low \SrSr mafic lithologies in I\&NG.
+15~Ma that has been attributed to exhumation of relatively low \SrSr Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the emergence of low \SrSr mafic lithologies in I\&NG.
 While paleogeographic changes elsewhere on Earth would have also modulated geological carbon sources and sinks, the onset of Northern hemisphere glaciation coincides with increased emergence in I\&NG following 5~Ma.
@@
 14~Ma that accelerated over the past 4~m.y.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling \citep{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \citep{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \citep{Haug1998a}, an increase in the strength of the Walker Circulation \citep{Molnar2015a}, feedbacks with terrestrial vegetation \citep{Cerling1997a}(MAYBE ONE OF THE DINEZIO REFS), uplift in the Himalayas \citep{Raymo1988a}, or a decrease in volcanic outgassing \citep{Berner1983a}(ROWAN+ROWLEY).
-Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \citep{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling \cite{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \cite{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \cite{Haug1998a}, an increase in the strength of the Walker Circulation \cite{Molnar2015a}, feedbacks with terrestrial vegetation \cite{Cerling1997a}(MAYBE ONE OF THE DINEZIO REFS), uplift in the Himalayas \cite{Raymo1988a}, or a decrease in volcanic outgassing \cite{Berner1983a}(ROWAN+ROWLEY).
+Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
 Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of I\&NG.
 
-%Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
+%Based on paleogeographic reconstructions of arc-continent collisions through time, \cite{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
 After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
 14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
-culminating with the onset of northern hemisphere glaciation \citep{Shackleton1984a, Zachos2001a}.
-The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a}.
+culminating with the onset of northern hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
+The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
 The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
@@
 Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
-FAM and NLS-H thank NSF FRES grants \#1926001 and \#1925990for support.
+FAM and NLS-H thank NSF FRES grants \#1926001 and \#1925990 for support.
 %%TC:endignore
@@
 Past shorelines are shown in A with associated area estimates summarized in B (where I\&NG refers to the sum of all the subdivided regions).
-A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \citep{Zachos2008a}.}
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \cite{Zachos2008a}.}
     \label{fig:shoreline_growth}
@@
 
-\bibliographystyle{naturemag}
+\bibliographystyle{nature}
 \bibliography{References}

### 2019-09-14 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-The values of four parameters within the silicate weathering component of GEOCLIM
-
-Of the parameters used in the equations that govern the silicate weathering component of GEOCLIM, the values of four parameters within the chemical weathering parameterization \cite{West2012a} that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age are poorly constrained (see SI).
+The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a} that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age are poorly constrained (see SI).
 Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
 For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
-We sum computed \COtwo consumption over large-scale watersheds from a global compilation\cite{Gaillardet1999a} and watersheds of the Amazon Basin \cite{Moquet2018a}.
-We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins, and eliminate all parameter combinations that result in low $r^{2}$.
-This approach leaves 3,339 parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
+We sum computed \COtwo consumption over watersheds within a global compilation \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
+After eliminating all parameter combinations that result in low $r^{2}$, this approach leaves 3,339 parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
 
 Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
-For the climate model component of GEOCLIM, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; see SI).
+For the climate model component of GEOCLIM, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
 The primary strength of using these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
-For reconstructions of the paleoshoreline  We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 20~m.y.
+We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 20~m.y.
 and include changes in areas of submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
@@
 
-Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is a distinct hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
+Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is another hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
 To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we test three additional scenarios using the lithologically-resolved implementation of GEOCLIM (Fig.
@@
 We replace the volcanics associated with the respective igneous provinces with the Ca+Mg concentration of bulk continental crust in the model.
-The resulting  \pCOtwo is \textless500~ppm indicating that, while the presence mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
+The resulting \pCOtwo is \textless500~ppm indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
 
@@
 15~Ma that has been attributed to exhumation of relatively low \SrSr Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the emergence of low \SrSr mafic lithologies in I\&NG.
-While paleogeographic changes elsewhere on Earth would have also modulated geological carbon sources and sinks, the onset of Northern hemisphere glaciation coincides with increased emergence in I\&NG following 5~Ma.
+While paleogeographic changes elsewhere on Earth would have also modulated geological carbon sources and sinks, the onset of Northern Hemisphere glaciation coincides with increased emergence in I\&NG following 5~Ma.
 Our coupled weathering-climate model indicates that \pCOtwo change associated with emergence of I\&NG is sufficient to explain climate change over the Neogene, suggesting that the global climate state is particularly sensitive to tropical paleogeography.
-
-%%% RELATIVELY COMPLETE TEXT ABOVE - BELOW ARE SNIPPETS THAT COULD BE INCORPORATED IN
-
-%After the MCO, a cooling trend began ca.
-14~Ma that accelerated over the past 4~m.y.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean circulation and upwelling \cite{Shevenell2004a, Holbourn2015a}, perhaps associated with increased organic carbon burial \cite{Vincent1985a, Flower1993a} or the closure of the Central American Seaway \cite{Haug1998a}, an increase in the strength of the Walker Circulation \cite{Molnar2015a}, feedbacks with terrestrial vegetation \cite{Cerling1997a}(MAYBE ONE OF THE DINEZIO REFS), uplift in the Himalayas \cite{Raymo1988a}, or a decrease in volcanic outgassing \cite{Berner1983a}(ROWAN+ROWLEY).
-Although changes in circulation and vegetation may have contributed to changes in \pCOtwo on hundreds of thousands to million-year timescales, the global extent and multi-million year timescale suggests that the long-term cooling trend since 14~Ma is the product of changes in the geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
-Global spreading rates, ridge length, and paleogeography elsewhere on Earth undoubtedly changed through the Neogene contributing to changes in geological carbon sources and sinks, however, here we show that the global climate state is particularly sensitive to the tropical paleogeograpahy of I\&NG.
-
-%Based on paleogeographic reconstructions of arc-continent collisions through time, \cite{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states.
-After the Miocene Climatic Optimum (MCO), a cooling trend began ca.
-14~Ma associated with the growth of the Antarctic ice sheet, and accelerated over the past 4~m.y.
-culminating with the onset of northern hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
-The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-The lower \pCOtwo threshold for northern hemisphere glaciation provides an explanation for southern hemisphere glaciation significantly preceding the development of ice sheets in both hemispheres.
-To evaluate the hypothesis that the increase of subaerially-exposed land area in I\&NG has played a significant role in lowering \pCOtwo and the development of northern hemisphere ice sheets, we use Neogene paleoshoreline reconstructions and couple a weathering model to climate model results to estimate the magnitude of changes in \pCOtwo that are coupled with changes in tropical paleogeography.
 

### 2019-09-19 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 Yuem Park\textsuperscript{1},
+Pierre Maffre\textsuperscript{1},
 Nicholas L. Swanson-Hysell\textsuperscript{1},
-Pierre Maffre\textsuperscript{1},
 Francis A. Macdonald\textsuperscript{2},
@@
 
-The silicate weathering component of the GEOCLIM model implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
+The silicate weathering component of the GEOCLIM model as been improved compare to the previous published version (Goddéris et al, 2017 natgeo).
+The new module implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
 We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
-In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering as a particle transits through the regolith or as a physically weathered particle once it reaches the top.
+In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, oras a physically weathered particle once it reaches the top.
 We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}.
-The transient time-varying version of this regolith model is described by the following system of three equations which form the basis of DynSoil:
+The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil module of GEOCLIM:
 
@@
 \noindent
-Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the time in the model (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
+Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the model time (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
 Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth (where $z$ is the height above the base of the regolith; m).
@@
 $P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
-In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the universal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
+In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
 $f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases.
@@
 $k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
-This formulation is similar to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}.
-This form and these exponent values are supported by various compilations (e.g.
-\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
+This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}.
+This formulation and these exponent values are supported by various compilations (e.g.
+\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is a variability in the proportionality constant that is difficult to capture at a global scale.
 
@@
 \noindent
-$x(z)$ gives the curve of primary phases varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
-Setting $z$ equal to regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the column of regolith.
+$x(z)$ is the abundance profile of primary phases inside the regolith, varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
+Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column.
 This value controls the overall integrated chemical weathering rate out of the regolith.
@@
 
-To calculate the flux of Ca and Mg associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
+We restricted the calculation of the weathering fluxes to the flux of dissolved Mg and Ca originating from continental silicate weathering, as they are the sole contributors to the long term consumption of atmospheric CO2.
+To calculate this flux associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
 \ref{fig:regolith_schematic}).
-Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land.
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we have implemented a lithologically-resolved version of the model.
+Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using a concept of diffuse lithology (it was assume that the weathering rates of shield rocks, basaltic rocks, and carbonate rocks was the same for each continental grid element, when this grid element was submitted to the same climatic conditions).
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, we have implemented a realistic lithologically-resolved version of the model.
 
@@
 \ref{fig:world_lithology}).
-However, to improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees.
+To improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees.
 To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel covered by each of the 6 lithologic categories (or water/ice) is captured by the extra dimension.
@@
 Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:8} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
-However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}.
+However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}).
 We use the value of 42~kJ/mol in our model runs.
 
-Of the various parameters used in the equations that govern the silicate weathering component of GEOCLIM (see SI), the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$)
-
-However, the values of four parameters within the chemical weathering parameterization of \cite{West2012a} are poorly constrained: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
+Within the equations that govern the silicate weathering component of GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM (Table \ref{tab:parameter_combinations}).
+We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM and test 78,000 unique parameter combinations (Table \ref{tab:parameter_combinations}).
 
@@
 80 watersheds in total are used in this study.
-We then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:
+We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:
 
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.pdf}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.jpg}
     \caption{\textbf{A)} Modeled global \COtwo consumption vs.
-the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of the watersheds.
+the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds.
 Each point represents model output using one of the 78,000 parameter combinations.
@@
     \caption{Modeled vs.
-observed \COtwo consumption in watersheds around the world.
+data-constrained \COtwo consumption in watersheds around the world.
 \textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
@@
 
-In our implementation, the mean global CO$_{2}$ consumption calculated from these 3,339 parameter combinations is 5.4$\times10^{12}$~mol/yr, with a standard deviation of 0.9$\times10^{12}$~mol/yr.
+The mean global CO$_{2}$ consumption calculated from these 3,339 parameter combinations is 5.4$\times10^{12}$~mol/yr, with a standard deviation of 0.9$\times10^{12}$~mol/yr.
 These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
@@
 For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf.
-This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, but it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
+This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
 
-\section*{Paleoshorelines}
+\section*{Limitations and Future Directions}
+
+As with any modeling study, there are limitations with the approach.
+We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
+While this model is effective in capturing the large-scale climatology, using an updated climate model at higher resolution may result in runoff and temperature fields with different spatial distribution and sensitivities to elevation.
+
+Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel.
+This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of any emerged pixel the same.
+Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
+However, implementing high-resolution climate models with both varying shorelines and change in paleogeographic position for these scenarios is valuable future direction, although it would require a better knowledge of the past continental topography, which is always challenging.
+Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.
+Given that precipitation in this region is significantly driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
+However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
+
+Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
+Relative to felsic lithologies, mafic lithologies contain minerals with faster dissolution kinetics (e.g.
+olivine and plagioclase), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
+We also do not include the effects of reverse weathering \citep{Michalopoulos1995a}, which could modulate the relationship between liberation of Ca and Mg during silicate weathering and long-term \COtwo consumption.
+
+\section*{Paleoshoreline reconstruction and geological synthesis}
 
 Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in I\&NG.
-Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$.
+Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$ (Fig.
+1 in main text).
 We also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
@@
 Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \cite{Harrington2017a}.
-Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable (Fig.
-1 in main text).
+Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable.
 
@@
 Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
-During the Eocene through Early Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}.
+Eocene through Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}.
 Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments.
-Extensive palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}.
+Palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}.
 Basement highs (e.g.
 the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}.
-Although most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, \citet{Molnar2015a} omit the Sunda Shelf from their paleogeographic model for 5~Ma.
+Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5 Ma paleogeographic model of \citet{Molnar2015a}.
 
@@
 However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
-The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are the products of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
+The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpretted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
 Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
@@
 Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms.
-These arms have high relief $\textgreater$3~km that are separated by deep basins and broadly define Sulawesi’s seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
+These arms have high relief ($\textgreater$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
 Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
@@
 The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, m\'elange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint \citep{Silver1983a, Monnier1995a, Parkinson1998a}.
-Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief towering over 3~km \citep{Kadarusman2004a}.
+Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief over 3~km \citep{Kadarusman2004a}.
 Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma (\citealp{Parkinson1998b} and references therein).
@@
 The Middle Miocene (16--14~Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt \citep{Visser1962a, Cloos2005a}.
-This precedes, by several million years, the beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement at ca.
+The beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement was later at ca.
 12~Ma \citep{vanUfford2005a}.
@@
 Exhumation of the Central Range accelerated over the past 4~m.y.
-due to slab-breakoff and buoyant uplift \citep{Cloos2005a}.
-This was accompanied with a change to left-lateral lateral motion on the northern coast of New Guinea, which caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a}and Halmahera \citep{Hall1988a, Ballantyne1992a}.
+which is interpretted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a} and Halmahera \citep{Hall1988a, Ballantyne1992a}.
 In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
 
-In the Bird’s Neck of western New Guinea, the Lengguru foldbelt formed during Middle to Late Miocene clockwise rotation and obduction of the Weyland Terrane \citep{Bailly2009a}.
+In the Bird’s Neck of western New Guinea, the Lengguru fold belt formed during Middle to Late Miocene clockwise rotation and obduction of the Weyland Terrane \citep{Bailly2009a}.
 Recent counter-clockwise rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge \citep{Sapin2009a}.
@@
 
-Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by paleogeographic reconstructions from Hall \citep{Hall1997a}, coupled with geologic maps of the archipelago \citep{GeologicalSurveyDivision1963a}.
+Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by the paleogeographic reconstructions of \citet{Hall1997a}, coupled with geologic maps of the archipelago \citep{GeologicalSurveyDivision1963a}.
 The Philippine ophiolites have been grouped in four distinct belts \citep{Yumul2007a}.
@@
 From ca.
-20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf, which was at least partially exposed throughout the past 20~Ma \citep{Hall2009a, Hall2013b}.
+20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf \citep{Hall2009a, Hall2013b}.
 Six major islands were subaerially exposed before 5~Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores.
@@
 Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}.
-In Timor and Sumba arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $\textgreater$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $\textgreater$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
 \citep{AudleyCharles1986a}.
-
-\section*{Limitations and Future Directions}
-
-As with any modeling study, there are limitations with the approach.
-We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
-Using an updated climate model at higher resolution may result in runoff and temperature fields with different spatial distribution and sensitivities to elevation.
-
-Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel.
-This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position.
-Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
-However, implementing high-resolution climate models with both varying shorelines and change in paleogeographic position for these scenarios is valuable future direction.
-Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.
-
-Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
-Relative to felsic lithologies, mafic lithologies contain minerals with faster dissolution kinetics (e.g.
-olivine and plagioclase), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
-We also do not include the effects of reverse weathering \citep{Michalopoulos1995a}, which could modulate the relationship between liberation of Ca and Mg during silicate weathering and long-term \COtwo consumption.
 

### 2019-09-19 Manuscript/Manuscript.tex

@@
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda Arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-14~Ma and accelerated over the past 4~m.y.
-culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
+15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation (REFS), a decrease in volcanic outgassing (REFS), uplift in the Himalaya (REFS), or an increase in basaltic watersheds in the tropics (REFS).
 Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
 (Fig.
-\ref{fig:shoreline_growth}) and quantify the increase in land area.
+\ref{fig:shoreline_growth}) and quantify the increase in land area in the tropics.
 We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence (Fig.
 \ref{fig:scenario_pCO2}).
-We find that the magnitude of change in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
+We find that the decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
 This result highlights the importance of highly weatherable regions in setting Earth's long-term climate state.
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate rocks \cite{Kump2000a}.
-However, assessing the effect of the emergence of I\&NG on \pCOtwo levels cannot be determined simply through assessing the present-day chemical weathering fluxes.
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment \cite{Kump2000a}.
 Steady-state \pCOtwo is set by the \pCOtwo level at which \COtwo sinks are equal to sources.
-For example, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
+As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
 Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.
@@
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
-Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance chemical weathering \cite{West2012a, Maher2014a}.
+Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can reduce or enhance chemical weathering \cite{West2012a, Maher2014a}.
 In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase in global weatherability associated with the increase of subaerially-exposed land area in I\&NG, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c} wherein a silicate weathering model is coupled to a global climate model run at varying \pCOtwo levels.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
+GEOCLIM consists of a silicate weathering model coupled to a 3D climate model and to a box-model describing the main biogeochemical cycles.
+GEOCLIM allows the coeval calculations of the CO2 level in the atmosphere and associated climate.
 Within the silicate weathering component of GEOCLIM, we implement a regolith development model \cite{Gabet2009a} that integrates a climatic dependence on chemical weathering \cite{West2012a}.
@@
 While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
-We therefore implement variable bedrock Ca+Mg concentration into the silicate weathering component of GEOCLIM.
+We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM.
 The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) \cite{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
@@
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells during calibration of the silicate weathering component of GEOCLIM.
+We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component of GEOCLIM.
 
-The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a} that modify the dependence of the silicate weathering rate on temperature, runoff, and particle age are poorly constrained (see SI).
+The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
 Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
 For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
-We sum computed \COtwo consumption over watersheds within a global compilation \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating all parameter combinations that result in low $r^{2}$, this approach leaves 3,339 parameter combinations that produce global and individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (see SI).
+We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available\cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
+After eliminating all parameter combinations that result in low $r^{2}$, 3,339 parameter combinations remain.
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux\cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
 
 Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
-For the climate model component of GEOCLIM, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
-The primary strength of using these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
+For the climate model component, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
+The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
 We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 20~m.y.
@@
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, it has been minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
+While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
 The ``reduced I\&NG'' scenario is identical to the ``reduced I\&NG + Sunda Shelf'' scenario, but excludes the Sunda Shelf.
-The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the area are removed.
+The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the region are removed.
 
-Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo's of 598--729~ppm (5--95 percentile values), the ``reduced I\&NG'' scenario resulted in 547--638~ppm \pCOtwo, and the ``reduced I\&NG + Sunda Shelf'' resulted in 533--611~ppm \pCOtwo (Fig.
+Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), the ``reduced I\&NG'' scenario resulted in 547--638~ppm \pCOtwo, and the ``reduced I\&NG + Sunda Shelf'' resulted in 533--611~ppm \pCOtwo (Fig.
 \ref{fig:scenario_pCO2}).
 These results suggest that without the Neogene increase in subaerially-exposed mountains associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain the long-term cooling of Earth's climate over the Neogene.
+These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 
-Enhanced weatherability associated with the movement of the Deccan Traps into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is another hypothesis to explain the draw-down of \COtwo leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
-To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we test three additional scenarios using the lithologically-resolved implementation of GEOCLIM (Fig.
+Enhanced weatherability associated with the movement of Deccan Traps flood basalts into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is another hypothesis to explain \COtwo draw-down leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
+To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we replace these mafic volcanics with the Ca+Mg concentration of bulk continental crust in the lithologically-resolved model (Fig.
 \ref{fig:scenario_pCO2}).
-We replace the volcanics associated with the respective igneous provinces with the Ca+Mg concentration of bulk continental crust in the model.
 The resulting \pCOtwo is \textless500~ppm indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
@@
 
-We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their contributions to the paleoshoreline reconstructions.
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
 Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
-FAM and NLS-H thank NSF FRES grants \#1926001 and \#1925990 for support.
+FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
 %%TC:endignore
@@
 Past shorelines are shown in A with associated area estimates summarized in B (where I\&NG refers to the sum of all the subdivided regions).
-A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record in C \cite{Zachos2008a}.}
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} in C.}
     \label{fig:shoreline_growth}
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the six scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched measured \COtwo consumption in watersheds across the world during calibration (see SI).
+For each of the six scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watersheds data (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
-the interquartile range), and the notch represents the median with its 95\% confidence interval.
+the interquartile range), and the notch represents the median.
 The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}

### 2019-09-20 Manuscript/Manuscript.tex

@@
 15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation (REFS), a decrease in volcanic outgassing (REFS), uplift in the Himalaya (REFS), or an increase in basaltic watersheds in the tropics (REFS).
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a}.
 Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment with a smaller sink being organic carbon burial \cite{Kump2000a}.
 Steady-state \pCOtwo is set by the \pCOtwo level at which \COtwo sinks are equal to sources.
@@
 To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
-GEOCLIM consists of a silicate weathering model coupled to a 3D climate model and to a box-model describing the main biogeochemical cycles.
-GEOCLIM allows the coeval calculations of the CO2 level in the atmosphere and associated climate.
+GEOCLIM consists of a silicate weathering model coupled to a 3D climate model, allowing for the coeval calculations of \pCOtwo and associated climate.
 Within the silicate weathering component of GEOCLIM, we implement a regolith development model \cite{Gabet2009a} that integrates a climatic dependence on chemical weathering \cite{West2012a}.
@@
 
-The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
+The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
 Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
 For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
-We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available\cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
+We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
 After eliminating all parameter combinations that result in low $r^{2}$, 3,339 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux\cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
 
 Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
-For the climate model component, we use temperature and runoff at three \pCOtwo levels (286, 572, and 1144~ppm) from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
+For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
 The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
-We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 20~m.y.
+We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 15~m.y.
 and include changes in areas of submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
@@
 
-To explore the sensitivity of global weatherability and climate state to the emergence of I\&NG, we test three scenarios using GEOCLIM (Fig.
+To explore the sensitivity of global weatherability and climate state to the emergence of I\&NG, we force GEOCLIM using our best estimates of the subaerial extent of islands in the area at ca.
+15, 10, and 5~Ma (``reduced I\&NG + Sunda Shelf'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
-The ``reduced I\&NG + Sunda Shelf'' scenario represents our best estimate of the subaerial extent of islands in the area at ca.
-15~Ma.
 Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations (see SI).
@@
 While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
-The ``reduced I\&NG'' scenario is identical to the ``reduced I\&NG + Sunda Shelf'' scenario, but excludes the Sunda Shelf.
-The ``removed I\&NG'' scenario represents an end-member scenario, in which all islands associated with arc-continent collision in the region are removed.
+We also test an end member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
+\ref{fig:scenario_pCO2}).
 
-Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), the ``reduced I\&NG'' scenario resulted in 547--638~ppm \pCOtwo, and the ``reduced I\&NG + Sunda Shelf'' resulted in 533--611~ppm \pCOtwo (Fig.
+Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), and the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
 \ref{fig:scenario_pCO2}).
-These results suggest that without the Neogene increase in subaerially-exposed mountains associated with arc-continent collision, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-These modeled \pCOtwo values resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
+These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene ($\sim$200--600~ppm) \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 
-Enhanced weatherability associated with the movement of Deccan Traps flood basalts into the tropical rain belt together with the low-latitude emplacement of the Ethiopian Traps is another hypothesis to explain \COtwo draw-down leading into Oligocene glaciation of Antarctica \cite{Kent2008a, Kent2013a}.
-To explore the effect of the modern-day surface exposure of basaltic lavas on steady-state \pCOtwo, we replace these mafic volcanics with the Ca+Mg concentration of bulk continental crust in the lithologically-resolved model (Fig.
+The results of our I\&NG scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth’s climate state.
+To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in the lithologically-resolved model (Fig.
 \ref{fig:scenario_pCO2}).
@@
 \ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+
 Uplift in the Himalayas \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic sea floor spreading to be relative uniform \cite{Rowan2016a}.
-Proponents of the Himalayan hypothesis point toward rising marine \SrSr as being consistent with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
-This rise in \SrSr since ca.
-35~Ma is punctuated with a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the emergence of low \SrSr mafic lithologies in I\&NG.
-While paleogeographic changes elsewhere on Earth would have also modulated geological carbon sources and sinks, the onset of Northern Hemisphere glaciation coincides with increased emergence in I\&NG following 5~Ma.
-Our coupled weathering-climate model indicates that \pCOtwo change associated with emergence of I\&NG is sufficient to explain climate change over the Neogene, suggesting that the global climate state is particularly sensitive to tropical paleogeography.
+Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
+This increase in \SrSr since ca.
+35~Ma is followed by a decrease in slope ca.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, or weathering of the Columbia River flood basalts \cite{}, but could also be partially driven by the exhumation of mafic and organic-rich lithologies in I\&NG forearcs during arc-continent collision.
+
+In addition to I\&NG, paleogeographic changes elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
+Yet, not only does the history of I\&NG emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the scale of cooling.
+These results highlight that the Earth's climate state is particularly sensitive to tropical paleogeography.
 
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the six scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watersheds data (see SI).
+For each of the seven scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
-the interquartile range), and the notch represents the median.
+the interquartile range), and the notch represents the median with its 95\% confidence interval.
 The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}

### 2019-09-20 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 
-The silicate weathering component of the GEOCLIM model as been improved compare to the previous published version (Goddéris et al, 2017 natgeo).
-The new module implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
+The silicate weathering component of the GEOCLIM model has been improved compared to the previously published version \citep{Godderis2017b}.
+The new component implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
 We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
-In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, oras a physically weathered particle once it reaches the top.
+In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
 We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}.
-The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil module of GEOCLIM:
+The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil component of GEOCLIM:
 
@@
 This formulation and these exponent values are supported by various compilations (e.g.
-\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is a variability in the proportionality constant that is difficult to capture at a global scale.
+\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
 
@@
 Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column.
-This value controls the overall integrated chemical weathering rate out of the regolith.
 
@@
 
-We restricted the calculation of the weathering fluxes to the flux of dissolved Mg and Ca originating from continental silicate weathering, as they are the sole contributors to the long term consumption of atmospheric CO2.
-To calculate this flux associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
+We restricted the calculation of the weathering fluxes to the flux of dissolved Ca+Mg originating from continental silicate weathering, as they are the sole contributors to the long-term consumption of atmospheric \COtwo.
+To calculate this flux, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
 \ref{fig:regolith_schematic}).
-Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using a concept of diffuse lithology (it was assume that the weathering rates of shield rocks, basaltic rocks, and carbonate rocks was the same for each continental grid element, when this grid element was submitted to the same climatic conditions).
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, we have implemented a realistic lithologically-resolved version of the model.
+Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using the concept of ``diffuse lithology'' (it was assumed that the weathering rates of all rocks was the same for each continental grid element, provided that the grid element was submitted to the same climatic conditions).
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, and instead we have implemented a more realistic lithologically-resolved version of the model.
 
@@
 However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
-We therefore eliminate all parameter combinations that produce a $r^{2}\textless0.4$, which leaves 3,339 unique parameter combinations (Figs.
+We therefore eliminate all parameter combinations that produce a $r^{2}<0.4$, which leaves 3,339 unique parameter combinations (Figs.
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
 
-The mean global CO$_{2}$ consumption calculated from these 3,339 parameter combinations is 5.4$\times10^{12}$~mol/yr, with a standard deviation of 0.9$\times10^{12}$~mol/yr.
+Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr.
 These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
@@
 We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
+Temperature and runoff fields for any arbitrary \pCOtwo value between 286 and 1144~ppm is obtained by interpolating between the three modeled \pCOtwo levels.
 
@@
     \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM.
+Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead.
 Each column represents a tested scenario.
@@
 
-We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshorelines}).
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
+    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
+For each of the ten scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
+The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
+the interquartile range), and the notch represents the median with its 95\% confidence interval.
+The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
+    \label{fig:scenario_pCO2_all}
+\end{figure}
+
+We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).
 To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig.
 \ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG.
-To generate the lithologic map used in the ``reduced I\&NG'' scenario (Fig.
-\ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15~Ma, then remove all land that falls outside of these bounds.
-For this scenario, we do not add any land (i.e.
-the Sunda Shelf) that is not exposed today to the lithologic map, and the slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for 15~Ma.
-
-The lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenario (Fig.
-\ref{fig:ING_fracs}) is identical to that used in the ``reduced I\&NG'' scenario, with the addition of the Sunda Shelf.
+To generate the lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
+\ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
+The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice.
+We then add land that is not exposed today (i.e.
+the Sunda Shelf).
 However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed.
-As discussed in \textit{Paleoshorelines}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
+As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
 Basement highs between Malaysia/Sumatra and Borneo are typically granitic and bounded by basins dominated by shallow marine clastic fill.
@@
 This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
+
+To explore the sensitivity of our results to the inclusion of the Sunda Shelf, we also tested ``reduced I\&NG'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``reduced I\&NG + Sunda Shelf'' scenarios without the addition of the Sunda Shelf.
+We find that the estimated steady state \pCOtwo's for the ``reduced I\&NG'' scenarios are only marginally higher than those of the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
+\ref{fig:scenario_pCO2_all}), suggesting that our results are largely insensitive to the inclusion of the Sunda Shelf.
+This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low slope values distributed across the shelf.
 
@@
 Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
-However, implementing high-resolution climate models with both varying shorelines and change in paleogeographic position for these scenarios is valuable future direction, although it would require a better knowledge of the past continental topography, which is always challenging.
+However, implementing high-resolution climate models with both varying shorelines and changes in paleogeographic position for these scenarios is a valuable future direction.
 Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.
-Given that precipitation in this region is significantly driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
+Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
@@
 
-\section*{Paleoshoreline reconstruction and geological synthesis}
+\section*{Paleoshoreline Reconstruction and Geological Synthesis}
 
@@
 the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}.
-Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5 Ma paleogeographic model of \citet{Molnar2015a}.
+Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5~Ma paleogeographic model of \citet{Molnar2015a}.
 
@@
 Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \citep{Hall2008a, Ilao2018a}.
-The Sarawak Basin and NW Borneo trough offshore Sabah host $\textgreater$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}.
+The Sarawak Basin and NW Borneo trough offshore Sabah host $>$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}.
 The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \citep{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
@@
 However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
-The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpretted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
+The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpreted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
 Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
@@
 Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms.
-These arms have high relief ($\textgreater$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
+These arms have high relief ($>$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
 Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
@@
 Exhumation of the Central Range accelerated over the past 4~m.y.
-which is interpretted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}.
+which is interpreted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}.
 A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a} and Halmahera \citep{Hall1988a, Ballantyne1992a}.
@@
 Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}.
-In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $\textgreater$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
+In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $>$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
 \citep{AudleyCharles1986a}.

### 2019-09-23 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
     1$\times$10$^{-4}$ & 1$\times$10$^{-2}$ & 0 & 5$\times$10$^{-3}$ & 3000 & 2000 \\
-    2$\times$10$^{-4}$ & 2$\times$10$^{-2}$ & 0.1 &  & 3500 & 2500 \\
+    2$\times$10$^{-4}$ & 2$\times$10$^{-2}$ & 0.1 & 1$\times$10$^{-2}$ & 3500 & 2500 \\
     5$\times$10$^{-4}$ & 5$\times$10$^{-2}$ & 0.3 &  & 4000 & 3000 \\
@@
 We then compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg fluxes) using present-day runoff, temperature, and slope fields.
+Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radat Topography Mission; \citealp{Farr2007a}) fields.
 As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
@@
 
+The decrease in \pCOtwo that we estimate associated with the emergence of I\&NG since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
+This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}).
+\pCOtwo is then estimated from $T$ using a simple energy balance equation.
+However, as discussed in the main text, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
+The linear relationship described in Equation \ref{eq:Molnar3} ignores these feedbacks in the global climate system.
+
 \section*{Limitations and Future Directions}
@@
 This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of any emerged pixel the same.
-Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
+Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor (Fig.
+XXX), we consider this simplification to be reasonable to first-order.
 However, implementing high-resolution climate models with both varying shorelines and changes in paleogeographic position for these scenarios is a valuable future direction.
@@
 Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
-Relative to felsic lithologies, mafic lithologies contain minerals with faster dissolution kinetics (e.g.
-olivine and plagioclase), and thus incorporating lithology-dependent kinetics of mineral dissolution into our simulations would likely increase the sensitivity of steady-state \pCOtwo to increases in subaerially-exposed land area and tectonic topography of I\&NG.
-We also do not include the effects of reverse weathering \citep{Michalopoulos1995a}, which could modulate the relationship between liberation of Ca and Mg during silicate weathering and long-term \COtwo consumption.
+Relative to felsic lithologies, mafic lithologies contain a higher concentration of minerals with faster dissolution kinetics (e.g.
+plagioclase).
+However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.
+We also do not include the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be relatively minor flux in the Cenozoic \citep{Isson2018a}.
+Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.
 
@@
 
-Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2013b, Hall2002a}.
+Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2002a, Hall2013b}.
 Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}.

### 2019-09-23 Manuscript/Manuscript.tex

@@
 % Word counts refer to the text of the paper. References, title, author list and acknowledgements do not have to be included in total word counts.
-
-%%%%%%%%%%%%
-% GAMEPLAN %
-%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%
-%% MAIN TEXT %%
-%%%%%%%%%%%%%%%
-
-% introduction to the question of IA
-% topography, climate, and lithology are important
-% GEOCLIM model
-% dependence on slope and lithology
-% calibration (various feasible combinations of parameters)
-% temperature and runoff from GFDL
-% tested scenarios, including explanation of the paleoshoreline data
-% IA is important, traps are less important
-% limitations with approach
-    % GFDL equilibrated by last 100 years?
-    % not changing paleogeography
-    % source term. Can cite: Rowan2016a related to spreading rate.
-    % low resolution
-    % GFDL model is quite old
-    % kinetics are not changing between lithologies
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% MAIN TEXT FIGURES %%
-%%%%%%%%%%%%%%%%%%%%%%%
-
-% paleoshorelines
-% steady state pCO2 for all scenarios
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPLEMENTARY INFORMATION %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% details of DynSoil and GEOCLIM
-% details of parameter exploration/calibration
-    % basin data used for parameter exploration
-% details of lithology implementation
-    % including Sunda Shelf
-% details of paleoshoreline reconstructions
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SUPPLMENTARY INFORMATION FIGURES %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% regolith model schematic
-% maps of all scenarios
 
@@
 We find that the decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
-This result highlights the importance of highly weatherable regions in setting Earth's long-term climate state.
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment with a smaller sink being organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
 Steady-state \pCOtwo is set by the \pCOtwo level at which \COtwo sinks are equal to sources.
@@
 
-Topography, climate, and lithology all play important roles in modulating chemical weathering.
-In low-relief regions, low erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \cite{Gabet2009a, West2012a, Maher2014a}.
+Topography, climate, and lithology all modulate chemical weathering.
+In low-relief regions, low erosion rates lead to supply-limited weathering regimes, in contrast to regions of high denudation \cite{Gabet2009a, West2012a, Maher2014a}.
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
@@
 To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
-GEOCLIM consists of a silicate weathering model coupled to a 3D climate model, allowing for the coeval calculations of \pCOtwo and associated climate.
+GEOCLIM consists of a silicate weathering model linked to climate model runs at variable \pCOtwo, allowing for estimates of \pCOtwo associated with coupled changes in weathering sinks and climatology.
 Within the silicate weathering component of GEOCLIM, we implement a regolith development model \cite{Gabet2009a} that integrates a climatic dependence on chemical weathering \cite{West2012a}.
@@
 The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell.
-In previous versions of the model, the silicate weathering rate was a function of temperature and runoff only, and assumed that the bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-However, more recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces an additional dependence on topographic slope (see SI).
-While this introduction of regolith development into GEOCLIM is an important piece of assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
+In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on topographic slope (see SI).
+While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
 We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM.
 The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) \cite{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
-Each continental grid cell is assigned one of these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic grid cells are assigned based on the mean of MgO and CaO data on rocks of each of these lithologic categories compiled from the EarthChem Portal (\url{www.earthchem.org/portal}).
+Each continental grid cell is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
+The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
-As a result, the Ca+Mg concentration of the mafic category is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative.
-The weathering of carbonate does not contribute to long-term \COtwo consumption and therefore its Ca+Mg concentration is ignored in this analysis.
+As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative.
+The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored.
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We therefore explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component of GEOCLIM.
-
+We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component of GEOCLIM.
 The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
-Rather than prescribing values for these parameters, we instead allow them and the Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies to vary within reasonable bounds, resulting in 78,000 unique combinations.
-For each of these combinations, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
+Rather than prescribing values for these parameters, we allow them and metamorphic and siliciclastic Ca+Mg concentrations to vary within reasonable bounds, resulting in 78,000 unique combinations.
+For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating all parameter combinations that result in low $r^{2}$, 3,339 parameter combinations remain.
+After eliminating parameter combinations that result in low $r^{2}$, 3,339 parameter combinations remain.
 The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
@@
 associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}.
-Many of the islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{Harris2006a, Hall2013b}.
 These tectonic drivers lead to progressive emergence over the past 15~m.y.
@@
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
+While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography.
 We also test an end member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
@@
 These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early Miocene ($\sim$200--600~ppm) \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a}.
+These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets as \pCOtwo subsequently decreased.
+These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.
 
 The results of our I\&NG scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth’s climate state.
-To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in the lithologically-resolved model (Fig.
+To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is \textless500~ppm indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
+The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
 
@@
 Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
-However, the prolonged timescale of the cooling trend since 14~Ma (Fig.
+However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
 \ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
 
-Uplift in the Himalayas \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic sea floor spreading to be relative uniform \cite{Rowan2016a}.
+Uplift in the Himalayas \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic seafloor spreading to be relative uniform \cite{Rowan2016a}.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, or weathering of the Columbia River flood basalts \cite{}, but could also be partially driven by the exhumation of mafic and organic-rich lithologies in I\&NG forearcs during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearcs sediments in I\&NG during arc-continent collision.
 
 In addition to I\&NG, paleogeographic changes elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
-Yet, not only does the history of I\&NG emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the scale of cooling.
+Yet, not only does the history of I\&NG emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
 These results highlight that the Earth's climate state is particularly sensitive to tropical paleogeography.

### 2019-09-23 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a}.
-These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene \cite{Foster2017a}, indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
+These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.

### 2019-09-23 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.jpg}
+    \caption{\textbf{A)} Modeled global \COtwo consumption vs.
+the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds.
+Each point represents model output using one of the 78,000 parameter combinations.
+\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.
+The black line represents the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
+The black box represents the preferred range of those estimates.}
+    \label{fig:W_vs_r2}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
+    \caption{Modeled vs.
+data-constrained \COtwo consumption in watersheds around the world.
+\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations.
+The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines.
+\textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.
+The chemical weathering and regolith thickness maps of this parameter combination are shown in Figure \ref{fig:weathering_regolith_maps}.}
+    \label{fig:r2_cross_plot}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
+    \caption{Chemical weathering and regolith thickness maps for the parameter combination that produced that highest $r^{2}$ (Fig.
+\ref{fig:r2_cross_plot}B).
+Parameter values used for this combination are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
+    \label{fig:weathering_regolith_maps}
+\end{figure}
+
 Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:8} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
@@
 $M_{i}$ is the modeled \COtwo consumption over watershed $i$, $O_{i}$ is the observed \COtwo consumption over watershed $i$, and $\overline{\log_{10}(O)}$ is the mean of the log of observed \COtwo consumption over all watersheds.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.jpg}
-    \caption{\textbf{A)} Modeled global \COtwo consumption vs.
-the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds.
-Each point represents model output using one of the 78,000 parameter combinations.
-\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.
-The black line represents the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
-The black box represents the preferred range of those estimates.}
-    \label{fig:W_vs_r2}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
-    \caption{Modeled vs.
-data-constrained \COtwo consumption in watersheds around the world.
-\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
-Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations.
-The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines.
-\textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.}
-    \label{fig:r2_cross_plot}
-\end{figure}
 
@@
 However, as discussed in the main text, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
-The linear relationship described in Equation \ref{eq:Molnar3} ignores these feedbacks in the global climate system.
+The linear relationship used in \citet{Molnar2015a} ignores these feedbacks in the global climate system.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/pCO2_proxies.pdf}
+    \caption{Proxy-based \pCOtwo estimates for the past 20~m.y.
+For the \citet{Foster2017a} data, error bars indicate standardized uncertainties.
+For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
+The \citet{Bereiter2015a} data comes from the EPICA Dome C ice core.}
+    \label{fig:pCO2_proxies}
+\end{figure}
+
+Proxy-based \pCOtwo estimates for the past 20~m.y.
+are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a}.
+Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain.
+Nevertheless, our modeled \pCOtwo values for 15~Ma (Fig.
+\ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene.
 

### 2019-09-24 Manuscript/Manuscript.tex
checkpoint in manuscript

@@
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a}.
-Here we use geologic data to reconstruct the shorelines of I\&NG over the past 20~m.y.
+Here we use geologic data to reconstruct the shorelines of I\&NG over the past 15~m.y.
 (Fig.
@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequesters carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
-Steady-state \pCOtwo is set by the \pCOtwo level at which \COtwo sinks are equal to sources.
+Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
 As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
-Conversely, the magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.
+Conversely, the magnitude of the \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.
 
 Topography, climate, and lithology all modulate chemical weathering.
-In low-relief regions, low erosion rates lead to supply-limited weathering regimes, in contrast to regions of high denudation \cite{Gabet2009a, West2012a, Maher2014a}.
+In contrast to regions of high denudation, in low-relief regions low erosion rates lead to supply-limited weathering regimes \cite{Gabet2009a, West2012a, Maher2014a}.
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
@@
 In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}.
-These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo and initiating glacial climate over the past 520~m.y.
+These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
 \cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019a}.
@@
 To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
-GEOCLIM consists of a silicate weathering model linked to climate model runs at variable \pCOtwo, allowing for estimates of \pCOtwo associated with coupled changes in weathering sinks and climatology.
-Within the silicate weathering component of GEOCLIM, we implement a regolith development model \cite{Gabet2009a} that integrates a climatic dependence on chemical weathering \cite{West2012a}.
+GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of \pCOtwo associated with coupled changes in weathering sinks and climatology.
 
@@
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on topographic slope (see SI).
+More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on topographic slope (see SI) \cite{Gabet2009a, West2012a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
 We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM.
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) \cite{Hartmann2012a} and represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
 Each continental grid cell is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
@@
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component of GEOCLIM.
-The values of four parameters within the silicate weathering component of GEOCLIM \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
+We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component.
+
+The values of four parameters within the silicate weathering component \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
 Rather than prescribing values for these parameters, we allow them and metamorphic and siliciclastic Ca+Mg concentrations to vary within reasonable bounds, resulting in 78,000 unique combinations.
@@
 We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 15~m.y.
-and include changes in areas of submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
+and include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
 \ref{fig:shoreline_growth}; see SI).
@@
 
-To explore the sensitivity of global weatherability and climate state to the emergence of I\&NG, we force GEOCLIM using our best estimates of the subaerial extent of islands in the area at ca.
+We force GEOCLIM using our best estimates of the subaerial extent of islands in the area at ca.
 15, 10, and 5~Ma (``reduced I\&NG + Sunda Shelf'' scenarios; Fig.
@@
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography.
-We also test an end member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
+While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
+We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
 \ref{fig:scenario_pCO2}).
@@
 These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
-Estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between proxy records and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
-These modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
+The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 Past shorelines are shown in A with associated area estimates summarized in B (where I\&NG refers to the sum of all the subdivided regions).
-A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} in C.}
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in C.}
     \label{fig:shoreline_growth}

### 2019-09-24 Manuscript/Supplementary_Information/Supplementary_Information.tex
checkpoint in manuscript

@@
 $k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
-This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}.
+This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but with the replacement of elevation with slope and without a temperature dependence, resulting in a functional form that more closely resembles the stream power law \citep{Davy2000a}.
 This formulation and these exponent values are supported by various compilations (e.g.
@@
 
-The spatial distribution of lithologies is sourced from the global lithologic map (GLiM) of \citet{Hartmann2012a}.
+The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) of \citet{Hartmann2012a}.
 The raw data takes the form of polygon vectors, where each polygon is assigned one of 16 lithologic categories.
@@
 In this way, we calculate an area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
+
+The Ca+Mg concentrations of felsic (1,521~mol/m$^{3}$), intermediate (4,759~mol/m$^{3}$), and mafic (10,317~mol/m$^{3}$) lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
+For metamorphic and siliciclastic lithologies, we explore a range of feasible Ca+Mg concentrations during calibration of the silicate weathering component of GEOCLIM (\textit{GEOCLIM Calibration}).
 
@@
 \ref{fig:W_vs_r2}).
-Given that the coefficient of determination ($r^{2}$) is calculated using equation \ref{eq:10} rather than fitting a linear model, many of the combinations associated with particularly poor fits result in negative $r^{2}$ values (Fig.
+Given that the coefficient of determination ($r^{2}$) is calculated using Equation \ref{eq:10} rather than fitting a linear model, many of the combinations associated with particularly poor fits result in negative $r^{2}$ values (Fig.
 \ref{fig:W_vs_r2}).
@@
 These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
-1860 at a 2.0\degrees $\times$ 2.5\degrees resolution.
+1860 at 2.0\degrees $\times$ 2.5\degrees resolution.
 In the ``1860 control'' experiment, forcing agents representative of conditions ca.
@@
 We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
-Temperature and runoff fields for any arbitrary \pCOtwo value between 286 and 1144~ppm is obtained by interpolating between the three modeled \pCOtwo levels.
+Temperature and runoff fields for any arbitrary \pCOtwo value between 286 and 1144~ppm are obtained by linearly interpolating between the three modeled \pCOtwo levels.
 
@@
 
-We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).
+We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y.
+(described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).
 To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig.
@@
 As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
-Basement highs between Malaysia/Sumatra and Borneo are typically granitic and bounded by basins dominated by shallow marine clastic fill.
-However, basement highs are less common in the Gulf of Thailand.
-We therefore assign all pixels of the Sunda Shelf to be 50\% felsic and 50\% sediment, except in the Gulf of Thailand where pixels are assigned to be 100\% sediment.
+Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \citep{Darmadi2007a, Hall2009a, Hall2013b}.
+However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \citep{Darmadi2007a}.
+We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment.
 For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.
@@
 We find that the estimated steady state \pCOtwo's for the ``reduced I\&NG'' scenarios are only marginally higher than those of the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
-\ref{fig:scenario_pCO2_all}), suggesting that our results are largely insensitive to the inclusion of the Sunda Shelf.
-This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low slope values distributed across the shelf.
+\ref{fig:scenario_pCO2_all}), suggesting that our results are insensitive to the inclusion of the Sunda Shelf.
+This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low relief across the shelf.
 
@@
 As with any modeling study, there are limitations with the approach.
-We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0) has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
+We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0), which has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
 While this model is effective in capturing the large-scale climatology, using an updated climate model at higher resolution may result in runoff and temperature fields with different spatial distribution and sensitivities to elevation.
@@
 However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.
-We also do not include the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be relatively minor flux in the Cenozoic \citep{Isson2018a}.
+We also do not include the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
 Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.

### 2019-10-21 Manuscript/Manuscript.tex

@@
 Pierre Maffre\textsuperscript{1},
-Nicholas L. Swanson-Hysell\textsuperscript{1},
+Yves Godd\'eris\textsuperscript{3},
 Francis A. Macdonald\textsuperscript{2},
 Eliel A. Anttila\textsuperscript{2},
-Yves Godd\'eris\textsuperscript{3}
+Nicholas L. Swanson-Hysell\textsuperscript{1}
 
@@
 (Fig.
-\ref{fig:shoreline_growth}) and quantify the increase in land area in the tropics.
+\ref{fig:shoreline_growth}) and quantify the increase in land area.
 We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence (Fig.
 \ref{fig:scenario_pCO2}).
-We find that the decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
+We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
 
-Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean-atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
@@
 In contrast to regions of high denudation, in low-relief regions low erosion rates lead to supply-limited weathering regimes \cite{Gabet2009a, West2012a, Maher2014a}.
-In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
+In warm and wet regions, mineral dissolution occurs more readily leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
 Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can reduce or enhance chemical weathering \cite{West2012a, Maher2014a}.
@@
 To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
-GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of \pCOtwo associated with coupled changes in weathering sinks and climatology.
+GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of long-term steady-state \pCOtwo associated with coupled changes in weathering sinks and climatology.
 
@@
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on topographic slope (see SI) \cite{Gabet2009a, West2012a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope see SI) \cite{Gabet2009a, West2012a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}.
-These tectonic drivers lead to progressive emergence over the past 15~m.y.
+These tectonic drivers and other throughout the region lead to progressive emergence over the past 15~m.y.
 that accelerated following 5~Ma (Fig.
@@
 
-We force GEOCLIM using our best estimates of the subaerial extent of islands in the area at ca.
+We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca.
 15, 10, and 5~Ma (``reduced I\&NG + Sunda Shelf'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
-Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations (see SI).
+Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, we also extrapolate the runoff field from surrounding lands (see SI).
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, it has been relatively minimal, and this fixed scenario is a good approximation of the paleogeography (see SI).
+While there has been motion of these tectonic blocks since 15~Ma, they have remianed within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
@@
 
-Using the 3,339 unique parameter combinations, the ``removed I\&NG'' scenario resulted in steady-state \pCOtwo values of 598--729~ppm (5--95 percentile), and the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
+Using the 3,339 unique parameter combinations, the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
 \ref{fig:scenario_pCO2}).
-These results indicate progressive cooling over the Neogene associated with the emergence of I\&NG, and suggest that without it, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of I\&NG, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
@@
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-These modeled values are therefore consistent with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.
+These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.
 
-The results of our I\&NG scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth’s climate state.
+The results of our I\&NG scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
 To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo, its contribution is less than that of the wetter and higher-relief I\&NG.
+The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling\cite{Kent2008a, Kent2013a}, its contribution is less than that of the wetter and higher-relief I\&NG.
 
@@
 
-We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
 Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
 FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
 %%TC:endignore

### 2019-10-21 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 $k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
-This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but with the replacement of elevation with slope and without a temperature dependence, resulting in a functional form that more closely resembles the stream power law \citep{Davy2000a}.
-This formulation and these exponent values are supported by various compilations (e.g.
+This formulation is directly inspired from the stream power law \citep{Davy2000a}.
+This law and these exponent values are supported by various compilations (e.g.
 \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
@@
 
-We restricted the calculation of the weathering fluxes to the flux of dissolved Ca+Mg originating from continental silicate weathering, as they are the sole contributors to the long-term consumption of atmospheric \COtwo.
+We restricted the calculation of the weathering fluxes to the flux of dissolved Ca+Mg originating from continental silicate weathering, as they are essential for the long-term consumption of atmospheric \COtwo through silicate weathering.
 To calculate this flux, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
 \ref{fig:regolith_schematic}).
-Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using the concept of ``diffuse lithology'' (it was assumed that the weathering rates of all rocks was the same for each continental grid element, provided that the grid element was submitted to the same climatic conditions).
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, and instead we have implemented a more realistic lithologically-resolved version of the model.
+Previous implementations of GEOCLIM have used a ``diffuse lithology'' where all exposed land is assigned the composition of bulk upper continental crust.
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology approach cannot be used, and instead we have implemented a more realistic lithologically-resolved version of the model.
 
@@
 
-The Ca+Mg concentrations of felsic (1,521~mol/m$^{3}$), intermediate (4,759~mol/m$^{3}$), and mafic (10,317~mol/m$^{3}$) lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
+The Ca+Mg concentrations of felsic (1,521~mol/m$^{3}$), intermediate (4,759~mol/m$^{3}$), and mafic (10,317~mol/m$^{3}$) lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}; calculations made in the code with the repository).
 For metamorphic and siliciclastic lithologies, we explore a range of feasible Ca+Mg concentrations during calibration of the silicate weathering component of GEOCLIM (\textit{GEOCLIM Calibration}).
@@
 Each point represents model output using one of the 78,000 parameter combinations.
-\textbf{B)} Same as A, but zoomed to the plotting space with positive $r^{2}$'s.
-The black line represents the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
-The black box represents the preferred range of those estimates.}
+\textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values.
+The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
+The black box represents the preferred range of these estimates preferred by \citet{Gerlach2011a}.}
     \label{fig:W_vs_r2}
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
-    \caption{Chemical weathering and regolith thickness maps for the parameter combination that produced that highest $r^{2}$ (Fig.
+    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced that highest $r^{2}$ (Fig.
 \ref{fig:r2_cross_plot}B).
-Parameter values used for this combination are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
+Parameter values used for this model run are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
     \label{fig:weathering_regolith_maps}
@@
 We use the value of 42~kJ/mol in our model runs.
+While we implement lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
+Relative to felsic lithologies, mafic lithologies contain a higher concentration of minerals with faster dissolution kinetics (e.g.
+plagioclase).
+However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.
 
@@
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM and test 78,000 unique parameter combinations (Table \ref{tab:parameter_combinations}).
+We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM.
+In total, we test 78,000 unique parameter combinations (Fig.
+\ref{fig:r2_cross_plot}; Table \ref{tab:parameter_combinations}).
 
@@
 We then compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radat Topography Mission; \citealp{Farr2007a}) fields.
+Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields.
 As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+Those lasts are nested watersheds, which means we have to subtract upstream weathering fluxes from downstream fluxes.
+Watersheds for which this subtraction yields an aberrant value are not considered.
 80 watersheds in total are used in this study.
@@
 Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr.
-These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
+These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
 \ref{fig:W_vs_r2}B).
+Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state.
+Hence, in the GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing.
+In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
+Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.
 
@@
 We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
-Temperature and runoff fields for any arbitrary \pCOtwo value between 286 and 1144~ppm are obtained by linearly interpolating between the three modeled \pCOtwo levels.
+Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.
+Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
+However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
 
@@
 The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice.
+This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same.
+Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
 We then add land that is not exposed today (i.e.
@@
 We find that the estimated steady state \pCOtwo's for the ``reduced I\&NG'' scenarios are only marginally higher than those of the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
-\ref{fig:scenario_pCO2_all}), suggesting that our results are insensitive to the inclusion of the Sunda Shelf.
+\ref{fig:scenario_pCO2_all}), suggesting that our results are relatively insensitive to the inclusion of the Sunda Shelf.
 This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low relief across the shelf.
@@
     \caption{Proxy-based \pCOtwo estimates for the past 20~m.y.
-For the \citet{Foster2017a} data, error bars indicate standardized uncertainties.
+For data that was compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties.
 For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
@@
 \ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene.
-
-\section*{Limitations and Future Directions}
-
-As with any modeling study, there are limitations with the approach.
-We consider one of the current limitations to be the climate model component of GEOCLIM that was used in this study (GFDL CM2.0), which has a lower spatial resolution (2.0\degrees $\times$ 2.5\degrees) than what can be achieved with the latest climate models and computational resources.
-While this model is effective in capturing the large-scale climatology, using an updated climate model at higher resolution may result in runoff and temperature fields with different spatial distribution and sensitivities to elevation.
-
-Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel.
-This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of any emerged pixel the same.
-Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor (Fig.
-XXX), we consider this simplification to be reasonable to first-order.
-However, implementing high-resolution climate models with both varying shorelines and changes in paleogeographic position for these scenarios is a valuable future direction.
-Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.
-Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
-However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
-
-Finally, although the silicate weathering component of GEOCLIM now implements lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution.
-Relative to felsic lithologies, mafic lithologies contain a higher concentration of minerals with faster dissolution kinetics (e.g.
-plagioclase).
-However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.
-We also do not include the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
-Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.
 

### 2019-10-21 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \begin{flushleft}
-{\Large \textbf{Emergence of Indonesia and New Guinea as a driver for Neogene cooling}}
+{\Large \textbf{Emergence of the SE Asian islands as a driver for Neogene cooling}}
 
@@
 
-Indonesia and New Guinea (I\&NG) have an outsized contribution to modern chemical weathering fluxes.
+The SE Asian islands (SEAI) have an outsized contribution on modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
@@
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a}.
-Here we use geologic data to reconstruct the shorelines of I\&NG over the past 15~m.y.
+Here we use geologic data to reconstruct the shorelines of SEAI over the past 15~m.y.
 (Fig.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in I\&NG and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
 GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of long-term steady-state \pCOtwo associated with coupled changes in weathering sinks and climatology.
@@
 
-Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of I\&NG.
+Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of SEAI.
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
 The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
-We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of I\&NG over the past 15~m.y.
+We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of SEAI over the past 15~m.y.
 and include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
@@
 We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca.
-15, 10, and 5~Ma (``reduced I\&NG + Sunda Shelf'' scenarios; Fig.
+15, 10, and 5~Ma (``reduced SEAI + Sunda Shelf'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
@@
 While there has been motion of these tectonic blocks since 15~Ma, they have remianed within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
-We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed I\&NG'' scenario; Fig.
+We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
 \ref{fig:scenario_pCO2}).
 
-Using the 3,339 unique parameter combinations, the ``reduced I\&NG + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
+Using the 3,339 unique parameter combinations, the ``reduced SEAI + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of I\&NG, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
-The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of I\&NG is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 
-The results of our I\&NG scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
+The results of our SEAI scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
 To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling\cite{Kent2008a, Kent2013a}, its contribution is less than that of the wetter and higher-relief I\&NG.
+The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling\cite{Kent2008a, Kent2013a}, its contribution is less than that of the wetter and higher-relief SEAI.
 
-Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with I\&NG emergence \cite{Molnar2015a}.
+Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
 Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearcs sediments in I\&NG during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearcs sediments in SEAI during arc-continent collision.
 
-In addition to I\&NG, paleogeographic changes elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
-Yet, not only does the history of I\&NG emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
+In addition to SEAI, paleogeographic changes elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
+Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
 These results highlight that the Earth's climate state is particularly sensitive to tropical paleogeography.
@@
     \includegraphics[width=0.9\textwidth]{Figures/shoreline_growth.pdf}
-    \caption{The emergence of Indonesia and New Guinea (I\&NG) from the mid-Miocene to present in the region referred to in climate science literature as the Maritime Continent.
-Past shorelines are shown in A with associated area estimates summarized in B (where I\&NG refers to the sum of all the subdivided regions).
+    \caption{The emergence of the SE Asian islands (SEAI) from the mid-Miocene to present in the region referred to in climate science literature as the Maritime Continent.
+Past shorelines are shown in A with associated area estimates summarized in B.
 A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in C.}

### 2019-10-22 Manuscript/Manuscript.tex
fixing overleaf

@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean-atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
@@
 These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
-\cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019a}.
+\cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019b}.
 
@@
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope see SI) \cite{Gabet2009a, West2012a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope, see SI) \cite{Gabet2009a, West2012a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of SEAI.
-For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI).
+For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI).
 The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
@@
 \ref{fig:scenario_pCO2}).
-Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, we also extrapolate the runoff field from surrounding lands (see SI).
+Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding lands (see SI).
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, they have remianed within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
+While there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
@@
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.
+These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.
 

### 2019-10-22 Manuscript/Supplementary_Information/Supplementary_Information.tex
fixing overleaf

@@
 $k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
-This formulation is directly inspired from the stream power law \citep{Davy2000a}.
-This law and these exponent values are supported by various compilations (e.g.
+This formulation is directly inspired by the stream power law \citep{Davy2000a}.
+This formulation and these exponent values are supported by various compilations (e.g.
 \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
-    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced that highest $r^{2}$ (Fig.
+    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced the highest $r^{2}$ (Fig.
 \ref{fig:r2_cross_plot}B).
@@
 As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
-Those lasts are nested watersheds, which means we have to subtract upstream weathering fluxes from downstream fluxes.
+The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes.
 Watersheds for which this subtraction yields an aberrant value are not considered.
@@
 \ref{fig:W_vs_r2}B).
+
 Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state.
@@
 the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
+The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.
+The only difference with Fig.
+X in the main text....}
     \label{fig:scenario_pCO2_all}
@@
 This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same.
-Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order.
+Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order (Fig.
+XXX).
 We then add land that is not exposed today (i.e.

### 2019-10-22 Manuscript/Manuscript.tex

@@
 
-The SE Asian islands (SEAI) have an outsized contribution on modern chemical weathering fluxes.
+The SE Asian islands (SEAI) have an out-sized contribution on modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area of SEAI and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
 GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of long-term steady-state \pCOtwo associated with coupled changes in weathering sinks and climatology.
@@
 The values of four parameters within the silicate weathering component \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
-Rather than prescribing values for these parameters, we allow them and metamorphic and siliciclastic Ca+Mg concentrations to vary within reasonable bounds, resulting in 78,000 unique combinations.
+Rather than prescribing values for these parameters, we allow them and metamorphic and siliciclastic Ca+Mg concentrations to vary within reasonable bounds, resulting in 93,600 unique combinations.
 For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating parameter combinations that result in low $r^{2}$, 3,339 parameter combinations remain.
+After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain.
 The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
@@
 We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca.
-15, 10, and 5~Ma (``reduced SEAI + Sunda Shelf'' scenarios; Fig.
+15, 10, and 5~Ma (``paleo-SEAI'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
-Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding lands (see SI).
+Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land (see SI).
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
@@
 
-Using the 3,339 unique parameter combinations, the ``reduced SEAI + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
+Using the 5,381 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 493--709~ppm, 435--532~ppm, and 378--447~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$493--709~ppm.
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
+For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
 the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
+The whiskers extend to the 2.5 and 97.5 percentile values.}
     \label{fig:scenario_pCO2}

### 2019-10-22 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 \begin{flushleft}
-{\Large \textbf{Supplementary Information for ``Emergence of Indonesia and New Guinea as a driver for Neogene cooling''}}
+{\Large \textbf{Supplementary Information for ``Emergence of the SE Asian islands as a driver for Neogene cooling''}}
 
@@
 
-The Ca+Mg concentrations of felsic (1,521~mol/m$^{3}$), intermediate (4,759~mol/m$^{3}$), and mafic (10,317~mol/m$^{3}$) lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}; calculations made in the code with the repository).
+The Ca+Mg concentrations of felsic (1,521~mol/m$^{3}$), intermediate (4,759~mol/m$^{3}$), and mafic (10,317~mol/m$^{3}$) lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}; calculations made in the code within the repository).
 For metamorphic and siliciclastic lithologies, we explore a range of feasible Ca+Mg concentrations during calibration of the silicate weathering component of GEOCLIM (\textit{GEOCLIM Calibration}).
@@
 the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds.
-Each point represents model output using one of the 78,000 parameter combinations.
+Each point represents model output using one of the 93,600 parameter combinations.
 \textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values.
 The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
-The black box represents the preferred range of these estimates preferred by \citet{Gerlach2011a}.}
+The black box represents the range of these estimates preferred by \citet{Gerlach2011a}.}
     \label{fig:W_vs_r2}
@@
 data-constrained \COtwo consumption in watersheds around the world.
-\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 3,339 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
-Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 3,339 parameters combinations.
+\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 5,381 parameters combinations.
 The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines.
@@
     2$\times$10$^{-4}$ & 2$\times$10$^{-2}$ & 0.1 & 1$\times$10$^{-2}$ & 3500 & 2500 \\
-    5$\times$10$^{-4}$ & 5$\times$10$^{-2}$ & 0.3 &  & 4000 & 3000 \\
+    5$\times$10$^{-4}$ & 5$\times$10$^{-2}$ & 0.3 & 1.5$\times$10$^{-2}$ & 4000 & 3000 \\
     1$\times$10$^{-3}$ & 1$\times$10$^{-1}$ &  &  &  &  \\
@@
 \caption{Values tested for poorly constrained parameters in the silicate weathering component of GEOCLIM.
-Every permutation of the listed values were tested (except those permutations where the Ca+Mg concentration of the sediments are higher than that of the metamorphics), resulting in 78,000 unique parameter combinations.}
+Every permutation of the listed values were tested (except those permutations where the Ca+Mg concentration of the sediments are higher than that of the metamorphics), resulting in 93,600 unique parameter combinations.}
 \label{tab:parameter_combinations}
@@
 
-The majority of the original 78,000 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig.
+The majority of the original 93,600 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig.
 \ref{fig:W_vs_r2}).
@@
 However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
-We therefore eliminate all parameter combinations that produce a $r^{2}<0.4$, which leaves 3,339 unique parameter combinations (Figs.
+We therefore eliminate all parameter combinations that produce a $r^{2}<0.4$, which leaves 5,381 unique parameter combinations (Figs.
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
 
-Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr.
+Global CO$_{2}$ consumption calculated from these 5,381 parameter combinations ranges from 2.9$\times10^{12}$~mol/yr to 7.7$\times10^{12}$~mol/yr, with a mean of 5.7$\times10^{12}$~mol/yr.
 These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
@@
 Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.
-Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG.
+Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for SEAI.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
 
-\section*{Indonesia and New Guinea Scenarios}
+\section*{SE Asian Islands Scenarios}
 
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/ING_fracs.jpg}
-    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM.
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/SEAI_fracs.jpg}
+    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the SE Asian islands (SEAI) used to force GEOCLIM.
 Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead.
@@
 Solid black lines show present-day shorelines.}
-    \label{fig:ING_fracs}
+    \label{fig:SEAI_fracs}
 \end{figure}
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the ten scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
+For each of the ten scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
 the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.
-The only difference with Fig.
-X in the main text....}
+The whiskers extend to the 2.5 and 97.5 percentile values.
+This figure is identical to Figure 2 in the main text, with the additional scenarios used to test the sensitivity of the results to the inclusion of Greenland and the Sunda Shelf.}
     \label{fig:scenario_pCO2_all}
@@
 
-We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
+    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
+    \label{fig:paleogeographic_reconstructions}
+\end{figure}
+
+We use geological data to quantify the changes in the area of SEAI over the past 20~m.y.
 (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).
-To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig.
-\ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG.
-To generate the lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
-\ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
-The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice.
+To generate the lithologic map used in the ``removed SEAI'' scenario (Fig.
+\ref{fig:SEAI_fracs}), we remove all land associated with arc-continent collision in SEAI.
+To generate the lithologic map used in the ``reduced SEAI + Sunda Shelf'' scenarios (Fig.
+\ref{fig:SEAI_fracs}), we take the estimated paleoshorelines of SEAI for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
+The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of SEAI for their respective time slice.
 This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same.
 Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order (Fig.
-XXX).
+\ref{fig:paleogeographic_reconstructions}).
 We then add land that is not exposed today (i.e.
@@
 
-To explore the sensitivity of our results to the inclusion of the Sunda Shelf, we also tested ``reduced I\&NG'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``reduced I\&NG + Sunda Shelf'' scenarios without the addition of the Sunda Shelf.
-We find that the estimated steady state \pCOtwo's for the ``reduced I\&NG'' scenarios are only marginally higher than those of the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig.
+To explore the sensitivity of our results to the inclusion of the Sunda Shelf, we also tested ``reduced SEAI'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``reduced SEAI + Sunda Shelf'' scenarios without the addition of the Sunda Shelf.
+We find that the estimated steady state \pCOtwo's for the ``reduced SEAI'' scenarios are only marginally higher than those of the ``reduced SEAI + Sunda Shelf'' scenarios (Fig.
 \ref{fig:scenario_pCO2_all}), suggesting that our results are relatively insensitive to the inclusion of the Sunda Shelf.
@@
 
-The decrease in \pCOtwo that we estimate associated with the emergence of I\&NG since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
+The decrease in \pCOtwo that we estimate associated with the emergence of SEAI since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
 This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}).
@@
 
-Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in I\&NG.
+Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in SEAI.
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$ (Fig.

### 2019-10-23 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Chemcal weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
-As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
-Conversely, the magnitude of the \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.
+As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo, and vice versa \cite{Kump1997a}.
 
-Topography, climate, and lithology all modulate chemical weathering.
-In contrast to regions of high denudation, in low-relief regions low erosion rates lead to supply-limited weathering regimes \cite{Gabet2009a, West2012a, Maher2014a}.
-In warm and wet regions, mineral dissolution occurs more readily leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
-Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can reduce or enhance chemical weathering \cite{West2012a, Maher2014a}.
-In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}.
+Topography, climate, and lithology all effect chemical weathering.
+High-relief regions lead to reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a, Maher2014a}.
+In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
+Additionally, mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
 These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
@@
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope, see SI) \cite{Gabet2009a, West2012a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and therefore topographic slope, see SI) \cite{Gabet2009a, West2012a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 \ref{fig:shoreline_growth}; see SI).
+The paleoshoreline data indicate that the Sunda-Banda Arc and New Guinea are primarily responsible for the increase in area since 15~Ma.
+Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \cite{Harris2006a}.
+Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma.
+However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{AudleyCharles2004a, Hall2009a, Hall2013b}.
 In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
@@
 associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}.
-These tectonic drivers and other throughout the region lead to progressive emergence over the past 15~m.y.
+These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y.
 that accelerated following 5~Ma (Fig.
@@
 \ref{fig:scenario_pCO2}).
-Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land (see SI).
+Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land pixels (see SI).
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
@@
 
-Uplift in the Himalayas \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling, although recent analyses show Cenozoic seafloor spreading to be relative uniform \cite{Rowan2016a}.
+Himalayan uplift \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Myrow2015a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearcs sediments in SEAI during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Quade1997a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
 
-In addition to SEAI, paleogeographic changes elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
+Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
 Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.

### 2019-10-23 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 
-The silicate weathering component of the GEOCLIM model has been improved compared to the previously published version \citep{Godderis2017b}.
-The new component implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile.
+The silicate weathering component of the GEOCLIM model has been modified from the previously published version \citep{Godderis2017b}.
+The new component implements the model of \citet{Gabet2009a} for the development of a chemically weathered profile.
 We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
 In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
-We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}.
-The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil component of GEOCLIM:
+We use the DynSoil implementation of the \citet{Gabet2009a} model, which integrates chemical weathering within the regolith \citep{West2012a}.
+The transient time-varying version of this regolith model is described by three equations:
 
@@
 
-Our implementation of the erosion rate follows one of the approaches taken in \citet{Godderis2017c} and \citet{Maffre2018a}, where it is parameterized based on runoff and slope ($s$; m/m):
+Our implementation of the erosion rate is parameterized based on runoff and slope ($s$; m/m):
 
@@
 \noindent
-$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}.
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \citep{Maffre2018a}.
 This formulation is directly inspired by the stream power law \citep{Davy2000a}.
-This formulation and these exponent values are supported by various compilations (e.g.
-\citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.
+This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \citep{Lague2013a}.
 
@@
 Previous implementations of GEOCLIM have used a ``diffuse lithology'' where all exposed land is assigned the composition of bulk upper continental crust.
-Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology approach cannot be used, and instead we have implemented a more realistic lithologically-resolved version of the model.
+These previous studies assumed that the weathering rates of all rocks was the same for each continental grid element, provided that the grid element was submitted to the same climatic conditions.
+Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, we instead implement a more realistic lithologically-resolved version of the model.
 
@@
 
-We then compute spatially-resolved long-term \COtwo consumption (i.e.
+We compute spatially-resolved long-term \COtwo consumption (i.e.
 Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields.
@@
 
-We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) for the climate model component of GEOCLIM.
-These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
+For the climate model component of GEOCLIM, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
 1860 at 2.0\degrees $\times$ 2.5\degrees resolution.
@@
 The whiskers extend to the 2.5 and 97.5 percentile values.
-This figure is identical to Figure 2 in the main text, with the additional scenarios used to test the sensitivity of the results to the inclusion of Greenland and the Sunda Shelf.}
+This figure is identical to Figure 2 in the main text, with the additional scenarios used to test the sensitivity of the results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.}
     \label{fig:scenario_pCO2_all}
@@
 \ref{fig:SEAI_fracs}), we remove all land associated with arc-continent collision in SEAI.
-To generate the lithologic map used in the ``reduced SEAI + Sunda Shelf'' scenarios (Fig.
+To generate the lithologic map used in the ``paleo-SEAI'' scenarios (Fig.
 \ref{fig:SEAI_fracs}), we take the estimated paleoshorelines of SEAI for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
@@
 \ref{fig:paleogeographic_reconstructions}).
-We then add land that is not exposed today (i.e.
-the Sunda Shelf).
-However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed.
+We then add land in the region that is not exposed today.
+The assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed.
 As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
@@
 We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment.
-For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.
+For the slope, we take the mean slope value (0.0043~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.
 For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf.
 This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
+For other minor islands in the region that were exposed 15, 10, or 5~m.y.
+ago but are currently submerged, we take the same approach as that described for the Sunda Shelf to generate the runoff field.
+For the slope and lithology, we take the mean slope and Ca+Mg concentrations from neighbouring pixels.
+We also remove the Greenland ice sheet in the ``paleo-SEAI'' scenarios, to be consistent with Northern Hemisphere ice sheets developing after 5~Ma \citep{Haug2005a}.
+The runoff field over Greenland is generated using the same method as that for the Sunda Shelf.
+For the slope, we take the mean slope value (0.0529~m/m) of land for which slope data exists around the edges of Greenland, and assign this value to all pixels of Greenland covered by the ice sheet.
+We also assign the Ca+Mg concentration of bulk continental crust to these pixels.
 
-To explore the sensitivity of our results to the inclusion of the Sunda Shelf, we also tested ``reduced SEAI'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``reduced SEAI + Sunda Shelf'' scenarios without the addition of the Sunda Shelf.
-We find that the estimated steady state \pCOtwo's for the ``reduced SEAI'' scenarios are only marginally higher than those of the ``reduced SEAI + Sunda Shelf'' scenarios (Fig.
-\ref{fig:scenario_pCO2_all}), suggesting that our results are relatively insensitive to the inclusion of the Sunda Shelf.
-This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low relief across the shelf.
+To explore the sensitivity of our results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet, we also tested ``paleo-SEAI - Greenland'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAI'' scenarios without the removal of the Greenland ice sheet.
+We also tested ``paleo-SEAI - Greenland - Sunda Shelf'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAI'' scenarios without the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.
+We find that the estimated steady state \pCOtwo's for the ``paleo-SEAI - Greenland'' and ``paleo-SEAI - Greenland - Sunda Shelf'' scenarios are only marginally higher than those of the ``paleo-SEAI'' scenarios (Fig.
+\ref{fig:scenario_pCO2_all}), suggesting that our results are relatively insensitive to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.
+This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low relief across the Sunda Shelf, and the low runoff values at high latitudes.
 
-The decrease in \pCOtwo that we estimate associated with the emergence of SEAI since 5~Ma ($\sim$120~ppm) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
+The decrease in \pCOtwo that we estimate associated with the emergence of SEAI since 5~Ma (92--161~ppm; Fig.
+\ref{fig:scenario_pCO2_all}) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
 This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}).
 \pCOtwo is then estimated from $T$ using a simple energy balance equation.
-However, as discussed in the main text, as \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo.
-The linear relationship used in \citet{Molnar2015a} ignores these feedbacks in the global climate system.
+However, the relationship between mean global temperature and weathering-rate-weighted land area is not linear.
+Furthermore, as acknowledged by the authors themselves, this simple linear relationship ignores spatial variability in runoff, temperature, and slope, and only crudely accounts for spatial variability in lithology.
+In fact, the 19~ppm value estimated by \citet{Molnar2015a} is closer to the decrease in \pCOtwo that we estimate if mafic volcanics associated with the Deccan Traps (a relatively flat area outside of the warm and wet tropics) are replaced with the Ca+Mg concentration of bulk continental crust (21--81~ppm; Fig.
+\ref{fig:scenario_pCO2_all}).
+The significant difference in steady-state \pCOtwo estimated between the ``removed Deccan Traps'' scenario and the SEAI scenarios (Fig.
+\ref{fig:scenario_pCO2_all}) demonstrates that considering changes in the spatial distribution of lithologies alone is not adequate for estimating changes in steady-state \pCOtwo.
+Instead, spatially-varying runoff, temperature, and slope significantly modulates silicate weathering rates, and must be accounted for.
 
@@
 \ref{fig:scenario_pCO2_all}) resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene.
+Given this scatter in the proxy-based \pCOtwo estimates, we instead infer the Neogene cooling trend from the Miocene benthic foram oxygen isotope record shown in Figure 1 in the main text.
 
@@
 We also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
-Larger islands were further divided into regions based upon their position on microplates as defined by \citep{Matthews2016a}.
+Larger islands were further divided into regions based upon their position on microplates as defined by \citet{Matthews2016a}.
 Miocene to present stratigraphic columns were compiled from each region to develop an age model for paleoenvironmental indicators.
@@
 Several new sources for paleoshoreline data have become available since \citet{Molnar2015a}.
-Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \cite{Harrington2017a}.
+Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \citet{Harrington2017a}.
 Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable.

### 2019-10-24 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-The SE Asian islands (SEAI) have an out-sized contribution on modern chemical weathering fluxes.
+The SE Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
@@
 \ref{fig:shoreline_growth}) and quantify the increase in land area.
-We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence (Fig.
+We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with SEAI emergence (Fig.
 \ref{fig:scenario_pCO2}).
@@
 Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump2000a}.
-Chemcal weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
+Chemical weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
@@
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
-Additionally, mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
+Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
 These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area of SEAI and enhanced global weatherability, we use the spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
-GEOCLIM consists of a silicate weathering model linked to climate model runs at different \pCOtwo levels, allowing for estimates of long-term steady-state \pCOtwo associated with coupled changes in weathering sinks and climatology.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
+GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in chemical weathering and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
 
-The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell.
-In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and therefore topographic slope, see SI) \cite{Gabet2009a, West2012a}.
+The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each subaerially-exposed land pixel.
+In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all land pixels had identical chemical compositions \cite{Godderis2014a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding, which introduces a dependence on erosion (and therefore topographic slope) \cite{Gabet2009a, West2012a, Maffre2018a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
-Each continental grid cell is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
+Each land pixel is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
+The Ca+Mg concentrations of felsic, intermediate, and mafic pixels are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
@@
 The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment pixels are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
+We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment pixels during calibration of the silicate weathering component.
 
-The values of four parameters within the silicate weathering component \cite{Gabet2009a, West2012a, Maffre2018a} that modify the dependence of silicate weathering on temperature, runoff, and particle age are poorly constrained (see SI).
-Rather than prescribing values for these parameters, we allow them and metamorphic and siliciclastic Ca+Mg concentrations to vary within reasonable bounds, resulting in 93,600 unique combinations.
+The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
+Rather than prescribing values for these parameters, we allow them and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI).
+This results in 93,600 unique combinations of these six parameters.
 For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
@@
 After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (see SI).
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state.
 
-Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of SEAI.
+Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with SEAI emergence.
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI).
-The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated.
-We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of SEAI over the past 15~m.y.
-and include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig.
+The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
 \ref{fig:shoreline_growth}; see SI).
@@
 However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{AudleyCharles2004a, Hall2009a, Hall2013b}.
-In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt.
+In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
 Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
@@
 \ref{fig:shoreline_growth}C).
+We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed.
+Given that spatially-resolved surface data are lacking for the Sunda Shelf, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land pixels (see SI).
 
-We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca.
+We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of SEAI at ca.
 15, 10, and 5~Ma (``paleo-SEAI'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
-Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land pixels (see SI).
 While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
 While there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
-We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
-\ref{fig:scenario_pCO2}).
+We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).
 
@@
 The results of our SEAI scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
-To independently explore the effect of the modern-day surface exposure of lower-slope basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
+To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
@@
 Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
-Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but also our coupled weathering-climate model indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
+Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
 These results highlight that the Earth's climate state is particularly sensitive to tropical paleogeography.
@@
     \includegraphics[width=0.9\textwidth]{Figures/shoreline_growth.pdf}
-    \caption{The emergence of the SE Asian islands (SEAI) from the mid-Miocene to present in the region referred to in climate science literature as the Maritime Continent.
-Past shorelines are shown in A with associated area estimates summarized in B.
+    \caption{The emergence of the SE Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
+Past shorelines are shown in A with associated land area summarized in B.
 A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in C.}
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
+For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.

### 2019-10-24 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 
-These supplementary information materials provide details on the model framework used in this study. Code associated with this study can be found at: \url{https://github.com/Swanson-Hysell-Group/XXX}.
+These supplementary information materials provide details on the model framework used in this study.
+Code associated with this study can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
 
@@
 Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the model time (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
-Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth (where $z$ is the height above the base of the regolith; m).
+Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth ($z$; m).
 $K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless) which implements a time-dependence.
@@
 the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds.
-Each point represents model output using one of the 93,600 parameter combinations.
+Each point represents model output using one of the 93,600 parameter combinations (Table \ref{tab:parameter_combinations}).
 \textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values.
@@
 data-constrained \COtwo consumption in watersheds around the world.
-\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day.
+\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (i.e.
+the orange points in Figure \ref{fig:W_vs_r2}).
 Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 5,381 parameters combinations.
@@
 However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
-We therefore eliminate all parameter combinations that produce a $r^{2}<0.4$, which leaves 5,381 unique parameter combinations (Figs.
+We therefore eliminate all parameter combinations that produce a $r^{2}\leq0.4$, which leaves 5,381 unique parameter combinations (Figs.
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
@@
 Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state.
-Hence, in the GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing.
+Hence, in the non-calibration GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing.
 In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
@@
 Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.
-Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for SEAI.
+Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for this region.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the ten scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI).
+For each of the 13 scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
@@
 The whiskers extend to the 2.5 and 97.5 percentile values.
-This figure is identical to Figure 2 in the main text, with the additional scenarios used to test the sensitivity of the results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.}
+This figure is identical to Figure 2 in the main text, except for the addition of the scenarios used to test the sensitivity of the results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.}
     \label{fig:scenario_pCO2_all}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
+    \includegraphics[width=0.7\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
     \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
@@
 
-We use geological data to quantify the changes in the area of SEAI over the past 20~m.y.
+We use geological data to quantify the changes in the area of SEAI over the past 15~m.y.
 (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).

### 2019-10-28 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 \begin{flushleft}
-{\Large \textbf{Supplementary Information for ``Emergence of the SE Asian islands as a driver for Neogene cooling''}}
+{\Large \textbf{Supplementary Information for ``Emergence of the Southeast Asian islands as a driver for Neogene cooling''}}
 
@@
 
-\section*{SE Asian Islands Scenarios}
+\section*{Southeast Asian Islands Scenarios}
 
@@
     \includegraphics[width=1\textwidth]{Manuscript/Figures/SEAI_fracs.jpg}
-    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the SE Asian islands (SEAI) used to force GEOCLIM.
+    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAI) used to force GEOCLIM.
 Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead.

### 2019-10-28 Manuscript/Manuscript.tex

@@
 \begin{flushleft}
-{\Large \textbf{Emergence of the SE Asian islands as a driver for Neogene cooling}}
+{\Large \textbf{Emergence of the Southeast Asian islands as a driver for Neogene cooling}}
 
@@
 
-The SE Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
+The Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a}.
+15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2001a}.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a}.
 Here we use geologic data to reconstruct the shorelines of SEAI over the past 15~m.y.
@@
 
-Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump1997a}.
 Chemical weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
-As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo, and vice versa \cite{Kump1997a}.
+As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo \cite{Kump1997a}.
+Conversely, as regions with high carbon sequestration potential emerge on Earth's surface, \pCOtwo decreases resulting in other weathering sinks to diminish leading to a new lower steady-state \pCOtwo concentration.
 
 Topography, climate, and lithology all effect chemical weathering.
-High-relief regions lead to reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a, Maher2014a}.
-In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
+High-relief regions lead to reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a}.
+In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{West2012a}.
 Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2014a, Godderis2017c}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2017c}.
 GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in chemical weathering and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
 
-The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each subaerially-exposed land pixel.
-In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all land pixels had identical chemical compositions \cite{Godderis2014a}.
-More recent versions of GEOCLIM implement regolith development and soil shielding, which introduces a dependence on erosion (and therefore topographic slope) \cite{Gabet2009a, West2012a, Maffre2018a}.
+The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for subaerially-exposed land.
+In previous versions of the model, silicate weathering was a function of temperature and runoff only, and all bedrock was assigned identical chemical compositions \cite{Godderis2017c}.
+More recent versions of GEOCLIM implement regolith development and soil shielding, which introduces a dependence on erosion (and therefore topographic slope) \cite{Maffre2018a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 Each land pixel is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic pixels are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
+The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
@@
 The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored.
-The Ca+Mg concentrations of metamorphic and siliciclastic sediment pixels are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment pixels during calibration of the silicate weathering component.
+The Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
+We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component.
 
@@
 After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state.
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (see SI).
 
@@
 Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma.
-However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{AudleyCharles2004a, Hall2009a, Hall2013b}.
-In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
+However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{Hall2013b}.
+In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
 Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
-associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}.
+associated with slab-breakoff and buoyant uplift while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval.
 These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y.
@@
 \ref{fig:shoreline_growth}C).
-We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed.
-Given that spatially-resolved surface data are lacking for the Sunda Shelf, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding land pixels (see SI).
+We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (see SI).
 
@@
 
-Using the 5,381 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 493--709~ppm, 435--532~ppm, and 378--447~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig.
+Using the 5,381 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 378--447~ppm \pCOtwo for 5~Ma, 435--532~ppm for 10~Ma, and 493--709~ppm for 15~Ma (Fig.
 \ref{fig:scenario_pCO2}).
 These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$493--709~ppm.
-Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}.
+Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
 The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
@@
 The results of our SEAI scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
-To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2008a, Kent2013a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
+To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2008a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling\cite{Kent2008a, Kent2013a}, its contribution is less than that of the wetter and higher-relief SEAI.
+The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling \cite{Kent2013a}, its contribution is less than that of the wetter and higher-relief SEAI.
 
@@
 However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
-\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Kump1997a}.
 
-Himalayan uplift \cite{Raymo1988a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
+Himalayan uplift \cite{Raymo1992a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Quade1997a, Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
 
@@
 Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
-These results highlight that the Earth's climate state is particularly sensitive to tropical paleogeography.
+These results highlight that the Earth's climate state is particularly sensitive to changes in tropical paleogeography.
 
@@
 We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
+The GEOCLIM code associated with this study is openly available: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}
 %%TC:endignore
@@
     \includegraphics[width=0.9\textwidth]{Figures/shoreline_growth.pdf}
-    \caption{The emergence of the SE Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
+    \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
 Past shorelines are shown in A with associated land area summarized in B.

### 2019-10-28 Manuscript/Manuscript.tex
round 3 checkpoint

@@
 The results of our SEAI scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
-To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2008a}, we replace mafic volcanics associated with the Deccan and Ethiopian Traps with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
+To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2013a}, we replace mafic volcanics associated with the Deccan Traps, Ethiopian Traps, and Columbia River Basalts with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
@@
 the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to the 2.5 and 97.5 percentile values.}
+The whiskers extend to the 2.5 and 97.5 percentile values.
+Glaciation thresholds \cite{DeConto2008a} are shown on the x-axis.}
     \label{fig:scenario_pCO2}

### 2019-10-30 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
-Rather than prescribing values for these parameters, we allow them and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI).
+Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI).
 This results in 93,600 unique combinations of these six parameters.
@@
 The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+
 To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
@@
 Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma.
-However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{Hall2013b}.
-In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
-Exhumation of the New Guinea Central Range accelerated over the past 4~m.y.
-associated with slab-breakoff and buoyant uplift while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval.
+However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma \cite{Hall2013b}.
+In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{Cloos2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
+Exhumation accelerated over the past 4~m.y.
+in the New Guinea Central Range due to slab-breakoff and buoyant uplift, and in eastern New Guinea due to jamming of the north-dipping subduction zone \cite{Cloos2005a}.
 These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y.
@@
 \ref{fig:scenario_pCO2}).
-While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography.
-While there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
+Because we use a clmate model forced with modern geography, the position of the tectonic blocks remain fixed.
+Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).

### 2019-10-30 Manuscript/Supplementary_Information/Supplementary_Information.tex

@@
 These supplementary information materials provide details on the model framework used in this study.
-Code associated with this study can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
+The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}.
+The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
 
@@
 
-Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:8} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
+Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
 However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}).
@@
 
-Within the equations that govern the silicate weathering component of GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
+Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}.
+However, there is uncertainty in the value of the proportionality constant ($k_e$).
+We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr).
+Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.
+This value is only dependent on the spatially-resolved land area, runoff, and slope.
+
+Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
@@
 Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.
-Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for this region.
+
+Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
+The climate model also assumes the present-day extent of ice-sheets for all \pCOtwo levels.
+However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop; \citealp{Pollard2005b}), but the amount of emergent land would be lower.
+We expect this bias to be minimal, as all of our \pCOtwo estimates fall below the glaciation threshold for Antarctica.
 

### 2019-11-05 Manuscript/Supplementary_Information/Supplementary_Information.tex
Updates from Overleaf

@@
 Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.
-This value is only dependent on the spatially-resolved land area, runoff, and slope.
 
-Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:8}).
+Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:6}).
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.

### 2019-11-13 Manuscript/Manuscript.tex
Updates from Overleaf

@@
-% Letters
-% Letters are short reports of original research focused on an outstanding finding whose importance means that it will be of interest to scientists in other fields.
+% more detailed formatting: https://www.sciencemag.org/authors/instructions-preparing-initial-manuscript
+% formatting: https://www.sciencemag.org/authors/science-information-authors
+% submit at: https://cts.sciencemag.org/
 
-% They do not normally exceed 4 pages of Nature and, as a guideline, allow up to 30 references.
-They begin with a fully referenced paragraph, ideally of about 200 words, but certainly no more than 300 words, aimed at readers in other disciplines.
-This paragraph starts with a 2-3 sentence basic introduction to the field; followed by a one-sentence statement of the main conclusions starting 'Here we show' or equivalent phrase; and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards.
+% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance.
+Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.
+Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
 
-% Please refer to our annotated example to see how the summary paragraph for a Letter should be constructed.
+% Titles should be no more than 96 characters (including spaces).
 
-% The rest of the text is typically about 1,500 words long.
-Any discussion at the end of the text should be as succinct as possible, not repeating previous summary/introduction material, to briefly convey the general relevance of the work.
+% Short titles should be no more than 40 characters (including spaces).
 
-% Letters typically have 3 or 4 small display items (figures or tables).
+% One-sentence summaries capturing the most important point should be submitted for Research Articles, Reports and Reviews.
+These should be a maximum of 125 characters and should complement rather than repeat the title
 
-% Word counts refer to the text of the paper. References, title, author list and acknowledgements do not have to be included in total word counts.
+% Abstracts of Research Articles and Reports should explain to the general reader why the research was done, what was found and why the results are important.
+They should start with some brief BACKGROUND information: a sentence giving a broad introduction to the field comprehensible to the general reader, and then a sentence of more detailed background specific to your study.
+This should be followed by an explanation of the OBJECTIVES/METHODS and then the RESULTS.
+The final sentence should outline the main CONCLUSIONS of the study, in terms that will be comprehensible to all our readers.
+The Abstract is distinct from the main body of the text, and thus should not be the only source of background information critical to understanding the manuscript.
+Please do not include citations or abbreviations in the Abstract.
+The abstract should be 125 words or less.
+
+% Main Text is not divided into sub-headings for Reports. Subheadings are used only in Research Articles, and Reviews.
+
+% References and Notes are numbered in the order in which they are cited, first through the text, then through the figure and table legends and finally through Supplementary Materials.
 
@@
 
+\noindent
+% MAXIMUM 125 CHARACTERS 
+\textbf{One Sentence Summary:} Arc-continent collision in the Southeast Asian islands increased global weatherability enough to drive Neogene cooling.
+
+% MAXIMUM 125 WORDS
+% NO CITATIONS OR ABBREVIATIONS
+\begin{abstract}
+The Southeast Asian islands bring together steep topography, a tropical climate, and mafic lithologies to efficiently weather rocks in the region and consume large amounts of \COtwo.
+Concurrent with global cooling since the mid-Miocene, ongoing arc-continent collision has led to a significant increase of subaerially-exposed land area in the region.
+Here we use geologic data to reconstruct the shorelines of the Southeast Asian islands over the past 15 million years and quantify the increase in land area.
+We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence.
+We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
+\end{abstract}
+
 The Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
@@
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a}.
-Here we use geologic data to reconstruct the shorelines of SEAI over the past 15~m.y.
-(Fig.
-\ref{fig:shoreline_growth}) and quantify the increase in land area.
-We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with SEAI emergence (Fig.
-\ref{fig:scenario_pCO2}).
-We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
+Here we instead propose that SEAI emergence was itself sufficient to explain long-term climatic cooling over the Neogene.
 
@@
 
+% 70 words
 \begin{figure}[h!]
@@
 
+% 100 words
 \begin{figure}[h!]

### 2019-11-20 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 % MAXIMUM 125 CHARACTERS 
-\textbf{One Sentence Summary:} Arc-continent collision in the Southeast Asian islands increased global weatherability enough to drive Neogene cooling.
+\textbf{One Sentence Summary:} The growth of tropical maritime islands led to efficient carbon consumption and drove cooling over the past 15 million years.
 
@@
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2001a}.
+15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2008a}.
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a}.
@@
 
+It continues to be argued that hypotheses that invoke enhanced weathering in regions of active uplift predict an increase in global alkalinity fluxes \cite{Si2019a}.
+However, total alkalinity delivery through silicate weathering needs to be nearly constant when considering the requirement of steady-state for the long-term carbon cycle.
+Rather, when weathering is enhanced in a region such as the emerging SEAI, the decrease in \pCOtwo results in a decrease in chemical weathering in other regions where there has not been tectonically-forced change.
+The same alkalinity delivery occurs more efficiently and thereby at lower \pCOtwo.
+Given that the weathering of carbonate is disconnected from long-term carbon-cycle mass balance, alkalinity delivery through carbonate weathering could change in an evolving climate and be observable in proxy records \cite{Si2019a}.
+
 Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.

### 2019-11-20 Manuscript/Supplementary_Information.tex

--- 2019-11-05 Manuscript/Supplementary_Information/Supplementary_Information.tex
@@
 \usepackage{setspace}
-\usepackage{}
 \usepackage{fancyhdr}
@@
 \usepackage{url}
-%\usepackage[hidelinks]{hyperref}
+\usepackage[hidelinks]{hyperref}
 \urlstyle{same}
 \doublespacing
+
+% this isn't working... come back to it later
+%\usepackage{xcite}
+%\externalcitedocument{Manuscript}
 
@@
 \setcounter{figure}{0}
-\renewcommand{\thefigure}{SI\arabic{figure}}
+\renewcommand{\thefigure}{S\arabic{figure}}
 \setcounter{table}{0}
-\renewcommand{\thetable}{SI\arabic{table}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 \begin{center}
-	\includegraphics[width=0.6\textwidth]{../Figures/regolith_schematic.pdf}
+	\includegraphics[width=0.6\textwidth]{Figures/regolith_schematic.pdf}
 	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state.
@@
 
-The silicate weathering component of the GEOCLIM model has been modified from the previously published version \citep{Godderis2017b}.
-The new component implements the model of \citet{Gabet2009a} for the development of a chemically weathered profile.
+The silicate weathering component of the GEOCLIM model has been modified from the previously published version \cite{Godderis2017b}.
+The new component implements the model of Gabet and Mudd (2009) \cite{Gabet2009a} for the development of a chemically weathered profile.
 We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
-In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
-We use the DynSoil implementation of the \citet{Gabet2009a} model, which integrates chemical weathering within the regolith \citep{West2012a}.
+In the model of Gabet and Mudd (2009) \cite{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
+We use the DynSoil implementation of the Gabet and Mudd (2009) \cite{Gabet2009a} model, which integrates chemical weathering within the regolith \cite{West2012a}.
 The transient time-varying version of this regolith model is described by three equations:
@@
 $P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
-In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
+In Equation \ref{eq:6}, we parameterize this value using the model of Carretier et al.
+(2004) \cite{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
 $f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases.
-It takes an exponential form as is commonly applied in the literature (e.g.
-\citealp{Gabet2009a}).
-In Equation \ref{eq:7}, we follow \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
+It takes an exponential form as is commonly applied in the literature \cite{Gabet2009a}.
+In Equation \ref{eq:7}, we follow Heimsath et al.
+(1997) \cite{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
 
@@
 \noindent
-$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \citep{Maffre2018a}.
-This formulation is directly inspired by the stream power law \citep{Davy2000a}.
-This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \citep{Lague2013a}.
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \cite{Maffre2018a}.
+This formulation is directly inspired by the stream power law \cite{Davy2000a}.
+This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \cite{Lague2013a}.
 
@@
 \noindent
-Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
+Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \cite{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
 
@@
 \vspace{10pt}
-\caption{Grouping of 16 lithologic categories in GLiM \citep{Hartmann2012a} to 6 broader categories for GEOCLIM.}
+\caption{Grouping of 16 lithologic categories in GLiM \cite{Hartmann2012a} to 6 broader categories for GEOCLIM.}
 \label{tab:GLiM_to_GEOCLIM}
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/world_lithology.jpg}
+    \includegraphics[width=1\textwidth]{Figures/world_lithology.jpg}
     \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM.}
@@
 
-The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) of \citet{Hartmann2012a}.
+The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) of Hartmann and Moosdorf (2012) \cite{Hartmann2012a}.
 The raw data takes the form of polygon vectors, where each polygon is assigned one of 16 lithologic categories.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.jpg}
+    \includegraphics[width=1\textwidth]{Figures/W_vs_r2.jpg}
     \caption{\textbf{A)} Modeled global \COtwo consumption vs.
@@
 \textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values.
-The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}.
-The black box represents the range of these estimates preferred by \citet{Gerlach2011a}.}
+The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in Gerlach (2011) \cite{Gerlach2011a}.
+The black box represents the range of these estimates preferred by Gerlach (2011) \cite{Gerlach2011a}.}
     \label{fig:W_vs_r2}
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
+    \includegraphics[width=1\textwidth]{Figures/r2_cross_plot.pdf}
     \caption{Modeled vs.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
+    \includegraphics[width=1\textwidth]{Figures/weathering_regolith_maps.jpg}
     \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced the highest $r^{2}$ (Fig.
@@
 
-Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
-However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}).
+Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \cite{Brantley2003a}.
+However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol \cite{Li2016a}; 42.3~kJ/mol \cite{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol \cite{Knauss1993a}) and for labradorite (42.1~kJ/mol \cite{Carroll2005a}).
 We use the value of 42~kJ/mol in our model runs.
@@
 
-Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}.
+Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \cite{Lague2013a}.
 However, there is uncertainty in the value of the proportionality constant ($k_e$).
-We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr).
+We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in Milliman and Farnsworth (2013) \cite{Milliman2013a} ($20\times10^{12}$~kg/yr).
 Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.
@@
 We compute spatially-resolved long-term \COtwo consumption (i.e.
-Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields.
-As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0 \cite{Fekete1999a}), temperature (CRU TS v.4.03 \cite{Harris2013a}), and slope (Shuttle Radar Topography Mission \cite{Farr2007a}) fields.
+As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of Gaillardet et al.
+(1999) \cite{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of Moquet et al.
+(2011) \cite{Moquet2011a}, Moquet et al.
+(2016) \cite{Moquet2016a}, and Moquet et al.
+(2018) \cite{Moquet2018a}.
 The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes.
@@
 Global CO$_{2}$ consumption calculated from these 5,381 parameter combinations ranges from 2.9$\times10^{12}$~mol/yr to 7.7$\times10^{12}$~mol/yr, with a mean of 5.7$\times10^{12}$~mol/yr.
-These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
+These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in Gerlach (2011) \cite{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
 \ref{fig:W_vs_r2}B).
@@
 Hence, in the non-calibration GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing.
-In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
+In these estimates we are not including the effects of reverse weathering \cite{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \cite{Isson2018a}.
 Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.
@@
 
-For the climate model component of GEOCLIM, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
+For the climate model component of GEOCLIM, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca.
 1860 at 2.0\degrees $\times$ 2.5\degrees resolution.
@@
 
-Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
+Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \cite{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
 The climate model also assumes the present-day extent of ice-sheets for all \pCOtwo levels.
-However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop; \citealp{Pollard2005b}), but the amount of emergent land would be lower.
+However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop \cite{Pollard2005b}), but the amount of emergent land would be lower.
 We expect this bias to be minimal, as all of our \pCOtwo estimates fall below the glaciation threshold for Antarctica.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/SEAI_fracs.jpg}
+    \includegraphics[width=1\textwidth]{Figures/SEAI_fracs.jpg}
     \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAI) used to force GEOCLIM.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
+    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
-    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
+    \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
+    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of Matthews et al. (2016) \cite{Matthews2016a}.}
     \label{fig:paleogeographic_reconstructions}
@@
 As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust.
-Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \citep{Darmadi2007a, Hall2009a, Hall2013b}.
-However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \citep{Darmadi2007a}.
+Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \cite{Darmadi2007a, Hall2009a, Hall2013b}.
+However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \cite{Darmadi2007a}.
 We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment.
@@
 For the slope and lithology, we take the mean slope and Ca+Mg concentrations from neighbouring pixels.
-We also remove the Greenland ice sheet in the ``paleo-SEAI'' scenarios, to be consistent with Northern Hemisphere ice sheets developing after 5~Ma \citep{Haug2005a}.
+We also remove the Greenland ice sheet in the ``paleo-SEAI'' scenarios, to be consistent with Northern Hemisphere ice sheets developing after 5~Ma \cite{Haug2005a}.
 The runoff field over Greenland is generated using the same method as that for the Sunda Shelf.
@@
 The decrease in \pCOtwo that we estimate associated with the emergence of SEAI since 5~Ma (92--161~ppm; Fig.
-\ref{fig:scenario_pCO2_all}) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
-This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}).
+\ref{fig:scenario_pCO2_all}) is greater than that proposed in Molnar and Cronin (2015) \cite{Molnar2015a} ($\sim$19~ppm).
+This 19~ppm value in Molnar and Cronin (2015) \cite{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in Molnar and Cronon, 2015 \cite{Molnar2015a}).
 \pCOtwo is then estimated from $T$ using a simple energy balance equation.
@@
 Furthermore, as acknowledged by the authors themselves, this simple linear relationship ignores spatial variability in runoff, temperature, and slope, and only crudely accounts for spatial variability in lithology.
-In fact, the 19~ppm value estimated by \citet{Molnar2015a} is closer to the decrease in \pCOtwo that we estimate if mafic volcanics associated with the Deccan Traps (a relatively flat area outside of the warm and wet tropics) are replaced with the Ca+Mg concentration of bulk continental crust (21--81~ppm; Fig.
+In fact, the 19~ppm value estimated by Molnar and Cronin (2015) \cite{Molnar2015a} is closer to the decrease in \pCOtwo that we estimate if mafic volcanics associated with the Deccan Traps (a relatively flat area outside of the warm and wet tropics) are replaced with the Ca+Mg concentration of bulk continental crust (21--81~ppm; Fig.
 \ref{fig:scenario_pCO2_all}).
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/pCO2_proxies.pdf}
+    \includegraphics[width=1\textwidth]{Figures/pCO2_proxies.pdf}
     \caption{Proxy-based \pCOtwo estimates for the past 20~m.y.
-For data that was compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties.
-For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
-The \citet{Bereiter2015a} data comes from the EPICA Dome C ice core.}
+For data that was compiled within Foster et al.
+(2017) \cite{Foster2017a}, error bars indicate standardized uncertainties.
+For the Ji et al.
+(2018) \cite{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
+The Bereiter et al.
+(2015) \cite{Bereiter2015a} data comes from the EPICA Dome C ice core.}
     \label{fig:pCO2_proxies}
@@
 Proxy-based \pCOtwo estimates for the past 20~m.y.
-are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a}.
+are shown in Figure \ref{fig:pCO2_proxies} \cite{Bereiter2015a, Foster2017a, Ji2018a}.
 Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain.
@@
 Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in SEAI.
-Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$ (Fig.
+Following Molnar and Cronin (2015) \cite{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$ (Fig.
 1 in main text).
 We also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda Shelf.
-Larger islands were further divided into regions based upon their position on microplates as defined by \citet{Matthews2016a}.
+Larger islands were further divided into regions based upon their position on microplates as defined by Matthews et al.
+(2016) \cite{Matthews2016a}.
 Miocene to present stratigraphic columns were compiled from each region to develop an age model for paleoenvironmental indicators.
@@
 
-Several new sources for paleoshoreline data have become available since \citet{Molnar2015a}.
-Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \citet{Harrington2017a}.
-Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable.
+Several new sources for paleoshoreline data have become available since Molnar and Cronin (2015) \cite{Molnar2015a}.
+Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by Nugraha et al.
+(2018) \cite{Nugraha2018a}, and in New Guinea, we followed Gold et al.
+(2017) \cite{Gold2017a} and Harrington et al.
+(2017) \cite{Harrington2017a}.
+Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by Molnar and Cronin (2015) \cite{Molnar2015a} our calculated area is comparable.
 
@@
 
-Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2002a, Hall2013b}.
-Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}.
-Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}.
-Eocene through Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}.
+Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \cite{Hall2002a, Hall2013b}.
+Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \cite{Hall2002a}.
+Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \cite{Moss1998a}.
+Eocene through Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \cite{Morley2013a}.
 Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments.
-Palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}.
+Palynological analysis suggests widespread swamp environments persisting through the Late Miocene \cite{Morley2013a}.
 Basement highs (e.g.
-the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}.
-Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5~Ma paleogeographic model of \citet{Molnar2015a}.
+the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \cite{Darmadi2007a}.
+Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \cite{Hall2013b, Madon2013a}, although it is omitted in the 5~Ma paleogeographic model of Molnar and Cronin (2015) \cite{Molnar2015a}.
 
@@
 
-Our reconstructions of the paleoshorelines of Borneo and Palawan are largely informed by extant paleogeographic constructions detailed in \citet{Hall2001a, Hall2013a, Hall2013b},  as well as geologic maps and local shoreline reconstructions found in \citet{vandeWeerd1992a, Witts2012a, Madon2013a, Kessler2015a}.
-Comprised of Paleozoic to Mesozoic crustal components that were largely accreted to Sundaland by the late Cretaceous \citep{Metcalfe2013a}, the southern and western portions of Borneo (SW Kalimantan) have been subaerially exposed throughout the Neogene \citep{Hall2013a, Hall2013b}.
-A collision along the northern margin of Borneo associated with the initiation of rifting in the South China Sea resulted in the late Eocene uplift of the Central Range mountains \citep{Hutchison1996a}, which provide sediment to basins along the Southern (Kalimantan) and northern (Sabah and Sarawak) coasts of the island.
-In the east, the Barito, Kutei, and Tarakan basins developed as a single area of subsidence (associated with the opening of the Makassar Straits in the Eocene) before the basins were isolated by Oligocene faulting and Miocene uplift \citep{Witts2012a}.
-The Kutei Basin is characterized by eastward-prograding deltaic and shallow shelf deposits that have been steadily supplied with sediment from the Schwaner Mountains of SW Kalimantan and the Central Range \citep{vandeWeerd1992a}.
-North of the Kutei Basin, the deltaic deposits found in the Tarakan Basin similarly prograde eastward from the mid Miocene onward, fed by the northern drainages of the Central Range \citep{Satyana1999a}.
+Our reconstructions of the paleoshorelines of Borneo and Palawan are largely informed by extant paleogeographic constructions detailed in Hall (2001, 2013a, 2013b) \cite{Hall2001a, Hall2013a, Hall2013b},  as well as geologic maps and local shoreline reconstructions found in van de Weerd and Armin (1992) \cite{vandeWeerd1992a}, Witts et al.
+(2012) \cite{Witts2012a}, Madon et al.
+(2013) \cite{Madon2013a}, and Kessler and Jong (2015) \cite{Kessler2015a}.
+Comprised of Paleozoic to Mesozoic crustal components that were largely accreted to Sundaland by the late Cretaceous \cite{Metcalfe2013a}, the southern and western portions of Borneo (SW Kalimantan) have been subaerially exposed throughout the Neogene \cite{Hall2013a, Hall2013b}.
+A collision along the northern margin of Borneo associated with the initiation of rifting in the South China Sea resulted in the late Eocene uplift of the Central Range mountains \cite{Hutchison1996a}, which provide sediment to basins along the Southern (Kalimantan) and northern (Sabah and Sarawak) coasts of the island.
+In the east, the Barito, Kutei, and Tarakan basins developed as a single area of subsidence (associated with the opening of the Makassar Straits in the Eocene) before the basins were isolated by Oligocene faulting and Miocene uplift \cite{Witts2012a}.
+The Kutei Basin is characterized by eastward-prograding deltaic and shallow shelf deposits that have been steadily supplied with sediment from the Schwaner Mountains of SW Kalimantan and the Central Range \cite{vandeWeerd1992a}.
+North of the Kutei Basin, the deltaic deposits found in the Tarakan Basin similarly prograde eastward from the mid Miocene onward, fed by the northern drainages of the Central Range \cite{Satyana1999a}.
 In SE Kalimantan, the Barito Basin is bounded to the east by the ophiolite-bearing Meratus Mountains.
-Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene \citep{Witts2012a}.
+Sedimentological data suggests that the Meratus Mountains were not emergent until the Late Miocene \cite{Witts2012a}.
 
-Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \citep{Hall2008a, Ilao2018a}.
-The Sarawak Basin and NW Borneo trough offshore Sabah host $>$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}.
-The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \citep{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
-Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities \citep{Hall2008a}.
-Early Miocene sediments in northern Sabah have a provenance from Palawan \citep{Suggate2014a}, indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny.
-However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
-The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpreted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}.
+Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \cite{Hall2008a, Ilao2018a}.
+The Sarawak Basin and NW Borneo trough offshore Sabah host $>$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \cite{Hall2008a}.
+The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \cite{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah.
+Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities \cite{Hall2008a}.
+Early Miocene sediments in northern Sabah have a provenance from Palawan \cite{Suggate2014a}, indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny.
+However, initiation of back-arc extension in the Sulu Sea around 19~Ma \cite{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough.
+The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpreted to be the result of crustal thinning related to pulses of regional backarc extension \cite{Hall2013a}.
 Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt.
-Kinabalu, the tallest peak in Borneo \citep{Cottam2013a}.
+Kinabalu, the tallest peak in Borneo \cite{Cottam2013a}.
 
@@
 
-For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of \citet{Nugraha2018a}.
+For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of Nugraha et al.
+(2018) \cite{Nugraha2018a}.
 Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms.
-These arms have high relief ($>$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
-Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
-After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.
-The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (\citealp{Monnier1995a}, but see \citealp{Kadarusman2004a} for an alternative view), and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block of \citealp{Villeneuve2001a}).
-During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.
+These arms have high relief ($>$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \cite{Hamilton1979a, Katili1978a}.
+Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \cite{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
+After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \cite{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \cite{Hall2002b}, also referred to as the Great Indonesian arc \cite{Harris2006a}.
+The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting \cite{Monnier1995a} (but see Kadarusman et al.,2004 \cite{Kadarusman2004a} for an alternative view) and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma \cite{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block \cite{Villeneuve2001a}).
+During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \cite{Smith1991a, Bergman1996a}.
 
-The North Arm consists of Late Pliocene and younger volcanic rocks built upon Eocene to Early Miocene oceanic basalt, basaltic andesite, pelagic sediments, and metamorphic rocks \citep{Elburg1998a}.
-Eocene to Early Miocene volcanic rocks formed above a NW dipping slab in an arc system that extended to West and South Sulawesi \citep{vanLeeuwen2005a}.
+The North Arm consists of Late Pliocene and younger volcanic rocks built upon Eocene to Early Miocene oceanic basalt, basaltic andesite, pelagic sediments, and metamorphic rocks \cite{Elburg1998a}.
+Eocene to Early Miocene volcanic rocks formed above a NW dipping slab in an arc system that extended to West and South Sulawesi \cite{vanLeeuwen2005a}.
 Gorontalo Bay is an extensional basin that formed in the Pliocene, and prior to that time, the East Sulawesi was attached to the North Arm of Sulawesi.
-After the Oligocene to Early Miocene accretion of the East Sulawesi ophiolite, the arc system stepped out to the SE forming a Neogene (23--16~Ma) volcanic belt on the East arm \citep{Kadarusman2004a}.
+After the Oligocene to Early Miocene accretion of the East Sulawesi ophiolite, the arc system stepped out to the SE forming a Neogene (23--16~Ma) volcanic belt on the East arm \cite{Kadarusman2004a}.
 In the Late Miocene, SE-dipping subduction was initiated below the North Arm, which continues today at the North Sulawesi Trough.
-North Sulawesi was not substantially emergent until the Pliocene \citep{vanLeeuwen2005a}.
+North Sulawesi was not substantially emergent until the Pliocene \cite{vanLeeuwen2005a}.
 Active volcanoes extend from the North Arm through the Sangihe Arc into the southern Philippines.
 
-The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000~km$^{2}$ \citep{Monnier1995a}.
-The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, m\'elange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint \citep{Silver1983a, Monnier1995a, Parkinson1998a}.
-Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief over 3~km \citep{Kadarusman2004a}.
-Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma (\citealp{Parkinson1998b} and references therein).
-K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36--23~Ma \citep{Parkinson1998a, Villeneuve2001a}, which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) \citep{Silver1983a, Parkinson1998a}; however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
+The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000~km$^{2}$ \cite{Monnier1995a}.
+The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, m\'elange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint \cite{Silver1983a, Monnier1995a, Parkinson1998a}.
+Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief over 3~km \cite{Kadarusman2004a}.
+Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma \cite{Parkinson1998b}.
+K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36--23~Ma \cite{Parkinson1998a, Villeneuve2001a}, which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) \cite{Silver1983a, Parkinson1998a}; however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
 
-Collision of the Sula-Banggai block with the East margin of Sulawesi began in the latest Miocene and uplift accelerated in the early Pliocene (5.2~Ma to 3.8~Ma) and is associated with a major pulse of sedimentation in adjacent basins \citep{Davies1990a, Villeneuve2000a}.
-Off the northeast margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes m\'elange \citep{Davies1990a}.
+Collision of the Sula-Banggai block with the East margin of Sulawesi began in the latest Miocene and uplift accelerated in the early Pliocene (5.2~Ma to 3.8~Ma) and is associated with a major pulse of sedimentation in adjacent basins \cite{Davies1990a, Villeneuve2000a}.
+Off the northeast margin of the east arm of Sulawesi, Miocene platform carbonates on the Sula-Banggai block are overlain by Late Miocene to Early Pliocene ophiolite detritus in the Celebes m\'elange \cite{Davies1990a}.
 Thus, although the East Sulawesi ophiolite was trapped/emplaced onto the composite Sundaland margin (i.e.
@@
 
-West Sulawesi rifted from Borneo during the Eocene forming the Makassar Straits back arc basin behind a southwest-facing arc \citep{Polve1997a}.
-Eroded fragments of ophiolite and extensive belts of volcanic rocks are preserved on the West and South Arms of Sulawesi \citep{Bergman1996a, vanLeeuwen2010a}.
+West Sulawesi rifted from Borneo during the Eocene forming the Makassar Straits back arc basin behind a southwest-facing arc \cite{Polve1997a}.
+Eroded fragments of ophiolite and extensive belts of volcanic rocks are preserved on the West and South Arms of Sulawesi \cite{Bergman1996a, vanLeeuwen2010a}.
 During the Middle Miocene (ca.
-15--13~Ma), extensional faults in the Bone basin reversed, which was accompanied by uplift and erosion of the Bone Mountain ophiolite and Lamasi complex in West Sulawesi \citep{Bergman1996a, vanLeeuwen2010a}.
-Shortening was likely due to the collision of the leading edge of the Buton-Tukang Besi block, which collided with Buton and the SE Arm of Sulawesi \citep{Smith1991a}.
-Uplift and erosion is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins \citep{Bergman1996a, vanLeeuwen2010a}.
+15--13~Ma), extensional faults in the Bone basin reversed, which was accompanied by uplift and erosion of the Bone Mountain ophiolite and Lamasi complex in West Sulawesi \cite{Bergman1996a, vanLeeuwen2010a}.
+Shortening was likely due to the collision of the leading edge of the Buton-Tukang Besi block, which collided with Buton and the SE Arm of Sulawesi \cite{Smith1991a}.
+Uplift and erosion is recorded by the presence of a major Middle Miocene unconformity and sedimentary breccias in marginal basins \cite{Bergman1996a, vanLeeuwen2010a}.
 Uplift was diachronous, not effecting units below and to the west of the Lamasai ophiolite until the Middle Miocene (ca.
-13~Ma) \citep{vanLeeuwen2010a}.
+13~Ma) \cite{vanLeeuwen2010a}.
 Alkali volcanism ensued at ca.
 11~Ma and is associated with a second phase of extension and exhumation.
-From the Late Miocene to Pliocene fission track ages imply deep exhumation \citep{Smith1991a, Bergman1996a, vanLeeuwen2010a}.
+From the Late Miocene to Pliocene fission track ages imply deep exhumation \cite{Smith1991a, Bergman1996a, vanLeeuwen2010a}.
 
-Fission track ages from granitoids in central Sulawesi indicate rapid uplift of Central Sulawesi (200--700~m/m.y.) starting at about 5~Ma associated with movement on the Palo-Koro fault \citep{Bellier2006a}.
-The fault system also shows a normal component with rapid exhumation of rock west of the fault in western Sulawesi -- all fission track dates are younger than 5~Ma \citep{Bellier2006a}.
+Fission track ages from granitoids in central Sulawesi indicate rapid uplift of Central Sulawesi (200--700~m/m.y.) starting at about 5~Ma associated with movement on the Palo-Koro fault \cite{Bellier2006a}.
+The fault system also shows a normal component with rapid exhumation of rock west of the fault in western Sulawesi -- all fission track dates are younger than 5~Ma \cite{Bellier2006a}.
 Just east of the Palo-Koro fault, the Palu Metamorphic Complex was exhumed in the Late Miocene to early Pliocene in the north (ca.
@@
 3.1--2.7~Ma) at rates of up to 400~m/m.y.
-\citep{Hennig2017b}.
+\cite{Hennig2017b}.
 
@@
 
-Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of \citet{Nichols1991a}, \citet{Cloos2005a}, \citet{Gold2017a}, and \citet{Harrington2017a}.
+Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of Nichols and Hall (1991) \cite{Nichols1991a}, Cloos et al.
+(2005) \cite{Cloos2005a}, Gold et al.
+(2017) \cite{Gold2017a}, and Harrington et al.
+(2017) \cite{Harrington2017a}.
 These studies were all based to varying degrees on lithological distributions, biostratigraphic, borehole data, and tectonic models.
@@
 
-Northern New Guinea, including the Melanesian Arc, was emplaced above the Australian plate during the Miocene \citep{Hamilton1979a, Cloos2005a, vanUfford2005a, Baldwin2012a}.
+Northern New Guinea, including the Melanesian Arc, was emplaced above the Australian plate during the Miocene \cite{Hamilton1979a, Cloos2005a, vanUfford2005a, Baldwin2012a}.
 Two major ophiolite belts marking the suture -- the Irian-Marum ophiolite belt (including the April ultramafics), and the Papuan Ultramafic Belt (PUB) -- are preserved along the Central Range and Peninsular Range, respectively.
@@
 
-The Middle Miocene (16--14~Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt \citep{Visser1962a, Cloos2005a}.
+The Middle Miocene (16--14~Ma) basal Makats Formation contains siliciclastic sediment that was transported from the south into the forearc basin associated with the Irian-Marum ophiolite belt \cite{Visser1962a, Cloos2005a}.
 The beginning of widespread synorogenic sedimentation to the south and on the Australian continental basement was later at ca.
-12~Ma \citep{vanUfford2005a}.
+12~Ma \cite{vanUfford2005a}.
 Mountain building began in Late Miocene time, ca.
-8--7~Ma \citep{vanUfford2005a, Baldwin2012a}, but major relief was not generated until the Pliocene \citep{Weiland1996a}.
-Similarly, the Marum ophiolite was uplifted between 8--5~Ma with 3--4~km of denudation \citep{Crowhurst1996a}.
-An estimated 80--100~km of shortening has been accommodated by deformation on the south side of the Central Range \citep{Hill1989a, Cloos2005a}.
+8--7~Ma \cite{vanUfford2005a, Baldwin2012a}, but major relief was not generated until the Pliocene \cite{Weiland1996a}.
+Similarly, the Marum ophiolite was uplifted between 8--5~Ma with 3--4~km of denudation \cite{Crowhurst1996a}.
+An estimated 80--100~km of shortening has been accommodated by deformation on the south side of the Central Range \cite{Hill1989a, Cloos2005a}.
 
-Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene \citep{Davies1971a, Davies1984a, vanUfford2005a}.
+Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene \cite{Davies1971a, Davies1984a, vanUfford2005a}.
 Miocene to present arc-continent collision in New Guinea has progressed from west to east.
 Exhumation of the Central Range accelerated over the past 4~m.y.
-which is interpreted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}.
-A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a} and Halmahera \citep{Hall1988a, Ballantyne1992a}.
-In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.
+which is interpreted to be the result of slab-breakoff and buoyant uplift \cite{Cloos2005a}.
+A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \cite{Monnier1999a} and on the islands Obi \cite{Ali2001a} and Halmahera \cite{Hall1988a, Ballantyne1992a}.
+In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \cite{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \cite{Abbott1994a} and thermochronology \cite{Hill1989a}.
 
-In the Bird’s Neck of western New Guinea, the Lengguru fold belt formed during Middle to Late Miocene clockwise rotation and obduction of the Weyland Terrane \citep{Bailly2009a}.
-Recent counter-clockwise rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge \citep{Sapin2009a}.
+In the Bird’s Neck of western New Guinea, the Lengguru fold belt formed during Middle to Late Miocene clockwise rotation and obduction of the Weyland Terrane \cite{Bailly2009a}.
+Recent counter-clockwise rotation has exhumed several core complexes and has produced Pliocene compressional deformation to the southwest in the Misool-Onin-Kumawa Ridge \cite{Sapin2009a}.
 
@@
 
-Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by the paleogeographic reconstructions of \citet{Hall1997a}, coupled with geologic maps of the archipelago \citep{GeologicalSurveyDivision1963a}.
-The Philippine ophiolites have been grouped in four distinct belts \citep{Yumul2007a}.
+Our reconstructions of the paleoshorelines of the Phillipines were primarily informed by the paleogeographic reconstructions of Hall (1997) \cite{Hall1997a}, coupled with geologic maps of the archipelago \cite{GeologicalSurveyDivision1963a}.
+The Philippine ophiolites have been grouped in four distinct belts \cite{Yumul2007a}.
 We combine belts 1 and 2 in an eastern suture zone and belts 3 and 4 in a western suture zone.
-Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments \citep{Pubellier1991a}.
-The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present \citep{Yumul2013a}.
-In addition to collision, arc magmatism contributed significantly to crustal growth in the Philippines \citep{Dimalanta2006a}.
+Belts 1 and 2 were juxtaposed in the eastern suture zone during sinistral transpression, which resulted in Early Miocene uplift and deposition of coarse clastic sediments \cite{Pubellier1991a}.
+The Palawan microcontinental block and Philippine mobile belt collided during late Early Miocene to early Middle Miocene resulting into the emplacement of the western ophiolites, with exhumation extending from the Late Miocene to the present \cite{Yumul2013a}.
+In addition to collision, arc magmatism contributed significantly to crustal growth in the Philippines \cite{Dimalanta2006a}.
 
@@
 
-Paleoshoreline reconstructions of the Sunda-Banda Arc system relied largely on the work of \citet{Hall2013a}.
-The Sunda-Banda Arc system is composed of thick sequences of mixed sedimentary rocks and basement intruded and overlain by igneous rocks derived from continual arc magmatism through much of the Cenozoic \citep{Hall2017a}.
+Paleoshoreline reconstructions of the Sunda-Banda Arc system relied largely on the work of Hall (2013) \cite{Hall2013a}.
+The Sunda-Banda Arc system is composed of thick sequences of mixed sedimentary rocks and basement intruded and overlain by igneous rocks derived from continual arc magmatism through much of the Cenozoic \cite{Hall2017a}.
 Sumatra is the largest island within the Sunda-Banda Arc system that stretches from the Andaman Sea in the northwest to the Banda Sea in the east.
-The western portion of the Sumatra is underlain by the Wolya ophiolite, which was obducted in the Cretaceous and exhumed in the Eocene to Oligocene \citep{Allen2008a}, whereas the southeast corner on the islands of Bangka and Belitung is composed of granite basement of the Sunda Shelf \citep{Hall2009a, Hall2013b}.
-Java is composed of a complex of E-W-striking deformational and magmatic belts \citep{AudleyCharles2004a}.
-The belts have a volcanic arc in the south and a continental shelf to the north with an adjacent sedimentary basin \citep{AudleyCharles2004a}.
-Bali and Flores are mostly volcanic with some sedimentary cover \citep{Hall2009a}.
+The western portion of the Sumatra is underlain by the Wolya ophiolite, which was obducted in the Cretaceous and exhumed in the Eocene to Oligocene \cite{Allen2008a}, whereas the southeast corner on the islands of Bangka and Belitung is composed of granite basement of the Sunda Shelf \cite{Hall2009a, Hall2013b}.
+Java is composed of a complex of E-W-striking deformational and magmatic belts \cite{AudleyCharles2004a}.
+The belts have a volcanic arc in the south and a continental shelf to the north with an adjacent sedimentary basin \cite{AudleyCharles2004a}.
+Bali and Flores are mostly volcanic with some sedimentary cover \cite{Hall2009a}.
 
-Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \citep{Harris2006a}.
+Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \cite{Harris2006a}.
 From ca.
-20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf \citep{Hall2009a, Hall2013b}.
+20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf \cite{Hall2009a, Hall2013b}.
 Six major islands were subaerially exposed before 5~Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores.
-However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5~Ma \citep{Hall2009a, Hall2013b}.
-Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}.
+However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5~Ma \cite{Hall2009a, Hall2013b}.
+Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{AudleyCharles2004a, Harris2006a, Hall2013b}.
 In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $>$1~km above sea level with estimated average uplift rates of 1.5~km/m.y.
-\citep{AudleyCharles1986a}.
+\cite{AudleyCharles1986a}.
 
@@
 
-\bibliographystyle{gsabull}
+\bibliographystyle{nature}
 \bibliography{References}

### 2019-11-20 Manuscript/Manuscript.tex
Updates from Overleaf

--- 2019-11-13 Manuscript/Manuscript.tex
@@
 
+It is sometimes argued that hypotheses that invoke enhanced weathering in regions of active uplift predict an increase in global alkalinity fluxes.
+However, global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state.
+Enhanced weathering in a region such as the SEAI is compensated by a decrease in weathering in other regions where there has not been tectonically-forced change.
+Global alkalinity delivery from silicate weathering does not change, but occurs more efficiently and thereby at lower \pCOtwo.
+However, given that carbonate weathering is disconnected from long-term carbon-cycle mass balance, alkalinity delivery from carbonate weathering could change in an evolving climate and be observable in proxy records \cite{Si2019a}.
+
 Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.

### 2019-11-26 Manuscript/Manuscript.tex
Updates from Overleaf

--- 2019-11-13 Manuscript/Manuscript.tex
@@
 % MAXIMUM 125 CHARACTERS 
-\textbf{One Sentence Summary:} Arc-continent collision in the Southeast Asian islands increased global weatherability enough to drive Neogene cooling.
+\textbf{One Sentence Summary:} Growth of tropical islands led to efficient carbon consumption and drove cooling over the past 15 million years.
 
@@
 \begin{abstract}
-The Southeast Asian islands bring together steep topography, a tropical climate, and mafic lithologies to efficiently weather rocks in the region and consume large amounts of \COtwo.
-Concurrent with global cooling since the mid-Miocene, ongoing arc-continent collision has led to a significant increase of subaerially-exposed land area in the region.
-Here we use geologic data to reconstruct the shorelines of the Southeast Asian islands over the past 15 million years and quantify the increase in land area.
-We use these paleoshorelines as boundary conditions within a coupled weathering-climate model to quantify the change in steady-state \pCOtwo associated with this emergence.
+The steep topography, tropical climate, and mafic lithologies of the Southeast Asian islands contribute to efficient chemical weathering and carbon sequestration.
+Concurrent with global cooling since the mid-Miocene, ongoing collision between the Sunda-Banda arc and Australia has increased the area of subaerially-exposed land in the region.
+Here we integrate lithology into a coupled weathering-climate model to assess how the global climate system responds to changes in paleogeography.
+We then use paleoshoreline data as boundary conditions to quantify the change in steady-state \pCOtwo associated with the tectonic emergence of the Southeast Asian islands since the mid-Miocene.
 We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
@@
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2001a}.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a}.
+15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2008a}.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a, Galy2007a}.
 Here we instead propose that SEAI emergence was itself sufficient to explain long-term climatic cooling over the Neogene.
@@
 
-Himalayan uplift \cite{Raymo1992a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
+Himalayan uplift \cite{Raymo1992a, Galy2007a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalaya lithologies \cite{Colleps2018a}, but could also be partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
+Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.
 
-Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have effected geological carbon sources and sinks.
-Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated \pCOtwo change is sufficient to explain the magnitude of cooling.
+It has been argued that enhanced weathering in regions of active uplift predicts an increase in global alkalinity fluxes.
+However, global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
+Enhanced weathering in a region such as SEAI is compensated by a decrease in weathering elsewhere.
+Global alkalinity delivery from silicate weathering does not change, but occurs more efficiently and thereby at lower \pCOtwo.
+Given that carbonate weathering is disconnected from the long-term carbon-cycle mass balance, changes in carbonate accumulation through time \cite{Si2019a} could be driven by changes in carbonate weathering.
+
+Coupled geological constraints and modeling experiments demonstrate that SEAI has been a growing hot spot for carbon sequestration from the Miocene to present.
+Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
+Such changes are likely to have induced complex feedbacks on many Earth surface processes.
+Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain the magnitude of cooling.
 These results highlight that the Earth's climate state is particularly sensitive to changes in tropical paleogeography.

### 2020-03-31 Manuscript/Manuscript.tex
convert to PNAS template - formatting/revisions to come

@@
-% more detailed formatting: https://www.sciencemag.org/authors/instructions-preparing-initial-manuscript
-% formatting: https://www.sciencemag.org/authors/science-information-authors
-% submit at: https://cts.sciencemag.org/
+% PNAS website: https://www.pnas.org/page/authors/submission
 
-% Reports (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance.
-Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.
-Materials and Methods should be included in supplementary materials, which should also include information needed to support the paper's conclusions.
+% Only a single PDF file containing all text, figures, tables, and supporting information (SI) is required for initial submissions; high-resolution files are not required.
 
-% Titles should be no more than 96 characters (including spaces).
+% The preferred length of articles is 6 pages, but PNAS will allow articles up to a maximum of 12 pages.
 
-% Short titles should be no more than 40 characters (including spaces).
+% Proof pages - Word count - Maximum no. of figures and tables - Maximum no. of references
+% 6   4,000   4   50
+% 7   4,500   5   55
+% 8   5,000   6   60
+% 9   5,500   7   70
+% 10  6,000   8   75
+% 11  6,500   9   80
+% 12  7,000   10  85
 
-% One-sentence summaries capturing the most important point should be submitted for Research Articles, Reports and Reviews.
-These should be a maximum of 125 characters and should complement rather than repeat the title
+% Include department, institution, and complete address, with the ZIP/postal code, for each author.
+Use lower case letters to match authors with institutions, as shown in the example.
+Authors with an ORCID ID may supply this information at submission.
 
-% Abstracts of Research Articles and Reports should explain to the general reader why the research was done, what was found and why the results are important.
-They should start with some brief BACKGROUND information: a sentence giving a broad introduction to the field comprehensible to the general reader, and then a sentence of more detailed background specific to your study.
-This should be followed by an explanation of the OBJECTIVES/METHODS and then the RESULTS.
-The final sentence should outline the main CONCLUSIONS of the study, in terms that will be comprehensible to all our readers.
-The Abstract is distinct from the main body of the text, and thus should not be the only source of background information critical to understanding the manuscript.
-Please do not include citations or abbreviations in the Abstract.
-The abstract should be 125 words or less.
+% All authors must submit their articles at http://www.pnascentral.org/cgi-bin/main.plex.
+If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
 
-% Main Text is not divided into sub-headings for Reports. Subheadings are used only in Research Articles, and Reviews.
+% Images must be provided at final size, preferably 1 column width (8.7cm).
+Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide.
+Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.
 
-% References and Notes are numbered in the order in which they are cited, first through the text, then through the figure and table legends and finally through Supplementary Materials.
+% Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references.
+PNAS will publish SI uncomposed, as the authors have provided it.
+Additional details can be found here: https://www.pnas.org/page/authors/format#Supporting_Information.
+Refer to the SI Appendix in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.
 
-\documentclass[11pt,letterpaper]{article}
+% Use the lineno option to display guide line numbers if required.
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
 
-%\usepackage{fontspec}
-%\usepackage[utf8]{inputenc}
-\usepackage{textcomp,marvosym}
-\usepackage{amsmath,amssymb}
-\usepackage[normalem]{ulem}
-\usepackage[left]{lineno}
-\usepackage{booktabs}
-\usepackage{changepage}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{natbib}
-\usepackage{setspace}
-\usepackage{array}
-\usepackage{fancyhdr}
-\usepackage{graphicx}
+% other packages
 \usepackage{xspace}
-\usepackage[hidelinks]{hyperref}
-\urlstyle{same}
-\usepackage{threeparttable}
-\doublespacing
 
-\raggedright
-\textwidth = 6.5 in
-\textheight = 8.25 in
-\oddsidemargin = 0.0 in
-\evensidemargin = 0.0 in
-\topmargin = 0.0 in
-\headheight = 0.0 in
-\headsep = 0.5 in
-\parskip = 0.1 in
-\parindent = 0.2in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-
-% Remove brackets from numbering in List of References
-%\makeatletter
-%\renewcommand{\@biblabel}[1]{\quad#1.}
-%\makeatother
-
-% Self defined commands
+% self-defined commands
 \newcommand{\degreesC}{\textdegree C\xspace}
@@
 \newcommand{\permil}{\textperthousand\xspace}
-\newcommand{\UPb}{$^{206}$Pb/$^{238}$U\xspace}
 \newcommand{\pCOtwo}{\textit{p}CO$_{2}$\xspace}
 \newcommand{\COtwo}{CO$_{2}$\xspace}
-%
 
-\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{Park et al., in preparation}
-\rhead{\thepage}
+\templatetype{pnasresearcharticle}
 
-\begin{document}
+\title{Emergence of the Southeast Asian islands as a driver for Neogene cooling}
 
-\begin{flushleft}
-{\Large \textbf{Emergence of the Southeast Asian islands as a driver for Neogene cooling}}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Yuem Park}
+\author[a]{Pierre Maffre} 
+\author[b]{Yves Godd\'eris}
+\author[c]{Francis A. Macdonald}
+\author[c]{Eliel A. Anttila}
+\author[a]{Nicholas L. Swanson-Hysell}
 
-Yuem Park\textsuperscript{1},
-Pierre Maffre\textsuperscript{1},
-Yves Godd\'eris\textsuperscript{3},
-Francis A. Macdonald\textsuperscript{2},
-Eliel A. Anttila\textsuperscript{2},
-Nicholas L. Swanson-Hysell\textsuperscript{1}
+\affil[a]{Department of Earth and Planetary Science, University of California, Berkeley, CA 94720, USA}
+\affil[b]{G\'eosciences Environnement Toulouse, CNRS--Universit\'e Paul Sabatier - IRD, Toulouse 31400, France}
+\affil[c]{Department of Earth Science, University of California, Santa Barbara, CA 93106, USA}
 
-\bigskip
-\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA
+% Please give the surname of the lead author for the running footer
+\leadauthor{Park}
 
-\textsuperscript{2} Department of Earth Science, University of California, Santa Barbara, CA, USA
+% Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
+The significance statement appears in the paper itself and is required for all research papers.
+\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
+The significance statement appears in the paper itself and is required for all research papers.}
 
-\textsuperscript{3} G\'eosciences Environnement Toulouse, CNRS--Universit\'e Paul Sabatier - IRD, Toulouse, France
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{No competing interests to declare.}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: yuempark@berkeley.edu}
 
-\bigskip
+% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
+\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
 
-\end{flushleft}
-
-\linenumbers
-
-\noindent
-% MAXIMUM 125 CHARACTERS 
-\textbf{One Sentence Summary:} Growth of tropical islands led to efficient carbon consumption and drove cooling over the past 15 million years.
-
-% MAXIMUM 125 WORDS
-% NO CITATIONS OR ABBREVIATIONS
+% Please provide an abstract of no more than 250 words in a single paragraph.
+Abstracts should explain to the general reader the major contributions of the article.
+References in the abstract must be cited in full within the abstract itself and cited in the text.
 \begin{abstract}
@@
 
-The Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+\dropcap{T}he Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a}.
@@
 
-%%TC:ignore
-\section*{ACKNOWLEDGEMENTS \label{sec:ACKNOWLEDGEMENTS}}
+\section*{Guide to using this template on Overleaf}
 
-Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
-FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
-We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
-The GEOCLIM code associated with this study is openly available: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}
-%%TC:endignore
-
-\clearpage
-\newpage
-
-\section*{FIGURES}
-
-% 70 words
-\begin{figure}[h!]
+\begin{figure}
     \centering
-    \includegraphics[width=0.9\textwidth]{Figures/shoreline_growth.pdf}
+    \includegraphics[width=0.9\linewidth]{Figures/shoreline_growth.pdf}
     \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
@@
 
-% 100 words
-\begin{figure}[h!]
+\begin{figure}
     \centering
-    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2.pdf}
+    \includegraphics[width=1\linewidth]{Figures/scenario_pCO2.pdf}
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
@@
 
-\clearpage
-\newpage
-\footnotesize
+\subsection*{Author Affiliations}
 
-\singlespacing
+x
 
-\bibliographystyle{nature}
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=.8\linewidth]{Figures/regolith_schematic.pdf}
+\caption{Placeholder image of a frog with a long example legend to show justification setting.}
+\label{fig:frog}
+\end{figure}
+
+\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\centering
+\includegraphics[width=11.4cm,height=11.4cm]{Figures/regolith_schematic.pdf}
+\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
+\end{SCfigure*}
+
+\begin{figure*}[bt!]
+\begin{align*}
+(x+y)^3&=(x+y)(x+y)^2\\
+       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
+       &=x^3+3x^2y+3xy^3+x^3. 
+\end{align*}
+\end{figure*}
+
+\begin{table}%[tbhp]
+\centering
+\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
+\begin{tabular}{lrrr}
+Species & CBS & CV & G3 \\
+\midrule
+1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
+2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
+3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+\bottomrule
+\end{tabular}
+
+\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
+\end{table}
+
+\matmethods{Please describe your materials and methods here.
+This can be more than one paragraph, and may contain subsections and equations as required.
+Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
+
+\subsection*{Subsection for Method}
+Example text for subsection.
+}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
+FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
+The GEOCLIM code associated with this study is openly available: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}}
+
+\showacknow{} % Display the acknowledgments section
+
+% Bibliography
 \bibliography{References}

### 2020-04-01 Code/Notebooks/weatherability_curves/weatherability_SEAI.py
Add script and outputs to draw weatherability plot

@@ -1,0 +1,57 @@
+import netCDF4 as nc
+import numpy as np
+from matplotlib import pyplot as plt
+from cycler import cycler
+
+####################################
+##########   PARAMETERS   ##########
+RECORD = True
+
+# COLOR CYCLE - 5 colors
+COLOR = ['royalblue','purple','firebrick','coral','gold']
+plt.rc( 'axes', prop_cycle=cycler('color',COLOR) )
+
+# FILES
+f = list()
+f.append( nc.Dataset('GDSS_GFDL_3lvl_ctrl.nc') )
+f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA5.nc') )
+f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA10.nc') )
+f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA.nc') )
+f.append( nc.Dataset('GDSS_GFDL_3lvl_noIA.nc') )
+####################################
+####################################
+
+####################################################
+
+##  Plotting
+
+fig, ax = plt.subplots(figsize=(9,6))
+fig.delaxes(ax)
+ax = fig.add_axes([0.11,0.13,0.87,0.85])
+
+#--------------------------------------------------#
+
+for fk in f:
+    ax.plot( fk['atm_CO2_level'][:], 1e-12*fk['volcanic_degassing'][:], linestyle='-', marker='o', linewidth=3 )
+
+ylim = [3.25, 5]
+ax.set_ylim(ylim)
+
+for k in range(len(f)):
+    Slopemean = 1e-12 * (f[k]['volcanic_degassing'][-1]-f[k]['volcanic_degassing'][0])/2
+    ax.text(900, ylim[0]+(0.4-(0.35/len(f))*k)*(ylim[1]-ylim[0]),
+            '~{0:.2f} Tmol per 2xCO$_2$'.format(Slopemean), color=COLOR[k], fontsize=14)
+
+ax.legend(['PD ctrl','5 Myr','10 Myr','15 Myr','no SEAI'], fontsize=15)
+ax.set_xlabel('Atm. CO$_2$ (ppm)', fontsize=18)
+ax.set_ylabel('Sil. weathering flux (Tmol of CO$_2$ / yr)', fontsize=18)
+ax.tick_params(axis='both', labelsize=14)
+
+####################################################
+
+##  Drawing
+
+if RECORD:
+    fig.savefig('weatherability_SEAI.pdf', format='pdf', transparent=True)
+else:
+    plt.show()

### 2020-04-07 Code/Notebooks/weatherability_curves/weatherability_SEAI.py
update weatherability curves figure

@@ -1,57 +1,0 @@
-import netCDF4 as nc
-import numpy as np
-from matplotlib import pyplot as plt
-from cycler import cycler
-
-####################################
-##########   PARAMETERS   ##########
-RECORD = True
-
-# COLOR CYCLE - 5 colors
-COLOR = ['royalblue','purple','firebrick','coral','gold']
-plt.rc( 'axes', prop_cycle=cycler('color',COLOR) )
-
-# FILES
-f = list()
-f.append( nc.Dataset('GDSS_GFDL_3lvl_ctrl.nc') )
-f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA5.nc') )
-f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA10.nc') )
-f.append( nc.Dataset('GDSS_GFDL_3lvl_redIA.nc') )
-f.append( nc.Dataset('GDSS_GFDL_3lvl_noIA.nc') )
-####################################
-####################################
-
-####################################################
-
-##  Plotting
-
-fig, ax = plt.subplots(figsize=(9,6))
-fig.delaxes(ax)
-ax = fig.add_axes([0.11,0.13,0.87,0.85])
-
-#--------------------------------------------------#
-
-for fk in f:
-    ax.plot( fk['atm_CO2_level'][:], 1e-12*fk['volcanic_degassing'][:], linestyle='-', marker='o', linewidth=3 )
-
-ylim = [3.25, 5]
-ax.set_ylim(ylim)
-
-for k in range(len(f)):
-    Slopemean = 1e-12 * (f[k]['volcanic_degassing'][-1]-f[k]['volcanic_degassing'][0])/2
-    ax.text(900, ylim[0]+(0.4-(0.35/len(f))*k)*(ylim[1]-ylim[0]),
-            '~{0:.2f} Tmol per 2xCO$_2$'.format(Slopemean), color=COLOR[k], fontsize=14)
-
-ax.legend(['PD ctrl','5 Myr','10 Myr','15 Myr','no SEAI'], fontsize=15)
-ax.set_xlabel('Atm. CO$_2$ (ppm)', fontsize=18)
-ax.set_ylabel('Sil. weathering flux (Tmol of CO$_2$ / yr)', fontsize=18)
-ax.tick_params(axis='both', labelsize=14)
-
-####################################################
-
-##  Drawing
-
-if RECORD:
-    fig.savefig('weatherability_SEAI.pdf', format='pdf', transparent=True)
-else:
-    plt.show()

### 2020-04-08 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \usepackage{xspace}
+\urlstyle{same}
 
@@
 \newcommand{\COtwo}{CO$_{2}$\xspace}
+\newcommand{\SI}{\textit{SI Appendix}\xspace}
+\newcommand{\MM}{\textit{Materials and Methods}\xspace}
 
@@
 \authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{No competing interests to declare.}
+\authordeclaration{The authors declare no competing interest.}
 \correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: yuempark@berkeley.edu}
@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
+\keywords{silicate weathering $|$ weatherability $|$ arc-continent collision $|$ Neogene cooling $|$ Southeast Asian islands}
 
@@
 As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo \cite{Kump1997a}.
-Conversely, as regions with high carbon sequestration potential emerge on Earth's surface, \pCOtwo decreases resulting in other weathering sinks to diminish leading to a new lower steady-state \pCOtwo concentration.
+Conversely, as regions with high carbon sequestration potential emerge on Earth's surface, \pCOtwo decreases resulting in other weathering sinks to diminish leading to a new lower steady-state \pCOtwo.
 
 Topography, climate, and lithology all effect chemical weathering.
-High-relief regions lead to reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a}.
+High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a}.
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{West2012a}.
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2017c}.
-GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in chemical weathering and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
+Previous work has estimated that the decrease in \pCOtwo associated with the emergence of the SEAI since 5~Ma is $\sim$19~ppm \cite{Molnar2015a}.
+This 19~ppm estimate was obtained using a 0-D equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
+\pCOtwo was then estimated from the calculated temperature using a simple energy balance equation.
+However, the relationship between mean global temperature (or \pCOtwo) and weathering-rate-weighted land area is not linear.
+As regions with high carbon sequestration potential, such as the SEAI, emerge on Earth's surface, the way in which silicate weathering rates evolve around the globe to adjust to the higher global weatherability depends on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
+For example, if the spatial distribution of topography, climate, and lithology were such that Earth's surface overall had a relatively low carbon sequestration potential at the time of SEAI emergence, then a larger \pCOtwo change would be required to balance the sinks and sources of \COtwo than if Earth's surface overall had a relatively high carbon sequestration potential.
+However, silicate weathering within the emerging SEAI themselves would be adapting to changing \pCOtwo, ultimately resulting in a complex, non-linear, and spatially-varying system of feedbacks between climate and silicate weathering.
+Therefore, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must account for the spatial distribution of topography, climate, and lithology.
+
+\section*{Methods}
+
+\subsection*{GEOCLIM model}
+
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in the SEAI, we use the globally spatially-resolved GEOCLIM model \cite{Godderis2017c}.
+GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in erosion, chemical weathering, and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/regolith_schematic.pdf}
+    \caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state.
+A rock particle leaves the unweathered bedrock with production rate $P_{r}$, and transits vertically through a regolith of height $h$.
+Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
+As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
+Details of the formulation for the silicate weathering component of GEOCLIM can be found in \MM.}
+    \label{fig:regolith_schematic}
+\end{figure}
 
@@
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and all bedrock was assigned identical chemical compositions \cite{Godderis2017c}.
-More recent versions of GEOCLIM implement regolith development and soil shielding, which introduces a dependence on erosion (and therefore topographic slope) \cite{Maffre2018a}.
+More recent versions of GEOCLIM implement regolith development and soil shielding (Fig.
+\ref{fig:regolith_schematic}), which introduces a dependence on erosion (and therefore topographic slope) \cite{Maffre2018a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
-We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM.
-The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI).
+We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM (\SI).
+The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment.
 Each land pixel is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
@@
 
+\subsection*{Calibration}
+
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
-Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI).
+Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (\SI).
 This results in 93,600 unique combinations of these six parameters.
@@
 After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (see SI).
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (\SI).
 
 Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with SEAI emergence.
-For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI).
+For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (\SI).
 The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
 
+\subsection*{Paleoshorelines}
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=11.4cm]{Figures/shoreline_growth.pdf}
+    \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
+Past shorelines are shown in \textit{A} with associated land area summarized in \textit{B}.
+A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in \textit{C}.}
+    \label{fig:shoreline_growth}
+\end{figure*}
+
 To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
-\ref{fig:shoreline_growth}; see SI).
+\ref{fig:shoreline_growth}; \SI).
 The paleoshoreline data indicate that the Sunda-Banda Arc and New Guinea are primarily responsible for the increase in area since 15~Ma.
@@
 \ref{fig:shoreline_growth}C).
-We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (see SI).
+We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI).
 
-We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of SEAI at ca.
+We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of the SEAI at ca.
 15, 10, and 5~Ma (``paleo-SEAI'' scenarios; Fig.
 \ref{fig:scenario_pCO2}).
-Because we use a clmate model forced with modern geography, the position of the tectonic blocks remain fixed.
-Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI).
+Because we use a climate model forced with modern geography, the position of the tectonic blocks remain fixed.
+Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (\SI).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).
+
+\section*{Results and Discussion}
+
+\subsection*{Neogene cooling}
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/scenario_pCO2.pdf}
+    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
+For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
+The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
+the interquartile range), and the notch represents the median with its 95\% confidence interval.
+The whiskers extend to the 2.5 and 97.5 percentile values.
+Glaciation thresholds \cite{DeConto2008a} are shown on the x-axis.}
+    \label{fig:scenario_pCO2}
+\end{figure*}
 
@@
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$493--709~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$493--709~ppm.
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
-The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI), indicating that the increase in subaerially-exposed land area and tectonic topography of the SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 
+\subsection*{Marine isotopes}
+
 Himalayan uplift \cite{Raymo1992a, Galy2007a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in SEAI during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
 Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.
@@
 However, global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
-Enhanced weathering in a region such as SEAI is compensated by a decrease in weathering elsewhere.
+Enhanced weathering in a region such as the SEAI is compensated by a decrease in weathering elsewhere.
 Global alkalinity delivery from silicate weathering does not change, but occurs more efficiently and thereby at lower \pCOtwo.
@@
 
-Coupled geological constraints and modeling experiments demonstrate that SEAI has been a growing hot spot for carbon sequestration from the Miocene to present.
+\section*{Conclusions}
+
+Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration from the Miocene to present.
 Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
@@
 
-\section*{Guide to using this template on Overleaf}
+\matmethods{
+The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}.
+The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{Figures/shoreline_growth.pdf}
-    \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
-Past shorelines are shown in A with associated land area summarized in B.
-A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in C.}
-    \label{fig:shoreline_growth}
-\end{figure}
+\subsection*{GEOCLIM silicate weathering component}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=1\linewidth]{Figures/scenario_pCO2.pdf}
-    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI).
-The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
-the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to the 2.5 and 97.5 percentile values.
-Glaciation thresholds \cite{DeConto2008a} are shown on the x-axis.}
-    \label{fig:scenario_pCO2}
-\end{figure}
+The silicate weathering component of the GEOCLIM model has been modified from the previously published version \cite{Godderis2017b}.
+The new component implements the model of Gabet and Mudd (2009) \cite{Gabet2009a} for the development of a chemically weathered profile.
+We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
+In the model of Gabet and Mudd (2009), material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
+We use the DynSoil implementation of the Gabet and Mudd (2009) model, which integrates chemical weathering within the regolith \cite{West2012a}.
+The transient time-varying version of this regolith model is described by three equations:
 
-\subsection*{Author Affiliations}
+\begin{equation}
+    \frac{dh}{dt} = P_{r} - E_{p}
+    \label{eq:1}
+\end{equation}
 
-x
+\begin{equation}
+    \frac{\partial x}{\partial t} = -P_{r} \frac{\partial x}{\partial z} - K \tau^{\sigma}x
+    \label{eq:2}
+\end{equation}
 
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=.8\linewidth]{Figures/regolith_schematic.pdf}
-\caption{Placeholder image of a frog with a long example legend to show justification setting.}
-\label{fig:frog}
-\end{figure}
+\begin{equation}
+    \frac{\partial \tau}{\partial t} = -P_{r} \frac{\partial \tau}{\partial z} + 1
+    \label{eq:3}
+\end{equation}
 
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
-\centering
-\includegraphics[width=11.4cm,height=11.4cm]{Figures/regolith_schematic.pdf}
-\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
-\end{SCfigure*}
+\noindent
+Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the model time (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
+Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth ($z$, m).
+$K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless) which implements a time-dependence.
+Equation \ref{eq:3} describes how the time that a given rock particle has spent in the regolith changes as time in the model progresses.
 
-\begin{figure*}[bt!]
-\begin{align*}
-(x+y)^3&=(x+y)(x+y)^2\\
-       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-       &=x^3+3x^2y+3xy^3+x^3. 
-\end{align*}
-\end{figure*}
+The net weathering rate in the regolith column ($W$, m/yr) can then be calculated with:
 
-\begin{table}%[tbhp]
-\centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
-\midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-\bottomrule
-\end{tabular}
+\begin{equation}
+    W = \int_{0}^{h} K \tau^{\sigma} x\;dz
+    \label{eq:4}
+\end{equation}
 
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-\end{table}
+The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and a soil production function ($f(h)$):
 
-\matmethods{Please describe your materials and methods here.
-This can be more than one paragraph, and may contain subsections and equations as required.
-Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
+\begin{equation}
+    P_{r} = P_{0}\;f(h)
+    \label{eq:5}
+\end{equation}
 
-\subsection*{Subsection for Method}
-Example text for subsection.
+\begin{equation}
+    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:6}
+\end{equation}
+
+\begin{equation}
+    f(h) = e^{-\frac{-h}{d_{0}}}
+    \label{eq:7}
+\end{equation}
+
+\noindent
+$P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
+In Equation \ref{eq:6}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K), we parameterize the `optimal' regolith production rate \cite{Carretier2014a}.
+$f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases.
+It takes an exponential form as is commonly applied in the literature \cite{Gabet2009a}.
+In Equation \ref{eq:7}, $d_{0}$ is a reference regolith thickness (m) \cite{Heimsath1997a}.
+
+Our implementation of the erosion rate is parameterized based on runoff and slope ($s$; m/m):
+
+\begin{equation}
+    E_{p} = k_{e}\;q^{m}\;s^{n}
+    \label{eq:8}
+\end{equation}
+
+\noindent
+$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \cite{Maffre2018a}.
+This formulation is directly inspired by the stream power law \cite{Davy2000a}.
+This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \cite{Lague2013a}.
+
+The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the chemical weathering on climate:
+
+\begin{equation}
+    K = k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
+    \label{eq:9}
+\end{equation}
+
+\noindent
+Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \cite{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
+
+In this study, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
+The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations:
+
+\begin{equation}
+    h = \max\left(0,\;d_{0} \log\left(\frac{P_{0}}{E_{p}}\right)\right)
+    \label{eq:10}
+\end{equation}
+
+\begin{equation}
+    x(z) = \exp\left(-\frac{K}{\sigma+1}\left(\frac{z}{E_{p}}\right)^{\sigma+1}\right)
+    \label{eq:11}
+\end{equation}
+
+\begin{equation}
+    W = E_{p}(1-x(h)) = E_{p}(1-x_s)
+    \label{eq:12}
+\end{equation}
+
+\noindent
+$x(z)$ is the abundance profile of primary phases inside the regolith, varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
+Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column.
+
+\subsection*{SI Appendix}
+
+A detailed description of the implementation of lithology into the silicate weathering component of GEOCLIM, the calibration of the silicate weathering component of GEOCLIM, the GFDL CM2.0 climate model, and the paleoshoreline reconstruction can be found in the \SI.
 }
@@
 FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
-We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.
-The GEOCLIM code associated with this study is openly available: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}}
+We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.}
 

### 2020-04-08 Manuscript/Supplementary_Information.tex
Updates from Overleaf

--- 2019-11-05 Manuscript/Supplementary_Information/Supplementary_Information.tex
@@
 
-\section*{Silicate Weathering Component}
-
-\begin{SCfigure}[0.6][h!]
-\begin{center}
-	\includegraphics[width=0.6\textwidth]{../Figures/regolith_schematic.pdf}
-	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state.
-A rock particle leaves the unweathered bedrock with production rate $P_{r}$, and transits vertically through a regolith of height $h$.
-Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
-As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).}
-	\label{fig:regolith_schematic}
-\end{center}
-\end{SCfigure}
-
-The silicate weathering component of the GEOCLIM model has been modified from the previously published version \citep{Godderis2017b}.
-The new component implements the model of \citet{Gabet2009a} for the development of a chemically weathered profile.
-We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock.
-In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top.
-We use the DynSoil implementation of the \citet{Gabet2009a} model, which integrates chemical weathering within the regolith \citep{West2012a}.
-The transient time-varying version of this regolith model is described by three equations:
-
-\begin{equation}
-    \frac{dh}{dt} = P_{r} - E_{p}
-    \label{eq:1}
-\end{equation}
-
-\begin{equation}
-    \frac{\partial x}{\partial t} = -P_{r} \frac{\partial x}{\partial z} - K \tau^{\sigma}x
-    \label{eq:2}
-\end{equation}
-
-\begin{equation}
-    \frac{\partial \tau}{\partial t} = -P_{r} \frac{\partial \tau}{\partial z} + 1
-    \label{eq:3}
-\end{equation}
-
-\noindent
-Equation \ref{eq:1} is a statement of material conservation, where $h$ is the total height of the regolith (m), $t$ is the model time (yr), $P_{r}$ is the regolith production rate (m/yr), and $E_{p}$ is the physical erosion rate (m/yr).
-Equation \ref{eq:2} describes how the residual fraction of weatherable phases ($x$, unitless) changes as a function of time ($t$, yr) and depth ($z$; m).
-$K \tau^{\sigma}$ is the dissolution rate constant, which depends on the local climate (captured by $K$, yr$^{-1-\sigma}$) and the time that a given rock particle has spent in the regolith ($\tau$, yr) to some power $\sigma$ (unitless) which implements a time-dependence.
-Equation \ref{eq:3} describes how the time that a given rock particle has spent in the regolith changes as time in the model progresses.
-
-The net weathering rate in the regolith column ($W$, m/yr) can then be calculated with:
-
-\begin{equation}
-    W = \int_{0}^{h} K \tau^{\sigma} x\;dz
-    \label{eq:4}
-\end{equation}
-
-The regolith production rate can be expressed as the product of the optimal production rate ($P_{0}$) and a soil production function ($f(h)$):
-
-\begin{equation}
-    P_{r} = P_{0}\;f(h)
-    \label{eq:5}
-\end{equation}
-
-\begin{equation}
-    P_{0} = k_{rp}\;q\;e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
-    \label{eq:6}
-\end{equation}
-
-\begin{equation}
-    f(h) = e^{-\frac{-h}{d_{0}}}
-    \label{eq:7}
-\end{equation}
-
-\noindent
-$P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith.
-In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K).
-$f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases.
-It takes an exponential form as is commonly applied in the literature (e.g.
-\citealp{Gabet2009a}).
-In Equation \ref{eq:7}, we follow \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).
-
-Our implementation of the erosion rate is parameterized based on runoff and slope ($s$; m/m):
-
-\begin{equation}
-    E_{p} = k_{e}\;q^{m}\;s^{n}
-    \label{eq:8}
-\end{equation}
-
-\noindent
-$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \citep{Maffre2018a}.
-This formulation is directly inspired by the stream power law \citep{Davy2000a}.
-This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \citep{Lague2013a}.
-
-The $K$ in the dissolution rate constant in Equation \ref{eq:2} describes the dependence of the chemical weathering on climate:
-
-\begin{equation}
-    K = k_{d}\left(1-e^{-k_{w}q}\right)e^{-\frac{E_{a}}{R}\left(\frac{1}{T}-\frac{1}{T_{0}}\right)}
-    \label{eq:9}
-\end{equation}
-
-\noindent
-Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).
-
-In this study, we are interested in obtaining the steady-state solution rather than the transient time-varying solution.
-The steady-state solution for DynSoil can be calculated analytically by setting the time derivatives equal to zero resulting in the following set of equations:
-
-\begin{equation}\label{eq:dynsoil_ss}
-\left\{\ 
-\begin{aligned}
-& h   \ =\    \max\left(  \;  0  \;,\;  d_{0} \log\left(\frac{P_{0}}{E_{p}}\right)  \right)                                                 \\
-& x(z)      \ =\   \exp\left(   - \frac{K}{\sigma+1}  \left(\frac{z}{E_{p}}\right)^{\sigma+1}   \right)   \\
-& W        \ =\   (1-x(h)) \, E_{p}  \ =\   (1-x_s) \, E_{p}                                                                         \\
-\end{aligned}
-\right.
-\end{equation}
-
-\noindent
-$x(z)$ is the abundance profile of primary phases inside the regolith, varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}.
-Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column.
-
 \section*{Implementation of Lithology}
@@
 \vspace{10pt}
-\caption{Grouping of 16 lithologic categories in GLiM \citep{Hartmann2012a} to 6 broader categories for GEOCLIM.}
+\caption{Grouping of 16 lithologic categories in GLiM \citet{Hartmann2012a} to 6 broader categories for GEOCLIM.}
 \label{tab:GLiM_to_GEOCLIM}
@@
 To calculate this flux, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
-\ref{fig:regolith_schematic}).
+1).
 Previous implementations of GEOCLIM have used a ``diffuse lithology'' where all exposed land is assigned the composition of bulk upper continental crust.
@@
 
-Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
+Experimental determinations of the activation energy ($E_a$; Equations 6 and 9) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}.
 However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}).
@@
 
-Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}.
+Within the equation that governs the physical erosion rate in GEOCLIM (Equation 8), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}.
 However, there is uncertainty in the value of the proportionality constant ($k_e$).
@@
 
-Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation \ref{eq:9}), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation \ref{eq:9}), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation \ref{eq:4}), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation \ref{eq:6}).
+Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation 9), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation 9), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation 4), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation 6).
 Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
@@
 Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields.
-As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}.
+As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a}, \citet{Moquet2016a}, and \citet{Moquet2018a}.
 The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes.
@@
 Each row represents a lithologic category.
-Solid black lines show present-day shorelines.}
+Solid black lines show present-day shorelines.
+Total areas of each lithologic category within the SEAI for each of these tested scenarios is shown in Figure \ref{fig:lithology_areas}.}
     \label{fig:SEAI_fracs}
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
+    \includegraphics[width=1\textwidth]{Manuscript/Figures/lithology_areas.pdf}
+    \caption{Total area of the lithologic categories within the SEAI for each tested scenario.
+Note that the grey bars in the middle panel indicate the fractional area of each lithologic category for the entire Earth.}
+    \label{fig:lithology_areas}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
+    \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
     \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
@@
 
-The decrease in \pCOtwo that we estimate associated with the emergence of SEAI since 5~Ma (92--161~ppm; Fig.
-\ref{fig:scenario_pCO2_all}) is greater than that proposed in \citet{Molnar2015a} ($\sim$19~ppm).
-This 19~ppm value in \citet{Molnar2015a} is obtained from an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor $\alpha$ that is intended to account for the influence of both runoff and temperature (Equation 3 in \citealp{Molnar2015a}).
-\pCOtwo is then estimated from $T$ using a simple energy balance equation.
-However, the relationship between mean global temperature and weathering-rate-weighted land area is not linear.
-Furthermore, as acknowledged by the authors themselves, this simple linear relationship ignores spatial variability in runoff, temperature, and slope, and only crudely accounts for spatial variability in lithology.
-In fact, the 19~ppm value estimated by \citet{Molnar2015a} is closer to the decrease in \pCOtwo that we estimate if mafic volcanics associated with the Deccan Traps (a relatively flat area outside of the warm and wet tropics) are replaced with the Ca+Mg concentration of bulk continental crust (21--81~ppm; Fig.
-\ref{fig:scenario_pCO2_all}).
-The significant difference in steady-state \pCOtwo estimated between the ``removed Deccan Traps'' scenario and the SEAI scenarios (Fig.
-\ref{fig:scenario_pCO2_all}) demonstrates that considering changes in the spatial distribution of lithologies alone is not adequate for estimating changes in steady-state \pCOtwo.
-Instead, spatially-varying runoff, temperature, and slope significantly modulates silicate weathering rates, and must be accounted for.
-
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/pCO2_proxies.pdf}
+    \includegraphics[width=1\textwidth]{Figures/pCO2_proxies.pdf}
     \caption{Proxy-based \pCOtwo estimates for the past 20~m.y.
@@
 After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.
-The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (\citealp{Monnier1995a}, but see \citealp{Kadarusman2004a} for an alternative view), and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block of \citealp{Villeneuve2001a}).
+The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting \citep{Monnier1995a} (but see \citealp{Kadarusman2004a} for an alternative view) and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block; \citealp{Villeneuve2001a}).
 During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.
@@
 Locally, the structural thickness of the ophiolite exceeds 15~km with surface relief over 3~km \citep{Kadarusman2004a}.
-Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma (\citealp{Parkinson1998b} and references therein).
+Seventeen igneous K-Ar dates from the East Sulawesi ophiolite range from 93--32~Ma, clustering between 60~Ma and 40~Ma \citep{Parkinson1998b}.
 K-Ar dates on hornblende from the metamorphic sole yielded cooling ages between 36--23~Ma \citep{Parkinson1998a, Villeneuve2001a}, which dates the initial emplacement of the East Sulawesi ophiolite as the north Sulawesi volcanic arc was underthrusted by the Sula spur (Australian crust) \citep{Silver1983a, Parkinson1998a}; however, despite these relatively old ages of emplacement, it appears that the ophiolites on the East Arm were not substantially subaerially exposed and exhumed until the Miocene.
@@
 
-Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of \citet{Nichols1991a}, \citet{Cloos2005a}, \citet{Gold2017a}, and \citet{Harrington2017a}.
+Paleoshoreline maps were georeferenced from several previous tectonic and paleogeographic analyses, most notably those of \citet{Nichols1991a, Cloos2005a, Gold2017a, Harrington2017a}.
 These studies were all based to varying degrees on lithological distributions, biostratigraphic, borehole data, and tectonic models.

### 2020-04-08 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 
-\section*{Methods}
+\section*{Model and Boundary Conditions}
 
-\subsection*{GEOCLIM model}
+\subsection*{GEOCLIM}
 
@@
 
-\subsection*{Neogene cooling}
+\subsection*{SEAI Emergence}
 
@@
 
-\subsection*{Marine isotopes}
+\subsection*{Volcanic Degassing}
 
-Himalayan uplift \cite{Raymo1992a, Galy2007a} or a decrease in volcanic outgassing \cite{Berner1983a} have also been proposed as drivers for Neogene cooling.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
+    \caption{Weatherability curves for the ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
+Details on how these curves were generated are described in \MM.
+Each of the 4 curves represent a different tectonic boundary condition (i.e.
+the reconstructed paleoshorelines of the SEAI; Fig.
+\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (and vice versa).
+Point B represents the present day, in which \pCOtwo is 286~ppm.
+The arrow from Point A$_{1}$ to B represents the ``weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
+In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm).
+The arrow from Point A$_{2}$ to B instead represents the ``degassing only'' scenario, in which global weatherability remains the same as the present day, but the same change in \pCOtwo as the ``weatherability only'' scenario is achieved by decreasing the volcanic degassing flux.
+In this scenario, the volcanic degassing flux decreases from the value dictated by the present day weatherability curve ($\sim$13\% greater than the present day).
+The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value 10\% greater than the present day.}
+    \label{fig:weatherability_curves}
+\end{figure}
+
+A decrease in volcanic degassing \cite{Berner1983a} has also been proposed as a driver for Neogene cooling.
+Proxy-based estimates of the evolution of volcanic degassing fluxes throughout the Neogene are inconsistent with each other, such that not even the sign of the change in volcanic degassing over the past $\sim$15~m.y.
+is without ambiguity \cite{Godderis2017c}.
+For example, it has been estimated that the volcanic degassing flux was $\sim$25\% lower \cite{Cogne2006a} and $\sim$10\% higher \cite{Van-Der-Meer2014a} at 15~Ma relative to the present day.
+
+Nevertheless, our model framework provides us with the opportunity to estimate the magnitude of volcanic degassing flux decrease necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
+
+\subsection*{Marine Isotopes}
+
+Himalayan uplift \cite{Raymo1992a, Galy2007a} has also been proposed as drivers for Neogene cooling.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 
-\subsection*{GEOCLIM silicate weathering component}
+\subsection*{GEOCLIM Silicate Weathering Component}
 
@@
 
+\subsection*{Weatherability Curves}
+
+To create the 15~Ma ``paleo-SEAI'' curve shown in Figure \ref{fig:weatherability_curves}, we use the reconstructed paleoshorelines of the SEAI at 15~Ma (Fig.
+\ref{fig:shoreline_growth}A).
+We then select the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI).
+We then run GEOCLIM at each of these \pCOtwo levels until steady state is achieved (i.e.
+until the volcanic degassing flux is equal to the silicate weathering flux).
+We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' scenarios and the present day to generate the 3 other weatherability curves presented in Figure \ref{fig:weatherability_curves}.
+
 \subsection*{SI Appendix}

### 2020-04-08 Manuscript/Manuscript.tex

@@
 \ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (and vice versa).
-Point B represents the present day, in which \pCOtwo is 286~ppm.
+Point B represents the pre-Industrial, in which \pCOtwo is 286~ppm.
 The arrow from Point A$_{1}$ to B represents the ``weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
 In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm).
-The arrow from Point A$_{2}$ to B instead represents the ``degassing only'' scenario, in which global weatherability remains the same as the present day, but the same change in \pCOtwo as the ``weatherability only'' scenario is achieved by decreasing the volcanic degassing flux.
-In this scenario, the volcanic degassing flux decreases from the value dictated by the present day weatherability curve ($\sim$13\% greater than the present day).
-The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value 10\% greater than the present day.}
+The arrow from Point A$_{2}$ to B instead represents the ``degassing only'' scenario, in which global weatherability remains the same as the pre-Industrial, but the same change in \pCOtwo as the ``weatherability only'' scenario is achieved by decreasing the volcanic degassing flux.
+In this scenario, the volcanic degassing flux decreases from the value dictated by the pre-Industrial weatherability curve ($\sim$13\% greater than the pre-Industrial).
+The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value $\sim$4\% greater than the pre-Industrial.
+This change in the volcanic degassing flux corresponds to the \pCOtwo threshold estimated for Antarctic glaciation \cite{DeConto2008a} on a planet with our 15~Ma ``paleo-SEAI'' paleoshorelines.}
     \label{fig:weatherability_curves}
@@
 Nevertheless, our model framework provides us with the opportunity to estimate the magnitude of volcanic degassing flux decrease necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
+If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
+If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``weatherability only'' scenario in Fig.
+\ref{fig:weatherability_curves}).
+If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-Industrial to drive the same $\sim$280~ppm change in \pCOtwo (``degassing only'' scenario in Fig.
+\ref{fig:weatherability_curves}).
+
+This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}, suggesting that reasonable changes in the volcanic degassing flux alone are insufficient to drive the majority of cooling observed over the Neogene.
+However, changes in the volcanic degassing flux would have modulated changes in global weatherability associated with the emergence of the SEAI.
+In a hypothetical scenario in which \pCOtwo was at the threshold for Antarctic glaciation \cite{DeConto2008a} at 15~Ma and the SEAI are emerging, the volcanic degassing flux would only have had to have been $\sim$4\% higher relative to the pre-Industrial (``weatherability and degassing'' scenario in Fig.
+\ref{fig:weatherability_curves}).
 
@@
 until the volcanic degassing flux is equal to the silicate weathering flux).
-We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' scenarios and the present day to generate the 3 other weatherability curves presented in Figure \ref{fig:weatherability_curves}.
+We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' paleoshorelines and the present day paleoshorelines to generate the 3 other weatherability curves.
 

### 2020-04-13 Manuscript/Supplementary_Information.tex
Updates from Overleaf

@@
 data-constrained \COtwo consumption in watersheds around the world.
-\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (i.e.
+\textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 573 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (i.e.
 the orange points in Figure \ref{fig:W_vs_r2}).
-Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 5,381 parameters combinations.
-The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines.
+Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 573 parameters combinations.
+Watersheds within the compilation of \citet{Gaillardet1999a} that are in the SEAI (Fly, Kikori, Purari, and Sepik watersheds, all in New Guinea) are indicated in green.
 \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.
@@
 We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM.
-In total, we test 78,000 unique parameter combinations (Fig.
+In total, we test 93,600 unique parameter combinations (Fig.
 \ref{fig:r2_cross_plot}; Table \ref{tab:parameter_combinations}).
@@
 However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
-We therefore eliminate all parameter combinations that produce a $r^{2}\leq0.4$, which leaves 5,381 unique parameter combinations (Figs.
+We therefore eliminate all parameter combinations that produce a $r^{2}\leq0.5$, which leaves 573 unique parameter combinations (Figs.
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
 
-Global CO$_{2}$ consumption calculated from these 5,381 parameter combinations ranges from 2.9$\times10^{12}$~mol/yr to 7.7$\times10^{12}$~mol/yr, with a mean of 5.7$\times10^{12}$~mol/yr.
+Global CO$_{2}$ consumption calculated from these 573 parameter combinations ranges from 3.7$\times10^{12}$~mol/yr to 6.3$\times10^{12}$~mol/yr, with a mean of 5.2$\times10^{12}$~mol/yr.
 These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the 13 scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI).
+For each of the 13 scenarios, each point represents an estimate from one of the 573 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world.
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.

### 2020-04-13 Manuscript/Manuscript.tex

@@
 The significance statement appears in the paper itself and is required for all research papers.
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
-The significance statement appears in the paper itself and is required for all research papers.}
+\significancestatement{The Southeast Asian islands are a hotspot of \COtwo consumption via silicate weathering in the present day.
+Since $\sim$15 million years ago, these islands have been increasing in size at the same time that Earth's climate has been cooling.
+Here we test the hypothesis that global cooling over the past 15 million years has been driven by tectonic emergence of the Southeast Asian islands.
+To do so, we improve a silicate weathering model, then couple it to a climate model.
+We find that the majority of cooling over the past 15 million years can be explained by emergence of the Southeast Asian islands.}
 
@@
 \begin{abstract}
-The steep topography, tropical climate, and mafic lithologies of the Southeast Asian islands contribute to efficient chemical weathering and carbon sequestration.
-Concurrent with global cooling since the mid-Miocene, ongoing collision between the Sunda-Banda arc and Australia has increased the area of subaerially-exposed land in the region.
-Here we integrate lithology into a coupled weathering-climate model to assess how the global climate system responds to changes in paleogeography.
+Steep topography, a tropical climate, and mafic lithologies contribute to efficient chemical weathering and carbon sequestration in the Southeast Asian islands.
+Ongoing arc-continent collision between the Sunda-Banda arc and Australia has increased the area of subaerially-exposed land in the region since the mid-Miocene.
+Concurrently, Earth's climate has been cooling, marked with the onset of Northern Hemisphere glaciation 3 million years ago.
+Here we integrate lithology into a coupled weathering-climate model to assess how the global climate system responds to changes in weatherability associated with paleogeographic changes.
 We then use paleoshoreline data as boundary conditions to quantify the change in steady-state \pCOtwo associated with the tectonic emergence of the Southeast Asian islands since the mid-Miocene.
-We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain long-term climatic cooling over the Neogene.
+We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain the majority of long-term climatic cooling over the Neogene.
 \end{abstract}
@@
 \dropcap{T}he Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Macdonald2019a}.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
+There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Zachos2008a}.
-Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic outgassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1992a, Galy2007a}.
+15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
+Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic degassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a, Galy2007a}.
 Here we instead propose that SEAI emergence was itself sufficient to explain long-term climatic cooling over the Neogene.
 
-Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump1997a}.
+Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump2000a}.
 Chemical weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
@@
 Topography, climate, and lithology all effect chemical weathering.
-High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a}.
-In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{West2012a}.
+High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a, Maher2014a}.
+In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
 Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
@@
 Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
-As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
+As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca+Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
 Details of the formulation for the silicate weathering component of GEOCLIM can be found in \MM.}
@@
 The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for subaerially-exposed land.
+We assume that Ca and Mg are the only cations that consume \COtwo over geologic time-scales, such that each mole of Ca or Mg that is dissolved by silicate weathering consumes one mole of \COtwo.
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and all bedrock was assigned identical chemical compositions \cite{Godderis2017c}.
@@
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (\SI).
+After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain.
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo degassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (\SI).
 
@@
     \centering
-    \includegraphics[width=11.4cm]{Figures/shoreline_growth.pdf}
+    \includegraphics[width=0.60\linewidth]{Figures/shoreline_growth.pdf}
     \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
+For each of the seven scenarios, each point represents an estimate from one of the 573 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
@@
 
-Using the 5,381 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 378--447~ppm \pCOtwo for 5~Ma, 435--532~ppm for 10~Ma, and 493--709~ppm for 15~Ma (Fig.
+Using the 573 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 391--434~ppm \pCOtwo for 5~Ma, 457--516~ppm for 10~Ma, and 526--678~ppm for 15~Ma (Fig.
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$493--709~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
@@
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling \cite{Kent2013a}, its contribution is less than that of the wetter and higher-relief SEAI.
+The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling \cite{Kent2013a}, its contribution is less than that of the SEAI, which have high-relief and a wet tropical climate in conjunction with high Ca+Mg lithologies.
 
@@
 However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
-\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Kump1997a}.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
 
@@
 the reconstructed paleoshorelines of the SEAI; Fig.
-\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (and vice versa).
+\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (or vice versa).
 Point B represents the pre-Industrial, in which \pCOtwo is 286~ppm.
@@
 In this scenario, the volcanic degassing flux decreases from the value dictated by the pre-Industrial weatherability curve ($\sim$13\% greater than the pre-Industrial).
-The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value $\sim$4\% greater than the pre-Industrial.
-This change in the volcanic degassing flux corresponds to the \pCOtwo threshold estimated for Antarctic glaciation \cite{DeConto2008a} on a planet with our 15~Ma ``paleo-SEAI'' paleoshorelines.}
+The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value $\sim$4\% greater than the pre-Industrial.}
     \label{fig:weatherability_curves}
@@
 A decrease in volcanic degassing \cite{Berner1983a} has also been proposed as a driver for Neogene cooling.
-Proxy-based estimates of the evolution of volcanic degassing fluxes throughout the Neogene are inconsistent with each other, such that not even the sign of the change in volcanic degassing over the past $\sim$15~m.y.
+However, proxy-based estimates of the evolution of volcanic degassing fluxes throughout the Neogene are inconsistent with each other, such that not even the sign of the change in volcanic degassing over the past $\sim$15~m.y.
 is without ambiguity \cite{Godderis2017c}.
-For example, it has been estimated that the volcanic degassing flux was $\sim$25\% lower \cite{Cogne2006a} and $\sim$10\% higher \cite{Van-Der-Meer2014a} at 15~Ma relative to the present day.
+For example, it has both been estimated that the volcanic degassing flux was $\sim$25\% lower \cite{Cogne2006a} and $\sim$10\% higher \cite{Van-Der-Meer2014a} at 15~Ma relative to the present day.
 
-Nevertheless, our model framework provides us with the opportunity to estimate the magnitude of volcanic degassing flux decrease necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
+Nevertheless, our model framework provides us with the opportunity to estimate the decrease in volcanic degassing flux necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
@@
 This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}, suggesting that reasonable changes in the volcanic degassing flux alone are insufficient to drive the majority of cooling observed over the Neogene.
-However, changes in the volcanic degassing flux would have modulated changes in global weatherability associated with the emergence of the SEAI.
-In a hypothetical scenario in which \pCOtwo was at the threshold for Antarctic glaciation \cite{DeConto2008a} at 15~Ma and the SEAI are emerging, the volcanic degassing flux would only have had to have been $\sim$4\% higher relative to the pre-Industrial (``weatherability and degassing'' scenario in Fig.
+However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
+In a hypothetical scenario in which \pCOtwo was at the threshold for Antarctic glaciation \cite{DeConto2008a} at 15~Ma and the SEAI are emerging, the volcanic degassing flux would only have had to have been $\sim$4\% higher relative to the pre-Industrial to drive \pCOtwo from 750~ppm to 286~ppm (``weatherability and degassing'' scenario in Fig.
 \ref{fig:weatherability_curves}).
@@
 35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
+15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Quade1997a, Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
 Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.
@@
 Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration from the Miocene to present.
-Changes in volcanic outgassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
+Changes in volcanic degassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
 Such changes are likely to have induced complex feedbacks on many Earth surface processes.
-Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain the magnitude of cooling.
+Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain the majority of this cooling.
 These results highlight that the Earth's climate state is particularly sensitive to changes in tropical paleogeography.

### 2020-04-21 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 Steep topography, a tropical climate, and mafic lithologies contribute to efficient chemical weathering and carbon sequestration in the Southeast Asian islands.
-Ongoing arc-continent collision between the Sunda-Banda arc and Australia has increased the area of subaerially-exposed land in the region since the mid-Miocene.
-Concurrently, Earth's climate has been cooling, marked with the onset of Northern Hemisphere glaciation 3 million years ago.
-Here we integrate lithology into a coupled weathering-climate model to assess how the global climate system responds to changes in weatherability associated with paleogeographic changes.
-We then use paleoshoreline data as boundary conditions to quantify the change in steady-state \pCOtwo associated with the tectonic emergence of the Southeast Asian islands since the mid-Miocene.
-We find that the modeled decrease in \pCOtwo due to these paleogeographic changes is sufficient to explain the majority of long-term climatic cooling over the Neogene.
+Ongoing arc-continent collision between the Sunda-Banda arc system and Australia has increased the area of subaerially-exposed land in the region since the mid-Miocene.
+Concurrently, Earth's climate has cooled since the Miocene Climatic Optimum, leading to growth of the Antarctic ice sheet and onset of Northern Hemisphere glaciation.
+We seek to evaluate the hypothesis that the emergence of the Southeast Asian islands played a significant role in driving this cooling trend through increasing global weatherability.
+To do so, we have compiled paleoshoreline data and incorporated them into GEOCLIM, which couples a global climate model to a silicate weathering model with spatially resolved lithology.
+We find that without the increase in area of the Southeast Asian islands over the Neogene, atmospheric \pCOtwo would have been significantly higher than present day values, remaining above the levels necessary for initiating Northern Hemisphere ice sheets.
 \end{abstract}
@@
 The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
-There has been a significant increase of subaerially-exposed land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
+There has been a significant increase in the area of subaerially-exposed land within the region since the mid-Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
-15~Ma and accelerated over the past 4 million years (m.y.) culminating with the onset of Northern Hemisphere glaciation \cite{Shackleton1984a, Zachos2001a}.
+15~Ma and accelerated over the past 4 million years (m.y.) leading to the development of Northern Hemisphere ice sheets \cite{Shackleton1984a, Zachos2001a}.
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic degassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a, Galy2007a}.
-Here we instead propose that SEAI emergence was itself sufficient to explain long-term climatic cooling over the Neogene.
+Here we seek to evaluate the hypothesis that SEAI emergence was a significant factor in driving long-term climatic cooling over the Neogene.
 
@@
 High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a, Maher2014a}.
+The high physical erosion rates contribute to high chemical weathering fluxes in these high-relief regions \citep{Godderis2017b}.
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
-Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
-These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
+Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus have the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
+These factors have lead to the proposal that arc-continent collisions, which create steep landscapes of mafic lithologies, within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
 \cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019b}.
 
-Previous work has estimated that the decrease in \pCOtwo associated with the emergence of the SEAI since 5~Ma is $\sim$19~ppm \cite{Molnar2015a}.
-This 19~ppm estimate was obtained using a 0-D equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
-\pCOtwo was then estimated from the calculated temperature using a simple energy balance equation.
-However, the relationship between mean global temperature (or \pCOtwo) and weathering-rate-weighted land area is not linear.
-As regions with high carbon sequestration potential, such as the SEAI, emerge on Earth's surface, the way in which silicate weathering rates evolve around the globe to adjust to the higher global weatherability depends on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
-For example, if the spatial distribution of topography, climate, and lithology were such that Earth's surface overall had a relatively low carbon sequestration potential at the time of SEAI emergence, then a larger \pCOtwo change would be required to balance the sinks and sources of \COtwo than if Earth's surface overall had a relatively high carbon sequestration potential.
-However, silicate weathering within the emerging SEAI themselves would be adapting to changing \pCOtwo, ultimately resulting in a complex, non-linear, and spatially-varying system of feedbacks between climate and silicate weathering.
-Therefore, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must account for the spatial distribution of topography, climate, and lithology.
+Quantitatively estimating the magnitude of decrease in steady-state \pCOtwo associated with the emergence of a region with a high carbon sequestration potential, such as the SEAI, is not straight-forward.
+As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline.
+However, the sensitivity of the silicate weathering flux in any particular location to this change in \pCOtwo is highly variable and dependent on the specific topography, climate, and lithology at that location.
+Furthermore, how the local climate responds to this change in \pCOtwo is itself highly spatially-variable.
+Therefore, the magnitude of \pCOtwo change that is required to balance the total global silicate weathering flux with the volcanic degassing flux will depend on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
+Ultimately, there is a complex, non-linear, and spatially-varying system of feedbacks between climate and silicate weathering.
+Therefore, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must model spatially-resolved climatology and silicate weathering fluxes in tandem and account for the spatial distribution of the factors that affect these two systems.
 
-\section*{Model and Boundary Conditions}
-
-\subsection*{GEOCLIM}
+\section*{GEOCLIM Model}
 
@@
 GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in erosion, chemical weathering, and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
+
+\subsection*{Silicate Weathering Component}
 
@@
 Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
-As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca+Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
+As a particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$), with the flux of Ca+Mg being $W$ multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
 Details of the formulation for the silicate weathering component of GEOCLIM can be found in \MM.}
@@
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
+
 We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM (\SI).
@@
 Each land pixel is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees.
-The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}).
+The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data of these lithologic categories compiled in EarthChem (\url{www.earthchem.org/portal}).
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
@@
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component.
+We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component of GEOCLIM.
 
@@
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain.
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo degassing flux \cite{Gerlach2011a} as they should with the long-term carbon cycle being at steady-state (\SI).
+After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain (\SI).
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a} as they should given that the long-term carbon cycle is in steady state (\SI).
+
+\subsection*{Climate Model Component}
 
@@
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (\SI).
-The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+These experiments are useful for this analysis because all non-\COtwo forcings are held constant at values representative of pre-Industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+Furthermore, the experiments were run long enough for the final system to approximate steady-state.
 
-\subsection*{Paleoshorelines}
+\section*{Paleoshorelines}
 
-\begin{figure*}[h]
+\begin{SCfigure*}[\sidecaptionrelwidth][t]
     \centering
-    \includegraphics[width=0.60\linewidth]{Figures/shoreline_growth.pdf}
+    \includegraphics[width=13.0cm]{Figures/shoreline_growth.pdf}
     \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
@@
     \label{fig:shoreline_growth}
-\end{figure*}
+\end{SCfigure*}
 
@@
 However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma \cite{Hall2013b}.
-In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{Cloos2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
+In New Guinea, emergence in the mid-Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{Cloos2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
 Exhumation accelerated over the past 4~m.y.
@@
 
-\section*{Results and Discussion}
-
-\subsection*{SEAI Emergence}
+\section*{\pCOtwo Estimates}
 
@@
 
-Using the 573 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 391--434~ppm \pCOtwo for 5~Ma, 457--516~ppm for 10~Ma, and 526--678~ppm for 15~Ma (Fig.
+Using the 573 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 526--678~ppm for 15~Ma, 457--516~ppm for 10~Ma, and 391--434~ppm \pCOtwo for 5~Ma (Fig.
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-Industrial \pCOtwo would have been $\sim$526--678~ppm.
+
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
@@
 
-Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
-Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
-However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
-\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+Previous work has estimated that the decrease in \pCOtwo since 5~Ma associated with the emergence of the SEAI and enhanced silicate weathering is $\sim$19~ppm \cite{Molnar2015a}, in which case SEAI emergence would be a relatively minor contributor to Neogene cooling.
+This 19~ppm estimate was obtained using a 0-D equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
+\pCOtwo was then estimated from the calculated temperature using a simple energy balance equation.
+However, the relationship between mean global temperature (or \pCOtwo) and weathering-rate-weighted land area is not linear.
+Furthermore, this simple linear relationship ignores spatial variability in topography and climatology, and only crudely accounts for spatial variability in lithology.
+In fact, the 19~ppm estimate is closer in magnitude to the decrease in \pCOtwo that we estimate if mafic volcanics associated with the Deccan Traps (a relatively flat area outside of the warm and wet tropics) are replaced with the Ca+Mg concentration of bulk continental crust (22--70 ppm; Fig.
+\ref{fig:scenario_pCO2}).
+The significant difference in steady-state \pCOtwo estimated between the ``removed Deccan Traps'' scenario and the ``paleo-SEAI'' scenarios (Fig.
+\ref{fig:scenario_pCO2}) demonstrates that considering changes in the spatial distribution of lithologies alone is not adequate for estimating changes in steady-state \pCOtwo.
+Instead, spatially-varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for.
+
+\section*{Alternative Mechanisms}
 
@@
 Details on how these curves were generated are described in \MM.
-Each of the 4 curves represent a different tectonic boundary condition (i.e.
+Each of the 4 curves represent a different tectonic boundary condition and therefore a different global weatherability (i.e.
 the reconstructed paleoshorelines of the SEAI; Fig.
-\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo (or vice versa).
+\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo and global weatherability.
 Point B represents the pre-Industrial, in which \pCOtwo is 286~ppm.
-The arrow from Point A$_{1}$ to B represents the ``weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
+The arrow from Point A$_{1}$ to B represents the ``change in weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
 In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm).
-The arrow from Point A$_{2}$ to B instead represents the ``degassing only'' scenario, in which global weatherability remains the same as the pre-Industrial, but the same change in \pCOtwo as the ``weatherability only'' scenario is achieved by decreasing the volcanic degassing flux.
-In this scenario, the volcanic degassing flux decreases from the value dictated by the pre-Industrial weatherability curve ($\sim$13\% greater than the pre-Industrial).
-The arrow from Point A$_{3}$ to B represents the ``weatherability and degassing'' scenario, in which global weatherability changes as the SEAI emerge and the volcanic degassing flux decreases from a value $\sim$4\% greater than the pre-Industrial.}
+The arrow from Point A$_{2}$ to B instead represents the ``change in degassing only'' scenario, in which global weatherability remains the same as the pre-Industrial, but the same change in \pCOtwo as the ``change in weatherability only'' scenario is achieved by decreasing the volcanic degassing flux from a value $\sim$13\% greater than the pre-Industrial.
+The arrow from Point A$_{3}$ to B represents the ``change in weatherability and degassing'' scenario, in which the same change in \pCOtwo as the other two scenarios is achieved by increasing global weatherability from our 5~Ma tectonic boundary condition and decreasing the volcanic degassing flux from a value $\sim$7\% greater than the pre-Industrial.}
     \label{fig:weatherability_curves}
@@
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
-If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``weatherability only'' scenario in Fig.
+If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``change in weatherability only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
-If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-Industrial to drive the same $\sim$280~ppm change in \pCOtwo (``degassing only'' scenario in Fig.
+If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-Industrial to drive the same $\sim$280~ppm change in \pCOtwo (``change in degassing only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
@@
 However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
-In a hypothetical scenario in which \pCOtwo was at the threshold for Antarctic glaciation \cite{DeConto2008a} at 15~Ma and the SEAI are emerging, the volcanic degassing flux would only have had to have been $\sim$4\% higher relative to the pre-Industrial to drive \pCOtwo from 750~ppm to 286~ppm (``weatherability and degassing'' scenario in Fig.
+For example, let us assume that we have overestimated the increase in global weatherability associated with SEAI emergence over the past 15~m.y., and instead explore a scenario in which this increase is more accurately reflected by the difference in global weatherability between our 5~Ma and pre-Industrial tectonic boundary conditions (``change in weatherability and degassing'' scenario in Fig.
 \ref{fig:weatherability_curves}).
+In this scenario, if the volcanic degassing flux decreased from a value $\sim$7\% greater than the pre-Industrial, we can achieve the same $\sim$280~ppm change in \pCOtwo.
 
-\subsection*{Marine Isotopes}
+\subsection*{Himalayan Uplift}
 
@@
 Given that carbonate weathering is disconnected from the long-term carbon-cycle mass balance, changes in carbonate accumulation through time \cite{Si2019a} could be driven by changes in carbonate weathering.
+
+\subsection*{Ocean/Atmosphere Circulation}
+
+Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
+Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
+However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
 
@@
 until the volcanic degassing flux is equal to the silicate weathering flux).
-We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' paleoshorelines and the present day paleoshorelines to generate the 3 other weatherability curves.
+We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' paleoshorelines and the present day shorelines to generate the 3 other weatherability curves.
 
@@
 
-A detailed description of the implementation of lithology into the silicate weathering component of GEOCLIM, the calibration of the silicate weathering component of GEOCLIM, the GFDL CM2.0 climate model, and the paleoshoreline reconstruction can be found in the \SI.
+A detailed description of the implementation of lithology into the silicate weathering component of GEOCLIM, the calibration of the silicate weathering component of GEOCLIM, the GFDL CM2.0 climate model, and the paleoshoreline reconstructions can be found in the \SI.
 }

### 2020-04-21 Manuscript/Manuscript.tex
checkpoint in PNAS submission draft

@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
+\authorcontributions{Author contributions: Y.P., P.M., Y.G., and N.L.S-H.
+designed and implemented the GEOCLIM model.
+F.A.M.
+and E.A.A.
+constructed the paleoshoreline database.
+Y.P.
+and P.M.
+executed the model and analyzed the data.
+Y.P., P.M., Y.G., F.A.M., and N.L.S-H.
+wrote the manuscript.}
 \authordeclaration{The authors declare no competing interest.}
@@
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
-Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (\SI).
+Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (\SI; Table S2).
 This results in 93,600 unique combinations of these six parameters.
@@
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
-After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain (\SI).
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a} as they should given that the long-term carbon cycle is in steady state (\SI).
+After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain (\SI; Fig.
+S2).
+The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a} as they should given that the long-term carbon cycle is in steady state (\SI; Fig.
+S2).
 
@@
 \ref{fig:shoreline_growth}C).
-We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI).
+We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI; Fig.
+S5).
 
@@
 Because we use a climate model forced with modern geography, the position of the tectonic blocks remain fixed.
-Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (\SI).
+Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (\SI; Fig.
+S8).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).
@@
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
-The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI), indicating that the increase in subaerially-exposed land area and tectonic topography of the SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI; Fig.
+S9), indicating that the increase in subaerially-exposed land area and tectonic topography of the SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
@@
 Nevertheless, our model framework provides us with the opportunity to estimate the decrease in volcanic degassing flux necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
-If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
+If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
+S3), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
 If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``change in weatherability only'' scenario in Fig.
@@
 \ref{fig:shoreline_growth}A).
-We then select the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI).
+We then select the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
+S3), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI).
 We then run GEOCLIM at each of these \pCOtwo levels until steady state is achieved (i.e.

### 2020-04-30 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Author contributions: Y.P., P.M., Y.G., and N.L.S-H.
+\authorcontributions{Author contributions: Y.P., P.M., Y.G., and N.L.S.-H.
 designed and implemented the GEOCLIM model.
@@
 and P.M.
-executed the model and analyzed the data.
-Y.P., P.M., Y.G., F.A.M., and N.L.S-H.
+executed the model and analyzed the data in consultation with N.L.S.-H and Y.G..
+Y.P., P.M., Y.G., F.A.M., and N.L.S.-H.
 wrote the manuscript.}
@@
 Topography, climate, and lithology all effect chemical weathering.
-High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more sensitive to climate change \cite{Gabet2009a, West2012a, Maher2014a}.
-The high physical erosion rates contribute to high chemical weathering fluxes in these high-relief regions \citep{Godderis2017b}.
+High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more prone to adjust when climate changes \cite{Gabet2009a, West2012a, Maher2014a}.
+High physical erosion rates contribute to high chemical weathering fluxes in these high-relief regions \citep{Godderis2017b}.
 In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}.
@@
 As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline.
-However, the sensitivity of the silicate weathering flux in any particular location to this change in \pCOtwo is highly variable and dependent on the specific topography, climate, and lithology at that location.
-Furthermore, how the local climate responds to this change in \pCOtwo is itself highly spatially-variable.
+However, the sensitivity of the silicate weathering flux in any particular location to this change in \pCOtwo is variable and dependent on the specific topography, climate, and lithology at that location.
+Furthermore, how the local climate responds to this change in \pCOtwo is itself spatially-variable.
 Therefore, the magnitude of \pCOtwo change that is required to balance the total global silicate weathering flux with the volcanic degassing flux will depend on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
-Ultimately, there is a complex, non-linear, and spatially-varying system of feedbacks between climate and silicate weathering.
-Therefore, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must model spatially-resolved climatology and silicate weathering fluxes in tandem and account for the spatial distribution of the factors that affect these two systems.
+As a result, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must model spatially-resolved climatology and silicate weathering fluxes in tandem and account for the spatial distribution of the factors that affect these inter-connected systems.
 
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in the SEAI, we use the globally spatially-resolved GEOCLIM model \cite{Godderis2017c}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in the SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2017c}.
 GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in erosion, chemical weathering, and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
@@
 More recent versions of GEOCLIM implement regolith development and soil shielding (Fig.
-\ref{fig:regolith_schematic}), which introduces a dependence on erosion (and therefore topographic slope) \cite{Maffre2018a}.
+\ref{fig:regolith_schematic}), which introduces a dependence on erosion rate (and therefore topographic slope) \cite{Maffre2018a}.
 While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered.
@@
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
-Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (\SI; Table S2).
-This results in 93,600 unique combinations of these six parameters.
+Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic lithologies to vary within reasonable bounds (\SI; Table S2).
+This approach results in 93,600 unique combinations of these six parameters.
 For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
@@
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (\SI).
-These experiments are useful for this analysis because all non-\COtwo forcings are held constant at values representative of pre-Industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+These experiments are well suited for this analysis because all non-\COtwo forcings are held constant at values representative of pre-Industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
 Furthermore, the experiments were run long enough for the final system to approximate steady-state.
@@
     \includegraphics[width=13.0cm]{Figures/shoreline_growth.pdf}
-    \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in climate science literature) from the mid-Miocene to present.
-Past shorelines are shown in \textit{A} with associated land area summarized in \textit{B}.
+    \caption{The emergence of the Southeast Asian islands (also referred to as the Maritime Continent in the climate science literature) from the mid-Miocene to present.
+Past shorelines at 5, 10, and 15 Ma are shown in \textit{A} with associated land area summarized in \textit{B}.
 A significant increase in area over the past 5 million years is coincident with cooling and the onset of Northern Hemisphere glaciation as reflected in the benthic oxygen isotope record \cite{Zachos2008a} shown in \textit{C}.}
@@
 
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/scenario_pCO2.pdf}
+    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
+For each of the seven scenarios, each point represents an estimate from one of the 573 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
+The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
+the interquartile range), and the notch represents the median with its 95\% confidence interval.
+The whiskers extend to the 2.5 and 97.5 percentile values.
+Glaciation thresholds \cite{DeConto2008a} are shown on the x-axis.}
+    \label{fig:scenario_pCO2}
+\end{figure*}
+
 To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
@@
 The paleoshoreline data indicate that the Sunda-Banda Arc and New Guinea are primarily responsible for the increase in area since 15~Ma.
-Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \cite{Harris2006a}.
+Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the Australian Plate \cite{Harris2006a}.
 Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma.
@@
 
-\begin{figure*}[h]
-    \centering
-    \includegraphics[width=1\linewidth]{Figures/scenario_pCO2.pdf}
-    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 573 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
-The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
-the interquartile range), and the notch represents the median with its 95\% confidence interval.
-The whiskers extend to the 2.5 and 97.5 percentile values.
-Glaciation thresholds \cite{DeConto2008a} are shown on the x-axis.}
-    \label{fig:scenario_pCO2}
-\end{figure*}
-
 Using the 573 unique parameter combinations, the ``paleo-SEAI'' scenarios resulted in 526--678~ppm for 15~Ma, 457--516~ppm for 10~Ma, and 391--434~ppm \pCOtwo for 5~Ma (Fig.
@@
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.
+These modeled values of decreasing \pCOtwo associated with SEAI emergence are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.
 
@@
 \ref{fig:scenario_pCO2}) demonstrates that considering changes in the spatial distribution of lithologies alone is not adequate for estimating changes in steady-state \pCOtwo.
-Instead, spatially-varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for.
+Instead, spatially-varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for when estimating \pCOtwo change associated with paleogeographic change.
 
-\section*{Alternative Mechanisms}
+\section*{Alternative Mechanisms for Neogene Cooling}
 
@@
     \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
-    \caption{Weatherability curves for the ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
+    \caption{Weatherability curves for the modern and ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
 Details on how these curves were generated are described in \MM.
-Each of the 4 curves represent a different tectonic boundary condition and therefore a different global weatherability (i.e.
+Each of the 4 curves represent a different tectonic boundary condition (i.e.
 the reconstructed paleoshorelines of the SEAI; Fig.
-\ref{fig:shoreline_growth}A), and shows the volcanic degassing flux necessary to balance the silicate weathering flux at a given \pCOtwo and global weatherability.
+\ref{fig:shoreline_growth}A) and therefore a different global weatherability.
+The curves show the resulting \pCOtwo for a given volcanic degassing flux such that the input flux is balanced by the silicate weathering output flux.
 Point B represents the pre-Industrial, in which \pCOtwo is 286~ppm.
@@
 
-Nevertheless, our model framework provides us with the opportunity to estimate the decrease in volcanic degassing flux necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
+Our model framework provides an opportunity to estimate the decrease in volcanic degassing flux necessary to achieve the same change in \pCOtwo predicted for the increase in global weatherability associated with the emergence of the SEAI over the past 15~m.y.
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
-S3), GEOCLIM estimates a volcanic degassing flux for the pre-Industrial of 4.1$\times$10$^{12}$~mol/yr.
+S3), GEOCLIM estimates a pre-Industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
 If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``change in weatherability only'' scenario in Fig.
@@
 However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
-For example, let us assume that we have overestimated the increase in global weatherability associated with SEAI emergence over the past 15~m.y., and instead explore a scenario in which this increase is more accurately reflected by the difference in global weatherability between our 5~Ma and pre-Industrial tectonic boundary conditions (``change in weatherability and degassing'' scenario in Fig.
+For example, take a scenario where the increase in global weatherability associated with SEAI emergence over the past 15~m.y.
+is an overestimate and the increase is more accurately reflected by the difference in global weatherability between our 5~Ma and modern tectonic boundary conditions (``change in weatherability and degassing'' scenario in Fig.
 \ref{fig:weatherability_curves}).
-In this scenario, if the volcanic degassing flux decreased from a value $\sim$7\% greater than the pre-Industrial, we can achieve the same $\sim$280~ppm change in \pCOtwo.
+In this scenario, if the volcanic degassing flux decreased from a value $\sim$7\% greater than the pre-Industrial in conjunction with the smaller change in weatherability, we can achieve the same $\sim$280~ppm change in \pCOtwo.
 
@@
 
-Himalayan uplift \cite{Raymo1992a, Galy2007a} has also been proposed as drivers for Neogene cooling.
+Himalayan uplift \cite{Raymo1992a, Galy2007a} has also been proposed as a driver for Neogene cooling.
 Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
@@
 
-Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration from the Miocene to present.
+Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration due to silicate weathering from the Miocene to present.
 Changes in volcanic degassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
 Such changes are likely to have induced complex feedbacks on many Earth surface processes.
-Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain the majority of this cooling.
+Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain much of this cooling.
 These results highlight that the Earth's climate state is particularly sensitive to changes in tropical paleogeography.
@@
 We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' paleoshorelines and the present day shorelines to generate the 3 other weatherability curves.
+Each estimated \pCOtwo in Figure \ref{fig:scenario_pCO2} is the result of underlying weatherability curves that subtly change with the different chemical weathering parameters.
 
@@
 
-\acknow{Collaborative research between NLS-H and YG was initially supported by a grant from the France-Berkeley Fund.
-FAM and NLS-H gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
+\acknow{Collaborative research between N.L.S.-H.
+and Y.G.
+was initially supported by a grant from the France-Berkeley Fund.
+N.L.S.-H.
+and F.A.M.
+gratefully acknowledge support through NSF FRES grants \#1926001 and \#1925990.
 We thank Alec Brenner, Sam Lo Bianco, Mariana Lin, and Judy Pu for their data compilation contributions to the paleoshoreline reconstructions.}

### 2020-04-30 Manuscript/Supplementary_Information.tex
checkpoint in PNAS submission draft

@@
 To calculate this flux, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig.
-1).
+1 in the main text).
 Previous implementations of GEOCLIM have used a ``diffuse lithology'' where all exposed land is assigned the composition of bulk upper continental crust.
@@
 To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel covered by each of the 6 lithologic categories (or water/ice) is captured by the extra dimension.
-In this way, we calculate an area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
+In this way, we calculate an area-weighted mean Ca+Mg concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.
 
@@
 
-\begin{equation}
+\begin{equation*}
     r^{2} = 1 - \frac{\sum\left[ \log_{10}(M_{i}) - \log_{10}(O_{i}) \right]^{2}}{\sum\left[ \log_{10}(O_{i}) - \overline{\log_{10}(O)} \right]^{2}}
     \label{eq:10}
-\end{equation}
+\end{equation*}
 
@@
 
+\section*{Seawater Model}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{Figures/seawater_model.pdf}
+    \caption{xxx}
+    \label{fig:seawater_model}
+\end{figure}
+
+\begin{table}[h!]
+\caption{Geochemical properties of sources used in the seawater model presented in Figure \ref{fig:seawater_model}.} 
+\begin{center}
+    \begin{tabular}{lcccc}
+    \hline
+     & & \textbf{Mg} & \textbf{Ca} & \textbf{Sr} \\
+    \textbf{source} & \textbf{\SrSr} & (ppm) & (ppm) & (ppm) \\
+    \hline
+    non-Himalayan radiogenic subaerial weathering & calculated$^{1}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
+    Himalayan radiogenic subaerial weathering & varied$^{3}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
+    non-Himalayan juvenile subaerial weathering & 0.7035$^{4}$ & 45500$^{5}$ & 71600$^{5}$ & 465$^{5}$ \\
+    non-Himalayan carbonate subaerial weathering & 0.7080$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+    Himalayan carbonate subaerial weathering & varied$^{3}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+    hydrothermal exchange & 0.7030$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+    \hline
+    \end{tabular}
+
+    \scriptsize
+	\flushleft \emph{Notes:} \\
+    $^{1}$ Calculated to account for $^{87}$Rb decay following the equations of \citet{Papanastassiou1968a}.\\
+    $^{2}$ From \citet{Wedepohl1995a}\\
+    $^{3}$ Varied as shown in Figure \ref{fig:seawater_model}.\\
+    $^{4}$ From \citet{Allegre2010a}.\\
+    $^{5}$ Taking the mean of \citet{Turekian1961a} and \citet{Taylor1964a}.\\
+    $^{6}$ Not required for our model, since the Sr fluxes from these sources are set directly in our model and they do not contribute to \COtwo sequestration.\\
+\label{tab:seawater_model}
+\end{center}
+\end{table}
+
+If the weathering of juvenile lithologies of the SEAI played an important role in driving Neogene cooling, the intuitive expectation is that marine \SrSr should have decreased over the Neogene as the global weathering flux of low \SrSr juvenile lithologies increased (compensated by a decrease in the global weathering flux of high \SrSr radiogenic lithologies to maintain constant global \COtwo consumption).
+Contrary to this expectation, marine \SrSr has been rising since ca.
+35~Ma (Fig.
+\ref{fig:seawater_model}).
+However, the global weathering flux of isotopically-representative lithologies can be decoupled from marine \SrSr via the regional weathering of isotopically-unique source rocks.
+For example, the decrease in slope in the marine \SrSr record ca.
+15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \citep{Myrow2015a, Colleps2018a}.
+This change in slope could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
+
+To test whether the Neogene marine \SrSr record is compatible with an increase in the global weathering flux of low \SrSr juvenile lithologies over the Neogene, we implement a simple seawater model that simulates changes in marine \SrSr that result from changing fluxes from several sources \citep{Brass1975a}:
+
+\begin{equation*}
+    \frac{dR_{SW}}{dt} = \frac{1}{\tau} \sum_{i} \frac{F_{i}}{F_{tot}}(R_{i} - R_{SW})
+\end{equation*}
+
+$R_{SW}$ is the \SrSr composition of the ocean, $t$ is time in the model, $\tau$ is the residence time of strontium in the oceans, $F_{i}$ is the flux of strontium from source $i$, $F_{tot}$ is the total flux of strontium from all sources, and $R_{i}$ is the \SrSr composition of source $i$.
+We define 6 simplified sources of strontium to the oceans (Table \ref{tab:seawater_model}): the non-Himalayan subaerial weathering of radiogenic (older cratonic rocks such as gneiss and granite), juvenile (such as basalt), and carbonate lithologies, the Himalayan subaerial weathering of radiogenic and carbonate lithologies, and the weathering of ocean crust during hydrothermal circulation on or near mid-ocean ridges.
+
+At initialization, we set the Sr flux from each of these sources such that at 50~Ma marine \SrSr is $\sim$0.7078 (Fig.
+\ref{fig:seawater_model}).
+As a percentage of the total Sr input into the ocean, this model yields hydrothermal = 10\%, carbonate = 35\%, radiogenic lithologies = 13\%, and juvenile lithologies = 42\%.
+These ini- tial steady-state Sr fluxes are similar to those estimated for the present-day (hydrothermal = 10\%, carbonate = 35\%, radiogenic lithologies = 25\%, and juvenile lithologies = 30\%, from \citealt{Allegre2010a}).
+The residence time of strontium in the oceans is poorly constrained, primarily due to the lack of constraints on the input flux of strontium from hydrothermal systems.
+Recent estimates for this residence time range from $\sim$1.7--3.7~m.y.
+\citep{Krabbenhoft2010a}, so we use a value that falls within this range (2.4~m.y.; \citealp{Ravizza2003a}).
+
+The \SrSr of Himalayan lithologies is highly variable, with values ranging from $\sim$0.71--1.11 \citep{Oliver2003a}, and as orogenesis proceeded, the \SrSr of waters draining the Himalayas has changed as different tectonic units are exhumed, as evidenced by the \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal (Fig.
+\ref{fig:seawater_model}; \citealp{Derry1996a, Quade1997a}).
+In our model, we make the assumption that the \SrSr of waters draining the Himalayas has been constant at 0.720 prior to 16~Ma, when the Outer Lesser Himalaya began to be exhumed \citep{Myrow2015a, Colleps2018a}.
+We also assume that the ratio of Sr coming from carbonates vs.
+radiogenic lithologies has remained constant throughout the past 50 m.y.
+(i.e.
+70\% from carbonates and 30\% from radiogenic lithologies; \citealp{Oliver2003a}).
+Finally, we assume that the average \SrSr composition of Himalayan carbonates and radiogenic lithologies that are being weathered at any given time is identical.
+Although this final assumption is likely incorrect in detail, it has been shown that Sr in carbonates from the Lesser and Higher Himalaya have equilibrated with Sr in coexisting silicates \citep{Oliver2003a}, which at least partially justifies our simplifying assumption.
+
+We maintain initial conditions until 35~Ma, at which point we begin to change both global and Himalayan subaerial weathering fluxes (Fig.
+\ref{fig:seawater_model}).
+The two primary watersheds draining the Himalayas are the Indus and Ganges-Brahmaputra basins.
+Together, the Sr flux from these two watersheds in the present day is $\sim$9.5$\times$10$^{8}$~mol/yr.
+We therefore linearly increase the total Sr flux coming from the Himalayas to 1$\times$10$^{9}$~mol/yr from 35~Ma to present.
+We also linearly increase the Sr flux coming from non-Himalayan radiogenic lithologies 35--16~Ma.
+Importantly, we also linearly decrease the Sr flux coming from non-Himalayan juvenile lithologies over this same time period in order to maintain constant \COtwo consumption (i.e.
+constant global Ca+Mg flux from subaerial weathering).
+In order to match the observed increase in marine \SrSr ca.
+35--16~Ma, the Sr flux coming from non-Himalayan juvenile lithologies needs to be increased to a value 1.42$\times$ the initial value.
+
 \clearpage

### 2020-04-30 Manuscript/Supplementary_Information.tex
Updates from Overleaf

@@
 For example, the decrease in slope in the marine \SrSr record ca.
-15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \citep{Myrow2015a, Colleps2018a}.
+15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Outer Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}.
 This change in slope could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
@@
 
-The \SrSr of Himalayan lithologies is highly variable, with values ranging from $\sim$0.71--1.11 \citep{Oliver2003a}, and as orogenesis proceeded, the \SrSr of waters draining the Himalayas has changed as different tectonic units are exhumed, as evidenced by the \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal (Fig.
+The \SrSr of Himalayan lithologies is highly variable, with values ranging from $\sim$0.71--1.11 \citep{Oliver2003a}, and as orogenesis proceeded, the \SrSr of waters draining the Himalayas has changed as different tectonic units are exhumed, as evidenced by the \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal, which are interpreted to reflect the \SrSr of Himalayan rivers (Fig.
 \ref{fig:seawater_model}; \citealp{Derry1996a, Quade1997a}).
@@
 Importantly, we also linearly decrease the Sr flux coming from non-Himalayan juvenile lithologies over this same time period in order to maintain constant \COtwo consumption (i.e.
-constant global Ca+Mg flux from subaerial weathering).
+constant global Ca+Mg flux from subaerial silicate weathering).
 In order to match the observed increase in marine \SrSr ca.
 35--16~Ma, the Sr flux coming from non-Himalayan juvenile lithologies needs to be increased to a value 1.42$\times$ the initial value.
+
+At 16~Ma, we begin to change the subaerial weathering fluxes in accordance with existing constraints (Fig.
+\ref{fig:seawater_model}).
+First, the total Sr flux from Himalayan lithologies continues to linearly increase.
+Second, we linearly decrease the \SrSr of Himalayan lithologies from 0.720 to 0.712 16--12~Ma, then linearly increase it from 0.712 to 0.750 12--0~Ma.
+The former \SrSr decrease is intended to represent the exhumation of low \SrSr Outer Lesser Himalaya lithologies, and the latter \SrSr increase is intended to represent the exhumation of higher \SrSr Inner Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}.
+These Himalayan \SrSr changes are consistent with the scattered \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal \citep{Derry1996a, Quade1997a}.
+We do not model the decrease in Himalayan river \SrSr \textless5~Ma, since its timing and magnitude is poorly constrained.
 

### 2020-05-01 Manuscript/Supplementary_Information.tex

@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/lithology_CO2_consumption.pdf}
+    \caption{xxx}
+    \label{fig:lithology_CO2_consumption}
+\end{figure}
+
 \begin{table}[h!]
@@
 First, the total Sr flux from Himalayan lithologies continues to linearly increase.
-Second, we linearly decrease the \SrSr of Himalayan lithologies from 0.720 to 0.712 16--12~Ma, then linearly increase it from 0.712 to 0.750 12--0~Ma.
-The former \SrSr decrease is intended to represent the exhumation of low \SrSr Outer Lesser Himalaya lithologies, and the latter \SrSr increase is intended to represent the exhumation of higher \SrSr Inner Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}.
+Second, we linearly decrease the \SrSr of Himalayan lithologies from 0.720 to 0.715 16--12~Ma, then linearly increase it from 0.715 to 0.740 12--3~Ma, then linearly decrease it from 0.740 to 0.728 3--0~Ma.
+The first \SrSr decrease is intended to represent the exhumation of low \SrSr Outer Lesser Himalaya lithologies, the second \SrSr increase is intended to represent the exhumation of higher \SrSr Inner Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}, and the third \SrSr decrease is intended to represent a Pliocene-Quaternary fall in Himalayan riverine \SrSr to present-day values.
 These Himalayan \SrSr changes are consistent with the scattered \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal \citep{Derry1996a, Quade1997a}.
-We do not model the decrease in Himalayan river \SrSr \textless5~Ma, since its timing and magnitude is poorly constrained.
+
+In addition to these Himalayan changes, we also linearly decrease the Ca+Mg flux coming from non-Himalayan radiogenic lithologies to a value 0.96$\times$ the 16~Ma value.
+This 0.96 value was obtained from our GEOCLIM experiments, in which
 

### 2020-05-13 Manuscript/Manuscript.tex
PNAS submission checkpoint

@@
 Quantitatively estimating the magnitude of decrease in steady-state \pCOtwo associated with the emergence of a region with a high carbon sequestration potential, such as the SEAI, is not straight-forward.
-As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline.
+As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline until a new steady-state is established where the total magnitude of the \COtwo sinks is the same as before the change.
 However, the sensitivity of the silicate weathering flux in any particular location to this change in \pCOtwo is variable and dependent on the specific topography, climate, and lithology at that location.
-Furthermore, how the local climate responds to this change in \pCOtwo is itself spatially-variable.
+Furthermore, how regional climate responds to this change in \pCOtwo is itself spatially-variable.
 Therefore, the magnitude of \pCOtwo change that is required to balance the total global silicate weathering flux with the volcanic degassing flux will depend on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
@@
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (\SI).
-These experiments are well suited for this analysis because all non-\COtwo forcings are held constant at values representative of pre-Industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
+These experiments are well-suited for this analysis because all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.
 Furthermore, the experiments were run long enough for the final system to approximate steady-state.
@@
 \ref{fig:scenario_pCO2}).
-These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-Industrial \pCOtwo would have been $\sim$526--678~ppm.
+These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
 
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
+The \pCOtwo values from the 5~Ma paleo-SEAI scenario overlap with many proxy-based estimates as well as values that emerge from hybrid climate model data assimilation approaches \cite{Berends2020a}.
 The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI; Fig.
@@
 The curves show the resulting \pCOtwo for a given volcanic degassing flux such that the input flux is balanced by the silicate weathering output flux.
-Point B represents the pre-Industrial, in which \pCOtwo is 286~ppm.
+Point B represents the pre-industrial, in which \pCOtwo is 286~ppm.
 The arrow from Point A$_{1}$ to B represents the ``change in weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
 In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm).
-The arrow from Point A$_{2}$ to B instead represents the ``change in degassing only'' scenario, in which global weatherability remains the same as the pre-Industrial, but the same change in \pCOtwo as the ``change in weatherability only'' scenario is achieved by decreasing the volcanic degassing flux from a value $\sim$13\% greater than the pre-Industrial.
-The arrow from Point A$_{3}$ to B represents the ``change in weatherability and degassing'' scenario, in which the same change in \pCOtwo as the other two scenarios is achieved by increasing global weatherability from our 5~Ma tectonic boundary condition and decreasing the volcanic degassing flux from a value $\sim$7\% greater than the pre-Industrial.}
+The arrow from Point A$_{2}$ to B instead represents the ``change in degassing only'' scenario, in which global weatherability remains the same as the pre-industrial, but the same change in \pCOtwo as the ``change in weatherability only'' scenario is achieved by decreasing the volcanic degassing flux from a value $\sim$13\% greater than the pre-industrial.
+The arrow from Point A$_{3}$ to B represents the ``change in weatherability and degassing'' scenario, in which the same change in \pCOtwo as the other two scenarios is achieved by increasing global weatherability from our 5~Ma tectonic boundary condition and decreasing the volcanic degassing flux from a value $\sim$7\% greater than the pre-industrial.}
     \label{fig:weatherability_curves}
@@
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
-S3), GEOCLIM estimates a pre-Industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
+S3), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
 If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``change in weatherability only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
-If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-Industrial to drive the same $\sim$280~ppm change in \pCOtwo (``change in degassing only'' scenario in Fig.
+If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-industrial to drive the same $\sim$280~ppm change in \pCOtwo (``change in degassing only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
 
-This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}, suggesting that reasonable changes in the volcanic degassing flux alone are insufficient to drive the majority of cooling observed over the Neogene.
+This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}, suggesting that changes in the volcanic degassing flux alone are insufficient to drive the majority of cooling observed over the Neogene.
 However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
@@
 \ref{fig:weatherability_curves}).
-In this scenario, if the volcanic degassing flux decreased from a value $\sim$7\% greater than the pre-Industrial in conjunction with the smaller change in weatherability, we can achieve the same $\sim$280~ppm change in \pCOtwo.
+In this scenario, a decrease in the volcanic degassing flux from a value $\sim$7\% greater than the pre-industrial in conjunction with the smaller change in weatherability can achieve the same $\sim$280~ppm decrease in \pCOtwo.
 
@@
 
-Himalayan uplift \cite{Raymo1992a, Galy2007a} has also been proposed as a driver for Neogene cooling.
-Rising marine \SrSr has been associated with increasing weathering of the high \SrSr Himalayan lithologies \cite{Raymo1992a}.
-This increase in \SrSr since ca.
-35~Ma is followed by a decrease in slope ca.
-15~Ma that has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \cite{Quade1997a, Colleps2018a}, but could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
-Enhanced organic matter burial in the Bay of Bengal due to high sedimentation rates could also have contributed to Neogene cooling \cite{Galy2007a}.
+Marine \SrSr has overall been increasing since ca.
+35~Ma \cite{McArthur2012a}.
+The traditional explanation for this trend is that it reflects increased weathering of radiogenic (i.e.
+high \SrSr) silicate rocks \cite{Raymo1988a}.
+Associated with this explanation is the proposal that increasing weathering of radiogenic silicate rocks in the Himalayas was the primary driver of Neogene cooling \cite{Raymo1992a}.
+It could be argued that increasing marine \SrSr is inconsistent with the hypothesis that increasing weathering of juvenile (i.e.
+low \SrSr) silicate rocks in the SEAI was an important driver of Neogene cooling.
+However, the ratio of silicate weathering fluxes from radiogenic cratonic rocks vs.
+juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically-unique carbonates.
+For example, highly radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
+As such, there are challenges in interpreting the marine \SrSr record as a proxy for silicate weathering fluxes in the Cenozoic, and potentially deeper in the geologic record as well.
+Nevertheless, steadily increasing marine \SrSr is interrupted ca.
+16~Ma when the slope of the \SrSr curve decreases \cite{McArthur2012a}.
+This decrease in slope has been attributed to coincident exhumation of relatively low \SrSr Outer Lesser Himalaya carbonates \cite{Myrow2015a, Colleps2018a}, but could also be at least partially driven by the emergence of low \SrSr lithologies in the SEAI during arc-continent collision.
 
-It has been argued that enhanced weathering in regions of active uplift predicts an increase in global alkalinity fluxes.
-However, global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
+Himalayan uplift would have affected geological carbon sinks, either via increased weathering of silicate rocks \cite{Raymo1992a} or enhanced burial of organic matter in the Bengal Fan \cite{Galy2007a}.
+Increased weathering of the SEAI as they emerged would have occurred in tandem with such changes in the Himalaya.
+Given the large uncertainty associated with changes in regional climatology across Asia due to Himalayan orogeny, developing quantitative estimates of the evolution of global silicate weathering fluxes associated with this orogeny remains a major challenge.
+
+Global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
 Enhanced weathering in a region such as the SEAI is compensated by a decrease in weathering elsewhere.
@@
 Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration due to silicate weathering from the Miocene to present.
-Changes in volcanic degassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also effected geological carbon sources and sinks.
-Such changes are likely to have induced complex feedbacks on many Earth surface processes.
+Changes in volcanic degassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also affected geological carbon sources and sinks.
 Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain much of this cooling.

### 2020-05-13 Manuscript/Supplementary_Information.tex
PNAS submission checkpoint

--- 2020-04-30 Manuscript/Supplementary_Information.tex
@@
     \includegraphics[width=1\textwidth]{Figures/world_lithology.jpg}
-    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM.}
+    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM modified from \cite{Hartmann2012a}.}
     \label{fig:world_lithology}
@@
 Within the equations that govern the chemical weathering rate in GEOCLIM, we identify the less constrained parameters: the proportionality constant that modifies the dependence of dissolution rate on runoff and temperature ($k_{d}$; Equation 9), the proportionality constant that modifies the dependence of dissolution rate on runoff only ($k_{w}$; Equation 9), the power constant that modifies the dependence of dissolution rate on the time that a rock particle has spent in the regolith ($\sigma$; Equation 4), and the proportionality constant that modifies the dependence of regolith production on runoff and temperature ($k_{rp}$; Equation 6).
-Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are also difficult to define.
-We therefore allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM.
+Furthermore, the Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are difficult to define.
+We allow these parameters to vary within reasonable bounds during the calibration stage of GEOCLIM.
 In total, we test 93,600 unique parameter combinations (Fig.
@@
 Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields.
-As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a}, \citet{Moquet2016a}, and \citet{Moquet2018a}.
+As described in the main text, we sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a}, \citet{Moquet2016a}, and \citet{Moquet2018a}.
 The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes.
@@
 However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well.
-We therefore eliminate all parameter combinations that produce a $r^{2}\leq0.5$, which leaves 573 unique parameter combinations (Figs.
+We eliminate all parameter combinations that produce a $r^{2}\leq0.5$, which leaves 573 unique parameter combinations (Figs.
 \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).
@@
 Global CO$_{2}$ consumption calculated from these 573 parameter combinations ranges from 3.7$\times10^{12}$~mol/yr to 6.3$\times10^{12}$~mol/yr, with a mean of 5.2$\times10^{12}$~mol/yr.
-These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
+These estimates fall within the range of independently estimated outgassing rates for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig.
 \ref{fig:W_vs_r2}B).
@@
 Hence, in the non-calibration GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing.
-In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
+In these estimates, we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}.
 Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.
@@
 \pCOtwo is then held constant for 160 years.
-We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.
+We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm \pCOtwo respectively.
 Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.
 
-Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
+Given that precipitation in the SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
@@
 However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop; \citealp{Pollard2005b}), but the amount of emergent land would be lower.
-We expect this bias to be minimal, as all of our \pCOtwo estimates fall below the glaciation threshold for Antarctica.
+We expect this bias to be minimal, as all of our modeled \pCOtwo estimates fall below the glaciation threshold for Antarctica.
 
@@
     \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
-    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
+    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
     \label{fig:paleogeographic_reconstructions}
@@
 Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \citep{Darmadi2007a, Hall2009a, Hall2013b}.
-However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \citep{Darmadi2007a}.
+However, basement highs of granite do not appear further to the northwest in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the Sunda Shelf was exposed \citep{Darmadi2007a}.
 We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment.
@@
 For other minor islands in the region that were exposed 15, 10, or 5~m.y.
-ago but are currently submerged, we take the same approach as that described for the Sunda Shelf to generate the runoff field.
+ago, but are currently submerged, we take the same approach as that described for the Sunda Shelf to generate the runoff field.
 For the slope and lithology, we take the mean slope and Ca+Mg concentrations from neighbouring pixels.
@@
     \caption{Proxy-based \pCOtwo estimates for the past 20~m.y.
-For data that was compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties.
+For data that were compiled within \citet{Foster2017a}, error bars indicate standardized uncertainties.
 For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
-The \citet{Bereiter2015a} data comes from the EPICA Dome C ice core.}
+The \citet{Bereiter2015a} data come from the EPICA Dome C ice core.}
     \label{fig:pCO2_proxies}
@@
 The paleoshoreline reconstructions within the paleogeographic models broadly coincide with sub-Epoch boundaries (i.e.
-Early Miocene = 23--16~Ma, Middle Miocene = 15.99--11.61~Ma, Late Miocene = 11.6--5.4~Ma, Pliocene =5.39--3.8 Ma) but we recognize that the biostratigraphic resolution is an additional source of uncertainty.
+Early Miocene = 23--16~Ma, Middle Miocene = 15.99--11.61~Ma, Late Miocene = 11.6--5.4~Ma, Pliocene =5.39--3.8 Ma), but we recognize that the biostratigraphic resolution is an additional source of uncertainty.
 
 Several new sources for paleoshoreline data have become available since \citet{Molnar2015a}.
-Particularly, for Sulawesi, we followed the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we followed \citet{Gold2017a} and \citet{Harrington2017a}.
+Particularly, for Sulawesi, we follow the recent stratigraphic compilation and paleoshorelines delineated by \citet{Nugraha2018a}, and in New Guinea, we follow \citet{Gold2017a} and \citet{Harrington2017a}.
 Although the outlines of our paleoshorelines are significantly different from those used for 5~Ma by \citet{Molnar2015a} our calculated area is comparable.
@@
 These arms have high relief ($>$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}.
-Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
+Like SW Borneo, the North and South Arms, and much of Central Sulawesi, are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}.
 After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.
-The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting \citep{Monnier1995a} (but see \citealp{Kadarusman2004a} for an alternative view) and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block; \citealp{Villeneuve2001a}).
+The ophiolites of Sulawesi were generated in this arc system in a back-arc to intra-arc setting \citep{Monnier1995a} (but see \citealp{Kadarusman2004a} for an alternative view) and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block; \citealp{Villeneuve2001a}).
 During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.
@@
 
-The East and Southeast Arms consist predominantly of mafic and ultramafic rock of the East Sulawesi Ophoilite that is exposed for more than 10,000~km$^{2}$ \citep{Monnier1995a}.
+The East and Southeast Arms consist predominantly of mafic and ultramafic rocks of the East Sulawesi Ophoilite that are exposed for more than 10,000~km$^{2}$ \citep{Monnier1995a}.
 The East Arm preserves a complete ophiolitic sequence underlain by a metamorphic sole, m\'elange, imbricate continental margin and crystalline basement with a blueschist metamorphic overprint \citep{Silver1983a, Monnier1995a, Parkinson1998a}.
@@
 Alkali volcanism ensued at ca.
-11~Ma and is associated with a second phase of extension and exhumation.
-From the Late Miocene to Pliocene fission track ages imply deep exhumation \citep{Smith1991a, Bergman1996a, vanLeeuwen2010a}.
+11~Ma and is associated with a second phase of extension and exhumation from the Late Miocene to Pliocene with fission track ages implying deep exhumation \citep{Smith1991a, Bergman1996a, vanLeeuwen2010a}.
 
@@
 
-\section*{Seawater Model}
+% \section*{Seawater Model}
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1\textwidth]{Figures/seawater_model.pdf}
-    \caption{xxx}
-    \label{fig:seawater_model}
-\end{figure}
-
-\begin{table}[h!]
-\caption{Geochemical properties of sources used in the seawater model presented in Figure \ref{fig:seawater_model}.} 
-\begin{center}
-    \begin{tabular}{lcccc}
-    \hline
-     & & \textbf{Mg} & \textbf{Ca} & \textbf{Sr} \\
-    \textbf{source} & \textbf{\SrSr} & (ppm) & (ppm) & (ppm) \\
-    \hline
-    non-Himalayan radiogenic subaerial weathering & calculated$^{1}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
-    Himalayan radiogenic subaerial weathering & varied$^{3}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
-    non-Himalayan juvenile subaerial weathering & 0.7035$^{4}$ & 45500$^{5}$ & 71600$^{5}$ & 465$^{5}$ \\
-    non-Himalayan carbonate subaerial weathering & 0.7080$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
-    Himalayan carbonate subaerial weathering & varied$^{3}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
-    hydrothermal exchange & 0.7030$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
-    \hline
-    \end{tabular}
-
-    \scriptsize
-	\flushleft \emph{Notes:} \\
-    $^{1}$ Calculated to account for $^{87}$Rb decay following the equations of \citet{Papanastassiou1968a}.\\
-    $^{2}$ From \citet{Wedepohl1995a}\\
-    $^{3}$ Varied as shown in Figure \ref{fig:seawater_model}.\\
-    $^{4}$ From \citet{Allegre2010a}.\\
-    $^{5}$ Taking the mean of \citet{Turekian1961a} and \citet{Taylor1964a}.\\
-    $^{6}$ Not required for our model, since the Sr fluxes from these sources are set directly in our model and they do not contribute to \COtwo sequestration.\\
-\label{tab:seawater_model}
-\end{center}
-\end{table}
-
-If the weathering of juvenile lithologies of the SEAI played an important role in driving Neogene cooling, the intuitive expectation is that marine \SrSr should have decreased over the Neogene as the global weathering flux of low \SrSr juvenile lithologies increased (compensated by a decrease in the global weathering flux of high \SrSr radiogenic lithologies to maintain constant global \COtwo consumption).
+% If the weathering of juvenile lithologies of the SEAI played an important role in driving Neogene cooling, the intuitive expectation is that marine \SrSr should have decreased over the Neogene as the global weathering flux of low \SrSr juvenile lithologies increased (compensated by a decrease in the global weathering flux of high \SrSr radiogenic lithologies to maintain constant global \COtwo consumption).
 Contrary to this expectation, marine \SrSr has been rising since ca.
@@
 For example, the decrease in slope in the marine \SrSr record ca.
-15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Lesser Himalayan lithologies \citep{Myrow2015a, Colleps2018a}.
+15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Outer Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}.
 This change in slope could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
 
-To test whether the Neogene marine \SrSr record is compatible with an increase in the global weathering flux of low \SrSr juvenile lithologies over the Neogene, we implement a simple seawater model that simulates changes in marine \SrSr that result from changing fluxes from several sources \citep{Brass1975a}:
+% To test whether the Neogene marine \SrSr record is compatible with an increase in the global weathering flux of low \SrSr juvenile lithologies over the Neogene, we implement a simple seawater model that simulates changes in marine \SrSr that result from changing fluxes from several sources \citep{Brass1975a}:
 
-\begin{equation*}
-    \frac{dR_{SW}}{dt} = \frac{1}{\tau} \sum_{i} \frac{F_{i}}{F_{tot}}(R_{i} - R_{SW})
-\end{equation*}
+% \begin{equation*}
+%     \frac{dR_{SW}}{dt} = \frac{1}{\tau} \sum_{i} \frac{F_{i}}{F_{tot}}(R_{i} - R_{SW})
+% \end{equation*}
 
-$R_{SW}$ is the \SrSr composition of the ocean, $t$ is time in the model, $\tau$ is the residence time of strontium in the oceans, $F_{i}$ is the flux of strontium from source $i$, $F_{tot}$ is the total flux of strontium from all sources, and $R_{i}$ is the \SrSr composition of source $i$.
+% $R_{SW}$ is the \SrSr composition of the ocean, $t$ is time in the model, $\tau$ is the residence time of strontium in the oceans, $F_{i}$ is the flux of strontium from source $i$, $F_{tot}$ is the total flux of strontium from all sources, and $R_{i}$ is the \SrSr composition of source $i$.
 We define 6 simplified sources of strontium to the oceans (Table \ref{tab:seawater_model}): the non-Himalayan subaerial weathering of radiogenic (older cratonic rocks such as gneiss and granite), juvenile (such as basalt), and carbonate lithologies, the Himalayan subaerial weathering of radiogenic and carbonate lithologies, and the weathering of ocean crust during hydrothermal circulation on or near mid-ocean ridges.
 
-At initialization, we set the Sr flux from each of these sources such that at 50~Ma marine \SrSr is $\sim$0.7078 (Fig.
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=1\textwidth]{Figures/seawater_model.pdf}
+%     \caption{Seawater model forcings and results.
+In the seawater \SrSr panel, the LOWESS fit is from \citet{McArthur2012a}, and the Deep Sea Drilling Project data are from \citet{Hodell1991a}, \citet{Hodell1994a}, and \citet{Oslick1994a}.
+In the source \SrSr panel, Bengal Fan data are from \citet{Derry1996a}, and Nepal data are from \citet{Quade1997a}.
+See text for details.}
+%     \label{fig:seawater_model}
+% \end{figure}
+
+% \begin{table}[h!]
+% \caption{Geochemical properties of sources used in the seawater model presented in Figure \ref{fig:seawater_model}.} 
+% \begin{center}
+%     \begin{tabular}{lcccc}
+%     \hline
+%      & & \textbf{Mg} & \textbf{Ca} & \textbf{Sr} \\
+%     \textbf{source} & \textbf{\SrSr} & (ppm) & (ppm) & (ppm) \\
+%     \hline
+%     non-Himalayan radiogenic subaerial weathering & calculated$^{1}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
+%     Himalayan radiogenic subaerial weathering & varied$^{3}$ & 12800$^{2}$ & 23750$^{2}$ & 310$^{2}$ \\
+%     non-Himalayan juvenile subaerial weathering & 0.7035$^{4}$ & 45500$^{5}$ & 71600$^{5}$ & 465$^{5}$ \\
+%     non-Himalayan carbonate subaerial weathering & 0.7080$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+%     Himalayan carbonate subaerial weathering & varied$^{3}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+%     hydrothermal exchange & 0.7030$^{4}$ & n/a$^{6}$ & n/a$^{6}$ & n/a$^{6}$ \\
+%     \hline
+%     \end{tabular}
+
+%     \scriptsize
+% 	\flushleft \emph{Notes:} \\
+%     $^{1}$ Calculated to account for $^{87}$Rb decay following the equations of \citet{Papanastassiou1968a}.\\
+%     $^{2}$ From \citet{Wedepohl1995a}\\
+%     $^{3}$ Varied as shown in Figure \ref{fig:seawater_model}.\\
+%     $^{4}$ From \citet{Allegre2010a}.\\
+%     $^{5}$ Taking the mean of \citet{Turekian1961a} and \citet{Taylor1964a}.\\
+%     $^{6}$ Not required for our model, since the Sr fluxes from these sources are set directly in our model and they do not contribute to \COtwo sequestration.\\
+% \label{tab:seawater_model}
+% \end{center}
+% \end{table}
+
+% At initialization, we set the Sr flux from each of these sources such that at 50~Ma marine \SrSr is $\sim$0.7078 (Fig.
 \ref{fig:seawater_model}).
@@
 
-The \SrSr of Himalayan lithologies is highly variable, with values ranging from $\sim$0.71--1.11 \citep{Oliver2003a}, and as orogenesis proceeded, the \SrSr of waters draining the Himalayas has changed as different tectonic units are exhumed, as evidenced by the \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal (Fig.
+% The \SrSr of Himalayan lithologies is highly variable, with values ranging from $\sim$0.71--1.11 \citep{Oliver2003a}, and as orogenesis proceeded, the \SrSr of waters draining the Himalayas has changed as different tectonic units are exhumed, as evidenced by the \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal, which are interpreted to reflect the \SrSr of Himalayan rivers (Fig.
 \ref{fig:seawater_model}; \citealp{Derry1996a, Quade1997a}).
@@
 
-We maintain initial conditions until 35~Ma, at which point we begin to change both global and Himalayan subaerial weathering fluxes (Fig.
+% We maintain initial conditions until 35~Ma, at which point we begin to change both global and Himalayan subaerial weathering fluxes (Fig.
 \ref{fig:seawater_model}).
@@
 Importantly, we also linearly decrease the Sr flux coming from non-Himalayan juvenile lithologies over this same time period in order to maintain constant \COtwo consumption (i.e.
-constant global Ca+Mg flux from subaerial weathering).
+constant global Ca+Mg flux from subaerial silicate weathering).
 In order to match the observed increase in marine \SrSr ca.
 35--16~Ma, the Sr flux coming from non-Himalayan juvenile lithologies needs to be increased to a value 1.42$\times$ the initial value.
+
+% At 16~Ma, we begin to change the subaerial weathering fluxes in accordance with existing constraints (Fig.
+\ref{fig:seawater_model}).
+First, the total Sr flux from Himalayan lithologies continues to linearly increase.
+Second, we linearly decrease the \SrSr of Himalayan lithologies from 0.720 to 0.715 16--12~Ma, then linearly increase it from 0.715 to 0.740 12--3~Ma, then linearly decrease it from 0.740 to 0.728 3--0~Ma.
+The first \SrSr decrease is intended to represent the exhumation of low \SrSr Outer Lesser Himalaya lithologies, the second \SrSr increase is intended to represent the exhumation of higher \SrSr Inner Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}, and the third \SrSr decrease is intended to represent a Pliocene-Quaternary fall in Himalayan riverine \SrSr to present-day values.
+These Himalayan \SrSr changes are consistent with the scattered \SrSr of pedogenic clays in the Bengal Fan and soil carbonate in Nepal \citep{Derry1996a, Quade1997a}.
+
+% In addition to these Himalayan changes, we also linearly decrease the Ca+Mg flux coming from non-Himalayan radiogenic lithologies to a value 0.96$\times$ the 16~Ma value.
+This 0.96 value was obtained from our GEOCLIM simulations, in which the Ca+Mg flux coming from radiogenic lithologies using modern paleogeographic boundary conditions was $\sim$0.96$\times$ the Ca+Mg flux coming from radiogenic lithologies using 15~Ma paleogeographic boundary conditions.
+
+% By imposing the geologically reasonable parameters described above, we are able to broadly reproduce the marine \SrSr record (``scenario 1'' in Figure \ref{fig:seawater_model}).
+In particular, despite an increasing contribution of low \SrSr strontium from juvenile lithologies over the past 16~m.y., marine \SrSr continues to rise due to the weathering of highly radiogenic lithologies in the Himalayas.
+To illustrate the effect of the weathering of these high \SrSr Himalayan lithologies, we also run an experiment in which we keep the \SrSr of Himalayan lithologies constant throughout the entire simulation (``scenario 2'' in Figure \ref{fig:seawater_model}).
+Without the contribution from the Himalayas, marine \SrSr begins to decrease after 16~Ma.
+
+% While our model is forced by a simplistic representation of Sr fluxes to the ocean, it nevertheless demonstrates the notion that the marine \SrSr can be decoupled from globally-averaged weathering fluxes of isotopically-representative lithologies via the regional weathering of isotopically-unique source rocks.
+Therefore, an increase in the global weathering flux of low \SrSr juvenile lithologies over the Neogene is plausibly compatible with rising marine \SrSr over the same time period.
 

### 2020-05-20 Manuscript/Manuscript.tex

@@
 
-Quantitatively estimating the magnitude of decrease in steady-state \pCOtwo associated with the emergence of a region with a high carbon sequestration potential, such as the SEAI, is not straight-forward.
+Quantitatively estimating the magnitude of decrease in steady-state \pCOtwo associated with the emergence of a region with a high carbon sequestration potential, such as the SEAI, requires constraints on changing tectonic context and accounting of associated feedbacks.
 As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline until a new steady-state is established where the total magnitude of the \COtwo sinks is the same as before the change.
@@
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
-The \pCOtwo values from the 5~Ma paleo-SEAI scenario overlap with many proxy-based estimates as well as values that emerge from hybrid climate model data assimilation approaches \cite{Berends2020a}.
+The \pCOtwo values from the 5~Ma paleo-SEAI scenario overlap with many proxy-based estimates as well as values that emerge from hybrid climate model data assimilation approaches \cite{van-de-Wal2011a, Berends2020a}.
 The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI; Fig.
@@
 \section*{Alternative Mechanisms for Neogene Cooling}
+
+\subsection*{Ocean/Atmosphere Circulation}
+
+Some hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
+Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
+However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
+\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
 
@@
 
-This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}, suggesting that changes in the volcanic degassing flux alone are insufficient to drive the majority of cooling observed over the Neogene.
+This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}.
 However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
@@
 low \SrSr) silicate rocks in the SEAI was an important driver of Neogene cooling.
-However, the ratio of silicate weathering fluxes from radiogenic cratonic rocks vs.
-juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically-unique carbonates.
+However, the ratio of silicate weathering fluxes from radiogenic cratonic rocks versus juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically-unique lithologies.
 For example, highly radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
@@
 Himalayan uplift would have affected geological carbon sinks, either via increased weathering of silicate rocks \cite{Raymo1992a} or enhanced burial of organic matter in the Bengal Fan \cite{Galy2007a}.
-Increased weathering of the SEAI as they emerged would have occurred in tandem with such changes in the Himalaya.
-Given the large uncertainty associated with changes in regional climatology across Asia due to Himalayan orogeny, developing quantitative estimates of the evolution of global silicate weathering fluxes associated with this orogeny remains a major challenge.
+Increased weathering of the emerging SEAI would have occurred in tandem with such changes in the Himalaya, such that the effects of these paleogeographic changes on geochemical proxy records, like marine \SrSr, become difficult to disentangle.
+In addition, given the large uncertainty associated with changes in regional climatology across Asia due to Himalayan orogeny, developing quantitative estimates of the evolution of global silicate weathering fluxes associated with Himalayan orogeny remains a major challenge.
+
+\section*{The Global Carbon Cycle}
 
 Global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
-Enhanced weathering in a region such as the SEAI is compensated by a decrease in weathering elsewhere.
+Enhanced silicate weathering in a region such as the SEAI is compensated by a decrease in silicate weathering elsewhere.
 Global alkalinity delivery from silicate weathering does not change, but occurs more efficiently and thereby at lower \pCOtwo.
@@
 
-\subsection*{Ocean/Atmosphere Circulation}
-
-Other hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
-Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
-However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
-\ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
+However, the long-term carbon-cycle mass balance can be perturbed via mechanisms that are disconnected from changes in volcanic degassing and silicate weathering rates.
+For example, sulphide oxidation coupled to carbonate dissolution could act as a source of \COtwo on million year timescales \cite{Torres2014a}.
+Similarly, the weathering of sedimentary organic matter could serve as a source of \COtwo \cite{Hilton2014a}.
+On the other hand, enhanced burial of organic matter could be an important sink of \COtwo, as has been suggested in the Bengal Fan \cite{Galy2007a} and Taiwan \cite{Kao2014a}.
+The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been effected by SEAI emergence and/or Himalayan orogeny.
+\pCOtwo changes that result from these processes would be superimposed on \pCOtwo changes associated with evolving silicate weathering rates.
+However, our coupled weathering-climate model indicates that the \pCOtwo change associated with evolving silicate weathering rates is sufficient to explain the majority of Neogene cooling (Fig.
+\ref{fig:scenario_pCO2}).
 

### 2020-07-31 Manuscript/Manuscript.tex
update manuscript with the latest

@@
 \author[c]{Francis A. Macdonald}
-\author[c]{Eliel A. Anttila}
+\author[c]{Eliel S. C. Anttila}
 \author[a]{Nicholas L. Swanson-Hysell}
@@
 The significance statement appears in the paper itself and is required for all research papers.
-\significancestatement{The Southeast Asian islands are a hotspot of \COtwo consumption via silicate weathering in the present day.
+\significancestatement{The Southeast Asian islands are a modern-day hotspot of \COtwo consumption via silicate weathering.
 Since $\sim$15 million years ago, these islands have been increasing in size at the same time that Earth's climate has been cooling.
-Here we test the hypothesis that global cooling over the past 15 million years has been driven by tectonic emergence of the Southeast Asian islands.
-To do so, we improve a silicate weathering model, then couple it to a climate model.
-We find that the majority of cooling over the past 15 million years can be explained by emergence of the Southeast Asian islands.}
+Here we test the hypothesis this global cooling could have been driven by tectonic emergence of the Southeast Asian islands.
+Using a new compilation of paleoshorelines in conjunction with a coupled silicate weathering and climate model, we find that this emergence is associated with a large decrease in \pCOtwo.
+Without these changes in tropical island paleogeography, there would not have been large Northern Hemisphere ice sheets as a defining feature of Earth's climate over the past 3 million years.}
 
@@
 F.A.M.
-and E.A.A.
+and E.S.C.A.
 constructed the paleoshoreline database.
@@
 Steep topography, a tropical climate, and mafic lithologies contribute to efficient chemical weathering and carbon sequestration in the Southeast Asian islands.
-Ongoing arc-continent collision between the Sunda-Banda arc system and Australia has increased the area of subaerially-exposed land in the region since the mid-Miocene.
+Ongoing arc-continent collision between the Sunda-Banda arc system and Australia has increased the area of subaerially exposed land in the region since the mid-Miocene.
 Concurrently, Earth's climate has cooled since the Miocene Climatic Optimum, leading to growth of the Antarctic ice sheet and onset of Northern Hemisphere glaciation.
@@
 To do so, we have compiled paleoshoreline data and incorporated them into GEOCLIM, which couples a global climate model to a silicate weathering model with spatially resolved lithology.
-We find that without the increase in area of the Southeast Asian islands over the Neogene, atmospheric \pCOtwo would have been significantly higher than present day values, remaining above the levels necessary for initiating Northern Hemisphere ice sheets.
+We find that without the increase in area of the Southeast Asian islands over the Neogene, atmospheric \pCOtwo would have been significantly higher than pre-industrial values, remaining above the levels necessary for initiating Northern Hemisphere ice sheets.
 \end{abstract}
@@
 
-\dropcap{T}he Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes.
-The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
-There has been a significant increase in the area of subaerially-exposed land within the region since the mid-Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
+\dropcap{T}he Southeast Asian islands (SEAI) have an out-sized contribution to modern chemical weathering fluxes relative to its area.
+The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations in the dissolved load and associated \COtwo consumption \cite{Gaillardet1999a, Hartmann2009a, Milliman2013a, Hartmann2014a}.
+There has been a significant increase in the area of subaerially exposed land within the region since the mid-Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system \cite{Molnar2015a, Hall2017a, Macdonald2019a}.
 Concurrently, after the Miocene Climatic Optimum, a cooling trend began ca.
@@
 Many hypotheses have been proposed to explain this cooling trend including changes in ocean/atmosphere circulation \cite{Haug1998a, Shevenell2004a, Molnar2015a}, a decrease in volcanic degassing \cite{Berner1983a}, or uplift in the Himalaya \cite{Raymo1988a, Galy2007a}.
-Here we seek to evaluate the hypothesis that SEAI emergence was a significant factor in driving long-term climatic cooling over the Neogene.
+Here we seek to evaluate the hypothesis that emergence of the SEAI was a significant factor in driving long-term climatic cooling over the Neogene.
 
-Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, with a smaller sink through organic carbon burial \cite{Kump2000a}.
+Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks and through organic carbon burial \cite{Kump2000a}.
 Chemical weathering delivers alkalinity and cations to the ocean which drives carbon sequestration through carbonate precipitation.
 Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources.
-As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo \cite{Kump1997a}.
-Conversely, as regions with high carbon sequestration potential emerge on Earth's surface, \pCOtwo decreases resulting in other weathering sinks to diminish leading to a new lower steady-state \pCOtwo.
+As \COtwo sinks increase and \pCOtwo falls, temperature decreases and the hydrological cycle is weakened, causing the efficiency of the silicate weathering sink to decrease until a new steady-state is achieved at lower \pCOtwo \cite{Kump1997a}.
 
 Topography, climate, and lithology all effect chemical weathering.
-High-relief regions lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more prone to adjust when climate changes \cite{Gabet2009a, West2012a, Maher2014a}.
+High-relief regions generally lack extensive regolith development, and thus tend to have reaction-limited weathering regimes that are more prone to adjust when climate changes \cite{Gabet2009a, West2012a, Maher2014a}.
 High physical erosion rates contribute to high chemical weathering fluxes in these high-relief regions \citep{Godderis2017b}.
@@
 Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus have the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
-These factors have lead to the proposal that arc-continent collisions, which create steep landscapes of mafic lithologies, within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
-\cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019b}.
+These factors have led to the proposal that arc-continent collisions, which create steep landscapes of mafic lithologies, within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
+\cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2020a}.
 
 Quantitatively estimating the magnitude of decrease in steady-state \pCOtwo associated with the emergence of a region with a high carbon sequestration potential, such as the SEAI, requires constraints on changing tectonic context and accounting of associated feedbacks.
-As this region emerges, the total global silicate weathering flux will briefly exceed the volcanic degassing flux, causing \pCOtwo to initially decline until a new steady-state is established where the total magnitude of the \COtwo sinks is the same as before the change.
+As this region emerges, the total global silicate weathering flux will transiently exceed the volcanic degassing flux, causing \pCOtwo to initially decline until a new steady-state is established where the total magnitude of the \COtwo sinks is the same as before the change.
 However, the sensitivity of the silicate weathering flux in any particular location to this change in \pCOtwo is variable and dependent on the specific topography, climate, and lithology at that location.
-Furthermore, how regional climate responds to this change in \pCOtwo is itself spatially-variable.
+Furthermore, how regional climate responds to this change in \pCOtwo is itself spatially variable.
 Therefore, the magnitude of \pCOtwo change that is required to balance the total global silicate weathering flux with the volcanic degassing flux will depend on the specific spatial distribution of topography, climate, and lithology at the time of emergence.
-As a result, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must model spatially-resolved climatology and silicate weathering fluxes in tandem and account for the spatial distribution of the factors that affect these inter-connected systems.
+As a result, any attempt to meaningfully estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI must model spatially resolved climatology and silicate weathering fluxes in tandem and account for the spatial distribution of the factors that affect these inter-connected systems.
 
@@
 
-To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially-exposed land area in the SEAI, we use the global spatially-resolved GEOCLIM model \cite{Godderis2017c}.
+To estimate the decrease in steady-state \pCOtwo associated with the increase of subaerially exposed land area in the SEAI, we use the global spatially resolved GEOCLIM model \cite{Godderis2017c}.
 GEOCLIM estimates changes in steady-state \pCOtwo associated with coupled changes in erosion, chemical weathering, and climatology by linking a silicate weathering model to climate model runs at multiple \pCOtwo levels.
@@
 Regolith production and physical erosion ($E_{p}$) are equal at steady-state.
-As a particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$), with the flux of Ca+Mg being $W$ multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
+As a particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$), with the flux of dissolved Ca+Mg being $W$ multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).
 Details of the formulation for the silicate weathering component of GEOCLIM can be found in \MM.}
@@
 
-The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for subaerially-exposed land.
+The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for subaerially exposed land.
 We assume that Ca and Mg are the only cations that consume \COtwo over geologic time-scales, such that each mole of Ca or Mg that is dissolved by silicate weathering consumes one mole of \COtwo.
+While reverse weathering is another potential sink for Ca or Mg \citep{Michalopoulos1995a}, its parameterization is unclear and it has been interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}, and so we do not include it in our model.
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and all bedrock was assigned identical chemical compositions \cite{Godderis2017c}.
@@
 
-The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained.
+The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, erosion, and regolith thickness are poorly constrained.
 Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic lithologies to vary within reasonable bounds (\SI; Table S2).
 This approach results in 93,600 unique combinations of these six parameters.
-For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
+For each combination, we compute spatially resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
 We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds.
@@
 S2).
-The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a} as they should given that the long-term carbon cycle is in steady state (\SI; Fig.
+The resulting global \COtwo consumption of these filtered model runs all overlap with independently derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a}, as they should given that the long-term carbon cycle is in steady state (\SI; Fig.
 S2).
@@
 
-Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with SEAI emergence.
+Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with emergence of the SEAI.
 For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (\SI).
@@
     \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text.
-For each of the seven scenarios, each point represents an estimate from one of the 573 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
+For each of the seven scenarios, each point represents an estimate from one of the 573 unique parameter combinations that most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (\SI).
 The box encloses the middle 50\% of the \pCOtwo estimates (i.e.
@@
 
-To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
+To determine the position of paleoshorelines in the SEAI over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig.
 \ref{fig:shoreline_growth}; \SI).
@@
 Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the Australian Plate \cite{Harris2006a}.
-Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma.
-However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma \cite{Hall2013b}.
+Most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma \cite{Hall2013b}.
 In New Guinea, emergence in the mid-Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{Cloos2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt.
@@
 in the New Guinea Central Range due to slab-breakoff and buoyant uplift, and in eastern New Guinea due to jamming of the north-dipping subduction zone \cite{Cloos2005a}.
-These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y.
+We also include changes in areas of presently submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI; Fig.
+S5).
+These tectonic drivers and others throughout the region led to progressive emergence over the past 15~m.y.
 that accelerated following 5~Ma (Fig.
@@
 \ref{fig:shoreline_growth}C).
-We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI; Fig.
-S5).
 
@@
 S8).
-We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).
+We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
+\ref{fig:scenario_pCO2}).
 
@@
 These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
+We note that these ``paleo-SEAI'' scenarios do not account for Neogene changes outside of the SEAI (e.g.
+changes in ocean/atmosphere circulation, volcanic degassing, and weathering fluxes elsewhere on Earth, discussed in \textit{Alternative Mechanisms for Neogene Cooling}).
+Therefore, these results do not imply that \pCOtwo would have been observed to be 526--678~ppm at 15~Ma, but rather that emergence of the SEAI alone is responsible for decreasing \pCOtwo by 240--392~ppm (526--678~ppm minus 286~ppm) over the past 15~m.y..
 
-Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a}.
-The \pCOtwo values from the 5~Ma paleo-SEAI scenario overlap with many proxy-based estimates as well as values that emerge from hybrid climate model data assimilation approaches \cite{van-de-Wal2011a, Berends2020a}.
-The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (\SI; Fig.
-S9), indicating that the increase in subaerially-exposed land area and tectonic topography of the SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
+Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation (\SI; Fig.
+S9).
+The \pCOtwo values from the 5~Ma ``paleo-SEAI'' scenario overlap with many proxy-based estimates \cite{Foster2017a} as well as values that emerge from approaches that assimilate climate and ice sheet model output with benthic \dO data \cite{van-de-Wal2011a, Berends2020a}.
+The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early to mid-Miocene, indicating that the increase in subaerially exposed land area and tectonic topography of the SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene.
 The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}.
-These modeled values of decreasing \pCOtwo associated with SEAI emergence are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.
+These modeled values of decreasing \pCOtwo associated with emergence of the SEAI are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.
 
-The results of our SEAI scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
+The results of our ``paleo-SEAI'' scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
 To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2013a}, we replace mafic volcanics associated with the Deccan Traps, Ethiopian Traps, and Columbia River Basalts with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
@@
 
-Previous work has estimated that the decrease in \pCOtwo since 5~Ma associated with the emergence of the SEAI and enhanced silicate weathering is $\sim$19~ppm \cite{Molnar2015a}, in which case SEAI emergence would be a relatively minor contributor to Neogene cooling.
+Previous work has estimated that the decrease in \pCOtwo since 5~Ma associated with the emergence of the SEAI and enhanced silicate weathering is $\sim$19~ppm \cite{Molnar2015a}, in which case their emergence would be a relatively minor contributor to Neogene cooling.
 This 19~ppm estimate was obtained using a 0-D equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
@@
 \ref{fig:scenario_pCO2}) demonstrates that considering changes in the spatial distribution of lithologies alone is not adequate for estimating changes in steady-state \pCOtwo.
-Instead, spatially-varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for when estimating \pCOtwo change associated with paleogeographic change.
+Instead, spatially varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for when estimating \pCOtwo change associated with paleogeographic change.
 
@@
 
-Some hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with SEAI emergence \cite{Molnar2015a}.
+Some hypotheses to explain ice sheet growth over the Neogene invoke changes in ocean/atmosphere circulation including: further climatic isolation of Antarctica due to strengthening of the circumpolar current \cite{Shevenell2004a}; increased atmospheric moisture in the Northern Hemisphere due to intensified thermohaline circulation following Panama Isthmus emergence \cite{Haug1998a}; and cooling of North America resulting from a strengthened Walker Circulation associated with emergence of the SEAI \cite{Molnar2015a}.
 Such changes in ocean/atmosphere circulation are likely to modulate \pCOtwo thresholds for glacial initiation and ice sheet growth \cite{DeConto2008a}.
-However, the prolonged timescale of the cooling trend since 15~Ma (Fig.
+However, the prolonged time-scale of the cooling trend since 15~Ma (Fig.
 \ref{fig:shoreline_growth}C) is most readily attributable to decreasing \pCOtwo associated with evolving geological sources and sinks of carbon, modulated by the silicate weathering feedback \cite{Walker1981a, Raymo1991a, Berner1997a, Kump1997a, Berner2001a}.
@@
 \subsection*{Volcanic Degassing}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
-    \caption{Weatherability curves for the modern and ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
-Details on how these curves were generated are described in \MM.
-Each of the 4 curves represent a different tectonic boundary condition (i.e.
-the reconstructed paleoshorelines of the SEAI; Fig.
-\ref{fig:shoreline_growth}A) and therefore a different global weatherability.
-The curves show the resulting \pCOtwo for a given volcanic degassing flux such that the input flux is balanced by the silicate weathering output flux.
-Point B represents the pre-industrial, in which \pCOtwo is 286~ppm.
-The arrow from Point A$_{1}$ to B represents the ``change in weatherability only'' scenario, in which global weatherability changes as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
-In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve ($\sim$570~ppm).
-The arrow from Point A$_{2}$ to B instead represents the ``change in degassing only'' scenario, in which global weatherability remains the same as the pre-industrial, but the same change in \pCOtwo as the ``change in weatherability only'' scenario is achieved by decreasing the volcanic degassing flux from a value $\sim$13\% greater than the pre-industrial.
-The arrow from Point A$_{3}$ to B represents the ``change in weatherability and degassing'' scenario, in which the same change in \pCOtwo as the other two scenarios is achieved by increasing global weatherability from our 5~Ma tectonic boundary condition and decreasing the volcanic degassing flux from a value $\sim$7\% greater than the pre-industrial.}
-    \label{fig:weatherability_curves}
-\end{figure}
 
@@
 S3), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
-If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI lead to a change in \pCOtwo of $\sim$280~ppm (``change in weatherability only'' scenario in Fig.
+If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI led to a change in \pCOtwo of $\sim$280~ppm (``increase in weatherability only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
-If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-industrial to drive the same $\sim$280~ppm change in \pCOtwo (``change in degassing only'' scenario in Fig.
+If we instead assume that global weatherability did not change over the past 15~m.y., then we estimate that the volcanic degassing flux needs to have been $\sim$13\% greater at 15~Ma relative to the pre-industrial to drive the same $\sim$280~ppm change in \pCOtwo (``decrease in degassing only'' scenario in Fig.
 \ref{fig:weatherability_curves}).
+This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}.
 
-This $\sim$13\% value is higher than 10\%, the highest current estimate for the volcanic degassing flux at 15~Ma relative to the present day \cite{Van-Der-Meer2014a}.
 However, changes in the volcanic degassing flux would have modulated changes in \pCOtwo associated with changes in global weatherability.
-For example, take a scenario where the increase in global weatherability associated with SEAI emergence over the past 15~m.y.
-is an overestimate and the increase is more accurately reflected by the difference in global weatherability between our 5~Ma and modern tectonic boundary conditions (``change in weatherability and degassing'' scenario in Fig.
-\ref{fig:weatherability_curves}).
-In this scenario, a decrease in the volcanic degassing flux from a value $\sim$7\% greater than the pre-industrial in conjunction with the smaller change in weatherability can achieve the same $\sim$280~ppm decrease in \pCOtwo.
+For example, some proxy-based approaches as well as some model-data assimilation approaches estimate that mid-Miocene \pCOtwo was lower than 568~ppm (\SI; Fig.
+S9).
+Take a scenario in which \pCOtwo was 400~ppm at 15~Ma.
+If we assume that the \pCOtwo decrease to the pre-industrial value of 286~ppm was driven by the increase in global weatherability associated with emergence of the SEAI in conjunction with an increase in volcanic degassing which counteracts cooling by increasing the flux of \COtwo to the atmosphere (``increase in weatherability and degassing'' scenario in Fig.
+\ref{fig:weatherability_curves}), the volcanic degassing flux would have had to have been $\sim$7\% smaller than the pre-industrial.
+More robust constraints on \pCOtwo (\SI; Fig.
+S9) and/or volcanic degassing rates over the past 20~m.y.
+are needed to constrain which of the ``increase in weatherability only'' or ``increase in weatherability and degassing'' scenarios (Fig.
+\ref{fig:weatherability_curves}) is more representative of the mechanisms driving Neogene cooling.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
+    \caption{Weatherability curves for the modern and ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
+The lower panel expands the lower \pCOtwo range (x-axis) of the upper panel.
+Details on how these curves were generated are described in \MM.
+Each of the 4 curves represent a different tectonic boundary condition (i.e.
+the reconstructed paleoshorelines of the SEAI; Fig.
+\ref{fig:shoreline_growth}A) and therefore a different global weatherability.
+The curves show the resulting \pCOtwo for a given volcanic degassing flux such that the input flux is balanced by the silicate weathering output flux.
+Point B represents the pre-industrial, in which \pCOtwo is 286~ppm.
+The arrow from Point A$_{1}$ to B represents the ``increase in weatherability only'' scenario, in which global weatherability increases as the SEAI emerge, but the volcanic degassing flux does not change over the past 15~m.y.
+In this scenario, the \pCOtwo decreases from the value dictated by the 15~Ma ``paleo-SEAI'' weatherability curve (568~ppm).
+The arrow from Point A$_{2}$ to B instead represents the ``decrease in degassing only'' scenario, in which global weatherability remains the same as the pre-industrial, but the same change in \pCOtwo as the ``increase in weatherability only'' scenario is achieved by decreasing the volcanic degassing flux from a value $\sim$13\% greater than the pre-industrial.
+The arrow from Point A$_{3}$ to B represents the ``increase in weatherability and degassing'' scenario, in which a change in \pCOtwo from 400~ppm to 286~ppm is achieved by increasing both global weatherability from our 15~Ma tectonic boundary condition as well as the volcanic degassing flux from a value $\sim$7\% smaller than the pre-industrial flux.}
+    \label{fig:weatherability_curves}
+\end{figure}
 
@@
 low \SrSr) silicate rocks in the SEAI was an important driver of Neogene cooling.
-However, the ratio of silicate weathering fluxes from radiogenic cratonic rocks versus juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically-unique lithologies.
-For example, highly radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
-As such, there are challenges in interpreting the marine \SrSr record as a proxy for silicate weathering fluxes in the Cenozoic, and potentially deeper in the geologic record as well.
+However, the globally averaged ratio of silicate weathering fluxes from radiogenic cratonic rocks versus juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically unique lithologies.
+For example, unusually radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
+As such, there are challenges in interpreting the marine \SrSr record as a direct proxy for silicate weathering fluxes.
 Nevertheless, steadily increasing marine \SrSr is interrupted ca.
@@
 
-\section*{The Global Carbon Cycle}
+\section*{The Geologic Carbon Cycle}
 
-Global alkalinity delivery from silicate weathering needs to be nearly constant to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
+If geological carbon sources remain approximately constant, global alkalinity delivery from silicate weathering needs to be approximately constant as well to keep the long-term carbon cycle in steady-state \cite{Kump1997a}.
 Enhanced silicate weathering in a region such as the SEAI is compensated by a decrease in silicate weathering elsewhere.
@@
 
-However, the long-term carbon-cycle mass balance can be perturbed via mechanisms that are disconnected from changes in volcanic degassing and silicate weathering rates.
-For example, sulphide oxidation coupled to carbonate dissolution could act as a source of \COtwo on million year timescales \cite{Torres2014a}.
+Alkalinity generated by the weathering of silicate Na and K consumes \COtwo until that alkalinity is removed via reverse weathering reactions \cite{Sun2016a}.
+Therefore, an imbalance in the Na and K cycles could consume significant quantities of \COtwo over multi-million year time-scales.
+However, the biogeochemical cycles of Na and K are currently poorly constrained, such that the residence time of these two cations, and therefore the time-scale over which an imbalance in their cycles could drive \COtwo consumption, is not well known.
+Estimates for these values in the literature range from XX to XX~m.y.
+We have therefore made the assumption in GEOCLIM that alkalinity generated by the weathering of silicate Na and K is instantaneously removed by reverse weathering reactions, but acknowledge that further work is necessary to evaluate whether such an assumption is valid.
+
+The long-term carbon-cycle mass balance can be perturbed via mechanisms that are disconnected from changes in volcanic degassing and silicate weathering rates.
+For example, sulphide oxidation coupled to carbonate dissolution could act as a source of \COtwo on million year time-scales \cite{Torres2014a}.
 Similarly, the weathering of sedimentary organic matter could serve as a source of \COtwo \cite{Hilton2014a}.
 On the other hand, enhanced burial of organic matter could be an important sink of \COtwo, as has been suggested in the Bengal Fan \cite{Galy2007a} and Taiwan \cite{Kao2014a}.
-The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been effected by SEAI emergence and/or Himalayan orogeny.
+The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been affected by emergence of the SEAI and/or Himalayan orogeny.
 \pCOtwo changes that result from these processes would be superimposed on \pCOtwo changes associated with evolving silicate weathering rates.
-However, our coupled weathering-climate model indicates that the \pCOtwo change associated with evolving silicate weathering rates is sufficient to explain the majority of Neogene cooling (Fig.
+However, our coupled weathering-climate model indicates that the \pCOtwo change associated with evolving silicate weathering rates driven by emergence of the SEAI is sufficient to explain the majority of Neogene cooling (Fig.
 \ref{fig:scenario_pCO2}).
+Without this emergence, \pCOtwo would have remained above the level necessary for the growth of Northern Hemisphere ice sheets.
 
@@
 
-Coupled geological constraints and modeling experiments demonstrate that the SEAI has been a growing hot spot for carbon sequestration due to silicate weathering from the Miocene to present.
+Coupled geological constraints and modeling experiments demonstrate that the SEAI have been a growing hot spot for carbon sequestration due to silicate weathering from the Miocene to present.
 Changes in volcanic degassing and paleogeography elsewhere on Earth, particularly in the Himalaya and Central America, would have also affected geological carbon sources and sinks.
-Yet, not only does the history of SEAI emergence coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain much of this cooling.
-These results highlight that the Earth's climate state is particularly sensitive to changes in tropical paleogeography.
+Yet, not only does the history of emergence of the SEAI coincide with Neogene cooling and the onset of Northern Hemisphere glaciation, but our coupled weathering-climate model also indicates that the associated steady-state \pCOtwo change is sufficient to explain much of this cooling.
+These results highlight that the Earth's climate state is particularly sensitive to changes in tropical geography.
 
 \matmethods{
-The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}.
+The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state/releases/tag/v1.0}.
 The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
@@
 \begin{equation}
-    f(h) = e^{-\frac{-h}{d_{0}}}
+    f(h) = e^{\frac{-h}{d_{0}}}
     \label{eq:7}
@@
 We then repeat this process for the 10 and 5~Ma ``paleo-SEAI'' paleoshorelines and the present day shorelines to generate the 3 other weatherability curves.
-Each estimated \pCOtwo in Figure \ref{fig:scenario_pCO2} is the result of underlying weatherability curves that subtly change with the different chemical weathering parameters.
+Each estimated \pCOtwo in Figure \ref{fig:scenario_pCO2} is the result of underlying weatherability curves that change with the different chemical weathering parameters.
 

### 2020-07-31 Manuscript/Supplementary_Information.tex
update manuscript with the latest

@@
 \begin{flushleft}
-{\Large \textbf{Supplementary Information for ``Emergence of the Southeast Asian islands as a driver for Neogene cooling''}}
+{\Large \textbf{Supporting Information for ``Emergence of the Southeast Asian islands as a driver for Neogene cooling''}}
 
@@
 Francis A. Macdonald\textsuperscript{2},
-Eliel A. Anttila\textsuperscript{2},
+Eliel S. C. Anttila\textsuperscript{2},
 Nicholas L. Swanson-Hysell\textsuperscript{1}
@@
 These supplementary information materials provide details on the model framework used in this study.
-The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state}.
+The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state/releases/tag/v1.0}.
 The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
@@
 However, there is uncertainty in the value of the proportionality constant ($k_e$).
-We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr).
+Using the simplification of a uniform value, we therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr).
 Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/paleogeographic_reconstructions.pdf}
     \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
@@
 For the \citet{Ji2018a} data, error bars represent the 16-84$^{th}$ percentile of resampled estimates.
+Individual data points from \citet{Cui2020a} were not published with uncertainties.
 The \citet{Bereiter2015a} data come from the EPICA Dome C ice core.}
@@
 Proxy-based \pCOtwo estimates for the past 20~m.y.
-are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a}.
+are shown in Figure \ref{fig:pCO2_proxies} \citep{Bereiter2015a, Foster2017a, Ji2018a, Cui2020a}.
 Even when stringent quality criteria and the latest understanding of each of the \pCOtwo proxies have been applied to available \pCOtwo records, both significant uncertainty in the estimated \pCOtwo for any given data point as well as disagreement between techniques remain.
+Statistical methods that are often utilized in an attempt to extract trends from this data (e.g.
+locally-weighted scatter plot smoothing, LOWESS) evaluate the mean (with some weighting) of the highly scattered data in a given time interval.
+However, such methods are only appropriate for estimating the ``true'' value at any given time interval in datasets in which individual data points are being drawn from a single distribution about a mean value.
+The \pCOtwo proxy compilation shown in Figure \ref{fig:pCO2_proxies} consists of several distinct techniques and samples, each with their own set of assumptions and associated probability distributions that influence their reported values.
+Therefore, simply taking the mean of these values neglects the fact that some of these techniques/samples are more reliable others.
+Constraining how robust each \pCOtwo proxy technique is remains an important challenge, and as such the ``true'' \pCOtwo could plausibly lie anywhere within the full range of proxy-based \pCOtwo estimates for any given time interval.
 Nevertheless, our modeled \pCOtwo values for 15~Ma (Fig.

### 2020-07-31 Manuscript/Manuscript.tex

@@
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
-As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative.
+As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative \cite{Schopka2011a}.
 The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored.
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component of GEOCLIM.
+The siliciclastic sediment lithologic category also includes sedimentary sequences in which any carbonate is identified but is not the dominant lithology \cite{Hartmann2012a}.
+We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component of GEOCLIM, but assume that all pixels of a given lithologic category share the same Ca+Mg concentration.
 
@@
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, erosion, and regolith thickness are poorly constrained.
-Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic lithologies to vary within reasonable bounds (\SI; Table S2).
-This approach results in 93,600 unique combinations of these six parameters.
+Rather than prescribing single values, we test several reasonable values for each of these four parameters along with the Ca+Mg concentration of metamorphic and siliciclastic lithologies (\SI; Table S2).
+We then permute all possible combinations of these reasonable values for the six parameters, leading to 93,600 unique parameter combinations (i.e.
+permutations).
 For each combination, we compute spatially resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope.
@@
 After eliminating parameter combinations that result in low $r^{2}$, 573 parameter combinations remain (\SI; Fig.
-S2).
+S3).
 The resulting global \COtwo consumption of these filtered model runs all overlap with independently derived estimates of the global \COtwo degassing flux \cite{Gerlach2011a}, as they should given that the long-term carbon cycle is in steady state (\SI; Fig.
-S2).
+S3).
 
@@
 We also include changes in areas of presently submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI; Fig.
-S5).
+S6).
 These tectonic drivers and others throughout the region led to progressive emergence over the past 15~m.y.
@@
 Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (\SI; Fig.
-S8).
+S9).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
@@
 These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
+Increasing seawater Mg/Ca since ca.
+15~Ma \cite{Higgins2012a} is consistent with an increasing proportion of the global silicate weathering flux being derived from mafic and ultramafic sources.
 We note that these ``paleo-SEAI'' scenarios do not account for Neogene changes outside of the SEAI (e.g.
 changes in ocean/atmosphere circulation, volcanic degassing, and weathering fluxes elsewhere on Earth, discussed in \textit{Alternative Mechanisms for Neogene Cooling}).
-Therefore, these results do not imply that \pCOtwo would have been observed to be 526--678~ppm at 15~Ma, but rather that emergence of the SEAI alone is responsible for decreasing \pCOtwo by 240--392~ppm (526--678~ppm minus 286~ppm) over the past 15~m.y..
+Therefore, these results do not imply that \pCOtwo would have been observed to be 526--678~ppm at 15~Ma, but rather that emergence of the SEAI alone is responsible for decreasing \pCOtwo by 240--392~ppm (526--678~ppm minus 286~ppm) over the past 15~m.y.
 
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation (\SI; Fig.
-S9).
+S10).
 The \pCOtwo values from the 5~Ma ``paleo-SEAI'' scenario overlap with many proxy-based estimates \cite{Foster2017a} as well as values that emerge from approaches that assimilate climate and ice sheet model output with benthic \dO data \cite{van-de-Wal2011a, Berends2020a}.
@@
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
-S3), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
+S4), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
 If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI led to a change in \pCOtwo of $\sim$280~ppm (``increase in weatherability only'' scenario in Fig.
@@
 For example, some proxy-based approaches as well as some model-data assimilation approaches estimate that mid-Miocene \pCOtwo was lower than 568~ppm (\SI; Fig.
-S9).
+S10).
 Take a scenario in which \pCOtwo was 400~ppm at 15~Ma.
@@
 More robust constraints on \pCOtwo (\SI; Fig.
-S9) and/or volcanic degassing rates over the past 20~m.y.
+S10) and/or volcanic degassing rates over the past 20~m.y.
 are needed to constrain which of the ``increase in weatherability only'' or ``increase in weatherability and degassing'' scenarios (Fig.
@@
 Alkalinity generated by the weathering of silicate Na and K consumes \COtwo until that alkalinity is removed via reverse weathering reactions \cite{Sun2016a}.
-Therefore, an imbalance in the Na and K cycles could consume significant quantities of \COtwo over multi-million year time-scales.
-However, the biogeochemical cycles of Na and K are currently poorly constrained, such that the residence time of these two cations, and therefore the time-scale over which an imbalance in their cycles could drive \COtwo consumption, is not well known.
-Estimates for these values in the literature range from XX to XX~m.y.
-We have therefore made the assumption in GEOCLIM that alkalinity generated by the weathering of silicate Na and K is instantaneously removed by reverse weathering reactions, but acknowledge that further work is necessary to evaluate whether such an assumption is valid.
+Therefore, sustained imbalances in the Na and K cycles could consume/generate significant quantities of \COtwo over multi-million year time-scales.
+However, the biogeochemical cycles of Na and K are currently poorly constrained.
+As such, the residence time of these two cations, and therefore the time-scale over which an imbalance in their cycles could drive \COtwo consumption/generation, is not well known.
+Estimates for these values in the literature range from $\sim$45-80~m.y.
+for Na, and $\sim$8-10~m.y.
+for K \cite{Emerson2008a, Lecuyer2016a}.
+Furthermore, it is not clear how to parameterize the output fluxes of Na and K.
+We have therefore made the assumption in GEOCLIM that alkalinity generated by the weathering of silicate Na and K is instantaneously removed by reverse weathering reactions, but acknowledge that further work is necessary to evaluate how important imbalances in the Na and K cycles are for the geologic carbon cycle.
 
@@
 We then select the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
-S3), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI).
+S4), and fix \pCOtwo at the 3 \pCOtwo levels at which the GFDL CM2.0 climate model experiments were computed (\SI).
 We then run GEOCLIM at each of these \pCOtwo levels until steady state is achieved (i.e.

### 2020-07-31 Manuscript/Supplementary_Information.tex

@@
     \includegraphics[width=1\textwidth]{Figures/world_lithology.jpg}
-    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM modified from \cite{Hartmann2012a}.}
+    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM modified from \citet{Hartmann2012a}.}
     \label{fig:world_lithology}
@@
 For metamorphic and siliciclastic lithologies, we explore a range of feasible Ca+Mg concentrations during calibration of the silicate weathering component of GEOCLIM (\textit{GEOCLIM Calibration}).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Figures/resolution_sensitivity.pdf}
+    \caption{Histograms showing the difference in the area-weighted mean Ca+Mg concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel that results from starting with an 0.1\degrees $\times$ 0.1\degrees GLiM raster versus an 0.05\degrees $\times$ 0.05\degrees GLiM raster.
+The left panel is calculated over all land pixels, whereas the right panel is calculated over land pixels in the SEAI only.
+Both panels use the parameter values that produced the highest $r^{2}$ during calibration: metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$ and sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
+    \label{fig:resolution_sensitivity}
+\end{figure}
+
+We also explore the sensitivity of the area-weighted mean Ca+Mg concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel to the resolution of the underlying GLiM raster (Fig.
+\ref{fig:resolution_sensitivity}).
+We find that the difference in the area-weighted mean Ca+Mg concentration of the surface when using the 0.1\degrees $\times$ 0.1\degrees GLiM raster versus the 0.05\degrees $\times$ 0.05\degrees GLiM raster can be well explained by a tight Gaussian distribution ($\sigma$\textless200~mol/m$^{3}$) about a mean of $\sim$0~mol/m$^{3}$.
+Therefore, using the lower-resolution GLiM raster does not overall bias our results to higher or lower area-weighted mean Ca+Mg concentrations of the surface, and only introduces some noise that has a magnitude that is significantly smaller than the most Ca+Mg-poor lithologic category (i.e.
+felsics, at 1,521~mol/m$^{3}$).
 
@@
 The \pCOtwo proxy compilation shown in Figure \ref{fig:pCO2_proxies} consists of several distinct techniques and samples, each with their own set of assumptions and associated probability distributions that influence their reported values.
-Therefore, simply taking the mean of these values neglects the fact that some of these techniques/samples are more reliable others.
+Therefore, simply taking the mean of these values neglects the fact that some of these techniques/samples are likely more reliable than the others.
 Constraining how robust each \pCOtwo proxy technique is remains an important challenge, and as such the ``true'' \pCOtwo could plausibly lie anywhere within the full range of proxy-based \pCOtwo estimates for any given time interval.

### 2020-08-03 Manuscript/Manuscript.tex
revision checkpoint

@@
 The results of our ``paleo-SEAI'' scenarios highlight the importance of the combination of topography, runoff, and lithology in setting Earth's climate state.
-To independently explore the effect of the modern-day surface exposure of low-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2013a}, we replace mafic volcanics associated with the Deccan Traps, Ethiopian Traps, and Columbia River Basalts with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
+To independently explore the effect of the modern-day surface exposure of lower-relief basaltic lavas on steady-state \pCOtwo \cite{Kent2013a}, we replace mafic volcanics associated with the Deccan Traps, Ethiopian Traps, and Columbia River Basalts with the Ca+Mg concentration of bulk continental crust in GEOCLIM (Fig.
 \ref{fig:scenario_pCO2}).
-The resulting \pCOtwo is $<$500~ppm, indicating that, while the presence of mafic rocks in these igneous provinces modulates steady-state \pCOtwo as has been suggested to be important for Paleogene cooling \cite{Kent2013a}, its contribution is less than that of the SEAI, which have high-relief and a wet tropical climate in conjunction with high Ca+Mg lithologies.
+The resulting \pCOtwo is $\sim$300--500~ppm, indicating that the presence of mafic rocks in these igneous provinces affects steady-state \pCOtwo as has been suggested to be important for Paleogene cooling \cite{Kent2013a}.
+However, the higher 526--678~ppm values for the 15~Ma ``paleo-SEAI'' scenario illustrate that higher relief and a wet tropical climate significantly increase the efficiency of \COtwo consumption, especially when paired with high Ca+Mg lithologies.
+As such, arc-continent collisions in the tropics are likely more important for driving long-term changes in \pCOtwo than the eruption of flood basalts \cite{Macdonald2019a, Park2019a}.
 
@@
 high \SrSr) silicate rocks \cite{Raymo1988a}.
-Associated with this explanation is the proposal that increasing weathering of radiogenic silicate rocks in the Himalayas was the primary driver of Neogene cooling \cite{Raymo1992a}.
+Associated with this explanation is the proposal that increasing weathering of radiogenic silicate rocks in the Himalayas was the primary driver of Neogene cooling \cite{Raymo1992a, Edmond1992a}.
 It could be argued that increasing marine \SrSr is inconsistent with the hypothesis that increasing weathering of juvenile (i.e.
@@
 On the other hand, enhanced burial of organic matter could be an important sink of \COtwo, as has been suggested in the Bengal Fan \cite{Galy2007a} and Taiwan \cite{Kao2014a}.
+An increased flux of dissolved nutrients derived from silicate weathering could even strengthen the biological carbon pump and sequester \COtwo as organic carbon in deep marine sediments \cite{Hartmann2013a}.
 The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been affected by emergence of the SEAI and/or Himalayan orogeny.

### 2020-08-03 Manuscript/Supplementary_Information.tex
revision checkpoint

@@
 We expect this bias to be minimal, as all of our modeled \pCOtwo estimates fall below the glaciation threshold for Antarctica.
+
+\citet{Zhu2020a} reviewed three climate models (CESM2, CESM1, and CCSM4) more recent than the GFDL CM2.0 experiments, and found that they had equilibrium climate sensitivities (the long-term response of global mean surface temperature to a doubling of \pCOtwo) of 5.3, 4.2, and 3.2\degreesC respectively.
+Of these three climate models, they found that the equilibrium climate sensitivity of CESM1 (4.2\degreesC) was most in agreement with paleo-\pCOtwo and global mean surface temperature data.
+The GFDL CM2.0 experiments have a lower equilibrium climate sensitivity ($\sim$2\degreesC) than CESM1.
+This discrepancy indicates that our \pCOtwo estimates (Fig.
+\ref{fig:scenario_pCO2_all}) are biased toward higher values than what would have been obtained if we had used CESM1 as the climate model component.
+However, equilibrium climate sensitivity over the Neogene remains poorly constrained, with estimates ranging from 1.5--4.5\degreesC \citep{Zhu2020a}.
 

### 2020-08-12 Manuscript/Manuscript.tex
fix figure paths

@@
 Since $\sim$15 million years ago, these islands have been increasing in size at the same time that Earth's climate has been cooling.
-Here we test the hypothesis this global cooling could have been driven by tectonic emergence of the Southeast Asian islands.
+Here we test the hypothesis that this global cooling could have been driven by tectonic emergence of the Southeast Asian islands.
 Using a new compilation of paleoshorelines in conjunction with a coupled silicate weathering and climate model, we find that this emergence is associated with a large decrease in \pCOtwo.
@@
 Mafic rocks have higher Ca and Mg concentrations and dissolution rates than felsic rocks, and thus have the potential to more efficiently sequester carbon through silicate weathering \cite{Dessert2003a}.
-These factors have led to the proposal that arc-continent collisions, which create steep landscapes of mafic lithologies, within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
+These factors have led to the proposal that arc-continent collisions, which create steep landscapes that include mafic lithologies, within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y.
 \cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2020a}.
@@
 We assume that Ca and Mg are the only cations that consume \COtwo over geologic time-scales, such that each mole of Ca or Mg that is dissolved by silicate weathering consumes one mole of \COtwo.
-While reverse weathering is another potential sink for Ca or Mg \citep{Michalopoulos1995a}, its parameterization is unclear and it has been interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}, and so we do not include it in our model.
+While reverse weathering is another potential sink for Ca or Mg \citep{Michalopoulos1995a}, its parameterization is unclear and it has been interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}, and we do not include it in our model.
 In previous versions of the model, silicate weathering was a function of temperature and runoff only, and all bedrock was assigned identical chemical compositions \cite{Godderis2017c}.
@@
 Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks.
-As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative \cite{Schopka2011a}.
+As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo could be conservative \cite{Schopka2011a}.
 The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored.
 The Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion.
-The siliciclastic sediment lithologic category also includes sedimentary sequences in which any carbonate is identified but is not the dominant lithology \cite{Hartmann2012a}.
-We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component of GEOCLIM, but assume that all pixels of a given lithologic category share the same Ca+Mg concentration.
+We explore a range of feasible Ca+Mg concentrations for metamorphic rocks and siliciclastic sediment during calibration of the silicate weathering component of GEOCLIM.
 
@@
 These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of the SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$526--678~ppm.
-Increasing seawater Mg/Ca since ca.
-15~Ma \cite{Higgins2012a} is consistent with an increasing proportion of the global silicate weathering flux being derived from mafic and ultramafic sources.
-We note that these ``paleo-SEAI'' scenarios do not account for Neogene changes outside of the SEAI (e.g.
+These ``paleo-SEAI'' scenarios do not account for Neogene changes outside of the SEAI (e.g.
 changes in ocean/atmosphere circulation, volcanic degassing, and weathering fluxes elsewhere on Earth, discussed in \textit{Alternative Mechanisms for Neogene Cooling}).
-Therefore, these results do not imply that \pCOtwo would have been observed to be 526--678~ppm at 15~Ma, but rather that emergence of the SEAI alone is responsible for decreasing \pCOtwo by 240--392~ppm (526--678~ppm minus 286~ppm) over the past 15~m.y.
+Therefore, these results are not estimating \pCOtwo at 15~Ma, but rather are quantifying \pCOtwo change associated with emergence of the SEAI alone.
 
@@
 Previous work has estimated that the decrease in \pCOtwo since 5~Ma associated with the emergence of the SEAI and enhanced silicate weathering is $\sim$19~ppm \cite{Molnar2015a}, in which case their emergence would be a relatively minor contributor to Neogene cooling.
-This 19~ppm estimate was obtained using a 0-D equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
+This 19~ppm estimate was obtained using an equation that assumes a direct linear relationship between mean global temperature and changes in weathering-rate-weighted land area, scaled by a factor that is intended to account for the influence of both runoff and temperature.
 \pCOtwo was then estimated from the calculated temperature using a simple energy balance equation.
@@
 Instead, spatially varying topography and climatology significantly modulates silicate weathering rates, and must be accounted for when estimating \pCOtwo change associated with paleogeographic change.
+
+An important caveat for these estimates of \pCOtwo is that our modeling is determining the climatology in the GFDL CM2.0 model at which steady-state is achieved -- a climatology that has an associated \pCOtwo value in the model.
+However, climate models are variable in their response to changes in \pCOtwo.
+One way to summarize this variability is through the equilibrium climate sensitivity value -- the steady-state change in global mean surface air temperature associated with a doubling of \pCOtwo.
+A range of 1.5 to 4.5\degreesC per \pCOtwo doubling was proposed in the landmark Charney report \cite{Charney1979a} and this range was considered to be the credible interval (\textgreater66\% likelihood) in the last IPCC report \cite{Stocker2013a}.
+Integrating constraints both from understanding of climate feedback processes and the climate record, a recent comprehensive review estimates the 66\% probability range of climate sensitivity to be 2.6 to 3.9\degreesC per \pCOtwo doubling with a 5 to 95\% range of 2.3 to 4.7\degreesC per \pCOtwo doubling \cite{Sherwood2020a}.
+The equilibrium climate sensitivity associated with the CM2.0 climate models is 2.9\degreesC per \pCOtwo doubling which falls within these ranges although these ranges remain broad.
+An alternative way to consider the results from our analysis would be that an estimate of 572~ppm (2\times pre-industrial \pCOtwo) for the 15~Ma paleo-SEAI scenario is implying that Earth would be $\sim$2.9\degreesC warmer.
+If Earth's climate sensitivity is at the higher end of the probable range and higher than in the CM2.0 model, as it is in some climate models, this same amount of Neogene cooling resulting from the emergence of the SEAI could have been driven by a smaller change in \pCOtwo.
 
@@
 If we use the parameter combination that had the highest $r^{2}$ between computed and measured \COtwo consumption in watersheds around the world during calibration (\textit{Calibration} and \SI; Fig.
-S4), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm.
+S4), GEOCLIM estimates a pre-industrial volcanic degassing flux of 4.1$\times$10$^{12}$~mol/yr to balance the silicate weathering flux at 286~ppm \pCOtwo.
 If we then assume that this volcanic degassing flux did not change over the past 15~m.y., then GEOCLIM estimates that the increase in global weatherability associated with the emergence of the SEAI led to a change in \pCOtwo of $\sim$280~ppm (``increase in weatherability only'' scenario in Fig.
@@
     \centering
-    \includegraphics[width=1\linewidth]{Manuscript/Figures/weatherability_curves.pdf}
+    \includegraphics[width=1\linewidth]{Figures/weatherability_curves.pdf}
     \caption{Weatherability curves for the modern and ``paleo-SEAI'' scenarios shown in Figure \ref{fig:scenario_pCO2}.
@@
 This decrease in slope has been attributed to coincident exhumation of relatively low \SrSr Outer Lesser Himalaya carbonates \cite{Myrow2015a, Colleps2018a}, but could also be at least partially driven by the emergence of low \SrSr lithologies in the SEAI during arc-continent collision.
+Increasing seawater Mg/Ca since ca.
+15~Ma \cite{Higgins2012a} is consistent with an increasing proportion of the global silicate weathering flux being derived from mafic and ultramafic sources.
 
@@
 
-Alkalinity generated by the weathering of silicate Na and K consumes \COtwo until that alkalinity is removed via reverse weathering reactions \cite{Sun2016a}.
-Therefore, sustained imbalances in the Na and K cycles could consume/generate significant quantities of \COtwo over multi-million year time-scales.
-However, the biogeochemical cycles of Na and K are currently poorly constrained.
-As such, the residence time of these two cations, and therefore the time-scale over which an imbalance in their cycles could drive \COtwo consumption/generation, is not well known.
-Estimates for these values in the literature range from $\sim$45-80~m.y.
-for Na, and $\sim$8-10~m.y.
-for K \cite{Emerson2008a, Lecuyer2016a}.
-Furthermore, it is not clear how to parameterize the output fluxes of Na and K.
-We have therefore made the assumption in GEOCLIM that alkalinity generated by the weathering of silicate Na and K is instantaneously removed by reverse weathering reactions, but acknowledge that further work is necessary to evaluate how important imbalances in the Na and K cycles are for the geologic carbon cycle.
-
 The long-term carbon-cycle mass balance can be perturbed via mechanisms that are disconnected from changes in volcanic degassing and silicate weathering rates.
@@
 Similarly, the weathering of sedimentary organic matter could serve as a source of \COtwo \cite{Hilton2014a}.
-On the other hand, enhanced burial of organic matter could be an important sink of \COtwo, as has been suggested in the Bengal Fan \cite{Galy2007a} and Taiwan \cite{Kao2014a}.
-An increased flux of dissolved nutrients derived from silicate weathering could even strengthen the biological carbon pump and sequester \COtwo as organic carbon in deep marine sediments \cite{Hartmann2013a}.
+On the other hand, enhanced burial of organic matter enabled by higher sediment and nutrient delivery could be an important sink of \COtwo, as has been suggested in the Bengal Fan \cite{Galy2007a} and Taiwan \cite{Kao2014a}.
 The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been affected by emergence of the SEAI and/or Himalayan orogeny.

### 2020-08-12 Manuscript/Supplementary_Information.tex
fix figure paths

--- 2020-07-31 Manuscript/Supplementary_Information.tex
@@
 We first group these 16 categories into 6 broader categories (metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment; Table \ref{tab:GLiM_to_GEOCLIM}).
+Note that the siliciclastic sediment lithologic category also includes sedimentary sequences in which any carbonate is identified but is not the dominant lithology \cite{Hartmann2012a}.
 We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the greatest area in that pixel (i.e.
@@
 For metamorphic and siliciclastic lithologies, we explore a range of feasible Ca+Mg concentrations during calibration of the silicate weathering component of GEOCLIM (\textit{GEOCLIM Calibration}).
+This approach makes the simplifying assumption that all pixels of a given lithologic category share the same Ca+Mg concentration.
 
@@
 
+A related issue is that our analysis focuses on the weathering of Ca and Mg silicates.
+Sources of alkalinity to the ocean include the four major cations of Ca, Mg, Na and K.
+A common approach for the long-term carbon cycle follows that of \citet{Berner1984a} where an emphasis is placed on Ca and Mg with the weathering of Na and K silicates considered to not be a long-term sink of \COtwo.
+The reasoning behind this assumption is that while weathering of these silicates can lead to transient uptake of \COtwo, this uptake is balanced by release.
+The ultimate mechanism through which silicate weathering sequesters carbon is the incorporation of Ca+Mg carbonates into the lithosphere.
+While alkalinity generated by the weathering of silicate Na and K consumes \COtwo, that alkalinity is removed through processes such as reverse weathering reactions and exchange for Ca that ultimately precipitates as carbonate \cite{Sun2016a}.
+An alternative approach for estimating carbon sequestration associated with silicate weathering that sought to incorporate all four cations was taken by \citet{France-Lanord1997a} who used a formulation of $\Delta CO_2 = \Delta Mg + \Delta Ca + 0.15 \Delta Na + 0.1 \Delta K$.
+This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^2+$ and Mg$^2+$ eventually forms carbonates and applies an estimate that 20\% of K$^+$ and 30\% of Na$^+$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance).
+Overall, uncertainty in the sinks for K and Na and the processes that control them make them difficult to parameterize and model.
+Given that we are modeling steady-state solutions and the ultimate sequestration of carbon through carbonate formation, we make the simplification of restricting the modeling of carbon sequestration to that associated with Ca and Mg delivery.
+Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the overall results would be similar.
+
 \section*{Climate Model}
@@
     \centering
-    \includegraphics[width=1\textwidth]{Manuscript/Figures/lithology_areas.pdf}
+    \includegraphics[width=1\textwidth]{Figures/lithology_areas.pdf}
     \caption{Total area of the lithologic categories within the SEAI for each tested scenario.
@@
 The \pCOtwo proxy compilation shown in Figure \ref{fig:pCO2_proxies} consists of several distinct techniques and samples, each with their own set of assumptions and associated probability distributions that influence their reported values.
-Therefore, simply taking the mean of these values neglects the fact that some of these techniques/samples are likely more reliable than the others.
+Therefore, running a regression through these values neglects the fact that some of these techniques/samples are likely more reliable than the others.
 Constraining how robust each \pCOtwo proxy technique is remains an important challenge, and as such the ``true'' \pCOtwo could plausibly lie anywhere within the full range of proxy-based \pCOtwo estimates for any given time interval.

### 2020-08-12 Manuscript/Supplementary_Information.tex
revisions checkpoint

@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/lithology_composition.pdf}
+    \caption{Ca, Mg, Na, and K concentrations within mafic, intermediate, and felsic lithologies, as calculated based on the mean of data compiled from EarthChem.}
+    \label{fig:lithology_composition}
+\end{figure}
+
 A related issue is that our analysis focuses on the weathering of Ca and Mg silicates.
@@
 Given that we are modeling steady-state solutions and the ultimate sequestration of carbon through carbonate formation, we make the simplification of restricting the modeling of carbon sequestration to that associated with Ca and Mg delivery.
-Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the overall results would be similar.
+Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but given that mafic lithologies have significantly higher Ca+Mg concentrations than intermediate/felsic lithologies relative to the slightly higher Na+K concentration of intermediate/felsic lithologies compared to mafic lithologies (Fig.
+\ref{fig:lithology_composition}), the importance of the emergence of the SEAI on Neogene \pCOtwo would remain unchanged.
 

### 2020-08-13 Manuscript/Supplementary_Information.tex
revision checkpoint

--- 2020-08-12 Manuscript/Supplementary_Information.tex
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/lithology_composition.pdf}
+    \caption{Ca, Mg, Na, and K concentrations within mafic, intermediate, and felsic lithologies, as calculated based on the mean of data compiled from EarthChem.}
+    \label{fig:lithology_composition}
+\end{figure}
+
 A related issue is that our analysis focuses on the weathering of Ca and Mg silicates.
 Sources of alkalinity to the ocean include the four major cations of Ca, Mg, Na and K.
-A common approach for the long-term carbon cycle follows that of \citet{Berner1984a} where an emphasis is placed on Ca and Mg with the weathering of Na and K silicates considered to not be a long-term sink of \COtwo.
+A common approach for the long-term carbon cycle follows that of \citet{Berner1983a} where an emphasis is placed on Ca and Mg with the weathering of Na and K silicates considered to not be a long-term sink of \COtwo.
 The reasoning behind this assumption is that while weathering of these silicates can lead to transient uptake of \COtwo, this uptake is balanced by release.
 The ultimate mechanism through which silicate weathering sequesters carbon is the incorporation of Ca+Mg carbonates into the lithosphere.
-While alkalinity generated by the weathering of silicate Na and K consumes \COtwo, that alkalinity is removed through processes such as reverse weathering reactions and exchange for Ca that ultimately precipitates as carbonate \cite{Sun2016a}.
-An alternative approach for estimating carbon sequestration associated with silicate weathering that sought to incorporate all four cations was taken by \citet{France-Lanord1997a} who used a formulation of $\Delta CO_2 = \Delta Mg + \Delta Ca + 0.15 \Delta Na + 0.1 \Delta K$.
-This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^2+$ and Mg$^2+$ eventually forms carbonates and applies an estimate that 20\% of K$^+$ and 30\% of Na$^+$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance).
-Overall, uncertainty in the sinks for K and Na and the processes that control them make them difficult to parameterize and model.
-Given that we are modeling steady-state solutions and the ultimate sequestration of carbon through carbonate formation, we make the simplification of restricting the modeling of carbon sequestration to that associated with Ca and Mg delivery.
-Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the overall results would be similar.
+Given that Na and K are not precipitating carbonate, when simulating long term steady-state solutions, as we are here, it is valid to follow the approach of \citet{Berner1983a} and not consider the Na-K-\COtwo cycles.
+Complexity arises given the long residence time of Na and K in seawater and the lag between the generation of alkalinity and its consumption that could modulate the carbonate system.
+Dynamic modeling of the effects of the Na and K cycles on the long term carbon cycle is an area ripe for future research although uncertainty in their sinks and the timescale of the processes that control them make them difficult to parameterize.
+An alternative approach for estimating carbon sequestration associated with silicate weathering that sought to incorporate all four cations was taken by \citet{France-Lanord1997a} who used a formulation of $\Delta \textrm{CO}_2 = \Delta \textrm{Mg} + \Delta \textrm{Ca} + 0.15 \Delta \textrm{Na} + 0.1 \Delta \textrm{K}$.
+This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^{2+}$ and Mg$^{2+}$ eventually forms carbonates and applies an estimate that 20\% of K$^{+}$ and 30\% of Na$^{+}$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance).
+Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the importance of the emergence of the SEAI on Neogene \pCOtwo would remain unchanged.
+The relative importance of mafic lithologies also remains given that with significantly higher Ca+Mg concentrations and subequal Na+K concentrations to intermediate/felsic lithologies they have the highest Ca+Mg+Na+K concentrations (Fig.
+\ref{fig:lithology_composition}).
 

### 2020-08-13 Manuscript/Manuscript.tex
revisions checkpoint

@@
 We also include changes in areas of presently submerged continental shelves such as the Sunda Shelf that were previously exposed (\SI; Fig.
-S6).
+S7).
 These tectonic drivers and others throughout the region led to progressive emergence over the past 15~m.y.
@@
 Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (\SI; Fig.
-S9).
+S10).
 We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig.
@@
 Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation (\SI; Fig.
-S10).
+S11).
 The \pCOtwo values from the 5~Ma ``paleo-SEAI'' scenario overlap with many proxy-based estimates \cite{Foster2017a} as well as values that emerge from approaches that assimilate climate and ice sheet model output with benthic \dO data \cite{van-de-Wal2011a, Berends2020a}.
@@
 Integrating constraints both from understanding of climate feedback processes and the climate record, a recent comprehensive review estimates the 66\% probability range of climate sensitivity to be 2.6 to 3.9\degreesC per \pCOtwo doubling with a 5 to 95\% range of 2.3 to 4.7\degreesC per \pCOtwo doubling \cite{Sherwood2020a}.
-The equilibrium climate sensitivity associated with the CM2.0 climate models is 2.9\degreesC per \pCOtwo doubling which falls within these ranges although these ranges remain broad.
-An alternative way to consider the results from our analysis would be that an estimate of 572~ppm (2\times pre-industrial \pCOtwo) for the 15~Ma paleo-SEAI scenario is implying that Earth would be $\sim$2.9\degreesC warmer.
+The equilibrium climate sensitivity associated with the CM2.0 climate models is 2.9\degreesC per \pCOtwo doubling, which falls within these ranges although these ranges remain broad.
+An alternative way to consider the results from our analysis would be that an estimate of 572~ppm (2$\times$ pre-industrial \pCOtwo) for the 15~Ma ``paleo-SEAI'' scenario is implying that Earth would be $\sim$2.9\degreesC warmer.
 If Earth's climate sensitivity is at the higher end of the probable range and higher than in the CM2.0 model, as it is in some climate models, this same amount of Neogene cooling resulting from the emergence of the SEAI could have been driven by a smaller change in \pCOtwo.
@@
 For example, some proxy-based approaches as well as some model-data assimilation approaches estimate that mid-Miocene \pCOtwo was lower than 568~ppm (\SI; Fig.
-S10).
+S11).
 Take a scenario in which \pCOtwo was 400~ppm at 15~Ma.
@@
 More robust constraints on \pCOtwo (\SI; Fig.
-S10) and/or volcanic degassing rates over the past 20~m.y.
+S11) and/or volcanic degassing rates over the past 20~m.y.
 are needed to constrain which of the ``increase in weatherability only'' or ``increase in weatherability and degassing'' scenarios (Fig.

### 2020-08-13 Manuscript/Supplementary_Information.tex
revisions checkpoint

@@
 The ultimate mechanism through which silicate weathering sequesters carbon is the incorporation of Ca+Mg carbonates into the lithosphere.
-Given that Na and K are not precipitating carbonate, when simulating long term steady-state solutions, as we are here, it is valid to follow the approach of \citet{Berner1983a} and not consider the Na-K-\COtwo cycles.
-Complexity arises given the long residence time of Na and K in seawater and the lag between the generation of alkalinity and its consumption that could modulate the carbonate system.
+Given that Na and K are not precipitating carbonate, when simulating long term steady-state solutions, as we are here, it is valid to follow the approach of \citet{Berner1983a} and not consider the Na and K cycles.
+Complexity arises given the long residence time of Na and K in seawater ($\sim$45--80~m.y.
+for Na, and $\sim$8--10~m.y.
+for K; \citealp{Emerson2008a, Lecuyer2016a}) and the lag between the generation of alkalinity and its consumption that could modulate the carbonate system.
 Dynamic modeling of the effects of the Na and K cycles on the long term carbon cycle is an area ripe for future research although uncertainty in their sinks and the timescale of the processes that control them make them difficult to parameterize.

### 2020-08-14 Manuscript/Manuscript.tex
revision checkpoint

@@
 The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, erosion, and regolith thickness are poorly constrained.
-Rather than prescribing single values, we test several reasonable values for each of these four parameters along with the Ca+Mg concentration of metamorphic and siliciclastic lithologies (\SI; Table S2).
-We then permute all possible combinations of these reasonable values for the six parameters, leading to 93,600 unique parameter combinations (i.e.
+Rather than prescribing single values, we select multiple values for each of these four parameters along with the Ca+Mg concentration of metamorphic and siliciclastic lithologies from within reasonable ranges (\SI; Table S2).
+We then permute all possible combinations of these values for the six parameters, leading to 93,600 unique parameter combinations (i.e.
 permutations).
@@
 changes in ocean/atmosphere circulation, volcanic degassing, and weathering fluxes elsewhere on Earth, discussed in \textit{Alternative Mechanisms for Neogene Cooling}).
-Therefore, these results are not estimating \pCOtwo at 15~Ma, but rather are quantifying \pCOtwo change associated with emergence of the SEAI alone.
+Therefore, these results are not estimating \pCOtwo at 15~Ma, but rather are quantifying \pCOtwo change associated with emergence of the SEAI.
 
@@
 The traditional explanation for this trend is that it reflects increased weathering of radiogenic (i.e.
-high \SrSr) silicate rocks \cite{Raymo1988a}.
-Associated with this explanation is the proposal that increasing weathering of radiogenic silicate rocks in the Himalayas was the primary driver of Neogene cooling \cite{Raymo1992a, Edmond1992a}.
+high \SrSr) silicate rocks \cite{Raymo1988a, Edmond1992a}.
+Associated with this explanation is the proposal that increasing weathering of radiogenic silicate rocks in the Himalayas was the primary driver of Neogene cooling \cite{Raymo1992a}.
 It could be argued that increasing marine \SrSr is inconsistent with the hypothesis that increasing weathering of juvenile (i.e.
@@
 However, the globally averaged ratio of silicate weathering fluxes from radiogenic cratonic rocks versus juvenile arc lithologies can be at least partially decoupled from marine \SrSr via the regional weathering of isotopically unique lithologies.
-For example, unusually radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
+For example, in addition to highly radiogenic granites and gneisses \cite{Edmond1992a}, unusually radiogenic carbonates are abundant in Himalayan strata, and it is estimated that $\sim$75\% of Sr coming from the Himalayas can be attributed to carbonate rather than silicate weathering \cite{Jacobson2002a, Quade2003a, Oliver2003a}.
 As such, there are challenges in interpreting the marine \SrSr record as a direct proxy for silicate weathering fluxes.
@@
 The fluxes of \COtwo represented by these processes are not accounted for in our model framework, and could have been affected by emergence of the SEAI and/or Himalayan orogeny.
-\pCOtwo changes that result from these processes would be superimposed on \pCOtwo changes associated with evolving silicate weathering rates.
-However, our coupled weathering-climate model indicates that the \pCOtwo change associated with evolving silicate weathering rates driven by emergence of the SEAI is sufficient to explain the majority of Neogene cooling (Fig.
+\pCOtwo changes that result from these processes would be superimposed on \pCOtwo changes associated with evolving silicate weathering fluxes.
+However, our coupled weathering-climate model indicates that the \pCOtwo change associated with increased global weatherability driven by emergence of the SEAI is sufficient to explain the majority of Neogene cooling (Fig.
 \ref{fig:scenario_pCO2}).

### 2020-09-09 Manuscript/Supplementary_Information.tex
Updates from Overleaf
pluralize SEAI in SI

@@
 \fancyhf{}
-\lhead{Park et al., in preparation}
+\lhead{Park et al.}
 \rhead{\thepage}
@@
     \caption{Histograms showing the difference in the area-weighted mean Ca+Mg concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel that results from starting with an 0.1\degrees $\times$ 0.1\degrees GLiM raster versus an 0.05\degrees $\times$ 0.05\degrees GLiM raster.
-The left panel is calculated over all land pixels, whereas the right panel is calculated over land pixels in the SEAI only.
+The left panel is calculated over all land pixels, whereas the right panel is calculated over land pixels in the SEAIs only.
 Both panels use the parameter values that produced the highest $r^{2}$ during calibration: metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$ and sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
@@
 Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 573 parameters combinations.
-Watersheds within the compilation of \citet{Gaillardet1999a} that are in the SEAI (Fly, Kikori, Purari, and Sepik watersheds, all in New Guinea) are indicated in green.
+Watersheds within the compilation of \citet{Gaillardet1999a} that are in the SEAIs (Fly, Kikori, Purari, and Sepik watersheds, all in New Guinea) are indicated in green.
 \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$.
@@
 This relationship, also applied by \citet{Schopka2011a}, assumes that all Ca$^{2+}$ and Mg$^{2+}$ eventually forms carbonates and applies an estimate that 20\% of K$^{+}$ and 30\% of Na$^{+}$ exchanges for Ca and Mg and ultimately produces carbonate (and considers charge balance).
-Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the importance of the emergence of the SEAI on Neogene \pCOtwo would remain unchanged.
+Different parameterizations such as that of \citet{France-Lanord1997a} could modulate our estimates, but the importance of the emergence of the SEAIs on Neogene \pCOtwo would remain unchanged.
 The relative importance of mafic lithologies also remains given that with significantly higher Ca+Mg concentrations and subequal Na+K concentrations to intermediate/felsic lithologies they have the highest Ca+Mg+Na+K concentrations (Fig.
@@
 
-Given that precipitation in the SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region.
+Given that precipitation in the SEAIs is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAIs scenarios is a reasonable approximation for this region.
 However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.
@@
     \includegraphics[width=1\textwidth]{Figures/SEAI_fracs.jpg}
-    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAI) used to force GEOCLIM.
+    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAIs) used to force GEOCLIM.
 Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead.
@@
 Solid black lines show present-day shorelines.
-Total areas of each lithologic category within the SEAI for each of these tested scenarios is shown in Figure \ref{fig:lithology_areas}.}
+Total areas of each lithologic category within the SEAIs for each of these tested scenarios is shown in Figure \ref{fig:lithology_areas}.}
     \label{fig:SEAI_fracs}
@@
     \includegraphics[width=1\textwidth]{Figures/lithology_areas.pdf}
-    \caption{Total area of the lithologic categories within the SEAI for each tested scenario.
+    \caption{Total area of the lithologic categories within the SEAIs for each tested scenario.
 Note that the grey bars in the middle panel indicate the fractional area of each lithologic category for the entire Earth.}
@@
     \includegraphics[width=0.6\textwidth]{Figures/paleogeographic_reconstructions.pdf}
-    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
+    \caption{Paleogeographically-reconstructed paleoshorelines for the SEAIs at 5, 10, and 15~Ma, using the paleoshoreline reconstruction of this study coupled to the paleogeographic model of \citet{Matthews2016a}.}
     \label{fig:paleogeographic_reconstructions}
@@
 
-We use geological data to quantify the changes in the area of SEAI over the past 15~m.y.
+We use geological data to quantify the changes in the area of the SEAIs over the past 15~m.y.
 (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}).
-To generate the lithologic map used in the ``removed SEAI'' scenario (Fig.
-\ref{fig:SEAI_fracs}), we remove all land associated with arc-continent collision in SEAI.
-To generate the lithologic map used in the ``paleo-SEAI'' scenarios (Fig.
-\ref{fig:SEAI_fracs}), we take the estimated paleoshorelines of SEAI for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
-The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of SEAI for their respective time slice.
+To generate the lithologic map used in the ``removed SEAIs'' scenario (Fig.
+\ref{fig:SEAI_fracs}), we remove all land associated with arc-continent collision in the SEAIs.
+To generate the lithologic map used in the ``paleo-SEAIs'' scenarios (Fig.
+\ref{fig:SEAI_fracs}), we take the estimated paleoshorelines of the SEAIs for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds.
+The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of the SEAIs for their respective time slice.
 This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same.
@@
 For the slope and lithology, we take the mean slope and Ca+Mg concentrations from neighbouring pixels.
-We also remove the Greenland ice sheet in the ``paleo-SEAI'' scenarios, to be consistent with Northern Hemisphere ice sheets developing after 5~Ma \citep{Haug2005a}.
+We also remove the Greenland ice sheet in the ``paleo-SEAIs'' scenarios, to be consistent with Northern Hemisphere ice sheets developing after 5~Ma \citep{Haug2005a}.
 The runoff field over Greenland is generated using the same method as that for the Sunda Shelf.
@@
 
-To explore the sensitivity of our results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet, we also tested ``paleo-SEAI - Greenland'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAI'' scenarios without the removal of the Greenland ice sheet.
-We also tested ``paleo-SEAI - Greenland - Sunda Shelf'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAI'' scenarios without the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.
-We find that the estimated steady state \pCOtwo's for the ``paleo-SEAI - Greenland'' and ``paleo-SEAI - Greenland - Sunda Shelf'' scenarios are only marginally higher than those of the ``paleo-SEAI'' scenarios (Fig.
+To explore the sensitivity of our results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet, we also tested ``paleo-SEAIs - Greenland'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAIs'' scenarios without the removal of the Greenland ice sheet.
+We also tested ``paleo-SEAIs - Greenland - Sunda Shelf'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``paleo-SEAIs'' scenarios without the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.
+We find that the estimated steady state \pCOtwo's for the ``paleo-SEAIs - Greenland'' and ``paleo-SEAIs - Greenland - Sunda Shelf'' scenarios are only marginally higher than those of the ``paleo-SEAIs'' scenarios (Fig.
 \ref{fig:scenario_pCO2_all}), suggesting that our results are relatively insensitive to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.
@@
 
-Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in SEAI.
+Geological maps, stratigraphic data, and previous paleoshoreline compilations were used to calculate the changes in different types of subaerially exposed rocks in the SEAIs.
 Following \citet{Molnar2015a}, we analyzed the area changes of islands that are larger than $\sim$200~km$^{2}$ (Fig.
@@
 
-% If the weathering of juvenile lithologies of the SEAI played an important role in driving Neogene cooling, the intuitive expectation is that marine \SrSr should have decreased over the Neogene as the global weathering flux of low \SrSr juvenile lithologies increased (compensated by a decrease in the global weathering flux of high \SrSr radiogenic lithologies to maintain constant global \COtwo consumption).
+% If the weathering of juvenile lithologies of the SEAIs played an important role in driving Neogene cooling, the intuitive expectation is that marine \SrSr should have decreased over the Neogene as the global weathering flux of low \SrSr juvenile lithologies increased (compensated by a decrease in the global weathering flux of high \SrSr radiogenic lithologies to maintain constant global \COtwo consumption).
 Contrary to this expectation, marine \SrSr has been rising since ca.
@@
 15~Ma has been attributed to exhumation of relatively low \SrSr and high \OsOs Outer Lesser Himalaya lithologies \citep{Myrow2015a, Colleps2018a}.
-This change in slope could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAI during arc-continent collision.
+This change in slope could also be at least partially driven by the exhumation of mafic and organic-rich forearc sediments in the SEAIs during arc-continent collision.
 

### 2020-09-10 Manuscript/Supplementary_Information.tex
update links to GitHub and Zenodo

@@
 The code for the GEOCLIM model used in this study can be found at: \url{https://github.com/piermafrost/GEOCLIM-dynsoil-steady-state/releases/tag/v1.0}.
-The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/GEOCLIM_Modern}.
+The code that generated the inputs and analyzed the output of the GEOCLIM model can be found at: \url{https://github.com/Swanson-Hysell-Group/2020_Southeast_Asian_Islands} or \url{https://doi.org/10.5281/zenodo.4021653}.
 
